### 2012-11-06 doc/power/power_manual.tex
Add power manual
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1285 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\documentclass[letterpaper,twoside,10pt]{article}

\usepackage[letterpaper, margin=1in]{geometry}

\usepackage[USenglish]{babel} %francais, polish, spanish, ...
\usepackage[T1]{fontenc}
\usepackage[ansinew]{inputenc}
\usepackage{lmodern} %Type1-font for non-english texts and characters

\usepackage{graphicx} %%For loading graphic files
\usepackage{amsmath, xfrac, array, mdwlist, ulem}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{fancyvrb}
\usepackage[svgnames]{xcolor}
%\usepackage{natbib}
\usepackage[style=verbose, backend=bibtex]{biblatex}

\usepackage{hyperref}

\setlength{\parskip}{7pt}%
\setlength{\parindent}{0pt}%

\definecolor{LightGray}{gray}{0.9}

\bibliography{power_manual}

\makeatletter
\newif\ifFV@bgcolor
\newbox\FV@bgbox
\define@key{FV}{bgcolor}{\FV@bgcolortrue\def\FV@bgcolor{#1}}
\define@key{FV}{framecolor}{\FV@bgcolortrue\def\FV@framecolor{#1}}
\def\FV@framecolor{white}

\def\FV@BeginVBox{%
  \leavevmode\ifFV@bgcolor\setbox\FV@bgbox=\fi
  \hbox\ifx\FV@boxwidth\relax\else to\FV@boxwidth\fi\bgroup
  \ifcase\FV@baseline\vbox\or\vtop\or$\vcenter\fi\bgroup}
\def\FV@EndVBox{\egroup\ifmmode$\fi\hfil\egroup
  \ifFV@bgcolor\fcolorbox{\FV@framecolor}{\FV@bgcolor}{\box\FV@bgbox}\fi}
\makeatother

%\def\algorithmautorefname{Algorithm}

\begin{document}

\renewcommand*{\chapterautorefname}{Chapter}
\renewcommand*{\sectionautorefname}{Section}
\renewcommand*{\subsectionautorefname}{Section}
\renewcommand*{\subsubsectionautorefname}{Section}

\title{Power Estimation for VTR}
\author{Jeffrey Goeders}
%\date{} %%If commented, the current date is used.
\maketitle

\setcounter{tocdepth}{2}
\tableofcontents %Table of contents

\newpage
\section{Overview}
	TBD (\autoref{fig:flow}).
\begin{figure}[ht]
	\centering
	\includegraphics[scale=0.8]{images/flow.pdf}
	\caption{Power Estimation in the VTR Flow}
	\label{fig:flow}
\end{figure}

\newpage
\section{Running VTR with Power Estimation}
\subsection{VTR Flow}
\subsection{VTR Tasks}
\subsection{VPR 6.0}
The following command-line options are added to the VPR executable to facilitate power estimation (all are required): \newline
\texttt{\--\--power}:  Use this option to enable power estimation. \newline
\texttt{\--\--activity\_file <activities.act>}: The activity file, produce by ACE 2.0, or another tool. \newline
\texttt{\--\--tech\_properties <tech\_properties.xml>}: The technology properties XML file.

\newpage
\section{Supporting Tools}
\subsection{Technology Properties Generation}
Power estimation requires information detailing the properties of the CMOS technology.  
This information, which includes transistor capacitances, leakage currents, etc. is included in an \texttt{.xml} file, and provided as a parameter to VPR.
This XML file is generated using a script which automatically runs HSPICE, performs multiple circuit simulations, and extract the necessary values.

Some of these technology XML files are included with the release, and are located here:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth]
<vtr>/vtr_flow/tech/*
\end{BVerbatim}

If the user wishes to use a different CMOS technology file, they must run the following script.  HSPICE must be included in the \$PATH\$.

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<vtr>/vtr_flow/scripts/generate_cmos_tech_data.pl <tech_file> <tech_size> <Vdd> <temp>
\end{BVerbatim}

\texttt{<tech\_file>}: A SPICE technology file, containing a \texttt{pmos} and \texttt{nmos} models. \newline
\texttt{<tech\_size>}: The technology size, in meters. For example, a 90nm technology would have the value \texttt{90e-9}. \newline
\texttt{<Vdd>}: Supply voltage in Volts. \newline
\texttt{<temp>}: Operating temperature, in Celcius.

\subsection{ACE 2.0 Activity Estimation}
Power estimation requires activity information for the entire netlist.  This ativity information consists of two values:
\vspace{-10pt}
\begin{enumerate}
	\item  \underline{The Signal Probability}, $P_1$, is the long-term probability that a signal is logic-high.  For example, a clock signal with a 50\% duty cycle will have $P_1(clk) = 0.5$.
	\item \underline{The Transition Density} (or switching activity), $A_S$, is the average number of times the signal will switch during each clock cycle.  For example, a clock has $A_S(clk)=2$.
\end{enumerate}

The default tool used to perform activity estimation in VTR is ACE 2.0 \footnote{\cite{ace2}}.  
This tool was originally designed to work with the Berkeley SIS tool, which is now obsolte.  ACE 2.0 was modifed to use ABC, and is included in the VTR package here:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<vtr>/ace2
\end{BVerbatim}

The tool can be run using the following command-line arguments:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<vtr>/ace2/ace -b <abc.blif> -o <activities.act> -n <new.blif>
\end{BVerbatim}

\texttt{<abc.blif>}: The input BLIF file produced by ABC. \newline
\texttt{<activities.act>}: The activity file to be created. \newline
\texttt{<new.blif>}: The new BLIF file.  This will be identical in function to the ABC blif; however, since ABC does not maintain internal node names, a new BLIF must be produced with node names that match the activity file. \newline

User's may with to use their own activity estimation tool.  The produced activity file must contain one line for each net in the BLIF file, in the following format:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<net name> <signal probability> <transistion density>
\end{BVerbatim}

\newpage
\section{Architecture Modelling}
The following section describes the architectural assumptions made by the power model, and the related parameters in the architecture file.

\subsection{Complex Blocks} \label{sec:clbs}
The VTR architecture description language supports a hierarchichal description of blocks.  In the architecture file, each block is described as a \texttt{pb\_type}, which may includes one or more children of type \texttt{pb\_type}, and interconnect structures to connect them.  The power estimation algorithm traverses this hierarchy recursively, and performs power estimation for each \texttt{pb\_type}.  The power model supports multiple power estimation methods, and the user specifies the desired method in the architecture file: \newline
\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<pb_type>
	<power method="<est-method>">
</pb_type>
\end{BVerbatim}

The following is a list of valid estimation methods. Detailed descriptions of each type are provided in the following sections.  The methods are listed in order from most accurate to least accurate. 
\begin{enumerate}
\item \texttt{specify-size}: Detailed transistor level modelleling. The user supplies all buffer sizes and wire-lengths.  Any not provided by the user are ignored. 
\item \texttt{auto-size}: Detailed transistor level modelleling. The user can supply buffer sizes and wire-lengths; however, they will be automatically inserted when not provided. 
\item \texttt{pin-toggle}: Higher-level modelling.  The user specifies energy per toggle of the pins.  Static power provided as an absolute.
\item \texttt{C-internal}: Higher-level modelling.  The user supplies the internal capacitance of the block.  Static power provided as an absolute.
\item \texttt{absolute}: Highest-level modelling.  The user supplies both dynamic and static power as absolutes.
\end{enumerate}

If no estimation method is provided, it is inherited from the parent \texttt{pb\_type}. \textit{\textbf{If the top-level \texttt{pb\_type} has no estimation method, \texttt{auto-size} is assumed.}}

\subsubsection{\texttt{specify-size}}
This estimation method provides a detailed transistor level modelling of CLBs, and will provide the most accurate power estimations.  For each \texttt{pb\_type}, power estimation accounts for the following components (See \autoref{fig:sample_clb}).
\begin{itemize*}
	\item Interconnect multiplexers
	\item Buffers and wire capacitances
	\item Child \texttt{pb\_types}
\end{itemize*}

\begin{figure}[ht]
	\centering
		\includegraphics[scale=0.5]{images/sample_clb.pdf}
	\caption{Sample Block}
	\label{fig:sample_clb}
\end{figure}

\textbf{Multiplexers} Interconnect multiplexers are modelled as 2-level pass-transistor multiplexers, comprised of minimum-size NMOS transistors. Their size is determined automatically from the \texttt{<interconnect/>} structures in the architecture description file.

\textbf{Buffers and Wires} Buffers and wire capacitances are not defined in the architecture file, and must be explicitly added by the user.  They are assigned on a per port basis using the following construct:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<pb_type>
	<input name="my_input" num_pins="1">
		<power ...options.../>
	</input>
</pb_type>
\end{BVerbatim}

The wire and buffer attributes can be set using the following options.  If no options are set, it is assumed that the wire capacitance is zero, and there are no buffers present.  Keep in mind that the port construct allows for multiple pins per port.  These attributes will be applied to each pin in the port.  If necessary, the user can seperate a port into multiple ports with different wire/buffer properties.
\begin{itemize}
	\item \texttt{wire\_capacitance="1.0e-15"}: The absolute capacitance of the wire, in Farads.
	\item \texttt{wire\_length="1.0e-7"}: The absolute length of the wire, in meters.  The local interconnect capacitance option must be specified, as described in \autoref{sec:local_interc_cap}.
	\item \texttt{wire\_length="auto"}: The wirelength is automatically sized.  See \autoref{sec:local_wire_autosize}.
	\item \texttt{buffer\_size="2.0"}: The size of the buffer at this pin.  See \autoref{sec:buffer_sizing} for more information.
	\item \texttt{buffer\_size="auto"}: The size of the buffer is automatically sized, assuming it drives the above wire capacitance and a single multiplexer.  
	See \autoref{sec:buffer_sizing} for more information.
\end{itemize}

\textbf{Primitives}
For all child \texttt{pb\_types}, the algorithm performs a recursive call. Eventually \texttt{pb\_types} will be reached that have no children.  
These are primitives, such as flip-flops, LUTs, or other hard-blocks.  The power model includes functions to perform transistor-level power estimation for flip-flops and LUTs. 
If the user wishes to use a design with other primitive types (memories, multipliers, etc), they must provide an equivalent function.  If the user makes such a function, the \texttt{power\_calc\_primitive} function should be modified to call it.  Alternatively, these blocks can be configured to use higher-level power estimation methods.

\textbf{NOTE:} Default Leakage Mode TBD

\subsubsection{\texttt{auto-size}}
This estimation method also performs detailed transistor-level modelling.  It is almost identical to the \texttt{specify-size} method described above.  The only difference is that the local wire capacitance and buffers are automatically inserted for all pins, when necessary.  This is equivalent to using the \texttt{specify-size} method with the \texttt{wire\_length="auto"} and 
\texttt{buffer\_size="auto"} options for every port.

\textbf{This is the default power estimation method.}  Although not as accurate as user-provided buffer and wire sizes, it is capable of automatically capturing trends in power dissipation as architectures are modified.

\subsubsection{\texttt{pin-toggle}}
This method allows users to specify the dynamic power of a block in terms of the energy per toggle of each input, output or clock pin for the \texttt{pb\_type}.  The static power is provided as an absolute (in Watts). This is done using the following construct:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<pb_type>
	<input name="my_input" num_pins="1">
		<power energy_per_toggle="1.0e-16">
	</input>
	...
	<power method="pin-toggle">
		<static_power power_per_instance="1.0e-6"/>
	</power>
</pb_type>
\end{BVerbatim}

This method does not perform any transistor-level estimations; the entire power estimation is performed using the above values.  \texttt{energy\_per\_toggle} specifies the energy consumed (in Joules) for a single toggle of a pin.  Keep in mind that the port construct allows for multiple pins per port.  These energy per toggle will be applied to each pin in the port.  If necessary, the user can seperate a port into multiple ports with different energy per toggle values.

\subsubsection{\texttt{C-internal}}
This method allows the users to specify the dynamic power of a block in terms of the internal capacitance of the block.  The activity will be averaged across all of the input pins, and will be supplied with the internal capacitance to the standard equation $P_{dyn}=\sfrac{1}{2}\alpha CV^2$.  Again, the static power is provided as an absolute (in Watts).  This is done using the following construct:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<pb_type>
	<power method="c-internal">
		<dynamic_power C_internal="1.0e-16"/>
		<static_power power_per_instance="1.0e-6"/>
	</power>
</pb_type>
\end{BVerbatim}

\subsubsection{\texttt{absolute}}
This method is the most basic power estimation method, and allows users to specify both the dynamic and static power of a block as absolute values (in Watts).  This is done using the following construct:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<pb_type>
	<power method="absolute">
		<dynamic_power power_per_instance="1.0e-6"/>
		<static_power power_per_instance="1.0e-6"/>
	</power>
</pb_type>
\end{BVerbatim}

\subsection{Global Routing}	
Global routing consists of switch boxes and input connection boxes.

\subsubsection{Switch Boxes} \label{sec:sb}
Switch boxes are modelled as the following components (\autoref{fig:sb}):
\vspace{-10pt}
\begin{enumerate*}
	\item Multiplexer 
	\item Buffer
	\item Wire capacitance
\end{enumerate*}

\begin{figure}[ht]
	\centering
		\includegraphics[scale=0.7]{images/sb.pdf}
	\caption{Switch Box}
	\label{fig:sb}
\end{figure}

\textbf{Multiplexer} 
	The multiplexer is modelled as 2-level pass-transistor multiplexer, comprised of minimum-size NMOS transistors.  
	The number of inputs to the multiplexer is automatically determined.

\textbf{Buffer} 
The buffer is a multistage CMOS buffer.  
Typically, this buffer is sized to drive the capacitance of the wire, as well as fanouts to other switch boxes and conneciton boxes.  The size of the buffer is specified by the user in the architecture file, as shown below.

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<switchlist>
	<switch type="mux" ... power_buf_size="16.0"
</switchlist>
\end{BVerbatim}

Alternatively, the size can be replaced by the "\texttt{auto}" keyword, and the buffer will be automatically sized based on the load capacitance.  See \autoref{sec:buffer_sizing} for information about buffer sizings.

\textbf{Wire Capacitance} The wire capacitance is determined using the segment information in the architecture file:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<segmentlist>
	<segment type="unidir" ... power_options .../>
</segmentlist>
\end{BVerbatim}

The following options can be used to specify wire capacitance:
\vspace{-10pt}
\begin{itemize*}
\item \texttt{power\_wire\_c}: Absolute wire capacitance, in farads.
\item \texttt{power\_wire\_c\_per\_m}: Wire capacitance in farads per meter.  The capacitance is calculated using an automatically determined wirelength, based on the area of a tile in the FPGA (\autoref{sec:area_est}).
\end{itemize*}

\subsection{Input Connection Boxes} \label{sec:cb}
Input connection boxes are modelled as the following components (\autoref{fig:cb}):
\vspace{-10pt}
\begin{itemize}
\item One buffer per routing track, sized to drive the load of all input multiplexers to which the buffer is connected (For buffer sizing see \autoref{sec:buffer_sizing}).
\item One multiplexer per block input pin, sized according to the number of routing tracks that connect to the pin.
\end{itemize}

\begin{figure}[ht]
	\centering
		\includegraphics[scale=1.6]{images/cb.pdf}
	\caption{Conneciton Box}
	\label{fig:cb}
\end{figure}

\subsection{Clock Network}
The clock network modelled is a four quadrant spine and rib design, as illustrated in \autoref{fig:clock_network}. At this time, the power model only supports a single clock.
The model assumes that the entire spine and rib clock network will contain buffers separated in distance by the length of a grid tile.  
The buffer sizes and wire capacitances are specified in the architecture file using the following construct:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<clocks>
	<clock ... clock_options .../>
</clocks>
\end{BVerbatim}

The following clock options are supported:
\vspace{-10pt}
\begin{itemize}
\item \texttt{C\_wire="1e-16"}: The absolute capacitance, in fards, of the wire between each clock buffer.
\item \texttt{C\_wire\_per\_m="1e-12"}: The wire capacitance, in fards per m.  The capacitance is calculated using an automatically determined wirelength, based on the area of a tile in the FPGA.
\item \texttt{buffer\_size="2.0"}: The size of each clock buffer.  This can be replaced with the "auto" keyword.  See \autoref{sec:buffer_sizing} for more information on buffer sizing.
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[scale=0.7]{images/clock_network}
	\caption{The clock network.  Squares represent CLBs, and the wires represent the clock network.}
	\label{fig:clock_network}
\end{figure}

\clearpage
\section{Other Architecture Options \& Techniques}
\subsection{Area Estimations} \label{sec:area_est}
The power estimator performs area estimations of both the global routing structures and the complex blocks.  The CLB area is determined through a recursive function that counts the area of interconnect, buffers, and primitives (flip-flops and LUTs).

If an architecture uses new primitives in CLBs, it should include a function that returns the transistor count.  This function should be called from within \texttt{power\_count\_transistors\_primitive()}.

\subsection{Local Wire Auto-Sizing} \label{sec:local_wire_autosize}
The power estimator contains a basic algorithm to estimate local interconnect wirelengths.  The algorithms determines the area of the interconnect structures, and calculates the side-length of the interconnect based on a square layout.  This side length is referred to as $L_{interc}$, as illustrated in \autoref{fig:local_interc_wirelength}.

\begin{figure}[ht]
	\centering
		\includegraphics[scale=0.7]{images/local_interc_wirelength.pdf}
	\caption{Local interconnect wirelength}
	\label{fig:local_interc_wirelength}
\end{figure}

It is assumed that each wire connecting a pin of a \texttt{pb\_type} to an interconnect structure is of length $0.5 \cdot L_{interc}$.  In reality, this length depends on the actual transistor
layout, and may be much larger or much smaller than the estimated value.  If desired, the user can override the 0.5 constant in the architecture file:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<architecture>
	<power>
		<local_interconnect factor="0.5">
	</power>
</architecture>
\end{BVerbatim}

\subsection{Buffer Sizing} \label{sec:buffer_sizing}
In the power estimator, a buffer size refers to the size of the final stage of multi-stage buffer (if small, only a single stage is used).  The specified size is the (\sfrac{W}{L}) of the NMOS transistor.  The PMOS transistor will automatically be sized larger.    
Generally, buffers are sized depending on the load capacitance, using the following equation (logical effort model):

\begin{align}
	\text{Buffer Size}=\frac{1}{f_{LE}}*\frac{C_{Load}}{C_{INV}}
	\label{eq:autosize}
\end{align}	
In this equation, $C_{INV}$ is the input capacitance of a minimum-sized inverter, and $f_{LE}$ is the logical effort factor, which by default is 4.  This is standard design practise for high-performance devices.  The logical effort factor can be modified in the architecture file:

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<architecture>
	<power>
		<buffers logical_effort_factor="4">
	</power>
</architecture>
\end{BVerbatim}

The number of buffer stages is chosen such that each stage is close to $f_{LE}$ times larger than the previous stage.

\subsection{Local Interconnect Capacitance} \label{sec:local_interc_cap}
If using the \texttt{auto-size} or \texttt{wire-length} options (\autoref{sec:clbs}), the local interconnect capacitance must be specified.  This is specified in the units of Farads/meter.

\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
<architecture>
	<power>
		<local_interconnect C_wire="1.0e-15">
	</power>
</architecture>
\end{BVerbatim}

\subsection{SRAM}
TBD

\newpage
\section{Support}
For support, please check \url{http://code.google.com/p/vtr-verilog-to-routing/wiki/Power}, or email \href{mailto:vtr.power.estimation@gmail.com}{vtr.power.estimation@gmail.com}.

\end{document}

### 2012-11-26 ODIN_II/SRC/adders.c
Cleaning up the code for the new change of subtractor. Fixed the parameter issue: Cannot show the width properly if use the parameter in there.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1303 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@ -49,10 +49,11 @@ int *adder = NULL;
  *-------------------------------------------------------------------------*/
 void init_add_distribution()
 {
-	int i;
+	int i, j;
 	oassert(hard_adders != NULL);
-	adder = (int *)malloc(sizeof(int) * (hard_adders->inputs->size + 1));
-	for (i = 0; i <= (hard_adders->inputs->size + hard_adders->inputs->next->size); i++)
+	j = hard_adders->inputs->size + hard_adders->inputs->next->size;
+	adder = (int *)malloc(sizeof(int) * (j + 1));
+	for (i = 0; i <= j; i++)
 		adder[i] = 0;
 	return;
 }
@@ -198,15 +199,13 @@ void instantiate_hard_adder(nnode_t *node, short mark, netlist_t *netlist)
 	{
 		if (node->output_pins[i]->name == NULL)
 		{
-			len = strlen(node->name) + 6; /* 6 chars for pin idx */
+			len = strlen(node->name) + 20; /* 6 chars for pin idx */
 			new_name = (char*)malloc(len);
 			sprintf(new_name, "%s[%d]", node->name, node->output_pins[i]->pin_node_idx);
 			node->output_pins[i]->name = new_name;
 		}
 	}
 
-	free(node->name);
-	node->name = new_name;
 	node->traverse_visited = mark;
 	return;
 }
@@ -618,12 +617,12 @@ void split_adder(nnode_t *nodeo, int a, int b, int sizea, int sizeb, int cin, in
 	oassert(nodeo->input_port_sizes[0] == a);
 	oassert(nodeo->input_port_sizes[1] == b);
 
-	node  = (nnode_t**)malloc(sizeof(nnode_t*)*(count + 1));
+	node  = (nnode_t**)malloc(sizeof(nnode_t*)*(count));
 
 	for(i = 0; i < count; i++)
 	{
 		node[i] = allocate_nnode();
-		node[i]->name = (char *)malloc(strlen(nodeo->name) + 4);
+		node[i]->name = (char *)malloc(strlen(nodeo->name) + 20);
 		sprintf(node[i]->name, "%s-%d", nodeo->name, i);
 		init_split_adder(nodeo, node[i], a, sizea, b, sizeb, cin, cout, i);
 		add_list = insert_in_vptr_list(add_list, node[i]);

### 2012-11-26 ODIN_II/SRC/ast_util.c

@@ -355,16 +355,18 @@ void add_child_to_node(ast_node_t* node, ast_node_t *child)
 
 /*---------------------------------------------------------------------------------------------
  * (function: get_range)
+ *  Check the node range is legal. Will return the range if it's legal.
+ *  Node should have three children. Second and Third children's type should be NUMBERS.
  *-------------------------------------------------------------------------------------------*/
 int get_range(ast_node_t* first_node) 
 {
 	long temp_value;
 	/* look at the first item to see if it has a range */
-	if (first_node->children[1] != NULL && first_node->children[1]->type == NUMBERS)
+	if (first_node->children[1] != NULL && first_node->children[1]->type == NUMBERS && first_node->children[2] != NULL && first_node->children[2]->type == NUMBERS)
 	{
 		/* IF the first element in the list has a second element...that is the range */
-		oassert(first_node->children[2] != NULL); // the third element should be a value
-		oassert((first_node->children[1]->type == NUMBERS) && (first_node->children[2]->type == NUMBERS)); // should be numbers
+		//oassert(first_node->children[2] != NULL); // the third element should be a value
+		//oassert((first_node->children[1]->type == NUMBERS) && (first_node->children[2]->type == NUMBERS)); // should be numbers
 		if(first_node->children[1]->types.number.value < first_node->children[2]->types.number.value)
 		{
 			// Reversing the indicies doesn't produce correct code. We need to actually handle these correctly.

### 2012-11-26 ODIN_II/SRC/odin_ii_func.c

@@ -113,7 +113,7 @@ void do_high_level_synthesis()
 	#ifdef VPR6
 	find_hard_multipliers();
 	find_hard_adders();
-	find_hard_adders_for_sub();
+	//find_hard_adders_for_sub();
 	register_hard_blocks();
 	#endif
 	global_param_table_sc = sc_new_string_cache();

### 2012-11-26 ODIN_II/SRC/output_blif.c

@@ -215,7 +215,6 @@ void output_blif(char *file_name, netlist_t *netlist)
 #ifdef VPR6
 	add_the_blackbox_for_mults(out);
 	add_the_blackbox_for_adds(out);
-	//add_the_blackbox_for_subs(out);
 	output_hard_blocks(out);
 #endif
 
@@ -358,10 +357,10 @@ void output_node(nnode_t *node, short traverse_number, FILE *fp)
 			break;
 
 		case MINUS:
-			if (hard_subs == NULL)
+			if (hard_adders == NULL)
 				oassert(FALSE); /* should be soft logic! */
 			#ifdef VPR6
-			if(hard_subs != NULL)
+			if(hard_adders != NULL)
 				define_add_function(node, node->type, fp);
 			#endif
 			break;

### 2012-11-26 ODIN_II/SRC/parse_making_ast.c

@@ -402,18 +402,76 @@ ast_node_t *markAndProcessSymbolListWith(ids id, ast_node_t *symbol_list)
 {
 	int i;
 	long sc_spot;
+	long range_temp_max = 0;
+	long range_temp_min = 0;
 	ast_node_t *range_min = 0;
 	ast_node_t *range_max = 0;
+	ast_node_t *newNode = 0;
 
 	for (i = 0; i < symbol_list->num_children; i++)
 	{
-		/* checks range is legal */
+		/* checks range is legal.  */
 		get_range(symbol_list->children[i]);
 
-		if ((i == 0) && (symbol_list->children[0]->children[1] != NULL)&& (symbol_list->children[0]->children[2] != NULL) && (symbol_list->children[0]->children[1]->type == NUMBERS) && (symbol_list->children[0]->children[2]->type == NUMBERS))
+		if ((i == 0) && (symbol_list->children[0]->children[1] != NULL) && (symbol_list->children[0]->children[2] != NULL)
+				&& ((symbol_list->children[0]->children[1]->type == NUMBERS) || (symbol_list->children[0]->children[1]->type == IDENTIFIERS))
+				&& ((symbol_list->children[0]->children[2]->type == NUMBERS) || (symbol_list->children[0]->children[2]->type == IDENTIFIERS)))
 		{
-			range_max = symbol_list->children[0]->children[1];
-			range_min = symbol_list->children[0]->children[2];
+			/* Do lookup in sc_add_string */
+			/* Verify node->type.variables.is_parameter == TRUE */
+			/* ELSE REPORT ERROR */
+			if (symbol_list->children[0]->children[1]->type == IDENTIFIERS)
+			{
+				if ((sc_spot = sc_lookup_string(defines_for_module_sc[num_modules], symbol_list->children[0]->children[1]->types.identifier)) != -1)
+				{
+					newNode = defines_for_module_sc[num_modules]->data[sc_spot];
+					if (newNode->types.variable.is_parameter == TRUE)
+					{
+						range_max = symbol_list->children[0]->children[1];
+						range_temp_max = newNode->types.number.value;
+					}
+					else
+						error_message(PARSE_ERROR, symbol_list->children[0]->children[1]->line_number, current_parse_file,
+								"parameter %s don't match\n", symbol_list->children[0]->children[1]->types.identifier);
+				}
+				else
+					error_message(PARSE_ERROR, symbol_list->children[0]->children[1]->line_number, current_parse_file,
[…]

### 2012-11-26 ODIN_II/SRC/partial_map.c

@@ -201,12 +201,12 @@ void partial_map_node(nnode_t *node, short traverse_number, netlist_t *netlist)
 			break;
 		case MINUS:
 			#ifdef VPR6
-			if (hard_subs != NULL && node->num_input_port_sizes == 3)
+			if (hard_adders != NULL && node->num_input_port_sizes == 3)
 			{
 				if ((node->input_port_sizes[0] + node->input_port_sizes[1]) > 0)
 					instantiate_hard_adder_subtraction(node, traverse_number, netlist, 0);
 			}
-			else if(hard_subs != NULL && node->num_input_port_sizes == 2)
+			else if(hard_adders != NULL && node->num_input_port_sizes == 2)
 			{
 				if ((node->input_port_sizes[0]) > 0)
 					instantiate_hard_adder_subtraction(node, traverse_number, netlist, 1);

### 2012-11-26 ODIN_II/SRC/read_blif.c

@@ -326,7 +326,10 @@ void create_latch_node_and_driver(FILE *file, hashtable_t *output_nets_hash)
 	char buffer[READ_BLIF_BUFFER];
 	while ((ptr = my_strtok (NULL, TOKENS, file, buffer)) != NULL)
 	{
-		names = (char**)realloc(names, sizeof(char*)* (input_token_count + 1));
+		if(input_token_count == 0)
+			names = (char**)malloc((sizeof(char*)));
+		else
+			names = (char**)realloc(names, (sizeof(char*))* (input_token_count + 1));
 		names[input_token_count++] = strdup(ptr);
 	}
 
@@ -402,6 +405,7 @@ void create_latch_node_and_driver(FILE *file, hashtable_t *output_nets_hash)
 
 	/* Free the char** names */
 	free(names);
+	free(ptr);
 }
 
 /*---------------------------------------------------------------------------------------------
@@ -764,6 +768,8 @@ short read_bit_map_find_unknown_gate(int input_count, nnode_t *node, FILE *file)
 {
 	fpos_t pos;
 	int last_line = file_line_number;
+	char *One = "1";
+	char *Zero = "0";
 	fgetpos(file,&pos);
 
 	if(!input_count)
@@ -782,7 +788,7 @@ short read_bit_map_find_unknown_gate(int input_count, nnode_t *node, FILE *file)
 	}
 
 	char **bit_map = NULL;
-	char *output_bit_map = 0;// to distinguish whether for the bit_map output is 1 or 0
+	char *output_bit_map = NULL;// to distinguish whether for the bit_map output is 1 or 0
 	int line_count_bitmap = 0; //stores the number of lines in a particular bit map
 	char buffer[READ_BLIF_BUFFER];
 	while(1)
@@ -793,8 +799,19 @@ short read_bit_map_find_unknown_gate(int input_count, nnode_t *node, FILE *file)
 
 		bit_map = (char**)realloc(bit_map,sizeof(char*) * (line_count_bitmap + 1));
 		bit_map[line_count_bitmap++] = strdup(my_strtok(buffer,TOKENS, file, buffer));
-
+		if (output_bit_map != NULL) free(output_bit_map);
 		output_bit_map = strdup(my_strtok(NULL,TOKENS, file, buffer));
+	}
+
+	if (!strcmp(output_bit_map, One))
+	{
+		free(output_bit_map);
+		output_bit_map = One;
+	}
+	else
+	{
[…]

### 2012-11-26 ODIN_II/SRC/subtractions.c

@@ -35,7 +35,7 @@ OTHER DEALINGS IN THE SOFTWARE.
 #include "globals.h"
 #include "errors.h"
 
-t_model *hard_subs = NULL;
+//t_model *hard_adders = NULL;
 struct s_linked_vptr *subs_list = NULL;
 struct s_linked_vptr *sub_list = NULL;
 struct s_linked_vptr *sub_one_node_list = NULL;
@@ -47,20 +47,6 @@ int *sub = NULL;
 #ifdef VPR6
 
 /*---------------------------------------------------------------------------
- * (function: init_add_distribution)
- *  For adder, the output will only be the maxim input size + 1
- *-------------------------------------------------------------------------*/
-void init_sub_distribution()
-{
-	int i;
-	oassert(hard_subs != NULL);
-	sub = (int *)malloc(sizeof(int) * (hard_subs->inputs->size + 1));
-	for (i = 0; i <= (hard_subs->inputs->size + hard_subs->inputs->next->size); i++)
-		sub[i] = 0;
-	return;
-}
-
-/*---------------------------------------------------------------------------
  * (function: report_add_distribution)
  *-------------------------------------------------------------------------*/
 
@@ -70,7 +56,7 @@ void report_sub_distribution()
 	int totalsub = 0;
 	chain_information_t *sub_chain;
 
-	if(hard_subs == NULL)
+	if(hard_adders == NULL)
 		return;
 
 	printf("\nHard MINUS Distribution\n");
@@ -100,29 +86,6 @@ void report_sub_distribution()
 }
 
 /*---------------------------------------------------------------------------
- * (function: find_hard_adders_for_sub)
- *-------------------------------------------------------------------------*/
-void find_hard_adders_for_sub()
-{
-
-	hard_subs = Arch.models;
-
-	while (hard_subs != NULL)
-	{
-		if (strcmp(hard_subs->name, "adder") == 0)
-		{
-			init_sub_distribution();
-			return;
-		}
-		else
-		{
-			hard_subs = hard_subs->next;
-		}
-	}
-	return;
-}
-
-/*---------------------------------------------------------------------------
  * (function: declare_hard_adder_for_sub)
[…]

### 2012-11-26 ODIN_II/SRC/subtractions.h

@@ -27,8 +27,7 @@ OTHER DEALINGS IN THE SOFTWARE.
 #include "util.h"
 #include "adders.h"
 
-extern t_model *hard_subs;
-extern t_model *hard_subs_unary;
+//extern t_model *hard_subs;
 extern struct s_linked_vptr *subs_list;
 extern struct s_linked_vptr *sub_list;
 extern struct s_linked_vptr *sub_one_node_list;
@@ -40,11 +39,6 @@ extern void init_sub_distribution();
 extern void report_sub_distribution();
 extern void declare_hard_adder_for_sub(nnode_t *node);
 extern void instantiate_hard_adder_subtraction(nnode_t *node, short mark, netlist_t *netlist, int type);
-extern void find_hard_adders_for_sub();
-extern void add_the_blackbox_for_subs(FILE *out);
-extern void add_the_blackbox_for_unary_subs(FILE *out);
-extern void define_sub_function(nnode_t *node, short type, FILE *out);
-extern void define_unary_sub_function(nnode_t *node, short type, FILE *out);
 extern void split_adder_for_sub(nnode_t *node, int a, int b, int sizea, int sizeb, int cin, int cout, int count, netlist_t *netlist);
 extern void iterate_adders_for_sub(netlist_t *netlist);
 extern void clean_adders_for_sub();

### 2012-11-26 ODIN_II/SRC/verilog_preprocessor.c

@@ -152,14 +152,14 @@ int add_veri_define(char *symbol, char *value, int line, veri_include *defined_i
 #endif
 			else if (0 != strcmp(def_iterator->value, value))
 			{
-				fprintf(stderr, "\tWarning: The value of %s has been redefined to %s, the prvious value was %s\n\n",
+				fprintf(stderr, "\tWarning: The value of %s has been redefined to %s, the previous value was %s\n\n",
 					symbol, value, def_iterator->value);
 				free(def_iterator->value);
 				def_iterator->value = (char *)strdup(value);
 			}
 
 			free(new_def);
-			return 0;
+			return -2;
 		}
 	}
 	
@@ -456,7 +456,7 @@ void veri_preproc_bootstraped(FILE *original_source, FILE *preproc_producer, ver
 		
 		//fprintf(stderr, "%s:%d\t%s\n", current_include->path,line_number, line);
 
-		/* Preprocessor directives have a backtick on the first coloumn. */
+		/* Preprocessor directives have a backtick on the first column. */
 		if (line[0] == '`') 
 		{
 			token = trim((char *)strtok(line, " \t"));
@@ -602,7 +602,7 @@ void veri_preproc_bootstraped(FILE *original_source, FILE *preproc_producer, ver
 			{
 				pop(skip);
 			}
-			/* Leave unhandled preprcessor directives in place. */
+			/* Leave unhandled preprocessor directives in place. */
 			else if (top(skip) < 1)
 			{
 				fprintf(preproc_producer, "%s %s\n", line, line + 1 + strlen(line));

### 2012-11-26 libarchfpga/include/physical_types.h
Power: Added callibration
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1305 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@ -155,6 +155,7 @@ struct s_clock_network {
 
 	float prob; /* Static probability of net assigned to this clock */
 	float dens; /* Switching density of net assigned to this clock */
+	float period; /* Period of clock */
 };
 
 /* Power-related architecture information */

### 2012-11-26 libarchfpga/read_xml_arch_file.c

@@ -1819,6 +1819,14 @@ static void alloc_and_load_default_child_for_pb_type( INOUTP t_pb_type *pb_type,
 	copy->num_input_pins = pb_type->num_input_pins;
 	copy->num_output_pins = pb_type->num_output_pins;
 	copy->num_pb = 1;
+
+	/* Power */
+	copy->pb_type_power = (t_pb_type_power*) my_calloc(1,
+			sizeof(t_pb_type_power));
+	copy->pb_type_power->estimation_method = power_method_inherited(
+			pb_type->pb_type_power->estimation_method);
+
+	/* Ports */
 	copy->num_ports = pb_type->num_ports;
 	copy->ports = (t_port*) my_calloc(pb_type->num_ports, sizeof(t_port));
 	for (i = 0; i < pb_type->num_ports; i++) {
@@ -1829,8 +1837,17 @@ static void alloc_and_load_default_child_for_pb_type( INOUTP t_pb_type *pb_type,
 		copy->ports[i].parent_pb_type = copy;
 		copy->ports[i].name = my_strdup(pb_type->ports[i].name);
 		copy->ports[i].port_class = my_strdup(pb_type->ports[i].port_class);
+
 		copy->ports[i].port_power = (t_port_power*) my_calloc(1,
 				sizeof(t_port_power));
+		//Defaults
+		if (copy->pb_type_power->estimation_method == POWER_METHOD_AUTO_SIZES) {
+			copy->ports[i].port_power->wire_type = POWER_WIRE_TYPE_AUTO;
+			copy->ports[i].port_power->buffer_type = POWER_BUFFER_TYPE_AUTO;
+		} else if (copy->pb_type_power->estimation_method == POWER_METHOD_SPECIFY_SIZES) {
+			copy->ports[i].port_power->wire_type = POWER_WIRE_TYPE_IGNORED;
+			copy->ports[i].port_power->buffer_type = POWER_BUFFER_TYPE_NONE;
+		}
 	}
 
 	copy->max_internal_delay = pb_type->max_internal_delay;
@@ -1870,10 +1887,7 @@ static void alloc_and_load_default_child_for_pb_type( INOUTP t_pb_type *pb_type,
 					pb_type->annotations[i].value[j]);
 		}
 	}
-	copy->pb_type_power = (t_pb_type_power*) my_calloc(1,
-			sizeof(t_pb_type_power));
-	copy->pb_type_power->estimation_method = power_method_inherited(
[…]

### 2012-11-26 vpr/SRC/power/PowerSpicedComponent.c

@@ -1,0 +1,117 @@
+/************************* INCLUDES *********************************/
+#include <string.h>
+#include <assert.h>
+#include <cfloat>
+
+#include "util.h"
+#include "PowerSpicedComponent.h"
+
+PowerSpicedComponent::PowerSpicedComponent(float (*fn)(float size)) {
+	component_usage = fn;
+
+	/* Always pad with a high and low entry */
+	num_entries = 2;
+	size_factors = (t_power_callib_pair*) my_malloc(
+			num_entries * sizeof(t_power_callib_pair));
+	size_factors[0].size = 0;
+	size_factors[1].size = FLT_MAX;
+
+	done_callibration = false;
+	sorted = true;
+}
+
+void PowerSpicedComponent::add_entry(float size, float power) {
+	if (done_callibration) {
+		assert(0);
+	}
+
+	num_entries++;
+
+	size_factors = (t_power_callib_pair*) my_realloc(size_factors,
+			(num_entries) * sizeof(t_power_callib_pair));
+	assert(size_factors);
+
+	// Thre is always an extra entry at the end for max-value placeholder
+	size_factors[num_entries - 2].size = size;
+	size_factors[num_entries - 2].power = power;
+
+	sorted = false;
+}
+
+int search_callib_pair(const void * k, const void * e) {
+	const t_power_callib_pair * key = (const t_power_callib_pair*) k;
+	const t_power_callib_pair * elem = (const t_power_callib_pair*) e;
+
+	if (key->size < elem->size)
+		return -1;
+	else if (key->size >= elem->size && key->size < (elem + 1)->size)
+		return 0;
+	else
+		return 1;
+}
+
+float PowerSpicedComponent::scale_factor(float size) {
+	t_power_callib_pair * result;
+	t_power_callib_pair key;
+	float perc_upper;
+
+	if (!done_callibration)
+		assert(0);
+
+	if (num_entries <= 2)
+		return 1.0;
+
+	key.size = size;
+	result = (t_power_callib_pair*) bsearch(&key, size_factors, num_entries - 1,
+			sizeof(t_power_callib_pair), search_callib_pair);
+	assert(result);
+
+	perc_upper = (size - result->size) / ((result + 1)->size - result->size);
+
+	return perc_upper * (result + 1)->factor + (1 - perc_upper) * result->factor;
+}
+
+int compare_callib_pair(const void * a, const void * b) {
[…]

### 2012-11-26 vpr/SRC/power/PowerSpicedComponent.h

@@ -1,0 +1,53 @@
+/*********************************************************************
+ *  The following code is part of the power modelling feature of VTR.
+ *
+ * For support:
+ * http://code.google.com/p/vtr-verilog-to-routing/wiki/Power
+ *
+ * or email:
+ * vtr.power.estimation@gmail.com
+ *
+ * If you are using power estimation for your researach please cite:
+ *
+ * Jeffrey Goeders and Steven Wilton.  VersaPower: Power Estimation
+ * for Diverse FPGA Architectures.  In International Conference on
+ * Field Programmable Technology, 2012.
+ *
+ ********************************************************************/
+
+#ifndef __POWER_POWERSPICEDCOMPONENT_H__
+#define __POWER_POWERSPICEDCOMPONENT_H__
+
+/************************* STRUCTS **********************************/
+typedef struct s_power_callib_pair t_power_callib_pair;
+struct s_power_callib_pair {
+	float size;
+	float power;
+	float factor;
+};
+
+class PowerSpicedComponent {
+	int num_entries;
+	t_power_callib_pair * size_factors;
+
+	/* Estimation function for this component */
+	float (*component_usage)(float size);
+
+	bool sorted;
+	bool done_callibration;
+
+	//float slope;
+	//float yint;
+
+	void sort_sizes(void);
+
+public:
+	PowerSpicedComponent(float (*fn)(float size));
+	void add_entry(float size, float power);
+	void update_scale_factor(float (*fn)(float size));
+	float scale_factor(float size);
+	void callibrate(void);
+	bool is_done_callibration(void);
+};
+
+#endif

### 2012-11-26 vpr/SRC/power/power.c

@@ -34,7 +34,7 @@
 #include "power_util.h"
 #include "power_lowlevel.h"
 #include "power_sizing.h"
-#include "power_verify.h"
+#include "power_callibrate.h"
 #include "power_cmos_tech.h"
 
 #include "physical_types.h"
@@ -134,11 +134,12 @@ static void power_calc_primitive(t_power_usage * power_usage, t_pb * pb,
 			SRAM_values = alloc_SRAM_values_from_truth_table(LUT_size,
 					logical_block[pb->logical_block].truth_table);
 		} else {
-			SRAM_values = alloc_SRAM_values_from_truth_table(LUT_size, NULL );
+			SRAM_values = alloc_SRAM_values_from_truth_table(LUT_size, NULL);
 		}
-		power_calc_LUT(&sub_power_usage, LUT_size, SRAM_values,
-				input_probabilities, input_densities);
+		power_usage_lut(&sub_power_usage, LUT_size, SRAM_values,
+				input_probabilities, input_densities, g_solution_inf.T_crit);
 		power_add_usage(power_usage, &sub_power_usage);
+		power_component_add_usage(&sub_power_usage, POWER_COMPONENT_LUT);
 		free(SRAM_values);
 		free(input_probabilities);
 		free(input_densities);
@@ -163,9 +164,10 @@ static void power_calc_primitive(t_power_usage * power_usage, t_pb * pb,
 		clk_prob = g_clock_arch->clock_inf[0].prob;
 		clk_dens = g_clock_arch->clock_inf[0].dens;
 
-		power_calc_FF(&sub_power_usage, D_prob, D_dens, Q_prob, Q_dens,
-				clk_prob, clk_dens);
+		power_usage_ff(&sub_power_usage, D_prob, D_dens, Q_prob, Q_dens,
+				clk_prob, clk_dens, g_solution_inf.T_crit);
 		power_add_usage(power_usage, &sub_power_usage);
+		power_component_add_usage(&sub_power_usage, POWER_COMPONENT_FF);
 
 	} else {
 		char msg[BUFSIZE];
@@ -203,13 +205,14 @@ void power_usage_local_pin_buffer_and_wire(t_power_usage * power_usage,
 
 	/* Wire switching */
 	C_wire = pin->pin_power->C_wire;
-	power_usage_wire(&sub_power_usage, C_wire, pin_dens(pb, pin));
+	power_usage_wire(&sub_power_usage, C_wire, pin_dens(pb, pin),
+			g_solution_inf.T_crit);
[…]

### 2012-11-26 vpr/SRC/power/power.h

@@ -24,6 +24,7 @@
 
 /************************* INCLUDES *********************************/
 #include "vpr_types.h"
+#include "PowerSpicedComponent.h"
 
 /************************* DEFINES ***********************************/
 /* Maximum size of logs */
@@ -208,6 +209,9 @@ struct s_power_commonly_used {
 	float INV_1X_C_in;
 	float INV_1X_C;
 	float INV_2X_C;
+
+	/* Component Callibrations Array [0..POWER_CALLIB_COMPONENT_MAX-1] */
+	PowerSpicedComponent ** component_callibration;
 
 	/* Subthreshold leakages */
 	float NMOS_1X_st_leakage;

### 2012-11-26 vpr/SRC/power/power_callibrate.c

@@ -20,7 +20,7 @@
  */
 
 /************************* INCLUDES *********************************/
-#include "power_verify.h"
+#include "power_callibrate.h"
 #include "power_components.h"
 #include "power_lowlevel.h"
 #include "power_util.h"
@@ -60,7 +60,8 @@ void power_print_spice_comparison(void) {
 	if (0) {
 		fprintf(g_power_output->out, "Energy of INV (High Activity)\n");
 		for (i = 0; i < (sizeof(inv_sizes) / sizeof(float)); i++) {
-			power_calc_inverter(&sub_power_usage, 2, 0.5, inv_sizes[i]);
+			power_calc_inverter(&sub_power_usage, 2, 0.5, inv_sizes[i],
+					power_callib_period);
 			fprintf(g_power_output->out, "%g\t%g\n", inv_sizes[i],
 					(sub_power_usage.dynamic + sub_power_usage.leakage)
 							* g_solution_inf.T_crit);
@@ -68,7 +69,8 @@ void power_print_spice_comparison(void) {
 
 		fprintf(g_power_output->out, "Energy of INV (No Activity)\n");
 		for (i = 0; i < (sizeof(inv_sizes) / sizeof(float)); i++) {
-			power_calc_inverter(&sub_power_usage, 0, 1, inv_sizes[i]);
+			power_calc_inverter(&sub_power_usage, 0, 1, inv_sizes[i],
+					power_callib_period);
 			fprintf(g_power_output->out, "%g\t%g\n", inv_sizes[i],
 					(sub_power_usage.dynamic + sub_power_usage.leakage)
 							* g_solution_inf.T_crit);
@@ -87,8 +89,9 @@ void power_print_spice_comparison(void) {
 			dens[j] = 2;
 			prob[j] = 0.5;
 		}
-		power_calc_mux_multilevel(&mux_power_usage,
-				power_get_mux_arch(mux_sizes[i]), prob, dens, 0, FALSE);
+		power_usage_mux_multilevel(&mux_power_usage,
+				power_get_mux_arch(mux_sizes[i]), prob, dens, 0, FALSE,
+				power_callib_period);
 		fprintf(g_power_output->out, "%d\t%g\n", mux_sizes[i],
 				(mux_power_usage.dynamic + mux_power_usage.leakage)
 						* g_solution_inf.T_crit);
[…]

### 2012-11-26 vpr/SRC/power/power_callibrate.h

@@ -26,8 +26,26 @@
 #include "power.h"
 
 /************************* DEFINES **********************************/
+const float power_callib_period = 5e-9;
+
+/************************* STRUCTS **********************************/
+/************************* ENUMS ************************************/
+typedef enum {
+	POWER_CALLIB_COMPONENT_BUFFER = 0,
+	POWER_CALLIB_COMPONENT_BUFFER_WITH_LEVR,
+	POWER_CALLIB_COMPONENT_MUX,
+	POWER_CALLIB_COMPONENT_LUT,
+	POWER_CALLIB_COMPONENT_FF,
+	POWER_CALLIB_COMPONENT_MAX
+} e_power_callib_component;
 
 /************************* FUNCTION DECLARATIONS ********************/
 void power_print_spice_comparison(void);
+void power_callibrate(void);
+float power_usage_buf_for_callibration(float size);
+float power_usage_buf_levr_for_callibration(float size);
+float power_usage_mux_for_callibration(float size);
+float power_usage_lut_for_callibration(float size);
+float power_usage_ff_for_callibration(float size);
 
 #endif

### 2012-11-26 vpr/SRC/power/power_cmos_tech.c

@@ -32,6 +32,8 @@
 #include "ezxml.h"
 #include "util.h"
 #include "read_xml_util.h"
+#include "PowerSpicedComponent.h"
+#include "power_callibrate.h"
 
 /************************* GLOBALS **********************************/
 static t_transistor_inf * g_transistor_last_searched;
@@ -53,6 +55,7 @@ static int power_compare_buffer_strength(const void * key_void,
 		const void * elem_void);
 static int power_compare_buffer_sc_levr(const void * key_void,
 		const void * elem_void);
+static void power_tech_xml_load_components(ezxml_t parent);
 
 /************************* FUNCTION DEFINITIONS *********************/
 
@@ -128,7 +131,62 @@ void power_tech_load_xml_file(char * cmos_tech_behavior_filepath) {
 	power_tech_xml_load_sc(child);
 	FreeNode(child);
 
+	/* Components */
+	child = FindElement(cur, "components", TRUE);
+	power_tech_xml_load_components(child);
+	FreeNode(child);
+
 	FreeNode(cur);
+}
+
+static void power_tech_xml_load_component(ezxml_t parent,
+		PowerSpicedComponent ** component, char * name,
+		float (*usage_fn)(float size)) {
+	ezxml_t cur;
+	ezxml_t child, prev;
+
+	*component = new PowerSpicedComponent(usage_fn);
+
+	cur = FindElement(parent, name, TRUE);
+	child = FindFirstElement(cur, "instance", TRUE);
+	while (child) {
+		float size = GetFloatProperty(child, "size", TRUE, 0.);
+		float power = GetFloatProperty(child, "power", TRUE, 0.);
+		(*component)->add_entry(size, power);
+
+		prev = child;
+		child = child->next;
+		FreeNode(prev);
+	}
+	FreeNode(cur);
+}
+
+static void power_tech_xml_load_components(ezxml_t parent) {
+
+	g_power_commonly_used->component_callibration =
+			(PowerSpicedComponent**) my_malloc(
+					POWER_CALLIB_COMPONENT_MAX * sizeof(PowerSpicedComponent*));
+
+	power_tech_xml_load_component(parent,
+			&g_power_commonly_used->component_callibration[POWER_CALLIB_COMPONENT_BUFFER],
+			"buf", power_usage_buf_for_callibration);
+
+	power_tech_xml_load_component(parent,
[…]

### 2012-11-26 vpr/SRC/power/power_components.c

@@ -27,6 +27,7 @@
 #include "power_components.h"
 #include "power_lowlevel.h"
 #include "power_util.h"
+#include "power_callibrate.h"
 
 /************************* STRUCTS **********************************/
 typedef struct s_power_breakdown t_power_breakdown;
@@ -47,7 +48,7 @@ static void power_calc_mux_rec(t_power_usage * power_usage, float * out_prob,
 		float * out_dens, float * v_out, t_mux_node * mux_node,
 		t_mux_arch * mux_arch, int * selector_values,
 		float * primary_input_prob, float * primary_input_dens,
-		boolean v_out_restored);
+		boolean v_out_restored, float period);
 
 /************************* FUNCTION DEFINITIONS *********************/
 
@@ -97,7 +98,7 @@ void power_component_get_usage(t_power_usage * power_usage,
  * - component_idx: Type of component
  */
 float power_component_get_usage_sum(e_power_component_type component_idx) {
-	return power_usage_sum(&g_power_breakdown.components[component_idx]);
+	return power_sum_usage(&g_power_breakdown.components[component_idx]);
 }
 
 /**
@@ -130,15 +131,15 @@ static void power_component_print_usage_rec(FILE * fp, char * name,
 
 	fprintf(fp, "%-*s%-20.4g%-15.3g%-15.3g%-15.3g\n",
 			component_str_size - 2 * indent_level, name,
-			power_usage_sum(&g_power_breakdown.components[type]),
-			power_usage_sum(&g_power_breakdown.components[type])
-					/ power_usage_sum(
+			power_sum_usage(&g_power_breakdown.components[type]),
+			power_sum_usage(&g_power_breakdown.components[type])
+					/ power_sum_usage(
 							&g_power_breakdown.components[parent_type]),
-			power_usage_sum(&g_power_breakdown.components[type])
-					/ power_usage_sum(
+			power_sum_usage(&g_power_breakdown.components[type])
+					/ power_sum_usage(
 							&g_power_breakdown.components[POWER_COMPONENT_TOTAL]),
 			g_power_breakdown.components[type].dynamic
-					/ power_usage_sum(&g_power_breakdown.components[type]));
[…]

### 2012-11-26 vpr/SRC/power/power_components.h

@@ -58,9 +58,6 @@ typedef enum {
 	POWER_COMPONENT_LOCAL_BUFFERS_AND_WIRE, /* Local interconnect wires */
 	POWER_COMPONENT_FF, /* Flip-flops */
 	POWER_COMPONENT_LUT, /* LUTs */
-	POWER_COMPONENT_LUT_DRIVER, /* LUT input drivers */
-	POWER_COMPONENT_LUT_MUX, /* All 2-muxs in the LUTs */
-	POWER_COMPONENT_LUT_BUFFERS, /* All level-restoring buffers in LUTs */
 
 	POWER_COMPONENT_MAX_NUM
 } e_power_component_type;
@@ -77,17 +74,18 @@ float power_component_get_usage_sum(e_power_component_type component_idx);
 
 void power_component_print_usage(FILE * fp);
 
-void power_calc_FF(t_power_usage * power_usage, float D_prob, float D_dens,
-		float Q_prob, float Q_dens, float clk_prob, float clk_dens);
-void power_calc_LUT(t_power_usage * power_usage, int LUT_size,
+void power_usage_ff(t_power_usage * power_usage, float D_prob, float D_dens,
+		float Q_prob, float Q_dens, float clk_prob, float clk_dens,
+		float period);
+void power_usage_lut(t_power_usage * power_usage, int LUT_size,
 		char * SRAM_values, float * input_densities,
-		float * input_probabilities);
+		float * input_probabilities, float period);
 void power_calc_local_interc_mux(t_power_usage * power_usage, t_pb * pb,
 		t_interconnect_pins * interc_pins);
-void power_calc_mux_multilevel(t_power_usage * power_usage,
+void power_usage_mux_multilevel(t_power_usage * power_usage,
 		t_mux_arch * mux_arch, float * in_prob, float * in_dens,
-		int selected_input, boolean output_level_restored);
+		int selected_input, boolean output_level_restored, float period);
 void power_usage_buffer(t_power_usage * power_usage, float size, float in_prob,
-		float in_dens, boolean level_restored, int input_mux_size);
+		float in_dens, boolean level_restored, int input_mux_size, float period);
 
 #endif

### 2012-11-26 vpr/SRC/power/power_lowlevel.c

@@ -30,11 +30,10 @@
 
 /************************* FUNCTION DELCARATIONS ********************/
 static float power_calc_node_switching_v(float capacitance, float density,
-		float voltage);
+		float period, float voltage);
 static void power_calc_transistor_capacitance(float *C_d, float *C_s,
 		float *C_g, e_tx_type transistor_type, float size);
-static float power_calc_leakage_st(e_tx_type transistor_type,
-		float size);
+static float power_calc_leakage_st(e_tx_type transistor_type, float size);
 static float power_calc_leakage_st_pass_transistor(float size, float v_ds);
 static float power_calc_leakage_gate(e_tx_type transistor_type, float size);
 static float power_calc_buffer_sc_levr(
@@ -59,10 +58,10 @@ void power_lowlevel_init() {
 	g_power_commonly_used->PMOS_1X_C_g = C_g;
 	g_power_commonly_used->PMOS_1X_C_s = C_s;
 
-	g_power_commonly_used->NMOS_1X_st_leakage = power_calc_leakage_st(
-			NMOS, 1.0);
-	g_power_commonly_used->PMOS_1X_st_leakage = power_calc_leakage_st(
-			PMOS, 1.0 * g_power_tech->PN_ratio);
+	g_power_commonly_used->NMOS_1X_st_leakage = power_calc_leakage_st(NMOS,
+			1.0);
+	g_power_commonly_used->PMOS_1X_st_leakage = power_calc_leakage_st(PMOS,
+			1.0 * g_power_tech->PN_ratio);
 
 	g_power_commonly_used->INV_1X_C_in = g_power_commonly_used->NMOS_1X_C_g
 			+ g_power_commonly_used->PMOS_1X_C_g;
@@ -84,9 +83,10 @@ void power_lowlevel_init() {
  * - capacitance: The capacitance of the nodoe
  * - density: The transition density of the node
  */
-float power_calc_node_switching(float capacitance, float density) {
+float power_calc_node_switching(float capacitance, float density,
+		float period) {
 	return 0.5 * g_power_tech->Vdd * g_power_tech->Vdd * capacitance * density
-			/ g_solution_inf.T_crit;
+			/ period;
 }
 
 /**
@@ -96,9 +96,8 @@ float power_calc_node_switching(float capacitance, float density) {
  * - voltage: The voltage when the node is charged
  */
[…]

### 2012-11-26 vpr/SRC/power/power_lowlevel.h

@@ -31,34 +31,33 @@
 /************************* FUNCTION DECLARATION *********************/
 void power_lowlevel_init();
 
-void power_calc_inverter(t_power_usage * power_usage, float in_density,
-		float in_probability, float size);
+void power_calc_inverter(t_power_usage * power_usage, float in_dens,
+		float in_prob, float size, float period);
 void power_calc_inverter_with_input(t_power_usage * power_usage,
 		float * input_dynamic_power, float in_density, float in_prob,
 		float size);
 void power_calc_inverter_irregular(t_power_usage * power_usage,
 		float * dyn_power_input, float in_density, float in_probability,
-		float PMOS_size, float NMOS_size);
-
+		float PMOS_size, float NMOS_size, float period);
 
 void power_usage_wire(t_power_usage * power_usage, float capacitance,
-		float density);
+		float density, float period);
 
 void power_calc_mux_singlelevel_static(t_power_usage * power_usage,
 		float * out_prob, float * out_dens, float * V_out, int num_inputs,
 		int selected_idx, float * in_prob, float * in_dens, float * v_in,
-		float transistor_size, boolean v_out_restored);
+		float transistor_size, boolean v_out_restored, float period);
 
 void power_calc_MUX2_transmission(t_power_usage * power_usage, float * in_dens,
-		float * in_prob, float sel_dens, float sel_prob, float out_dens);
+		float * in_prob, float sel_dens, float sel_prob, float out_dens, float period);
 
 void power_calc_mux_singlelevel_dynamic(t_power_usage * power_usage,
 		int num_inputs, float out_density, float out_prob, float v_out,
 		float * in_prob, float * in_density, float * v_in, float sel_dens,
-		float sel_prob, float transistor_size);
+		float sel_prob, float transistor_size, float period);
 
 void power_calc_level_restorer(t_power_usage * power_usage,
-		float * dyn_power_in, float in_density, float in_probability);
+		float * dyn_power_in, float in_density, float in_probability, float period);
 
 float power_calc_pb_switching_from_c_internal(t_pb * pb,
 		t_pb_graph_node * pb_graph_node);
@@ -69,5 +68,5 @@ float power_calc_mux_v_out(int num_inputs, float transistor_size, float v_in,
 float power_calc_buffer_sc(int stages, float gain, boolean level_restored,
 		int input_mux_size);
 
-float power_calc_node_switching(float capacitance, float density);
+float power_calc_node_switching(float capacitance, float density, float period);
 #endif

### 2012-11-26 vpr/SRC/power/power_sizing.c

@@ -35,7 +35,8 @@ static double power_count_transistors_connectionbox(void);
 static double power_count_transistors_mux(t_mux_arch * mux_arch);
 static double power_count_transistors_mux_node(t_mux_node * mux_node,
 		float * transistor_sizes);
-static void power_mux_node_max_inputs(t_mux_node * mux_node, float * max_inputs);
+static void power_mux_node_max_inputs(t_mux_node * mux_node,
+		float * max_inputs);
 static double power_count_transistors_interc(t_interconnect * interc);
 static double power_count_transistors_pb_node(t_pb_graph_node * pb_node);
 static double power_count_transistors_switchbox(t_arch * arch);
@@ -149,10 +150,11 @@ static double power_count_transistors_mux(t_mux_arch * mux_arch) {
  *  This function is used recursively to determine the largest single-level
  *  multiplexer at each level of a multi-level multiplexer
  */
-static void power_mux_node_max_inputs(t_mux_node * mux_node, float * max_inputs) {
+static void power_mux_node_max_inputs(t_mux_node * mux_node,
+		float * max_inputs) {
 
 	max_inputs[mux_node->level] =
-	max(max_inputs[mux_node->level], static_cast<float>(mux_node->num_inputs));
+			max(max_inputs[mux_node->level], static_cast<float>(mux_node->num_inputs));
 
 	if (mux_node->level != 0) {
 		int child_idx;

### 2012-11-26 vpr/SRC/power/power_util.c

@@ -48,7 +48,12 @@ void power_add_usage(t_power_usage * dest, t_power_usage * src) {
 	dest->leakage += src->leakage;
 }
 
-float power_usage_sum(t_power_usage * power_usage) {
+void power_scale_usage(t_power_usage * power_usage, float scale_factor) {
+	power_usage->dynamic *= scale_factor;
+	power_usage->leakage *= scale_factor;
+}
+
+float power_sum_usage(t_power_usage * power_usage) {
 	return power_usage->dynamic + power_usage->leakage;
 }
 

### 2012-11-26 vpr/SRC/power/power_util.h

@@ -38,7 +38,8 @@ void pb_foreach_pin(t_pb_graph_node * pb_node,
 /* Power Usage */
 void power_zero_usage(t_power_usage * power_usage);
 void power_add_usage(t_power_usage * dest, t_power_usage * src);
-float power_usage_sum(t_power_usage * power_usage);
+void power_scale_usage(t_power_usage * power_usage, float scale_factor);
+float power_sum_usage(t_power_usage * power_usage);
 
 /* Message Logger */
 void power_log_msg(e_power_log_type log_type, char * msg);

### 2012-11-26 vtr_flow/scripts/arch_gen/arch_gen.py

@@ -550,8 +550,6 @@ def gen_arch(dir, k_LUT, N_BLE, I_CLB, I_BLE, fracture_level, num_FF, seg_length
     xprop("Tdel", switch_T_del)
     xprop("mux_trans_size", switch_mux_trans_size)
     xprop("buf_size", switch_buf_size)
-    if (do_power):
-        xprop("buf_last_stage_size", "auto")
     xcloseend()
     xend() #switchlist
     

### 2012-11-26 vtr_flow/scripts/run_vtr_flow.pl

@@ -434,7 +434,7 @@ if ( $ending_stage >= $stage_idx_vpr and !$error_code ) {
 
 	if ($do_power) {
 		push( @vpr_power_args, "--power" );
-		push( @vpr_power_args, "--cmos_tech_behavior_file" );
+		push( @vpr_power_args, "--tech_properties" );
 		push( @vpr_power_args, "$tech_file" );
 	}
 	if ( $min_chan_width < 0 ) {

### 2012-11-26 vtr_flow/scripts/run_vtr_task.pl
Fix bug when task name contains a '/'
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1306 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@ -86,7 +86,7 @@ while ( $token = shift(@ARGV) ) {
 
 	# must be a task name
 	else {
-		if ( $token =~ /(.*)\// ) {
+		if ( $token =~ /(.*)\/$/ ) {
 			$token = $1;
 		}
 		push( @tasks, $token );

### 2012-11-27 vtr_flow/scripts/parse_vtr_task.pl
Fix bug when task name contains a '/'
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1307 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@ -81,7 +81,7 @@ while ( $token = shift(@ARGV) ) {
 	}
 	# must be a task name
 	else {
-		if ( $token =~ /(.*)\// ) {
+		if ( $token =~ /(.*)\/$/ ) {
 			$token = $1;
 		}
 		push( @tasks, $token );

### 2012-12-10 doc/power/power_manual.tex
Power: Added callibration
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1305 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Update to power manual
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1335 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
+The latest version of this document can be obtained here: \url{http://vtr-verilog-to-routing.googlecode.com/svn/trunk/doc/power/power_manual.pdf}
+
 \setcounter{tocdepth}{2}
@@
 \section{Overview}
-	TBD (\autoref{fig:flow}).
+This document describes the power estimation tool for VTR.  This tool provides transistor-level dynamic and static power estimates for a given architecture and circuit.
+
+\autoref{fig:flow} illustrates how the VTR flow is modified to add power estimation.
+The actual power estimation is performed within the VPR executable; however, additional files must be provided.
+In addition to the circuit and architecture files, power estimation requires files detailing the signal activities and technology properties.
+
+\autoref{sec:running} details how to run power estimation for VTR.
+\autoref{sec:tools} provides details on the supporting tools that are used to generate the signal activities and technology properties files.
+\autoref{sec:arch_modelling} provides details about how the tool models architectures, including different options of modelling methods.
+\autoref{sec:options} provides more advanced configuration options.
+
 \begin{figure}[ht]
@@
 \newpage
-\section{Running VTR with Power Estimation}
+\section{Running VTR with Power Estimation} \label{sec:running}
 \subsection{VTR Flow}
-\subsection{VTR Tasks}
+	The easiest way to run the VTR flow is to use the \texttt{run\_vtr\_flow.pl} script.
+See \url{http://code.google.com/p/vtr-verilog-to-routing/wiki/Run_VTR_Flow} for instructions on how to use the script.
+
+In order to perform power estimation, you must add the following options:
+
+\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth]
+-power 
+-cmos_tech <cmos_technology_properties_file.xml>
+\end{BVerbatim}
+
+The CMOS technology properties file is an XML file that contains relevant process-dependent information needed for power estimation.
+XML files for 22\,nm, 45\,nm, and 130\,nm PTM models can be found here:
+
+\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth]
+<vtr>/vtr_flow/tech/.../*.xml
+\end{BVerbatim}
+
+See \autoref{sec:tpg} for information on how to generate an XML file for your own SPICE technology model.
+
+%\subsection{VTR Tasks}
 \subsection{VPR 6.0}
+The easiest way to perform power estimation is to use the \texttt{run\_vtr\_flow.pl} script.  However, if you wish, you may run VPR directly.
 The following command-line options are added to the VPR executable to facilitate power estimation (all are required): \newline
@@
 
+This requires a activity file, which can be generated as described in \autoref{sec:ace}.
+
 \newpage
-\section{Supporting Tools}
-\subsection{Technology Properties Generation}
+\section{Supporting Tools} \label{sec:tools}
+\subsection{Technology Properties Generation} \label{sec:tpg}
 Power estimation requires information detailing the properties of the CMOS technology.  
@@
 
-\subsection{ACE 2.0 Activity Estimation}
+\subsection{ACE 2.0 Activity Estimation} \label{sec:ace}
 Power estimation requires activity information for the entire netlist.  This ativity information consists of two values:
@@
 \newpage
-\section{Architecture Modelling}
+\section{Architecture Modelling} \label{sec:arch_modelling}
 The following section describes the architectural assumptions made by the power model, and the related parameters in the architecture file.
@@
 
-\textbf{NOTE:} Default Leakage Mode TBD
+%\textbf{NOTE:} Default Leakage Mode TBD
 
@@
 
+It is assumed that the power usage specified here includes power of all child \texttt{pb\_types}.  No further recursive power estimation will be performed.
+
 \subsubsection{\texttt{absolute}}
@@
 \end{BVerbatim}
+
+It is assumed that the power usage specified here includes power of all child \texttt{pb\_types}.  No further recursive power estimation will be performed.
 
@@
 \clearpage
-\section{Other Architecture Options \& Techniques}
+\section{Other Architecture Options \& Techniques} \label{sec:options}
 \subsection{Area Estimations} \label{sec:area_est}
@@
 
-\subsection{SRAM}
-TBD
+%\subsection{SRAM}
+%TBD
 

### 2013-04-19 doc/power/power_manual.tex
Updated vpr manual and power manual.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@1850 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \usepackage[style=verbose, backend=bibtex]{biblatex}
+\usepackage{enumitem}
+\usepackage{booktabs}
 
@@
 \end{enumerate}
+Other methods of estimation:
+\begin{enumerate}[resume]
+\item \texttt{ignore}: The power of the \texttt{pb\_type} is ignored, including any children.
+\item \texttt{sum-of-children}: Power of \texttt{pb\_type} is solely the sum of all children \texttt{pb\_types}; interconnect between the \texttt{pb\_type} and its children is ignored.
+\end{enumerate}
 
@@
 \subsubsection{\texttt{pin-toggle}}
-This method allows users to specify the dynamic power of a block in terms of the energy per toggle of each input, output or clock pin for the \texttt{pb\_type}.
+This method allows users to specify the dynamic power of a block in terms of the energy per toggle (in Joules) of each input, output or clock pin for the \texttt{pb\_type}.
 The static power is provided as an absolute (in Watts).
@@
 <pb_type>	
-	<input name="my_input" num_pins="1">
-		<power energy_per_toggle="1.0e-16">
-	</input>
 	...
 	<power method="pin-toggle">
+		<port name="A" energy_per_toggle="1.0e-12"/>
+		<port name="B[3:2]" energy_per_toggle="1.0e-12"/>
+		<port name="C" energy_per_toggle="1.0e-12" scaled_by_static_prob="en1"/>
+		<port name="D" energy_per_toggle="1.0e-12" scaled_by_static_prob_n="en2"/>
 		<static_power power_per_instance="1.0e-6"/>
@@
 
+Keep in mind that the port construct allows for multiple pins per port.
+Unless an subset index is provided, the energy per toggle will be applied to each pin in the port.
+The energy per toggle can be scaled by another signal using the \texttt{scaled\_by\_static\_prob}.
+For example, you could scale the energy of a memory block by the read enable pin.
+If the read enable were high 80\% of the time, then the energy would be scaled by the $signal\_probability$, 0.8.
+Alternatively \texttt{scaled\_by\_static\_prob\_n} can be used for active low signals, and the energy will be scaled by $(1-signal\_probability)$.
+
 This method does not perform any transistor-level estimations; the entire power estimation is performed using the above values.
-\texttt{energy\_per\_toggle} specifies the energy consumed (in Joules) for a single toggle of a pin.
-Keep in mind that the port construct allows for multiple pins per port.
-These energy per toggle will be applied to each pin in the port.
-If necessary, the user can seperate a port into multiple ports with different energy per toggle values.
+It is assumed that the power usage specified here includes power of all child \texttt{pb\_types}.
+No further recursive power estimation will be performed.
 
@@
 \textbf{Buffer} 
-The buffer is a multistage CMOS buffer.  
-Typically, this buffer is sized to drive the capacitance of the wire, as well as fanouts to other switch boxes and conneciton boxes.
-The size of the buffer is specified by the user in the architecture file, as shown below.
+The buffer is a multistage CMOS buffer.  The buffer size is determined based upon output capacitance provided in the architecture file:
+
+\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
+<switchlist>
+	<switch type="mux" ... C_out="XXX"
+</switchlist>
+\end{BVerbatim}
+
+The user may override this method by providing the buffer size as shown below:
 
@@
 
-Alternatively, the size can be replaced by the "\texttt{auto}" keyword, and the buffer will be automatically sized based on the load capacitance.
+The size is the drive strength of the buffer, relative to a minimum-sized inverter.
+
+%Alternatively, the size can be replaced by the "\texttt{auto}" keyword, and the buffer will be automatically sized based on the load capacitance.
 See \autoref{sec:buffer_sizing} for information about buffer sizings.
 
-\textbf{Wire Capacitance} The wire capacitance is determined using the segment information in the architecture file:
+%\textbf{Wire Capacitance} The wire capacitance is determined using the information in the architecture file:
+%
+%\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
+%<segmentlist>
+%	<segment type="unidir" ... Cmetal="XXX" .../>
+%</segmentlist>
+%\end{BVerbatim}
 
-\begin{BVerbatim}[bgcolor=LightGray, boxwidth=\textwidth] 
-<segmentlist>
-	<segment type="unidir" ... power_options .../>
-</segmentlist>
-\end{BVerbatim}
-
-The following options can be used to specify wire capacitance:
-\vspace{-10pt}
-\begin{itemize*}
-\item \texttt{power\_wire\_c}: Absolute wire capacitance, in farads.
-\item \texttt{power\_wire\_c\_per\_m}: Wire capacitance in farads per meter.
+%The following options can be used to specify wire capacitance:
+%\vspace{-10pt}
+%\begin{itemize*}
+%\item \texttt{power\_wire\_c}: Absolute wire capacitance, in farads.
+%\item \texttt{power\_wire\_c\_per\_m}: Wire capacitance in farads per meter.
 The capacitance is calculated using an automatically determined wirelength, based on the area of a tile in the FPGA (\autoref{sec:area_est}).
-\end{itemize*}
+%\end{itemize*}
 
@@
 \section{Other Architecture Options \& Techniques} \label{sec:options}
-\subsection{Area Estimations} \label{sec:area_est}
-The power estimator performs area estimations of both the global routing structures and the complex blocks.
-The CLB area is determined through a recursive function that counts the area of interconnect, buffers, and primitives (flip-flops and LUTs).
 
+\subsection{Local Wire Auto-Sizing} \label{sec:local_wire_autosize}
+Due to the significant user effort required to provide local buffer and wire sizes, we developed an algorithm to estimate them automatically.  
+This algorithm recursively calculates the area of all entities within a CLB, which consists of the area of primitives and the area of local interconnect multiplexers.  
 If an architecture uses new primitives in CLBs, it should include a function that returns the transistor count.
@@
 
-\subsection{Local Wire Auto-Sizing} \label{sec:local_wire_autosize}
-The power estimator contains a basic algorithm to estimate local interconnect wirelengths.
-The algorithms determines the area of the interconnect structures, and calculates the side-length of the interconnect based on a square layout.
-This side length is referred to as $L_{interc}$, as illustrated in \autoref{fig:local_interc_wirelength}.
+ In order to determine the wire length that connects a parent entity to its children, the following assumptions are made:
+
+\begin{itemize}
+	\item Assumption 1: All components (CLB entities, multiplexers, crossbars) are assumed to be contained in a square-shaped area.
+	\item Assumption 2: All wires connecting a parent entity to its child pass through the \emph{interconnect square}, which is the sum area of all interconnect multiplexers belonging to the parent entity.
+\end{itemize}
+
+\autoref{fig:local_interc} provides an illustration of a parent entity connected to its child entities, containing one of each interconnect type (direct, many-to-1, and complete).  In this figure,
+the square on the left represents the area used by the transistors of the interconnect multiplexers.
+It is assumed that all connections from parent to child will pass through this area.
+Real wire lengths could me more or less than this estimate; some pins in the parent may be directly adjacent to child entities, or they may have to
+traverse a distance greater than just the interconnect area.  Unfortuantely, a more rigorous estimation would require some information about the transistor layout.
 
@@
 	\centering
-		\includegraphics[scale=0.7]{images/local_interc_wirelength.pdf}
-	\caption{Local interconnect wirelength}
-	\label{fig:local_interc_wirelength}
+	\includegraphics[width=\textwidth]{images/local_interc}
+	\caption[D Flip-Flop]{Local interconnect wirelength.}
+	\label{fig:local_interc}
 \end{figure}
 
+\begin{table}
+		\centering
+		\begin{tabular}{lccc}\toprule
+		Connection from Entity Pin to:  & Estimated Wirelength & Transistor  Capacitance \\ \midrule
+		Direct (Input or Output) & $0.5 \cdot L_{interc}$ & 0\\ 
+		Many-to-1 (Input or Output)  & $0.5 \cdot L_{interc}$ & $C_{INV}$ \\ 
+		Complete $m$:$n$ (Input) &$0.5 \cdot L_{interc} + L_{crossbar}$ & $n\cdot C_{INV}$ \\
+		Complete $m$:$n$ (Output) & $0.5 \cdot L_{interc}$ & $C_{INV}$\\ \bottomrule
+		\end{tabular}
+	\caption{Local interconnect wirelength and capacitance.  ($C_{INV}$ is the input capacitance of a minimum-sized inverter.)}
+	\label{tab:local_interc_wirelength}	
+\end{table}
+
+\autoref{tab:local_interc_wirelength} details how local wire lengths are determined as a function of entity and interconnect areas.
 It is assumed that each wire connecting a pin of a \texttt{pb\_type} to an interconnect structure is of length $0.5 \cdot L_{interc}$.  In reality, this length depends on the actual transistor
@@
 The PMOS transistor will automatically be sized larger.
-Generally, buffers are sized depending on the load capacitance, using the following equation (logical effort model):
+Generally, buffers are sized depending on the load capacitance, using the following equation:
 
 \begin{align}
-	\text{Buffer Size}=\frac{1}{f_{LE}}*\frac{C_{Load}}{C_{INV}}
+	\text{Buffer Size}=\frac{1}{2\cdot f_{LE}}*\frac{C_{Load}}{C_{INV}}
 	\label{eq:autosize}
 \end{align}	
-In this equation, $C_{INV}$ is the input capacitance of a minimum-sized inverter, and $f_{LE}$ is the logical effort factor, which by default is 4.
-This is standard design practise for high-performance devices.
+In this equation, $C_{INV}$ is the input capacitance of a minimum-sized inverter, and $f_{LE}$ is the logical effort factor.
+The logical effort factor is the gain between stages of the multi-stage buffer, which by default is 4 (minimal delay).
+The term $(2\cdot f_{LE})$ is used so that the ratio of the final stage to the driven capacitance is smaller.
+This produces a much lower-area, lower-power buffer that is still close to the optimal delay, more representative of common design practises.
 The logical effort factor can be modified in the architecture file:
@@
 
-The number of buffer stages is chosen such that each stage is close to $f_{LE}$ times larger than the previous stage.
+%The number of buffer stages is chosen such that each stage is close to $f_{LE}$ times larger than the previous stage.
 
@@
 	<power>
-		<local_interconnect C_wire="1.0e-15">
+		<local_interconnect C_wire="2.5e-15">
 	</power>

### 2013-04-19 vpr/SRC/power/power_util.c

@@ -173,7 +173,8 @@ static void log_msg(t_log * log_ptr, char * msg) {
  * final_stage_size: Size of the final inverter in the buffer, relative to a min size
  * desired_stage_effort: The desired gain between stages, typically 4
  */
-int power_calc_buffer_num_stages(float final_stage_size, float desired_stage_effort) {
+int power_calc_buffer_num_stages(float final_stage_size,
+		float desired_stage_effort) {
 	int N = 1;
 
 	if (final_stage_size <= 1.0) {
@@ -467,7 +468,7 @@ void output_logs(FILE * fp, t_log * logs, int num_logs) {
 float power_buffer_size_from_logical_effort(float C_load) {
 	return std::max(1.0,
 			C_load / g_power_commonly_used->INV_1X_C_in
-					/ 8.0);
+					/ (2 * g_power_arch->logical_effort_factor));
 }
 
 void power_print_title(FILE * fp, char * title) {

### 2014-11-09 doc/journal/fpgapaperfinal/fpga2014paper.tex
adding conference paper source. will use as base for journal paper
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4499 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{sig-alternate-2013}
\usepackage{graphicx}
%\linespread{0.94}

\newfont{\mycrnotice}{ptmr8t at 7pt}
\newfont{\myconfname}{ptmri8t at 7pt}
\let\crnotice\mycrnotice%
\let\confname\myconfname%

\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than the author(s) must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org.}
\conferenceinfo{FPGA'14,}{February 26--28, 2014, Monterey, CA, USA. \\
{\mycrnotice{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}}}
\copyrightetc{ACM \the\acmcopyr}
\crdata{978-1-4503-2671-1/14/02\ ...\$15.00.\\
http://dx.doi.org/10.1145/2554688.2554776}

\clubpenalty=10000
\widowpenalty = 10000

\begin{document}

% --- Author Metadata here ---
%\conferenceinfo{FPGA}{'14 Monterey, California USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{CAD and Routing Architecture for Interposer-Based Multi-FPGA Systems}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.

\numberofauthors{2}
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Andre Hahn Pereira\\
	   \affaddr{Computer and Digital Systems Engineering Department - Escola Polit\'{e}cnica}\\
       \affaddr{University of S\~{a}o Paulo}\\
       \affaddr{S\~{a}o Paulo, SP}\\
       \affaddr{Brazil}\\
       \email{andre.hahn@usp.br}
% 2nd. author
\alignauthor
Vaughn Betz\\
       \affaddr{Department of Electrical and Computer Engineering}\\
       \affaddr{University of Toronto}\\
       \affaddr{Toronto, ON}\\
       \affaddr{Canada}\\
       \email{vaughn@eecg.utoronto.ca}
}
\date{18 December 2013}

\maketitle
\begin{abstract}
%
Interposer-based multi-FPGA systems are composed of multiple FPGA dice connected through a silicon interposer. Such devices allow larger FPGA systems to be built than one monolithic die can accomodate and are now commercially available. An open question, however, is how efficient such systems are compared to a monolithic FPGA, as the number of signals passing between dice is reduced and the signal delay between dice is increased in an interposer system vs. a monolithic FPGA.

We create a new version of VPR to investigate the architecture of such systems, and show that by modifying the placement cost function to minimize the number of signals that must cross between dice we can reduce routing demand by 18\% and delay by 2\%. We also show that the signal count between dice and the signal delay between dice are key architecture parameters for interposer-based FPGA systems. We find that if an interposer supplies (between dice) 60\% of the routing capacity that the normal (within-die) FPGA routing channels supply, there is little impact on the routability of circuits. Smaller routing capacities in the interposer do impact routability however: minimum channel width increases by 20\% and 50\% when an interposer supplies only 40\% and 30\% of the within-die routing, respectively. The interposer also impacts delay, increasing circuit delay by 34\% on average for a 1 ns interposer signal delay and a four-die system. Reducing the interposer delay has a greater benefit in improving circuit speed than does reducing the number of dice in the system.
\end{abstract}

\category{B.7.1}{Integrated Circuits}{Types and Design Style}[VLSI]
\category{B.7.2}{Integrated Circuits}{Design Aids}[Placement and Routing]
\terms{Algorithms, Design}
\keywords{FPGA; Silicon interposer; 2.5D ICs}

\section{Introduction}
\label{introSection} 
%   - Describe multi-chip
Interposer-based multi-FPGA systems are composed of multiple FPGA dice, which are connected through a silicon interposer. The interposer is fabricated with an older technology than that used for the dies, and links between dice include both a micro-bump on each FPGA die and a metal wire on the interposer~\cite{xilinxTSVperformance}. This results in a reduced connectivity between dice and increased delay for connections between dice, as compared to the total routing connectivity and the delay one can achieve within a single die. In this paper we present an architecture study of silicon interposer-based FPGAs that analyzes how the reduced connectivity and increased delay impact their performance.

%   - Motivate: yield, bigger FPGA, no previous study
The silicon interposer FPGA technology is interesting because it enables the creation of large FPGAs composed of small dies and also of very large FPGAs, with higher logic capacity than one can achieve with a single die. Such FPGA systems are sometimes called 2.5D FPGAs, since they make use of vertical stacking of dies on an interposer to enable higher integration levels.

% yield rate
Being able to make large FPGAs with multiple smaller dies is particularly interesting at the beginning of a new manufacturing process, when defect densities are high. In such a case, good-die yield drops dramatically as the die size increases, and this drastically impacts the availability of large FPGAs early in the process lifetime.

%   - Prior work (El Gamal, Xilinx WP, Xilinx paper) - TODO
The idea of using 3 dimensions to design FPGAs is not new. \cite{3dfpga1995} proposed the creation of 3D FPGAs by stacking 2D FPGAs and connecting them with solder bumps. Lin et al also studied 3D FPGAs, but using a different approach~\cite{3dfpga}, proposing multiple active layers, with different FPGA functions in each. While promising, both approaches have manufacturing challenges: \cite{3dfpga1995} requires a higher density of through-silicon vias than can currently be manufactured and the multiple active layers required by \cite{3dfpga} are not yet widely available. In contrast, silicon interposers linked to dies via microbumps are now manufacturable.

Chaware et al presented Xilinx's approach to silicon interposer FPGAs~\cite{xilinxTSV}. They describe the physical characteristics of their implementation, including the bump pitch and estimates of the amount of die-to-die connectivity and the die-to-die delay. However~\cite{xilinxTSV} does not analyze the architecture question of the routability of the resulting system, nor describe possible CAD optimizations, which are the questions we investigate.

%   - Contributions
The main contributions of our work are as follows:
\begin{itemize}
%     * Study on routability vs wires cut
	\item A detailed study of how the multi-FPGA system routability is impacted by the amount of connectivity between dice provided by the interposer.
%     * Timing vs wires cut + delay
	\item An Analysis of the impact of the interposer on timing.
%     * CAD flow optimizations for 2.5D FPGAs
	\item Modification of the VPR CAD tool to model and target 2.5D silicon interposer-based FPGAs, as well as CAD changes to optimize for 2.5D FPGAs.
%     * Analysis of existing Virtex-7 FPGAs
	\item Analysis of the commercially available Xilinx Virtex-7 silicon interposer FPGAs~\cite{xilinxWP}\cite{xilinx7series}. We measure the connectivity reduction and delay increase between dies caused by the interposer on such FPGAs in order to see where these devices lie within the architcture space we explore.
\end{itemize}

The paper is organized as follows. Section \ref{siliconSection} gives more information about silicon interposer technology. Section \ref{archSection} describes the changes made to the FPGA architecture description and to VPR to target 2.5D FPGAs. Section \ref{cadSection} describes optimizations made to the VPR placement algorithm to improve quality on such FPGAs and Section \ref{resultsSection} presents architectural results for 2.5D FPGAs.

\section{Silicon Interposer Background}
\label{siliconSection}

% 1 - FPGAs "More than Moore" - want(?) to combine chips more tightly / with more connectivity than on a board
As mentioned in the previous section, interposer-based FPGAs allow the creation of chips larger than a single die, making a "More than Moore" improvement on the size and number of logic elements possible, and with chips combined far more tightly and with more connectivity than if they were on separate boards connected through conventional means.

% 2 - a) - Can sell bigger chips
%        - Can we make bigger? -> 28nm 1.95m LE vs 979k (V7) / 952k (SV)
The improvement in the number of logic elements of 2.5D FPGAs over conventional ones is very significant. Xilinx's largest interposer-based FPGA, the Virtex-7 XC7V2000T, has 4 dies (which Xilinx calls Super Logic Regions) and 1.954 million logic elements~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die (the XC7VX980T), has 979k logic elements and Altera's largest FPGA, the Stratix V 5SGXBB, has 952k logic elements~\cite{stratixV}. Even though all these FPGAs use a 28 nm process, silicon interposer technology allows the creation of FPGAs with twice the resources possible on even an extremely large single die.

%     b) - Yield results
%               -> Ref paper / use lecture slides
%               -> Defect density of 1/cm^2 is typical at early days of new process
%               -> FPGA early adopters of new process
Another major advantage of interposer-based FPGAs comes at the beginning of a new manufacturing process, when the defect density is high~\cite{xilinxTSVperformance}. Bigger dice suffer a much reduced yield compared to smaller dice, and this greatly affects the supply and cost of top-of-the-line FPGAs to early adopters.

To illustrate this impact, consider a new process in which the defect density is 1/$cm^2$, which is a reasonable value early in the process lifecycle, and the die area is $6cm^2$, which roughly matches the size of the largest member of a high-end FPGA family such as Virtex 7. Using the Poisson Yield Model~\cite{yieldmodel}, the yield is only $0.25\%$ of die. If instead the chip is composed of four $1.5cm^2$ dies, the yield is $22\%$. This means that a 12 inch silicon wafer with $730cm^2$ of area would produce on average $0.3$ working $6cm^2$ dies, while the same wafer would produce on average $107$ working $1.5cm^2$ dies. Therefore, as a $6cm^2$ chip would be composed of four $1.5cm^2$ dies, the wafer would yield $26.75$ systems on average, as the "assembly yield" of placing these four die on an interposer is very high~\cite{xilinxTSV}. Hence the number of interposer-based FPGAs created from the same silicon wafer would be almost $100\times$ greater than a monolithic FPGA of the same size.

When the process matures and the defect density decreases this advantage drops significantly. Consider a mature process with defect density of $0.1/cm^2$. The yield for a $6cm^2$ die is $55\%$ and the yield for a $1.5cm^2$ die is $86\%$. Hence the number of single die FPGAs created from a $730cm^2$ silicon wafer would be $66.9$ and the number of interposer-based FPGAs created would be $104.6$. While the interposer-based FPGA still has a yield advantage it might not lead to a major cost advantage, particularly when the cost of the interposer and assembling the die to it are included. For the large, state-of-the-art FPGAs that are built early in a process cycle and heavily used for prototyping, however, there is clearly a compelling cost advantage to an interposer-based solution.

\subsection{Virtex-7 Interposer-based FPGAs}
\label{virtex7section}

The Xilinx 2.5D FPGAs from the Virtex-7 family are currently the only commercially available silicon interposer-based FPGAs. As described in Section~\ref{archSection} we are studying the impact of several key interposer parameters on the performance of the multi-FPGA system, including (i) the percentage of the wiring normally present between rows of the FPGA which are cut when crossing between dice, and (ii) the extra delay (vs. a normal connection between adjacent rows) added when one must traverse the interposer.  To locate where Virtex-7 lies in this architecture space, we combined published information on the implementation of Xilinx's interposer-based FPGAs~\cite{xilinxTSV} with a detailed analysis of the~\textit{XC7V2000T} FPGA routing resources visible in the \textit{Vivado Device View}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{interposer.eps}
\caption{Lateral view of an interposer-based FPGA\cite{xilinxWP}. The FPGA dice are at the top, and are connected to the silicon interposer through microbumps. The interposer is then connected to the substrate through C4 bumps.}
\label{fig:interposer}
\end{figure}

% Xilinx details:
%  -> 13440 wires crossing
%  -> ~63k vertical wires
%  -> Micro-bumps -> pitch is 45um
%  -> Spread out near edge, to get more bumps
%  -> But results in longer signal travel
%  -> Net result: more delay, reduced connectivity

The \textit{XC7V2000T} is composed of four identical dice arranged such that the vertical routing crosses between the dice. Each horizontal edge of each die has 280 groups of 48 length-12 wires crossing the interposer, which sums to a total of 13440 wires between dice. There are also 40 clock wires crossing the interposer. The average number of wires per vertical channel of this FPGA is 210 and there are approximately 280 vertical channels on the FPGA, resulting in approximately 58800 vertical wires crossing a horizontal cutline within a die. Hence the number of wires which cross the interposer is about 23\% of the total number of within-die vertical wires.

The 28nm dies are connected to the 65nm silicon interposer through microbumps with a 45$\mu$m pitch. Hence the area occupied by microbumps at one edge of one die is 13440 $\times$ (45$\mu$m)\textsuperscript{2} $=$ 27mm\textsuperscript{2}. If we assume each die is $7 \times 12 mm$, as presented by Chaware et al. in \cite{xilinxTSV}, the bumps have to be spread out near the edge and need to go as far as $2.25mm$ away from the edge of the die. This greater distance from the border increaes the length of the inter-die connections, and along with the presence of the micro-bumps and their capacitance, leads to an increased delay for these crossing wires vs. that of a typical on-die routing wire. Chaware et al. state that the latency to cross the interposer is approximately $1ns$. For comparison, a typical medium length 28 nm FPGA routing wire (e.g. spanning four logic blocks) has a delay of approximately 125 ps, while a longer wire (e.g. spanning 12 logic blocks) has a delay of approximately 250 ps.

Overall, these interposer-based FPGAs have increased delay and reduced connectivity between dies, with approximately $23\%$ of the usual number of vertical wires crossing between dies and approximately $1ns$ of increased delay to cross the interposer.

\section{Architecture models}
\label{archSection}
%   - Define parameters
To properly model a silicon interposer FPGA, the popular FPGA exploration toolset, Verilog-to-Routing (VTR)~\cite{vtr2012}, was used. The logic synthesis portions of the flow (ODIN II and  ABC) were left untouched, while the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) was modified to model and optimize for interposer-based FPGAs. The modifications were made in such a way that they require no changes to any of the input files, so one can experiment with interposer-based FPGAs with any existing benchmark circuits and any existing VPR-format FPGA architecture description simply by specifying appropriate command-line parameters.

Three parameters were added to VPR: \textit{\% wires cut}, \textit{increased delay} and \textit{number of cuts}. These three parameters describe the interposer portion of the 2.5D FPGA as detailed below.

%   - % wires cut
\subsection{\% wires cut}
This variable describes and models the reduced connectivity between different dice by specifying the fraction of routing wires that are removed at the border between dice. For example, if a channel had 200 wires and \textit{\% wires cut} was 70, 140 of them would be cut and only 60 would pass through the interposer. Higher values of \textit{\% wires cut} make an interposer easier to manufacture, and can reduce the interposer delay by allowing all the microbumps linking dice to be placed near the die edge. However, the higher \textit{\% wires cut} is, the less routable the multi-die system becomes. As described in \ref{virtex7section}, the Virtex 7 family has \textit{\% wires cut} $=$ 77\%.

%   - Added delay
\subsection{Increased delay}
Interposer wires are longer and wider than on-die wires and have microbumps on each end. \textit{Increased delay} models the resulting larger delay when compared with wires which are internal to a die. A reasonable estimate for this variable is around $1ns$, as presented by Chaware et al\cite{xilinxTSV}.

%	- number of cuts
\subsection{Number of cuts}
\textit{Number of cuts} describes how many cuts were made to the interposer-based FPGA versus a monolithic die. If \textit{number of cuts} equals 1 then there are 2 dies, and so on. We investigate values of this parameter between 1 and 3 (between 2 and 4 dies), reflecting commercial practice: the Virtex 7 family has members with \textit{number of cuts} $=$ 2 and 3. Figure \ref{fig:fpga} shows a sample architecture with one cutline.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{fpga.eps}
\caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
\label{fig:fpga}
\end{figure}

%   - (briefly) explain how it was put in VPR
\subsection{Implementation}
To model an interposer-based FPGA in VPR the Routing Resource Graph (\textit{rrgraph}) must be modified. The rrgraph is the data structure that defines all the available routing wires and switches in the FPGA, as well the delay of each. Given a suitable rrgraph, the VPR router can implement circuits in the desired FPGA, and the VPR timing analyzer can estimate their delay.

The presence of multiple dice in an interposer-based FPGA was modeled by creating horizontal cuts in the FPGA, which are equally spaced vertically. Every wire that crosses one of these cuts in fact passes through the interposer. In the experiments below, we use an architecture that has only unidirectional wires~\cite{unidirectional}, as these are the dominant routing architecture in modern FPGAs. Such wires can only be driven at their beginning. To control the wiring capacity between dies \textit{\% wires cut} wires of each channel segment crossing a cutline have their connections to wires and block inputs on the die opposite to their starting point removed, and the wires which aren't cut have \textit{increased delay} added to their connections which cross the cutline. Note that the combination of a unidirectional routing architecture and a silicon interposer results in some blocks near the cutline having reduced routing connectivity as some of the routing wires driving their inputs are disconnected from the (single) wire driver, when that driver is on the other side of the cutline and the wire is one of the cut wires.

Figure \ref{fig:implementation} illustrates the approach used.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{implementation.eps}
\caption{Illustration of the approach used to simulate crossing the interposer. The red dashed line indicates the interposer, the black arrows indicate the direction of the wires. Blue arrows indicate normal connections from the wires to the CLBs. The red arrows indicate removed connections and green arrows indicate connections with increased delay.}
\label{fig:implementation}
\end{figure}

For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the "flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}). The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices. It consists of logic clusters with 10 fracturable 6-LUTs per block (N=10, k=6), and also includes 32kb RAM blocks and DSP blocks configurable to perform 9x9, 18x18 or 36x36 multipliers. The delay values in the architecture are taken from 40 nm circuit simulations and 40 nm commercial FPGAs. It uses unidirectional routing, with all wire segments having length L = 4.

\section{CAD Enhancements}
\label{cadSection}

Once the routing resource graph was modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture. Placement, however, is crucial in mitigating the impact of the reduced wiring and increased delay when crossing between dice; a good placement should minimize the number of signals crossing between dice, particularly time-critical ones. We investigated several enhancements to VPR's placement cost function to improve result quality.

VPR uses 2 different costs as the metrics for its placer algorithm: the timing cost and the bounding box (wiring) cost. The usual VPR timing cost is a (criticality-weighted) summation of the estimated delay (given the current placement) of every connection required by the circuit~\cite{timing2000}:

\begin{equation} \label{eq:timing_eq_full}
\begin{split}
Timing\_Cost = \sum_{\forall i, j \subset circuit} delay(\Delta x_{ij},\Delta y_{ij}) \times \\ 
criticality(i,j)
\end{split}
\end{equation}

where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
The bounding box cost estimates the amount of wiring required for a net, based on the number of pins and size of the net's bounding box. VPR's original formulation is~\cite{betz1997vpr}:

\begin{equation} \label{eq:bbcost}
\begin{split}
wiring\_cost_{orig} = \sum_{n=1}^{N_{nets}} q(n) \times [\frac{bb_x(n)}{avg\_chanx\_W(n)} + \\
\frac{bb_y(n)}{avg\_chany\_W(n)}]
\end{split}
\end{equation}

where $bb_x(n)$ and $bb_y(n)$ are the dimensions of the bounding box of net \textit{n} in the x and y directions, respectively. $avg\_chanx\_W$ and $avg\_chany\_W$ are the average x-directed and y-directed channel widths over this bounding box. Finally, $q(n)$ is a function obtained from \cite{icann} which models the fact that bounding boxes underpredict the required routing for high fanout nets. $q(n)$ slowly increases with the fanout of net \textit{n}, rising from 1 for nets with 3 or fewer terminals to 2.79 for nets with 50 terminals.

We modified both the placer's bounding box and timing costs to account for the increased latency to cross the interposer and for the reduced wire capacity close to the cutlines, as the wire availability becomes more sparse.

%   - Delay
\subsection{Placer timing cost}
The standard placer timing cost in VPR assumes that the FPGA is homogeneous and consequently the delay between 2 points $(x_1,y_1)$ and $(x_2,y_2)$ only depends on $(\Delta x,\Delta y)$. This is obviously not true for interposer-based FPGAs, as the cutlines make them heterogeneous in the \textit{y} direction.

To solve this problem and improve the quality of the results, an extra term was added to the delay function. The delay function becomes:

\begin{equation} \label{eq:timing_cost}
\begin{split}
delay(i,j) =& delay(\Delta x_{ij}, \Delta y_{ij}) + \\
&times\_crossed(i,j) \times delay\_increase
\end{split}
\end{equation}

where $times\_crossed(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.

%   - Wiring / cut cost
\subsection{Placer wiring cost}
VPR's wiring cost also considers the FPGA to be homogeneous, and uses only the number of nets, the size of the net's bounding box and the average number of wires per channel to calculate the cost. Thus, to account for the reduced connectivity near the cutlines an extra cost term, \textit{cut\_cost}, was created. This new cost is added to (\ref{eq:bbcost}) to create the total wiring cost.

\begin{equation} \label{eq:total_wiring}
wiring\_cost = wiring\_cost_{orig} + cut\_cost
\end{equation}

We tested several different \textit{cut\_cost} formulations, as well as different weighting (C values) for each. For all of the formulations the variable $C'$ was defined as:

\begin{equation} \label{eq:cprime}
\begin{split}
C' = \frac{C \times ratio\_wires\_cut}{avg\_chany\_W(n)}
\end{split}
\end{equation}

where $ratio\_wires\_cut$ is the ratio of wires cut at the cutline. This formulation of $C'$ ensures that when we choose a \textit{C} value of 1, the new term cost term is of roughly the same magnitude as $wire\_cost_{orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.

The first \textit{cut\_cost} term we tested was:

\begin{equation} \label{eq:cost0}
cut\_cost = \sum_{n=1}^{N_{nets}} C' \times times\_crossed(n)
\end{equation}

which penalizes the net according to the number of cutlines the bounding box spans; this directly penalizes each signal crossing between dice. Surprisingly this extra term did not result in significant improvements to the quality of results and in some cases even made them worse. We believe this is due to the discontinuous, non-smooth, nature of this cost function -- it does not guide placement optimization well as a bounding box shrinks toward a cutline. Placement changes that make bounding boxes "almost not cross" a cutline are not given any gain; only a sudden change in the placement that moves a bounding box entirely to one side or the other of the cutline yields a cost reduction. This insight led us to the other proposed functions, which give gradual gains as bounding boxes change in their size and come closer to avoiding a cutline crossing.

The other tested terms were:

%add C' x times_crossed(i)

\begin{equation} \label{eq:cost1}
cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbWidth(n) \times times\_crossed(n)
\end{equation}

\begin{equation} \label{eq:cost2}
cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbHeight(i) \times times\_crossed(n)
\end{equation}

\begin{equation} \label{eq:cost4}
\begin{split}
cut\_cost = \sum_{n=1}^{N_{nets}} C' \times minDist(i) \times times\_crossed(n) +\\
C' \times times\_crossed(n)
\end{split}
\end{equation}

\begin{equation} \label{eq:cost5}
cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbHeight(n)
\end{equation}

where \textit{bbWidth(n)} and \textit{bbHeight(n)} are the width and height of the bounding box of the net \textit{n}, respectively, and \textit{minDist(n)} is the minimum distance from the top or bottom of the bounding box to a cutline.

%   - Effectiveness / results
\subsection{Effectiveness of the enhancements}
\label{sec:CADeffect}

We used VPR 7.0, with the enhancements we detailed above, and the architecture file \textit{k6\_frac\_N10\_mem32K\_40nm} in the experiments below. All experiments targeted the smallest FPGA (with number of rows equal to number of columns) which could accomodate a benchmark circuit; this represents a very full FPGA with little white space left, and hence presents a difficult case to an interposer-based FPGA as no die can be left mostly empty.

Figures \ref{fig:minW_constant_sweep} and \ref{fig:areadelay_constant_sweep} show the efficacy of the five placer wiring cost modifications as their weight, \textit{C} varies. Each point in the graphs is the geometric mean of the results with \textit{\% wires cut} = 60 and 80, and the results for 60\% and 80\% are themselves geometric means of the results of 6 circuits from the VTR benchmark suite~\cite{vtr2012}:  \textit{stereovision0}, \textit{stereovision1}, \textit{mkDelayWorker32B}, \textit{mkSMAdapter4B}, \textit{or1200} and \textit{blob\_merge}. These values for \textit{\% wires cut} were chosen because they in the likely range of wires which can cross the interposer without exceeding the microbump capacity, as described in Section~\ref{virtex7section}. The \textit{area-delay product} is the product of the \textit{minimum channel width} and the \textit{critical path delay}. To obtain the \textit{critical path delay} the circuits were run with a low stress routing with a channel width, $W = 1.3 \times minW$, where $minW$ is the minimum channel width for which the circuit is still routable.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{minW_constant_sweep_log.eps}
\caption{Minimum channel width vs. weighting for different \textit{cut\_cost} terms.}
\label{fig:minW_constant_sweep}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{areadelay_constant_sweep_log.eps}
\caption{Area-delay product vs. weighting for different \textit{cut\_cost} terms.}
\label{fig:areadelay_constant_sweep}
\end{figure}

\begin{table}[htbp]
\begin{tabular}{|r|r|r|r|r|}
\hline
\multicolumn{1}{|l|}{Term} & \multicolumn{1}{l|}{Best C} & \multicolumn{1}{l|}{minW} & \multicolumn{1}{l|}{crit\_path(ns)} & \multicolumn{1}{l|}{Area-delay} \\ \hline \hline
None & - & 124.636 & 9.844 & 1226.955 \\ \hline
\ref{eq:cost0} & 0.3 & 122.929 & 9.413 & 1157.09 \\ \hline
\ref{eq:cost1} & 5 & 114.444 & 9.565 & 1094.68 \\ \hline
\ref{eq:cost2} & 1 & 107.303 & 9.285 & 996.26 \\ \hline
\ref{eq:cost4} & 1 & 113.381 & 9.280 & 1052.16 \\ \hline
\ref{eq:cost5} & 5 & 112.703 & 9.841 & 1109.11 \\ \hline
\end{tabular}
\caption{Best weighting and performance for each \textit{cut\_cost} term.}
\label{table:constant_sweep}
\end{table}

The best \textit{C} value and the resulting performance for each term is summarized in Table~\ref{table:constant_sweep}. Note that cost term (\ref{eq:cost2}) with weighting $C = 1.0$ has the best performance; this is the configuration used in the rest of our experiments.

We believe the key to the good performance of this cost function is that it produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline. Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}. Note that bounding box (a) and (b) both cross the cutline and hence are penalized equally by (\ref{eq:cost0}), while bounding box (c) does not cross a cutline and hence is not penalized. However, bounding box (b) is mostly on the lower side of the cutline; it is more likely that later smaller placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand. Cost function terms (\ref{eq:cost2}) and (\ref{eq:cost4}) will both penalize bounding box (a) more than bounding box (b) to guide placement to gradually move bounding boxes to one side or the other of a cutline. Figure \ref{fig:areadelay_constant_sweep} shows that these are the two best performing \textit{cut\_cost} terms.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{bb.eps}
\caption{Illustration of three different scenarios for the wiring cost. The dashed green box shows a case where \textit{(a)} it crosses the interposer, the dotted black box shows a case where \textit{(b)} it barely crosses the cutline, and the solid blue one shows a case where \textit{(c)} the bounding box does not span the cutline.}
\label{fig:bb_illustration}
\end{figure}

Figure \ref{fig:comparison_minW} shows the minimum channel width for different values of \textit{\% wires cut} when there are no optimizations and when \textit{cut\_const} (\ref{eq:cost2}) is used. % comment about impact vs % wires cut
Figure \ref{fig:comparison_bars} compares result quality with both enhanced timing and wiring cost functions. These experiments were run with term (\ref{eq:cost2}), and are the geometric means of the runs with \textit{\% wires cut = 60, 70, 80}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{comparison_minW.eps}
\caption{Minimum channel width vs. \textit{\% wires cut}.}
\label{fig:comparison_minW}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{comparison_bars_relative.eps}
\caption{Performance of different placer optimizations, relative to the original VPR results. The area-delay product is calculated as the product $minW \times critical path delay.$}
\label{fig:comparison_bars}
\end{figure}

The graphs show that combining the timing with the bounding box optimizations led to an average of 20\% improvement on the area-delay product and the most significant contribution was made by incorporation of \textit{cut\_cost} into the wire cost function.

%----------------------------------------------------
\section{Architecture results}
\label{resultsSection}

Using the best CAD settings found in Section~\ref{cadSection} we analyzed the impact of three key architecture parameters:  \textit{\% wires cut}, \textit{delay increase} and \textit{number of cuts}. All the experiments were run with the eight largest circuits from the VTR benchmark~\cite{vtr2012}, namely: \textit{bgm}, \textit{LU8PEEng}, \textit{LU32PEEng}, \textit{mcml}, \textit{mkDelayWorker32B}, \textit{stereovision0}, \textit{stereovision1} and \textit{stereovision2}. The size of these circuits ranges from $9,100$ to $153,000$ primitives (LUTs, FFs, etc.), with an average size of $52,600$ primitives. All results are the geometric mean over all circuits for a given interposer architecture.

%   - Channel width vs % wires cut
\subsection{Interposer Wiring Capacity (\% wires cut)}

To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the number of cuts and the increased delay constant. The values used for these parameters were \textit{number of cuts} $= 3$ (corresponding to four dice) and \textit{increased delay} $= 1ns$, as these values are similar to those of the \textit{XC7V2000T} device.

Figure \ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}. It can be noted from this graph that the minimum channel width increases slowly up to 60\% of wires cut, indicating that the placement engine is able to avoid saturating the interposer routing until that point. When more than 60\% of the wires are cut however (i.e. the interposer provides less than 40\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor. The minW value at 60\% is 20\% greater than with no wires cut, while at 70\% it is 52\% above and at 80\% it is 125\% larger than the minW with no cut wires.

The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 60\% of the wires cut is essentially the same as at 0\% wires cut. At 80\% of the wires cut the critical path delay does rise somewhat, by 6\%; this is due to two factors: the placement is being modified to improve routability at the expense of timing, and some circuitous routes are occuring due to saturated interconnect across the interposer.
%Andre, what is the critical path delay with no interposer (i.e. 0% cut and 0 delay)?

Tables \ref{table:standard_minW} and \ref{table:standard_path} show the minimum channel width and critical path delay for each circuit. The trends for individual circuits follow those of the averages quite closely.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{standard_minW.eps}
\caption{Minimum channel width vs. \textit{\% wires cut} for 3 cuts and $1ns$ of \textit{delay increase}.}
\label{fig:standard_minW}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{standard_crit_path.eps}
\caption{Critical path delay vs. \textit{\% wires cut} for 3 cuts and $1ns$ of \textit{delay increase}.}
\label{fig:standard_crit}
\end{figure}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 114 & 128 & 124 & 132 & 152 & 226 \\ \hline
LU32PEEng & 172 & 178 & 182 & 218 & 304 & 426 \\ \hline
LU8PEEng & 110 & 116 & 118 & 128 & 158 & 236 \\ \hline
mcml & 108 & 124 & 128 & 116 & 152 & 218 \\ \hline
mkDelayWorker32B & 76 & 80 & 82 & 88 & 100 & 132 \\ \hline
stereovision0 & 62 & 60 & 70 & 88 & 110 & 166 \\ \hline
stereovision1 & 104 & 108 & 116 & 120 & 178 & 312 \\ \hline
stereovision2 & 162 & 164 & 168 & 194 & 244 & 360 \\ \hline
Geometric mean & 108 & 114 & 118 & 129 & 164 & 243 \\ \hline
\end{tabular}
\caption{Minimum channel width when \textit{number of cuts = 3} and \textit{increased delay = 1ns}.}
\label{table:standard_minW}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 35.1 & 33.6 & 32.2 & 32.9 & 37.8 & 40.3 \\ \hline
LU32PEEng & 115 & 114 & 114 & 114 & 115 & 113 \\ \hline
LU8PEEng & 118 & 117 & 119 & 116 & 117 & 124 \\ \hline
mcml & 82.3 & 82.9 & 84.0 & 83.4 & 88.3 & 85.7 \\ \hline
mkDelayWorker32B & 11.0 & 12.3 & 12.0 & 12.3 & 14.4 & 12.8 \\ \hline
stereovision0 & 6.9 & 7.8 & 7.0 & 7.4 & 7.6 & 7.0 \\ \hline
stereovision1 & 11.6 & 9.2 & 10.4 & 11.0 & 11.6 & 11.8 \\ \hline
stereovision2 & 23.3 & 25.4 & 23.9 & 24.9 & 25.6 & 24.8 \\ \hline
Geometric mean & 30.8 & 30.9 & 30.6 & 31.2 & 33.1 & 32.6 \\ \hline
\end{tabular}
\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 1ns}.}
\label{table:standard_path}
\end{table}

% TODO
% "extra FPGA routing channel width" vs. wires across the interposer
%\subsection{Intra-FPGA Routing  width vs. wires across the interposer}

%Can you take a crack at adding a section that tries to present the data as "extra FPGA routing channel width" vs. wires across the interposer?  The idea would be to address reviewer 4's comments:  the graph will show we need some minimum number of wires across the interposer, and it will also directly show how many more wires we need to add to the FPGA to maintain routability.  I think you could get this data simply by processing data you already have.
%So: 
%x-axis:  geomean wires crossing interposer
%y-axis:  geomean channel width required

Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability. This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 4-die system. When 108 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die. As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer. The increase is gradual as the interposer routing is reduced from 108 tracks per channel to 52 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 108 tracks per channel to 129 tracks per channel. As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate. At 49 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{243} tracks to successfully route the designs. Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (48\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{numberofcrossingwires.eps}
\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 3 cuts and $1ns$ of \textit{delay increase}.}
\label{fig:crossingwires}
\end{figure}

%   - Delay vs % wires cut & wire delay
\subsection{Circuit Speed vs. Interposer Delay}

To investigate the impact of the interposer delay (\textit{delay increase}), we kept the \textit{number of cuts} constant at $3$ while \textit{delay increase} was swept between $0$ and $1.5ns$.

Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}. The penalty in critical path delay is significant, ranging between $5$ and $9$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay. Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA. The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 20\%, while a $1ns$ interposer delay increases critical path delay by approximately 35\% vs. a monolithic FPGA. Once again, the critical path increase shows little correlation to the \textit{\% wires cut}, however. Tables \ref{table:delay0}, \ref{table:delay500}, \ref{table:standard_path} and \ref{table:delay1500} show the critical path delay for each circuit for each of the different values of \textit{delay increase}. Once again the results for individual circuits closely parallel the overall averages.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{delays_crit_path.eps}
\caption{Critical path delay vs. \textit{\% wires cut} for 3 cuts and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
\label{fig:delays_crit}
\end{figure}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline  \hline
bgm & 25.9 & 26.3 & 25.8 & 26.1 & 25.8 & 26.7 \\ \hline
LU32PEEng & 113 & 113 & 117 & 118 & 113 & 114 \\ \hline
LU8PEEng & 115 & 114 & 116 & 115 & 118 & 115 \\ \hline
mcml & 77.2 & 78.7 & 79.5 & 80.5 & 77.3 & 76.7 \\ \hline
mkDelayWorker32B & 7.8 & 7.6 & 7.5 & 7.7 & 7.5 & 7.6 \\ \hline
stereovision0 & 4.5 & 4.3 & 4.5 & 4.5 & 4.5 & 4.5 \\ \hline
stereovision1 & 6.0 & 6.2 & 5.9 & 6.3 & 6.2 & 6.2 \\ \hline
stereovision2 & 18.0 & 17.1 & 17.1 & 17.5 & 17.6 & 17.7 \\ \hline
Geometric mean & 23.7 & 23.5 & 23.6 & 23.9 & 23.6 & 23.7 \\ \hline
\end{tabular}
\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 0ns}.}
\label{table:delay0}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 31.6 & 30.3 & 29.2 & 29.2 & 29.9 & 31.6 \\ \hline
LU32PEEng & 119 & 114 & 115 & 114 & 115 & 115 \\ \hline
LU8PEEng & 118 & 116 & 119 & 117 & 114 & 115 \\ \hline
mcml & 82.6 & 82.7 & 79.6 & 81.9 & 84.5 & 82.6 \\ \hline
mkDelayWorker32B & 9.8 & 10.6 & 10.4 & 9.6 & 10.2 & 9.1 \\ \hline
stereovision0 & 5.7 & 5.3 & 4.8 & 4.9 & 4.9 & 4.8 \\ \hline
stereovision1 & 8.0 & 7.5 & 7.4 & 8.2 & 7.9 & 8.0 \\ \hline
stereovision2 & 20.1 & 19.7 & 20.3 & 19.9 & 20.2 & 21.6 \\ \hline
Geometric mean & 27.5 & 26.9 & 26.4 & 26.5 & 26.7 & 26.8 \\ \hline
\end{tabular}
\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 0.5ns}.}
\label{table:delay500}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 47.2 & 52.4 & 43.1 & 42.7 & 44.8 & 48.3 \\ \hline
LU32PEEng & 115 & 112 & 117 & 114 & 116 & 113 \\ \hline
LU8PEEng & 161 & 159 & 122 & 146 & 118 & 129 \\ \hline
mcml & 94.0 & 86.0 & 93.9 & 88.7 & 77.3 & 87.5 \\ \hline
mkDelayWorker32B & 15.3 & 13.8 & 14.3 & 15.3 & 17.9 & 15.4 \\ \hline
stereovision0 & 7.5 & 9.4 & 8.9 & 7.8 & 9.6 & 8.3 \\ \hline
stereovision1 & 14.6 & 12.0 & 13.1 & 12.7 & 14.6 & 14.8 \\ \hline
stereovision2 & 29.9 & 30.3 & 28.7 & 30.3 & 30.1 & 29.8 \\ \hline
Geometric mean & 37.7 & 37.3 & 35.9 & 36.1 & 37.0 & 36.9 \\ \hline
\end{tabular}
\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 1.5ns}.}
\label{table:delay1500}
\end{table}

%   - Impact of # of cutlines
\subsection{Impact of Number of Dice}

To examine the impact of the number of die used to construct an interposer-based FPGA, we varied the number of cuts from 1 to 3 (which varies the number of die from 2 to 4).  In these experiments, the \textit{delay increase} was kept constant at $1ns$.

As shown in Figure \ref{fig:cuts_minW}, the number of cuts does have an impact on the minimum channel width, but not a very large or constant one. At 80\% of wires cut the experiments with 2 and 3 cuts had the minimum channel width only 8\% and 10\%, respectively, greater than the scenario with only 1 cut. At lower values of \textit{\% wires cut} the difference was even smaller.

Figure \ref{fig:cuts_crit} shows that the number of dice in an interposer-based FPGA impacts circuit speed significantly, as the critical path delay rises significantly as systems have more cuts. Recall that a monolithic FPGA has a geometric average critical path delay of $23.7ns$ for our test architecture. With \textit{number of cuts} $= 1$ the critical path delay increases by 3 to 4 ns, while with \textit{number of cuts} $= 2$ the critical path delay increases by 5 to 6 ns. With \textit{number of cuts} $= 3$, the critical path delay is typically 7 ns higher than that of a monolithic FPGA.

Tables \ref{table:1cut_minW}, \ref{table:1cut_path}, \ref{table:2cut_minW}, \ref{table:2cut_path}, \ref{table:standard_minW} and \ref{table:standard_path} show the minimum channel width and critical path delay for each circuit for each of the different scenarios. The critical path results for individual circuits again follow the overall averages well. Some of the minimum channel widths for individual circuits show anomalous trends as the number of cutlines increase, however. For example, \textit{mcml} often has a smaller minimum channel width with 2 cuts than with 1 cut. This may be due to a circuit structure that leads to easier division into thirds than halves, but also likely indicates room for further CAD optimization.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{cuts_minW.eps}
\caption{Minimum channel width vs. \textit{\% wires cut} for 1, 2 and 3 cuts and $1ns$ of \textit{delay increase}.}
\label{fig:cuts_minW}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{cuts_crit_path.eps}
\caption{Critical path delay vs. \textit{\% wires cut} for 1, 2 and 3 cuts and $1ns$ of \textit{delay increase}.}
\label{fig:cuts_crit}
\end{figure}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 114 & 114 & 122 & 126 & 142 & 156 \\ \hline
LU32PEEng & 170 & 176 & 182 & 206 & 318 & 396 \\ \hline
LU8PEEng & 110 & 114 & 116 & 138 & 164 & 236 \\ \hline
mcml & 108 & 114 & 106 & 124 & 150 & 238 \\ \hline
mkDelayWorker32B & 74 & 78 & 80 & 88 & 98 & 138 \\ \hline
stereovision0 & 62 & 60 & 62 & 76 & 132 & 146 \\ \hline
stereovision1 & 102 & 104 & 106 & 108 & 144 & 186 \\ \hline
stereovision2 & 160 & 158 & 164 & 200 & 264 & 432 \\ \hline
Geometric mean & 107 & 109 & 111 & 126 & 165 & 221 \\ \hline
\end{tabular}
\caption{Minimum channel width when \textit{number of cuts = 1} and \textit{increased delay = 1ns}.}
\label{table:1cut_minW}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 26.5 & 32.1 & 28.3 & 30.2 & 27.5 & 30.0 \\ \hline
LU32PEEng & 113 & 113 & 113 & 115 & 114 & 111 \\ \hline
LU8PEEng & 117 & 118 & 116 & 114 & 114 & 116 \\ \hline
mcml & 82.5 & 82.0 & 83.0 & 84.5 & 83.2 & 77.9 \\ \hline
mkDelayWorker32B & 11.2 & 11.4 & 11.2 & 9.9 & 10.6 & 11.5 \\ \hline
stereovision0 & 4.5 & 5.1 & 5.0 & 4.9 & 4.9 & 5.4 \\ \hline
stereovision1 & 7.4 & 8.2 & 8.4 & 7.2 & 7.2 & 7.9 \\ \hline
stereovision2 & 21.3 & 23.1 & 20.9 & 22.3 & 21.6 & 20.9 \\ \hline
Geometric mean & 26.3 & 28.1 & 27.2 & 26.8 & 26.5 & 27.3 \\ \hline
\end{tabular}
\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 1} and \textit{increased delay = 1ns}.}
\label{table:1cut_path}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 114 & 116 & 130 & 140 & 172 & 272 \\ \hline
LU32PEEng & 170 & 180 & 186 & 198 & 264 & 426 \\ \hline
LU8PEEng & 110 & 114 & 116 & 124 & 158 & 246 \\ \hline
mcml & 108 & 110 & 126 & 138 & 114 & 150 \\ \hline
mkDelayWorker32B & 76 & 80 & 82 & 88 & 100 & 132 \\ \hline
stereovision0 & 62 & 58 & 60 & 88 & 114 & 166 \\ \hline
stereovision1 & 104 & 108 & 116 & 136 & 184 & 296 \\ \hline
stereovision2 & 160 & 156 & 162 & 188 & 246 & 362 \\ \hline
Geometric mean & 108 & 109 & 116 & 132 & 160 & 237 \\ \hline
\end{tabular}
\caption{Minimum channel width when \textit{number of cuts = 2} and \textit{increased delay = 1ns}.}
\label{table:2cut_minW}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
bgm & 31.0 & 30.3 & 28.3 & 36.1 & 35.9 & 37.5 \\ \hline
LU32PEEng & 113 & 113 & 119 & 113 & 117 & 112 \\ \hline
LU8PEEng & 118 & 118 & 117 & 116 & 118 & 114 \\ \hline
mcml & 82.9 & 86.3 & 81.3 & 80.9 & 82.4 & 82.8 \\ \hline
mkDelayWorker32B & 12.3 & 12.1 & 9.8 & 11.1 & 13.5 & 11.3 \\ \hline
stereovision0 & 6.4 & 5.7 & 6.3 & 5.3 & 5.3 & 5.9 \\ \hline
stereovision1 & 9.3 & 10.4 & 9.5 & 9.6 & 10.2 & 9.9 \\ \hline
stereovision2 & 22.9 & 24.2 & 24.0 & 21.6 & 24.2 & 25.5 \\ \hline
Geometric mean & 29.5 & 29.7 & 28.6 & 28.8 & 30.4 & 30.1 \\ \hline
\end{tabular}
\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 2} and \textit{increased delay = 1ns}.}
\label{table:2cut_path}
\end{table}

\section{Conclusions and Future Work}
\label{conclusionSection}

We have extended VPR to model and optimize for interposer-based multi-FPGA systems. While such systems are now a commercial reality, we do not know of any prior study of their key architectural parameters. We found that by modifying VPR's placement cost function we could improve the routability (reduce minW) by 18\%, while simultaneously improving speed by 2\%. Interestingly, we found that a smooth cost function that less precisely models the interposing wiring demand outperformed direct monitoring of the number of required interposer connections during annealing.

We defined three key architecture parameters for interposer-based FPGAs, and used this extended VPR to analyze their impact on minimum channel width and critical path delay. We found that the \textit{minimum channel width} increases steeply after more than 60\% of the within-die wires are cut at the interposer, and when 80\% of wires are cut the \textit{minimum channel width} is more than double that required by a monolithic FPGA. The Xilinx \textit{XC7V2000T} FPGA has 77\% of the normal wiring cut at the interposer, indicating that commercial interposer-based FPGAs will require high-quality CAD optimization to maintain good routability. The \textit{critical path delay} is not strongly influenced by the \textit{\% wires cut} but is strongly influenced by the interposer delay and the number of cuts. Our results show that the critical path usually crosses the interposer cutline more than once, as the \textit{critical path delay} is increased by more than the interposer \textit{delay increase}. Increasing the number of dice in an interposer-based FPGA does not significantly impact the \textit{minimum channel width}, but does lead to a larger \textit{critical path delay}.

There are many interesting CAD and architecture questions for interposer-based FPGAs which we plan to investigate.  Currently the wires that cross the interposer are accessed with the same switch structure as other wires in the FPGA. Since interposer wires are more scarce, possibly these wires should have larger multiplexers feeding them, which would make them easier to use; this may help routability of the system. Such changes will have to be carefully considered however, to make sure the FPGA can still be laid out as an array of regular tiles. Alternative CAD flows to improve the interposer routability are also possible. For example, instead of following the synthesize, place and route CAD flow of a conventional FPGA, we could add a partitioning step before placement which would divide the circuit into one partition per die. Such a flow may improve routability, as a partitioner's main focus is minimizing the number of cut signals. Enhanced partitioners will be required, as current partitioners such as hMetis~\cite{hMetis} cannot model the heterogeneous balance constraint (i.e. use no more than the available device logic, RAM, DSP and I/O blocks in any die) present in FPGAs. Additionally, by dividing the placement problem into two pieces -- partitioning and within-die placement -- we may increase the critical path delay as we can no longer globally optimize the placement of timing critical paths in one unified placement step. Nonetheless, this is a viable alternative approach and a comparison to the approach taken in this paper would be very interesting.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

This work was supported by a Ci\^{e}ncia sem Fronteiras scholarship from CNPq - Brazil and the NSERC/Altera Industrial Research Chair in Programmable Silicon.
%
\bibliographystyle{abbrv}
\bibliography{fpga2014paper}

\balancecolumns
% That's all folks!
\end{document}

### 2014-11-11 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
updating with new material in section 3. section 4.1 and 4.2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4501 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

--- 2014-11-09 doc/journal/fpgapaperfinal/fpga2014paper.tex
@@
 \usepackage{graphicx}
+\usepackage{subfigure}
+
+\usepackage{epstopdf}
+\usepackage{enumitem}
 %\linespread{0.94}
@@
 
-Three parameters were added to VPR: \textit{\% wires cut}, \textit{increased delay} and \textit{number of cuts}.
-These three parameters describe the interposer portion of the 2.5D FPGA as detailed below.
+The following parameters were added to VPR: \textit{number of cuts}, \textit{\% wires cut}, \textit{increased delay}, and \textit{interposer-routing interface}.
+These parameters describe the interposer portion of the 2.5D FPGA as detailed below.
+
+%	- number of cuts
+\subsection{Number of cuts}
+\textit{Number of cuts} describes how many cuts were made to the interposer-based FPGA versus a monolithic die.
+If \textit{number of cuts} equals 1 then there are 2 dies, and so on.
+We investigate values of 1 and 3 (2 and 4 dies).
+The Virtex 7 family has members with \textit{number of cuts} $=$ 2 and 3.
+Figure \ref{fig:fpga} shows a sample architecture with one cutline.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{fpga.eps}
+\caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
+\label{fig:fpga}
+\end{figure}
 
@@
 
-%	- number of cuts
-\subsection{Number of cuts}
-\textit{Number of cuts} describes how many cuts were made to the interposer-based FPGA versus a monolithic die.
-If \textit{number of cuts} equals 1 then there are 2 dies, and so on.
-We investigate values of this parameter between 1 and 3 (between 2 and 4 dies), reflecting commercial practice: the Virtex 7 family has members with \textit{number of cuts} $=$ 2 and 3.
-Figure \ref{fig:fpga} shows a sample architecture with one cutline.
+%   - Interposer-Routing interface
+\subsection{Interposer-Routing interface}
+In a simple architecture, the vertical routing wires that can not connect to an interposer wire are dangling.
+Since interposer wires are scarce, we want to be able to study architectures in which the interposer wires are more accessible.
+To do so, we have enhanced the router to understand the following interposer-routing interfaces.
+
+\begin{enumerate}[label=(\alph*)]
+\item Fan-in Transfer for Interposer Wires\\
+For an architecture with \textit{\% wires cut} $=0$, all vertical routing wires have a corresponding interposer wire that connects them to the adjacent die.
+Practical interposers, however, can only accommodate a fraction of all vertical routing wires.
+So, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
+This allows for additional flexibility in routing.
+As an example, if \textit{\% wires cut} $=75\%$, 3 of every 4 interposer wires are cut.
+Therefore, we place a 4-to-1 multiplexer at the input of the one interposer wire that is not cut, and we connect the 3 vertical routing wires that do not have a corresponding interposer wire to the multiplexer.
+
+\item Fan-out Transfer for Interposer Wires\\
+We allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
+
+\item Bidirectional Interposer Wires\\
+We allow interposer wires to be driven from one end or the other.
+\end{enumerate}
 
@@
 \centering
-\includegraphics[width=\linewidth]{fpga.eps}
-\caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
-\label{fig:fpga}
+\includegraphics[width=\linewidth]{architecture_params.eps}
+\caption{For channel width of 4 and 75\% wires cut, only 1 interposer wire exists (in green) and 3 interposer wires are cut (in red).
+(a)~Fan-in Transfer for Interposer Wires allows red routing wires in die 1  to cross the interposer layer.
+(b)~Fan-out Transfer for Interposer Wires allows the output of the interposer wire to drive wires in other tracks in die 2.}
+\label{fig:architecture_params}
 \end{figure}
@@
 \subsection{Implementation}
-To model an interposer-based FPGA in VPR the Routing Resource Graph (\textit{rrgraph}) must be modified.
-The rrgraph is the data structure that defines all the available routing wires and switches in the FPGA, as well the delay of each.
+To model an interposer-based multi-FPGA system in the CAD tool, we use the existing VPR infrastructure for monolithic FPGAs and make several modifications to it.
+
+First, The presence of multiple dice in an interposer-based FPGA was modeled by creating horizontal cuts in the FPGA.
+The \textit{number of cuts} is specified by the user, and we use that to determine the y-location of the cuts.
+The cutlines ``snap to the grid" such that the height of all dies are equal and no cutline cuts through a block.
+For instance, a RAM block of height 4 must remain as a unit and can not be sliced by a cutline.
+
+Second, knowing the location of the cutlines, the placer uses the \textit{(x,y)} coordinate of any pin or block to determine which die it belongs to.
+The placer uses this information as well as \textit{interposer delay} to better decide where to place each block in order to minimize the cost of inter-die signal transfer as described in Section \ref{cad_enh_placer_subsection}.
+
+Third, the Routing Resource Graph (\textit{rrgraph}) must be modified.
+The rrgraph is the data structure that defines all the available routing wires and switches in the FPGA, as well as the delay of each.
 Given a suitable rrgraph, the VPR router can implement circuits in the desired FPGA, and the VPR timing analyzer can estimate their delay.
-
-The presence of multiple dice in an interposer-based FPGA was modeled by creating horizontal cuts in the FPGA, which are equally spaced vertically.
-Every wire that crosses one of these cuts in fact passes through the interposer.
 In the experiments below, we use an architecture that has only unidirectional wires~\cite{unidirectional}, as these are the dominant routing architecture in modern FPGAs.
 Such wires can only be driven at their beginning.
-To control the wiring capacity between dies \textit{\% wires cut} wires of each channel segment crossing a cutline have their connections to wires and block inputs on the die opposite to their starting point removed, and the wires which aren't cut have \textit{increased delay} added to their connections which cross the cutline.
-Note that the combination of a unidirectional routing architecture and a silicon interposer results in some blocks near the cutline having reduced routing connectivity as some of the routing wires driving their inputs are disconnected from the (single) wire driver, when that driver is on the other side of the cutline and the wire is one of the cut wires.
 
-Figure \ref{fig:implementation} illustrates the approach used.
+Previous work by Andre et al ~\cite{hahn2014cad} modified the rrgraph by directly removing \textit{\% wires cut} percent of edges from the graph at nodes located at the cutline and increased the switch delay of the remaining edges of the nodes at the cutline.
+While this is a reasonable method to model such architectures, it makes it difficult to experiment with different routing-interposer interfaces such as placing input muxes for interposer wires.
+Therefore, we modify the rrgraph differently to more accurately reflect the interposer layer:
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{implementation.eps}
-\caption{Illustration of the approach used to simulate crossing the interposer.
-The red dashed line indicates the interposer, the black arrows indicate the direction of the wires.
-Blue arrows indicate normal connections from the wires to the CLBs.
-The red arrows indicate removed connections and green arrows indicate connections with increased delay.}
-\label{fig:implementation}
+For every vertical routing wire in the channel, we add a new routing node (``\textit{interposer node}") into the rrgraph.
+Then, we determine routing nodes (e.g.
+vertical routing wires) that cross a cutline in the rrgraph, and we replace them with 2 routing nodes: one that is above the cutline and one that is below the cutline.
+We properly transfer fan-ins and fan-outs of the original node onto the two newly created nodes and we connect the two nodes using an interposer node.
+Finally, we remove \textit{all} edges from \textit{\%wires cut} percent of interposer nodes and increase the switch delay of output edges of the remaining interposer nodes by \textit{increased delay}.
+Figure~\ref{fig:rrgraph_ops} shows an example of these operations.
+
+\begin{figure}[!t]
+\hfill
+\subfigure[Chip View]{\includegraphics[width=4.9cm]{a_b.eps}}
+\hfill
+\subfigure[``rrgraph" View]{\includegraphics[width=3.3cm]{a_b_prime.eps}}
+\hfill
+\caption{(a) Left: regular routing wire connectivity.
+Right: replacing the long wire with two wires and connecting them via an interposer wire.
+(b) Left: rrgraph for regular connectivity.
+Right: rrgraph after modifications}
+\label{fig:rrgraph_ops}
 \end{figure}
 
-For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the "flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}).
-The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices.
-It consists of logic clusters with 10 fracturable 6-LUTs per block (N=10, k=6), and also includes 32kb RAM blocks and DSP blocks configurable to perform 9x9, 18x18 or 36x36 multipliers.
-The delay values in the architecture are taken from 40 nm circuit simulations and 40 nm commercial FPGAs.
-It uses unidirectional routing, with all wire segments having length L = 4.
+As illustrated in Figure~\ref{fig:vpr_interposer_gui}, no vertical routing wires cross a cutline boundary and a fraction of them connect to other dies via interposer wires.
+
+%%%
+%\begin{figure}[!t]
+%\subfigure[Full Chip View]{\includegraphics[width=\textwidth]{vpr_gui_full.eps}}
+%\subfigure[Connectivity at the cutline]{\includegraphics[width=\textwidth]{vpr_gui_zoom.eps}}
+%\caption{(a) Left: regular routing wire connectivity.
+Right: replacing the long wire with two wires and connecting them via an interposer wire.
+(b) Left: rrgraph for regular connectivity.
+Right: rrgraph after modifications}
+%\label{fig:vpr_interposer_gui}
+%\end{figure}
+
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=\textwidth]{vpr_gui.eps}
+\caption{VPR Graphical User Interface.
+Left:Full chip view.
+Right:Connectivity at the cutline.
+Interposer wires are drawn in red.
+Due to limited number of interposer wires, some vertical routing wires are not able to connect to any interposer wire and are left dangling.
+Routing wires do not crosses the cutline.}
+\label{fig:vpr_interposer_gui}
+\end{figure*}
 
@@
 
-Once the routing resource graph was modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
-Placement, however, is crucial in mitigating the impact of the reduced wiring and increased delay when crossing between dice; a good placement should minimize the number of signals crossing between dice, particularly time-critical ones.
-We investigated several enhancements to VPR's placement cost function to improve result quality.
+\subsection{Placer Optimization}
+\label{cad_enh_placer_subsection}
+Placement is crucial in mitigating the impact of the reduced wiring and increased delay when crossing between dice; a good placement should minimize the number of signals crossing between dice, particularly time-critical ones.
+VPR uses 2 different costs as the metrics for its placer algorithm: the timing cost and the bounding box (wiring) cost.
+We modified the placer's bounding box and timing costs to account for the increased latency to cross the interposer and for the reduced wire capacity close to the cutlines, as the wire availability becomes more sparse.
 
-VPR uses 2 different costs as the metrics for its placer algorithm: the timing cost and the bounding box (wiring) cost.
+%   - Delay
+\subsubsection{Placer timing cost}
 The usual VPR timing cost is a (criticality-weighted) summation of the estimated delay (given the current placement) of every connection required by the circuit~\cite{timing2000}:
@@
 \end{equation}
+where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
 
-where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
+This cost function assumes that the FPGA is homogeneous and consequently the delay between 2 points $(x_1,y_1)$ and $(x_2,y_2)$ only depends on $(\Delta x,\Delta y)$.
+This is obviously not true for interposer-based FPGAs, as the cutlines make them heterogeneous in the \textit{y} direction.
+To solve this problem and improve the quality of the results, an extra term was added to the delay function.
+The delay function becomes:
+
+\begin{equation} \label{eq:timing_cost}
+\begin{split}
+delay(i,j) =& delay(\Delta x_{ij}, \Delta y_{ij}) + \\
+&times\_crossed(i,j) \times delay\_increase
+\end{split}
+\end{equation}
+
+where $times\_crossed(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
+
+%   - Wiring / cut cost
+\subsubsection{Placer wiring cost}
 The bounding box cost estimates the amount of wiring required for a net, based on the number of pins and size of the net's bounding box. VPR's original formulation is~\cite{betz1997vpr}:
@@
 
-We modified both the placer's bounding box and timing costs to account for the increased latency to cross the interposer and for the reduced wire capacity close to the cutlines, as the wire availability becomes more sparse.
-
-%   - Delay
-\subsection{Placer timing cost}
-The standard placer timing cost in VPR assumes that the FPGA is homogeneous and consequently the delay between 2 points $(x_1,y_1)$ and $(x_2,y_2)$ only depends on $(\Delta x,\Delta y)$.
-This is obviously not true for interposer-based FPGAs, as the cutlines make them heterogeneous in the \textit{y} direction.
-
-To solve this problem and improve the quality of the results, an extra term was added to the delay function. The delay function becomes:
-
-\begin{equation} \label{eq:timing_cost}
-\begin{split}
-delay(i,j) =& delay(\Delta x_{ij}, \Delta y_{ij}) + \\
-&times\_crossed(i,j) \times delay\_increase
-\end{split}
-\end{equation}
-
-where $times\_crossed(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
-
-%   - Wiring / cut cost
-\subsection{Placer wiring cost}
 VPR's wiring cost also considers the FPGA to be homogeneous, and uses only the number of nets, the size of the net's bounding box and the average number of wires per channel to calculate the cost.
@@
 
-We tested several different \textit{cut\_cost} formulations, as well as different weighting (C values) for each. For all of the formulations the variable $C'$ was defined as:
-
-\begin{equation} \label{eq:cprime}
-\begin{split}
-C' = \frac{C \times ratio\_wires\_cut}{avg\_chany\_W(n)}
-\end{split}
-\end{equation}
-
-where $ratio\_wires\_cut$ is the ratio of wires cut at the cutline.
-This formulation of $C'$ ensures that when we choose a \textit{C} value of 1, the new term cost term is of roughly the same magnitude as $wire\_cost_{orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
-
-The first \textit{cut\_cost} term we tested was:
-
-\begin{equation} \label{eq:cost0}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times times\_crossed(n)
-\end{equation}
-
-which penalizes the net according to the number of cutlines the bounding box spans; this directly penalizes each signal crossing between dice.
-Surprisingly this extra term did not result in significant improvements to the quality of results and in some cases even made them worse.
-We believe this is due to the discontinuous, non-smooth, nature of this cost function -- it does not guide placement optimization well as a bounding box shrinks toward a cutline.
-Placement changes that make bounding boxes "almost not cross" a cutline are not given any gain; only a sudden change in the placement that moves a bounding box entirely to one side or the other of the cutline yields a cost reduction.
-This insight led us to the other proposed functions, which give gradual gains as bounding boxes change in their size and come closer to avoiding a cutline crossing.
-
-The other tested terms were:
-
-%add C' x times_crossed(i)
-
-\begin{equation} \label{eq:cost1}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbWidth(n) \times times\_crossed(n)
-\end{equation}
+In~\cite{hahn2014cad}, Andre et al.
+tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
+They found that the following \textit{cut\_cost} results in the best quality of results in terms of minimum routable channel width and critical path delay.
 
@@
 
-\begin{equation} \label{eq:cost4}
+where 
+\begin{equation} \label{eq:cprime}
 \begin{split}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times minDist(i) \times times\_crossed(n) +\\
-C' \times times\_crossed(n)
+C' = \frac{ratio\_wires\_cut}{avg\_chany\_W(n)}
 \end{split}
@@
 
-\begin{equation} \label{eq:cost5}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbHeight(n)
+\textit{bbHeight(n)} is the height of the bounding box of the net \textit{n} and $ratio\_wires\_cut$ is the ratio of wires cut at the cutline.
+This formulation of $C'$ ensures that the new cost term is of roughly the same magnitude as $wire\_cost_{orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
+
+\subsection{Router Optimization}
+Once the Routing Resource Graph was modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
+However, the router's \textit{lookahead algorithm} which decides the front on which the router should expand the current search to the destination was not interposer-aware.
+Hence it took a long time to converge to a good routing decision.
+
+The VPR router uses the following to calculate a cost for any routing element on the route from \textit{i} to \textit{j}:
+\begin{equation} \label{eq:routing_cost}
+\begin{split}
+Cost(k) = Criticality(i,j) \times Delay_{Elmore}(k, topology) + \\ [1-Criticality(i,j)] \times Congestion(k)
+\end{split}
 \end{equation}
 
-where \textit{bbWidth(n)} and \textit{bbHeight(n)} are the width and height of the bounding box of the net \textit{n}, respectively, and \textit{minDist(n)} is the minimum distance from the top or bottom of the bounding box to a cutline.
+Given this cost function, router's lookahead algorithm determines the total cost of the route from \textit{i} to \textit{j} when routing node \textit{n} is used:
+
+\begin{equation} \label{eq:lookahead_cost}
+\begin{split}
+Total\_Cost(n) = \sum_{\forall l \in RT(i,n)} Cost(l) + \alpha \times ExpectedCost(n,j)
+\end{split}
+\end{equation}
+
+where \textit{RT(i,n)} is the set of all routing nodes used so far from \textit{i} to \textit{n} and $\alpha$ is a constant number that determines the aggressiveness of the directed search.
+$\alpha$ of 0 results in a purely Breadth First Search of the routing graph from \textit{i} to \textit{j}.
+VPR's original $ExpectedCost(n,j)$ function is~\cite{betz1997vpr}:
+
+\begin{equation} \label{eq:expected_routing_cost}
+\begin{split}
+ExpectedCost(n,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(n,j);
+\end{split}
+\end{equation}
+
+$T_{delay}$ is an estimate of the delay that the router predicts it will see on the route from \textit{n} to \textit{j}.
+Since (\ref{eq:expected_routing_cost}) assumes a homogeneous FPGA, $T_{delay}$ does not include the additional delay some routes experience by going through the interposer.
+Thus, for interposer-based architectures, we have replaced $T_{delay}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{delay}$ to include the delay of expected interposer hops that are on the way form \textit{n} to \textit{j}.
+\begin{equation} \label{eq:tdelay}
+\begin{split}
+T^\prime_{delay} = T_{delay} + T_{interposer} \times ExpectedInterposerHops
+\end{split}
+\end{equation}
+
+The interposer-aware router lookahead is able to make routing decisions quicker given the large delay difference of a path that crosses the interposer as opposed to a path that doesn't.
 
@@
 \label{sec:CADeffect}
+For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the "flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}).
+The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices.
+It consists of logic clusters with 10 fracturable 6-LUTs per block (N=10, k=6), and also includes 32kb RAM blocks and DSP blocks configurable to perform 9x9, 18x18 or 36x36 multipliers.
+The delay values in the architecture are taken from 40 nm circuit simulations and 40 nm commercial FPGAs.
+It uses unidirectional routing, with all wire segments having length L = 4.
 

### 2014-11-11 doc/journal/fpgapaperfinal/FPGAPaper/rrgraph_before.tex
updating with new material in section 3. section 4.1 and 4.2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4501 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

% Drawing a graph using the PG 3.0 graphdrawing library
% Author: Mark Wibrow
\documentclass[tikz,border=10pt]{standalone}
\usetikzlibrary{graphdrawing}
\usetikzlibrary{graphs}
\usegdlibrary{trees}
\begin{document}
\begin{tikzpicture}[>=stealth, every node/.style={circle, draw, minimum size=0.75cm}]
\graph [tree layout, grow=down, fresh nodes, level distance=0.5in, sibling distance=0.5in]
    {
        a -> { 
          b -> { c, d }
        } 
    };
\end{tikzpicture}

\begin{tikzpicture}[>=stealth, every node/.style={circle, draw, minimum size=0.75cm}]
\graph [tree layout, grow=down, fresh nodes, level distance=0.5in, sibling distance=0.5in]
    {
        a -> { 
          b -> { d, 
                 e -> { f -> {c} } }
        } 
    };
\end{tikzpicture}

\end{document}

### 2014-11-11 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
minor fixes
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4502 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-Previous work by Andre et al ~\cite{hahn2014cad} modified the rrgraph by directly removing \textit{\% wires cut} percent of edges from the graph at nodes located at the cutline and increased the switch delay of the remaining edges of the nodes at the cutline.
+Previous work by Pereira et al ~\cite{hahn2014cad} modified the rrgraph by directly removing \textit{\% wires cut} percent of edges from the graph at nodes located at the cutline and increased the switch delay of the remaining edges of the nodes at the cutline.
 While this is a reasonable method to model such architectures, it makes it difficult to experiment with different routing-interposer interfaces such as placing input muxes for interposer wires.
@@
 Interposer wires are drawn in red.
-Due to limited number of interposer wires, some vertical routing wires are not able to connect to any interposer wire and are left dangling.
-Routing wires do not crosses the cutline.}
+Due to limited capacity of the interposer, some vertical routing wires are left dangling.
+Routing wires do not cross the cutline unless via a connection to an interposer wire.}
 \label{fig:vpr_interposer_gui}
@@
 
-In~\cite{hahn2014cad}, Andre et al.
-tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
+In~\cite{hahn2014cad}, Pereira et al tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
 They found that the following \textit{cut\_cost} results in the best quality of results in terms of minimum routable channel width and critical path delay.

### 2014-11-15 src/timing_analyzer/ParallelTimingAnalyzer.cpp
Move locks into TA object.

@@ -6,11 +6,39 @@
 #define DEFAULT_CLOCK_PERIOD 1.0e-9
 
 void ParallelTimingAnalyzer::calculate_timing(TimingGraph& timing_graph) {
+    create_locks(timing_graph);
 #pragma omp parallel
     {
+
+        //TODO: Lock init and pre traversal can occur in parallel
+        init_locks();
         pre_traversal(timing_graph);
+
+        //TODO: While forward and backward tranversals work on the same edges
+        //and nodes, they work on independant data sets (i.e. arrival time for forward and required time for backward)
+        //so they can occur in parallel
         forward_traversal(timing_graph);
         backward_traversal(timing_graph);
+
+        cleanup_locks();
+    }
+}
+
+void ParallelTimingAnalyzer::create_locks(TimingGraph& tg) {
+    node_locks_ = std::vector<omp_lock_t>(tg.num_nodes());
+}
+
+void ParallelTimingAnalyzer::init_locks() {
+#pragma omp for
+    for(int i = 0; i < (int) node_locks_.size(); i++) {
+        omp_init_lock(&node_locks_[i]);
+    }
+}
+
+void ParallelTimingAnalyzer::cleanup_locks() {
+#pragma omp for
+    for(int i = 0; i < (int) node_locks_.size(); i++) {
+        omp_destroy_lock(&node_locks_[i]);
     }
 }
 
@@ -86,6 +114,11 @@ void ParallelTimingAnalyzer::pre_traverse_node(TimingGraph& tg, TimingGraph::Nod
 }
 
 void ParallelTimingAnalyzer::forward_traverse_node(TimingGraph& tg, TimingGraph::NodeId node_id) {
+    //From node was updated by the previous level (i.e. is read-only) so we don't need to synchronize
+    //access to it.
+    //
+    //However, multiple edges may connect to to_node, so we multiple threads may try to update it.
+    //To avoid a race condition we must synchronize access to it.
     const TimingNode& from_node = tg.node(node_id);
     float from_arrival_time = from_node.arrival_time().value();
 
@@ -99,7 +132,8 @@ void ParallelTimingAnalyzer::forward_traverse_node(TimingGraph& tg, TimingGraph:
 
         {
             //Lock the to_node
-            to_node.lock();
+            //to_node.lock();
+            omp_set_lock(&node_locks_[to_node_id]);
 
             //TODO: Generalize this to support a more general Time class (e.g. vector of timing corners)
[…]

### 2014-11-15 src/timing_analyzer/ParallelTimingAnalyzer.hpp

@@ -31,6 +31,11 @@ class ParallelTimingAnalyzer : public SerialTimingAnalyzer {
         void pre_traverse_node(TimingGraph& tg, TimingGraph::NodeId node_id, int level_idx);
         void forward_traverse_node(TimingGraph& tg, TimingGraph::NodeId node_id);
         void backward_traverse_node(TimingGraph& tg, TimingGraph::NodeId node_id);
+
+        void create_locks(TimingGraph& tg);
+        void init_locks();
+        void cleanup_locks();
+        std::vector<omp_lock_t> node_locks_;
 };
 
 #endif

### 2014-11-16 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Fixed all issues based on comments from Vaughn in the Nov 11 meeting. Everything is updated up to and including section 4.2 (router optimizations)
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4504 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 %   - Define parameters
-To properly model a silicon interposer FPGA, the popular FPGA exploration toolset, Verilog-to-Routing (VTR)~\cite{vtr2012}, was used.
-The logic synthesis portions of the flow (ODIN II and  ABC) were left untouched, while the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) was modified to model and optimize for interposer-based FPGAs.
-The modifications were made in such a way that they require no changes to any of the input files, so one can experiment with interposer-based FPGAs with any existing benchmark circuits and any existing VPR-format FPGA architecture description simply by specifying appropriate command-line parameters.
+To properly model a silicon interposer FPGA, we use version 7.0 of the popular FPGA exploration toolset, Verilog-to-Routing (VTR)~\cite{luu2014vtr}.
+The logic synthesis portions of the flow (ODIN II and  ABC) are left untouched, while we modify the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) to model and optimize for interposer-based FPGAs.
+The modifications are made in such a way that they require no changes to any of the input files, so one can experiment with interposer-based FPGAs with any existing benchmark circuits and any existing VPR-format FPGA architecture description simply by specifying appropriate command-line parameters.
 
-The following parameters were added to VPR: \textit{number of cuts}, \textit{\% wires cut}, \textit{increased delay}, and \textit{interposer-routing interface}.
+We add the following parameters to VPR: \textit{number of cuts}, \textit{\% wires cut}, \textit{increased delay}, and \textit{interposer-routing interface}.
 These parameters describe the interposer portion of the 2.5D FPGA as detailed below.
@@
 \subsection{Number of cuts}
-\textit{Number of cuts} describes how many cuts were made to the interposer-based FPGA versus a monolithic die.
-If \textit{number of cuts} equals 1 then there are 2 dies, and so on.
-We investigate values of 1 and 3 (2 and 4 dies).
+\textit{Number of cuts} describes how many cuts are made to the interposer-based FPGA versus a monolithic die.
+If \textit{number of cuts} equals 1 then there are 2 dice, and so on.
+We investigate values of 1 and 3 (2 and 4 dice).
 The Virtex 7 family has members with \textit{number of cuts} $=$ 2 and 3.
@@
 Since interposer wires are scarce, we want to be able to study architectures in which the interposer wires are more accessible.
-To do so, we have enhanced the router to understand the following interposer-routing interfaces.
+To do so, we have enhanced the router to support the following interposer-routing interface options.
 
@@
 Practical interposers, however, can only accommodate a fraction of all vertical routing wires.
-So, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
+Hence, when this variable is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
 This allows for additional flexibility in routing.
@@
 \item Fan-out Transfer for Interposer Wires\\
-We allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
+When this option is \textit{on}, we allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
 
 \item Bidirectional Interposer Wires\\
-We allow interposer wires to be driven from one end or the other.
+Dominant routing architectures in modern FPGAs use unidirectional wires~\cite{unidirectional, lewis2005stratix}.
+Such wires can only be driven at one point, normally their start point.
+When this option is \textit{on}, we allow interposer wires to be driven from one end or the other by adding tri-state buffers at both ends.
 \end{enumerate}
@@
 \includegraphics[width=\linewidth]{architecture_params.eps}
-\caption{For channel width of 4 and 75\% wires cut, only 1 interposer wire exists (in green) and 3 interposer wires are cut (in red).
-(a)~Fan-in Transfer for Interposer Wires allows red routing wires in die 1  to cross the interposer layer.
+\caption{For a channel width of 4 and 75\% wires cut, only 1 interposer wire exists (in green) and 3 interposer wires are cut (in the red).
+(a)~Fan-in Transfer for Interposer Wires allows red routing wires in die 1  to drive a mux that allows them to cross the interposer layer.
 (b)~Fan-out Transfer for Interposer Wires allows the output of the interposer wire to drive wires in other tracks in die 2.}
@@
 \end{figure}
+
+These options are independent and can be combined to achieve more flexible routing.
+Enabling any of these options requires adding specific circuitry (such as multiplexers) on the dice which causes very small area and delay increase that we neglect in our experiments.
 
@@
 
-First, The presence of multiple dice in an interposer-based FPGA was modeled by creating horizontal cuts in the FPGA.
+First, The presence of multiple dice in an interposer-based FPGA is modeled by creating horizontal cuts in the FPGA.
 The \textit{number of cuts} is specified by the user, and we use that to determine the y-location of the cuts.
-The cutlines ``snap to the grid" such that the height of all dies are equal and no cutline cuts through a block.
-For instance, a RAM block of height 4 must remain as a unit and can not be sliced by a cutline.
+The cutlines ``snap to the grid" such that the height of all dice are equal and no cutline cuts through a block.
+For instance, a RAM block of height 4 must remain a unit and can not be sliced by a cutline.
 
-Second, knowing the location of the cutlines, the placer uses the \textit{(x,y)} coordinate of any pin or block to determine which die it belongs to.
-The placer uses this information as well as \textit{interposer delay} to better decide where to place each block in order to minimize the cost of inter-die signal transfer as described in Section \ref{cad_enh_placer_subsection}.
+Second, knowing the location of the cutlines, the placer uses the \textit{(x,y)} coordinate of any pin or block to determine the die to which it belongs.
+The placer uses this information as well as \textit{interposer delay} to better decide where to place each netlist block in order to minimize the cost of inter-die signal transfer as described in Section \ref{cad_enh_placer_subsection}.
 
 Third, the Routing Resource Graph (\textit{rrgraph}) must be modified.
-The rrgraph is the data structure that defines all the available routing wires and switches in the FPGA, as well as the delay of each.
+The rrgraph is the data structure that defines all the available routing wires (\textit{rrnodes}) and switches in the FPGA, as well as the delay of each~\cite{betz1999architecture}.
 Given a suitable rrgraph, the VPR router can implement circuits in the desired FPGA, and the VPR timing analyzer can estimate their delay.
-In the experiments below, we use an architecture that has only unidirectional wires~\cite{unidirectional}, as these are the dominant routing architecture in modern FPGAs.
-Such wires can only be driven at their beginning.
+In the experiments below, we use an architecture that has only unidirectional wires.
 
-Previous work by Pereira et al ~\cite{hahn2014cad} modified the rrgraph by directly removing \textit{\% wires cut} percent of edges from the graph at nodes located at the cutline and increased the switch delay of the remaining edges of the nodes at the cutline.
+Previous work by Pereira et al \cite{hahn2014cad} modeled the interposer indirectly: they did not introduce new rrnodes in the rrgraph to represent the interposer wires.
+Rather, they found the rrnodes that cross a cutline.
+Then they either removed all outgoing edges of these rrnodes (if the routing node was part of the \textit{\% wires cut}), or, they increased the switch delay of the rrnode fan-outs (if the wire was not cut) to account for the additional interposer delay.
+
 While this is a reasonable method to model such architectures, it makes it difficult to experiment with different routing-interposer interfaces such as placing input muxes for interposer wires.
-Therefore, we modify the rrgraph differently to more accurately reflect the interposer layer:
+Therefore, we modify the rrgraph differently to more accurately reflect the interposer layer directly:
+
+\begin{figure}[!t]
+\subfigure[]{\includegraphics[width=0.24\linewidth]{a.eps}}
+\subfigure[]{\includegraphics[width=0.24\linewidth]{b.eps}}
+\subfigure[]{\includegraphics[width=0.24\linewidth]{a_prime.eps}}
+\subfigure[]{\includegraphics[width=0.24\linewidth]{b_prime.eps}}
+\caption{(a) Regular routing wire connectivity.
+(b) Replacing the long wire with two wires and connecting them via an interposer wire.
+(c) rrgraph for regular connectivity.
+(d) rrgraph after modifications.}
+\label{fig:rrgraph_ops}
+\end{figure}
+
+\begin{figure*}[!thpb]
+\subfigure[]{\includegraphics[width=0.6\textwidth]{gui_1.eps}}
+\subfigure[]{\includegraphics[width=0.4\textwidth]{gui_2.eps}}
+\caption{VPR Graphical User Interface. (a) Full chip view. (b) Connectivity at the cutline. Due to the limited signal capacity of the interposer, some vertical routing wires are left dangling.}
+\label{fig:vpr_interposer_gui}
+\end{figure*}
 
 For every vertical routing wire in the channel, we add a new routing node (``\textit{interposer node}") into the rrgraph.
-Then, we determine routing nodes (e.g.
-vertical routing wires) that cross a cutline in the rrgraph, and we replace them with 2 routing nodes: one that is above the cutline and one that is below the cutline.
+Then, we determine rrnodes (e.g.
+vertical routing wires) that cross a cutline in the rrgraph, and we replace them with 2 rrnodes: one that is above the cutline and one that is below the cutline.
 We properly transfer fan-ins and fan-outs of the original node onto the two newly created nodes and we connect the two nodes using an interposer node.
@@
 
-\begin{figure}[!t]
-\hfill
-\subfigure[Chip View]{\includegraphics[width=4.9cm]{a_b.eps}}
-\hfill
-\subfigure[``rrgraph" View]{\includegraphics[width=3.3cm]{a_b_prime.eps}}
-\hfill
-\caption{(a) Left: regular routing wire connectivity.
-Right: replacing the long wire with two wires and connecting them via an interposer wire.
-(b) Left: rrgraph for regular connectivity.
-Right: rrgraph after modifications}
-\label{fig:rrgraph_ops}
-\end{figure}
-
 As illustrated in Figure~\ref{fig:vpr_interposer_gui}, no vertical routing wires cross a cutline boundary and a fraction of them connect to other dies via interposer wires.
-
-%%%
-%\begin{figure}[!t]
-%\subfigure[Full Chip View]{\includegraphics[width=\textwidth]{vpr_gui_full.eps}}
-%\subfigure[Connectivity at the cutline]{\includegraphics[width=\textwidth]{vpr_gui_zoom.eps}}
-%\caption{(a) Left: regular routing wire connectivity.
-Right: replacing the long wire with two wires and connecting them via an interposer wire.
-(b) Left: rrgraph for regular connectivity.
-Right: rrgraph after modifications}
-%\label{fig:vpr_interposer_gui}
-%\end{figure}
-
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=\textwidth]{vpr_gui.eps}
-\caption{VPR Graphical User Interface.
-Left:Full chip view.
-Right:Connectivity at the cutline.
-Interposer wires are drawn in red.
-Due to limited capacity of the interposer, some vertical routing wires are left dangling.
-Routing wires do not cross the cutline unless via a connection to an interposer wire.}
-\label{fig:vpr_interposer_gui}
-\end{figure*}
 
@@
 This is obviously not true for interposer-based FPGAs, as the cutlines make them heterogeneous in the \textit{y} direction.
-To solve this problem and improve the quality of the results, an extra term was added to the delay function.
-The delay function becomes:
+To solve this problem and improve the quality of the results, we add an extra term to the delay function:
 
@@
 Finally, $q(n)$ is a function obtained from \cite{icann} which models the fact that bounding boxes underpredict the required routing for high fanout nets.
-$q(n)$ slowly increases with the fanout of net \textit{n}, rising from 1 for nets with 3 or fewer terminals to 2.79 for nets with 50 terminals.
+$q(n)$ slowly increases with the fanout of net \textit{n}.
 
 VPR's wiring cost also considers the FPGA to be homogeneous, and uses only the number of nets, the size of the net's bounding box and the average number of wires per channel to calculate the cost.
-Thus, to account for the reduced connectivity near the cutlines an extra cost term, \textit{cut\_cost}, was created.
+Thus, to account for the reduced connectivity near the cutlines we create an extra cost term, \textit{cut\_cost}.
 This new cost is added to (\ref{eq:bbcost}) to create the total wiring cost.
@@
 \subsection{Router Optimization}
-Once the Routing Resource Graph was modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
-However, the router's \textit{lookahead algorithm} which decides the front on which the router should expand the current search to the destination was not interposer-aware.
-Hence it took a long time to converge to a good routing decision.
+Once the Routing Resource Graph is modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
+The router starts at a net source node and expands by discovering adjacent routing nodes until it reaches the destination node.
+At any point, the router knows the cost of the path taken to the current point and it uses a \textit{lookahead cost function} to estimate the resource and delay cost from the current node to the destination.
+Using this lookahead cost, the router can properly sort routing alternatives.
+The lookahead cost function in VPR is not aware of the interposer, and therefore it takes a long time to converge to a routing decision.
+To make the lookahead aware of the interposer, we modify the lookahead cost function by adding terms that account for extra \textit{interposer delay} and extra \textit{interposer nodes} that may exist on the route from the current node to the destination node.
 
-The VPR router uses the following to calculate a cost for any routing element on the route from \textit{i} to \textit{j}:
+The VPR router uses the following to calculate a cost for a routing node \textit{r} on the route from \textit{i} to \textit{j}:
 \begin{equation} \label{eq:routing_cost}
 \begin{split}
-Cost(k) = Criticality(i,j) \times Delay_{Elmore}(k, topology) + \\ [1-Criticality(i,j)] \times Congestion(k)
+Cost(r) = Criticality(i,j) \times Delay_{Elmore}(r, topology) + \\ [1-Criticality(i,j)] \times Congestion(r)
 \end{split}
@@
 
-Given this cost function, router's lookahead algorithm determines the total cost of the route from \textit{i} to \textit{j} when routing node \textit{n} is used:
+Given this cost function, the router's lookahead algorithm determines the total cost of the route from \textit{i} to \textit{j} when routing node \textit{r} is used:
 
@@
 \begin{split}
-Total\_Cost(n) = \sum_{\forall l \in RT(i,n)} Cost(l) + \alpha \times ExpectedCost(n,j)
+Total\_Cost(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectedCost(r,j)
 \end{split}
@@
 
-where \textit{RT(i,n)} is the set of all routing nodes used so far from \textit{i} to \textit{n} and $\alpha$ is a constant number that determines the aggressiveness of the directed search.
-$\alpha$ of 0 results in a purely Breadth First Search of the routing graph from \textit{i} to \textit{j}.
-VPR's original $ExpectedCost(n,j)$ function is~\cite{betz1997vpr}:
+where \textit{RT(i,r)} is the set of all routing nodes used so far from \textit{i} to \textit{r} and $\alpha$ is a constant number that determines the aggressiveness of the directed search.
+$\alpha$ is 1.2 by default in VPR.
+VPR's original $ExpectedCost(n,j)$ function is~\cite{betz1999architecture}:
 
@@
 \begin{split}
-ExpectedCost(n,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(n,j);
+ExpectedCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(r,j);
 \end{split}
@@
 
-$T_{delay}$ is an estimate of the delay that the router predicts it will see on the route from \textit{n} to \textit{j}.
+$T_{delay}$ is an estimate of the delay that the router predicts it will see on the route from \textit{r} to \textit{j}.
 Since (\ref{eq:expected_routing_cost}) assumes a homogeneous FPGA, $T_{delay}$ does not include the additional delay some routes experience by going through the interposer.
-Thus, for interposer-based architectures, we have replaced $T_{delay}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{delay}$ to include the delay of expected interposer hops that are on the way form \textit{n} to \textit{j}.
+Thus, for interposer-based architectures, we replace $T_{delay}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{delay}$ to include the delay of expected interposer hops from \textit{r} to \textit{j}.
 \begin{equation} \label{eq:tdelay}
@@
 
-The interposer-aware router lookahead is able to make routing decisions quicker given the large delay difference of a path that crosses the interposer as opposed to a path that doesn't.
+In a similar way, we modify the $ExpectedCongestionCost(r,j)$ in (\ref{eq:expected_routing_cost}) to account for the new interposer routing nodes.
+
+The interposer-aware router lookahead is able to make routing decisions more quickly given the large delay difference of a path that crosses the interposer as opposed to a path that doesn't.
 

### 2014-11-16 src/timing_graph/TimingNode.hpp
Remove lock from TimingNode

@@ -29,12 +29,8 @@ std::istream& operator>>(std::istream& os, TN_Type& type);
 
 class TimingNode {
     public:
-        TimingNode() { omp_init_lock(&lock_); }
-        TimingNode(TN_Type new_type): type_(new_type) { omp_init_lock(&lock_); }
-        ~TimingNode() { omp_destroy_lock(&lock_); }
-
-        void lock() { omp_set_lock(&lock_); }
-        void unlock() { omp_unset_lock(&lock_); }
+        TimingNode() {}
+        TimingNode(TN_Type new_type): type_(new_type) {}
 
         int num_out_edges() const { return out_edges_.size(); }
 
@@ -63,8 +59,6 @@ class TimingNode {
         TN_Type type_;
 
         std::vector<TimingEdge> out_edges_; //Timing edges driven by this node
-
-        omp_lock_t lock_;
 
 };
 

### 2014-11-16 src/main.cpp
Rename initial parallel analyzer to ParallelLevelizedLockedTimingAnalyzer

@@ -3,7 +3,7 @@
 
 #include "TimingGraph.hpp"
 #include "SerialTimingAnalyzer.hpp"
-#include "ParallelTimingAnalyzer.hpp"
+#include "ParallelLevelizedLockedTimingAnalyzer.hpp"
 #include "vpr_timing_graph_common.hpp"
 
 #define NUM_SERIAL_RUNS 3
@@ -28,7 +28,7 @@ int main(int argc, char** argv) {
     std::vector<node_arr_req_t> expected_arr_req_times;
 
     SerialTimingAnalyzer serial_analyzer = SerialTimingAnalyzer();
-    ParallelTimingAnalyzer parallel_analyzer = ParallelTimingAnalyzer();
+    ParallelLevelizedLockedTimingAnalyzer parallel_analyzer = ParallelLevelizedLockedTimingAnalyzer();
 
     {
         clock_gettime(CLOCK_MONOTONIC, &load_start);

### 2014-11-16 src/timing_analyzer/ParallelLevelizedLockedTimingAnalyzer.cpp

@@ -1,11 +1,11 @@
-#include "ParallelTimingAnalyzer.hpp"
+#include "ParallelLevelizedLockedTimingAnalyzer.hpp"
 #include "TimingGraph.hpp"
 
 #include <omp.h>
 
 #define DEFAULT_CLOCK_PERIOD 1.0e-9
 
-void ParallelTimingAnalyzer::calculate_timing(TimingGraph& timing_graph) {
+void ParallelLevelizedLockedTimingAnalyzer::calculate_timing(TimingGraph& timing_graph) {
     create_locks(timing_graph);
 #pragma omp parallel
     {
@@ -24,25 +24,25 @@ void ParallelTimingAnalyzer::calculate_timing(TimingGraph& timing_graph) {
     }
 }
 
-void ParallelTimingAnalyzer::create_locks(TimingGraph& tg) {
+void ParallelLevelizedLockedTimingAnalyzer::create_locks(TimingGraph& tg) {
     node_locks_ = std::vector<omp_lock_t>(tg.num_nodes());
 }
 
-void ParallelTimingAnalyzer::init_locks() {
+void ParallelLevelizedLockedTimingAnalyzer::init_locks() {
 #pragma omp for
     for(int i = 0; i < (int) node_locks_.size(); i++) {
         omp_init_lock(&node_locks_[i]);
     }
 }
 
-void ParallelTimingAnalyzer::cleanup_locks() {
+void ParallelLevelizedLockedTimingAnalyzer::cleanup_locks() {
 #pragma omp for
     for(int i = 0; i < (int) node_locks_.size(); i++) {
         omp_destroy_lock(&node_locks_[i]);
     }
 }
 
-void ParallelTimingAnalyzer::pre_traversal(TimingGraph& timing_graph) {
+void ParallelLevelizedLockedTimingAnalyzer::pre_traversal(TimingGraph& timing_graph) {
     /*
      * The pre-traversal sets up the timing graph for propagating arrival
      * and required times.
@@ -62,7 +62,7 @@ void ParallelTimingAnalyzer::pre_traversal(TimingGraph& timing_graph) {
     }
 }
 
-void ParallelTimingAnalyzer::forward_traversal(TimingGraph& timing_graph) {
+void ParallelLevelizedLockedTimingAnalyzer::forward_traversal(TimingGraph& timing_graph) {
     //Forward traversal (arrival times)
 
     //Need synchronization on nodes, since each source node could propogate
[…]

### 2014-11-16 src/timing_analyzer/ParallelLevelizedLockedTimingAnalyzer.hpp

@@ -6,7 +6,7 @@
 #include "SerialTimingAnalyzer.hpp"
 
 
-class ParallelTimingAnalyzer : public SerialTimingAnalyzer {
+class ParallelLevelizedLockedTimingAnalyzer : public SerialTimingAnalyzer {
     public: 
         void calculate_timing(TimingGraph& timing_graph);
 

### 2014-11-16 src/main.cpp
Initial task implementaion.  slower than standard for

@@ -4,6 +4,7 @@
 #include "TimingGraph.hpp"
 #include "SerialTimingAnalyzer.hpp"
 #include "ParallelLevelizedLockedTimingAnalyzer.hpp"
+#include "ParallelLevelizedLockedTasksTimingAnalyzer.hpp"
 #include "vpr_timing_graph_common.hpp"
 
 #define NUM_SERIAL_RUNS 3
@@ -28,7 +29,7 @@ int main(int argc, char** argv) {
     std::vector<node_arr_req_t> expected_arr_req_times;
 
     SerialTimingAnalyzer serial_analyzer = SerialTimingAnalyzer();
-    ParallelLevelizedLockedTimingAnalyzer parallel_analyzer = ParallelLevelizedLockedTimingAnalyzer();
+    ParallelLevelizedLockedTimingAnalyzer parallel_analyzer = ParallelLevelizedLockedTimingAnalyzer(); 
 
     {
         clock_gettime(CLOCK_MONOTONIC, &load_start);

### 2014-11-16 src/timing_analyzer/ParallelLevelizedLockedTasksTimingAnalyzer.cpp

@@ -1,0 +1,184 @@
+#include "ParallelLevelizedLockedTasksTimingAnalyzer.hpp"
+#include "TimingGraph.hpp"
+
+#include <omp.h>
+
+#define DEFAULT_CLOCK_PERIOD 1.0e-9
+
+void ParallelLevelizedLockedTasksTimingAnalyzer::calculate_timing(TimingGraph& timing_graph) {
+    create_locks(timing_graph);
+#pragma omp parallel
+    {
+
+        //TODO: Lock init and pre traversal can occur in parallel
+        init_locks();
+        pre_traversal(timing_graph);
+
+        //TODO: While forward and backward tranversals work on the same edges
+        //and nodes, they work on independant data sets (i.e. arrival time for forward and required time for backward)
+        //so they can occur in parallel
+        forward_traversal(timing_graph);
+        backward_traversal(timing_graph);
+
+        cleanup_locks();
+    }
+}
+
+void ParallelLevelizedLockedTasksTimingAnalyzer::create_locks(TimingGraph& tg) {
+    node_locks_ = std::vector<omp_lock_t>(tg.num_nodes());
+}
+
+void ParallelLevelizedLockedTasksTimingAnalyzer::init_locks() {
+#pragma omp for
+    for(int i = 0; i < (int) node_locks_.size(); i++) {
+        omp_init_lock(&node_locks_[i]);
+    }
+}
+
+void ParallelLevelizedLockedTasksTimingAnalyzer::cleanup_locks() {
+#pragma omp for
+    for(int i = 0; i < (int) node_locks_.size(); i++) {
+        omp_destroy_lock(&node_locks_[i]);
+    }
+}
+
+void ParallelLevelizedLockedTasksTimingAnalyzer::pre_traversal(TimingGraph& timing_graph) {
+    /*
+     * The pre-traversal sets up the timing graph for propagating arrival
+     * and required times.
+     * Steps performed include:
+     *   - Initialize arrival times on primary inputs
+     *   - Propogating clock delay to all clock pins
+     *   - Initialize required times on primary outputs
+     */
+    //We perform a BFS from primary inputs to initialize the timing graph
+#pragma omp single
+    for(int level_idx = 0; level_idx < timing_graph.num_levels(); level_idx++) {
+        const std::vector<int>& level_ids = timing_graph.level(level_idx);
+        for(int i = 0; i < (int) level_ids.size(); i++) {
+            int level_id = level_ids[i];
[…]

### 2014-11-16 src/timing_analyzer/ParallelLevelizedLockedTasksTimingAnalyzer.hpp

@@ -1,12 +1,12 @@
-#ifndef ParallelLevelizedLockedTimingAnalyzer_hpp
-#define ParallelLevelizedLockedTimingAnalyzer_hpp
+#ifndef ParallelLevelizedLockedTasksTimingAnalyzer_hpp
+#define ParallelLevelizedLockedTasksTimingAnalyzer_hpp
 #include <omp.h>
 
 #include "TimingGraph.hpp"
 #include "SerialTimingAnalyzer.hpp"
 
 
-class ParallelLevelizedLockedTimingAnalyzer : public SerialTimingAnalyzer {
+class ParallelLevelizedLockedTasksTimingAnalyzer : public SerialTimingAnalyzer {
     public: 
         void calculate_timing(TimingGraph& timing_graph);
 

### 2014-11-16 src/timing_analyzer/ParallelLevelizedLockedTimingAnalyzer.cpp
Revert to default scheduling

@@ -7,9 +7,9 @@
 
 void ParallelLevelizedLockedTimingAnalyzer::calculate_timing(TimingGraph& timing_graph) {
     create_locks(timing_graph);
+
 #pragma omp parallel
     {
-
         //TODO: Lock init and pre traversal can occur in parallel
         init_locks();
         pre_traversal(timing_graph);

### 2014-11-18 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
added all architecture experiment results
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4507 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{Architecture models}
@@
 \subsection{Interposer-Routing interface}
+\label{sec:interposer-routing-interface}
 In a simple architecture, the vertical routing wires that can not connect to an interposer wire are dangling.
@@
 Practical interposers, however, can only accommodate a fraction of all vertical routing wires.
-Hence, when this variable is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
-This allows for additional flexibility in routing.
+When \textit{fanin transfer} option is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
+This allows for additional flexibility in routing by ensuring that \textit{every} vertical routing wire could \textit{potentially} drive an interposer wire through a multiplexer.
 As an example, if \textit{\% wires cut} $=75\%$, 3 of every 4 interposer wires are cut.
@@
 \item Fan-out Transfer for Interposer Wires\\
-When this option is \textit{on}, we allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
+When \textit{fanout transfer} option is \textit{on}, we allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
 
@@
 
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{CAD Enhancements}
@@
 
-\subsection{Placer Optimization}
+\subsection{Placer optimization}
 \label{cad_enh_placer_subsection}
@@
 
-\subsection{Router Optimization}
+\subsection{Router optimization}
+\label{sec:router_optimizations}
 Once the Routing Resource Graph is modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
@@
 \label{sec:CADeffect}
-For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the "flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}).
+For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the ``flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}).
 The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices.
@@
 We used VPR 7.0, with the enhancements we detailed above, and the architecture file \textit{k6\_frac\_N10\_mem32K\_40nm} in the experiments below.
-All experiments targeted the smallest FPGA (with number of rows equal to number of columns) which could accomodate a benchmark circuit; this represents a very full FPGA with little white space left, and hence presents a difficult case to an interposer-based FPGA as no die can be left mostly empty.
+All experiments targeted the smallest FPGA (with number of rows equal to number of columns) which could accommodate a benchmark circuit; this represents a very full FPGA with little white space left, and hence presents a difficult case to an interposer-based FPGA as no die can be left mostly empty.
 
-Figures \ref{fig:minW_constant_sweep} and \ref{fig:areadelay_constant_sweep} show the efficacy of the five placer wiring cost modifications as their weight, \textit{C} varies.
-Each point in the graphs is the geometric mean of the results with \textit{\% wires cut} = 60 and 80, and the results for 60\% and 80\% are themselves geometric means of the results of 6 circuits from the VTR benchmark suite~\cite{vtr2012}:  \textit{stereovision0}, \textit{stereovision1}, \textit{mkDelayWorker32B}, \textit{mkSMAdapter4B}, \textit{or1200} and \textit{blob\_merge}.
-These values for \textit{\% wires cut} were chosen because they in the likely range of wires which can cross the interposer without exceeding the microbump capacity, as described in Section~\ref{virtex7section}.
-The \textit{area-delay product} is the product of the \textit{minimum channel width} and the \textit{critical path delay}.
+We have used the eight largest circuits from the VTR benchmark~\cite{vtr2012}, namely: \textit{bgm}, \textit{LU8PEEng}, \textit{LU32PEEng}, \textit{mcml}, \textit{mkDelayWorker32B}, \textit{stereovision0}, \textit{stereovision1} and \textit{stereovision2}.
+The size of these circuits ranges from $9,100$ to $153,000$ primitives (LUTs, FFs, etc.), with an average size of $52,600$ primitives.
+All results are the geometric mean over all circuits for a given interposer architecture.
+
 To obtain the \textit{critical path delay} the circuits were run with a low stress routing with a channel width, $W = 1.3 \times minW$, where $minW$ is the minimum channel width for which the circuit is still routable.
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{minW_constant_sweep_log.eps}
-\caption{Minimum channel width vs. weighting for different \textit{cut\_cost} terms.}
-\label{fig:minW_constant_sweep}
-\end{figure}
+\subsubsection{Effectiveness of placer optimizations}
+\label{sec:effectiveness_of_placer_enhancements}
+Enabling the enhancements described in Section~\ref{cad_enh_placer_subsection} improves the results ....
+haven't finished running these yet.
+So, <TODO EHSAN>: fill this section with actual results.
+For now, let's use a graph from Andre as a placeholder!
+The graphs show that combining the timing with the bounding box optimizations led to an average of 20\% improvement on the area-delay product and the most significant contribution was made by incorporation of \textit{cut\_cost} into the wire cost function.
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{areadelay_constant_sweep_log.eps}
-\caption{Area-delay product vs. weighting for different \textit{cut\_cost} terms.}
-\label{fig:areadelay_constant_sweep}
-\end{figure}
+We believe the key to the good performance of the new placer cost functions is that it produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
+Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
+Note that bounding box (a) and (b) both cross the cutline, but (b) is penalized less by (\ref{eq:cost2}) because bounding box (b) is mostly on the lower side of the cutline; it is more likely that later smaller placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
+Cost function (\ref{eq:cost2}) penalize bounding box (a) more than (b) and (b) more than (c) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
 
-\begin{table}[htbp]
-\begin{tabular}{|r|r|r|r|r|}
-\hline
-\multicolumn{1}{|l|}{Term} & \multicolumn{1}{l|}{Best C} & \multicolumn{1}{l|}{minW} & \multicolumn{1}{l|}{crit\_path(ns)} & \multicolumn{1}{l|}{Area-delay} \\ \hline \hline
-None & - & 124.636 & 9.844 & 1226.955 \\ \hline
-\ref{eq:cost0} & 0.3 & 122.929 & 9.413 & 1157.09 \\ \hline
-\ref{eq:cost1} & 5 & 114.444 & 9.565 & 1094.68 \\ \hline
-\ref{eq:cost2} & 1 & 107.303 & 9.285 & 996.26 \\ \hline
-\ref{eq:cost4} & 1 & 113.381 & 9.280 & 1052.16 \\ \hline
-\ref{eq:cost5} & 5 & 112.703 & 9.841 & 1109.11 \\ \hline
-\end{tabular}
-\caption{Best weighting and performance for each \textit{cut\_cost} term.}
-\label{table:constant_sweep}
-\end{table}
-
-The best \textit{C} value and the resulting performance for each term is summarized in Table~\ref{table:constant_sweep}.
-Note that cost term (\ref{eq:cost2}) with weighting $C = 1.0$ has the best performance; this is the configuration used in the rest of our experiments.
-
-We believe the key to the good performance of this cost function is that it produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
-Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
-Note that bounding box (a) and (b) both cross the cutline and hence are penalized equally by (\ref{eq:cost0}), while bounding box (c) does not cross a cutline and hence is not penalized.
-However, bounding box (b) is mostly on the lower side of the cutline; it is more likely that later smaller placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
-Cost function terms (\ref{eq:cost2}) and (\ref{eq:cost4}) will both penalize bounding box (a) more than bounding box (b) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
-Figure \ref{fig:areadelay_constant_sweep} shows that these are the two best performing \textit{cut\_cost} terms.
-
-\begin{figure}[!htbp]
+\begin{figure}[!h]
 \centering
@@
 \end{figure}
-
-Figure \ref{fig:comparison_minW} shows the minimum channel width for different values of \textit{\% wires cut} when there are no optimizations and when \textit{cut\_const} (\ref{eq:cost2}) is used.
-% comment about impact vs % wires cut
-Figure \ref{fig:comparison_bars} compares result quality with both enhanced timing and wiring cost functions.
-These experiments were run with term (\ref{eq:cost2}), and are the geometric means of the runs with \textit{\% wires cut = 60, 70, 80}.
 
@@
 
-\begin{figure}[!htbp]
+\subsubsection{Effectiveness of router optimizations}
+\label{sec:effectiveness_of_router_enhancements}
+The router optimizations described in Section~\ref{sec:router_optimizations} extensively improve the runtime of the router lookahead algorithm.
+Figure~\ref{fig:lookahead_runtime} shows geomean of the router runtime for the 8 circuits for different \textit{\% wires cut} values.
+On average, the router runs 31 times faster and finds identical minimum channel widths for all circuits.
+
+\begin{figure}[!hb]
 \centering
-\includegraphics[width=\linewidth]{comparison_bars_relative.eps}
-\caption{Performance of different placer optimizations, relative to the original VPR results. The area-delay product is calculated as the product $minW \times critical path delay.$}
-\label{fig:comparison_bars}
+\includegraphics[width=\linewidth]{router_opt.eps}
+\caption{Router runtime improvement using updated lookahead function}
+\label{fig:lookahead_runtime}
 \end{figure}
 
-The graphs show that combining the timing with the bounding box optimizations led to an average of 20\% improvement on the area-delay product and the most significant contribution was made by incorporation of \textit{cut\_cost} into the wire cost function.
+The new router lookahead changes the priority of the routes discovered by the router.
+Using the new router lookahead function increases the critical path delay by 1.3\% on average as shown in Figure~\ref{fig:lookahead_fmax}.
 
-%----------------------------------------------------
+\begin{figure}[!hb]
+\centering
+\includegraphics[width=\linewidth]{router_opt_fmax.eps}
+\caption{The new router lookahead increases the critical path delay.}
+\label{fig:lookahead_fmax}
+\end{figure}
+
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{Architecture results}
@@
 
-Using the best CAD settings found in Section~\ref{cadSection} we analyzed the impact of three key architecture parameters:  \textit{\% wires cut}, \textit{delay increase} and \textit{number of cuts}.
-All the experiments were run with the eight largest circuits from the VTR benchmark~\cite{vtr2012}, namely: \textit{bgm}, \textit{LU8PEEng}, \textit{LU32PEEng}, \textit{mcml}, \textit{mkDelayWorker32B}, \textit{stereovision0}, \textit{stereovision1} and \textit{stereovision2}.
-The size of these circuits ranges from $9,100$ to $153,000$ primitives (LUTs, FFs, etc.), with an average size of $52,600$ primitives.
-All results are the geometric mean over all circuits for a given interposer architecture.
-
-%   - Channel width vs % wires cut
-\subsection{Interposer Wiring Capacity (\% wires cut)}
-
-To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the number of cuts and the increased delay constant.
-The values used for these parameters were \textit{number of cuts} $= 3$ (corresponding to four dice) and \textit{increased delay} $= 1ns$, as these values are similar to those of the \textit{XC7V2000T} device.
-
-Figure \ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}.
-It can be noted from this graph that the minimum channel width increases slowly up to 60\% of wires cut, indicating that the placement engine is able to avoid saturating the interposer routing until that point.
-When more than 60\% of the wires are cut however (i.e.
-the interposer provides less than 40\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.
-The minW value at 60\% is 20\% greater than with no wires cut, while at 70\% it is 52\% above and at 80\% it is 125\% larger than the minW with no cut wires.
-
-The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 60\% of the wires cut is essentially the same as at 0\% wires cut.
-At 80\% of the wires cut the critical path delay does rise somewhat, by 6\%; this is due to two factors: the placement is being modified to improve routability at the expense of timing, and some circuitous routes are occuring due to saturated interconnect across the interposer.
-%Andre, what is the critical path delay with no interposer (i.e. 0% cut and 0 delay)?
-
-Tables \ref{table:standard_minW} and \ref{table:standard_path} show the minimum channel width and critical path delay for each circuit.
-The trends for individual circuits follow those of the averages quite closely.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{standard_minW.eps}
-\caption{Minimum channel width vs. \textit{\% wires cut} for 3 cuts and $1ns$ of \textit{delay increase}.}
-\label{fig:standard_minW}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{standard_crit_path.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 3 cuts and $1ns$ of \textit{delay increase}.}
-\label{fig:standard_crit}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 114 & 128 & 124 & 132 & 152 & 226 \\ \hline
-LU32PEEng & 172 & 178 & 182 & 218 & 304 & 426 \\ \hline
-LU8PEEng & 110 & 116 & 118 & 128 & 158 & 236 \\ \hline
-mcml & 108 & 124 & 128 & 116 & 152 & 218 \\ \hline
-mkDelayWorker32B & 76 & 80 & 82 & 88 & 100 & 132 \\ \hline
-stereovision0 & 62 & 60 & 70 & 88 & 110 & 166 \\ \hline
-stereovision1 & 104 & 108 & 116 & 120 & 178 & 312 \\ \hline
-stereovision2 & 162 & 164 & 168 & 194 & 244 & 360 \\ \hline
-Geometric mean & 108 & 114 & 118 & 129 & 164 & 243 \\ \hline
-\end{tabular}
-\caption{Minimum channel width when \textit{number of cuts = 3} and \textit{increased delay = 1ns}.}
-\label{table:standard_minW}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 35.1 & 33.6 & 32.2 & 32.9 & 37.8 & 40.3 \\ \hline
-LU32PEEng & 115 & 114 & 114 & 114 & 115 & 113 \\ \hline
-LU8PEEng & 118 & 117 & 119 & 116 & 117 & 124 \\ \hline
-mcml & 82.3 & 82.9 & 84.0 & 83.4 & 88.3 & 85.7 \\ \hline
-mkDelayWorker32B & 11.0 & 12.3 & 12.0 & 12.3 & 14.4 & 12.8 \\ \hline
-stereovision0 & 6.9 & 7.8 & 7.0 & 7.4 & 7.6 & 7.0 \\ \hline
-stereovision1 & 11.6 & 9.2 & 10.4 & 11.0 & 11.6 & 11.8 \\ \hline
-stereovision2 & 23.3 & 25.4 & 23.9 & 24.9 & 25.6 & 24.8 \\ \hline
-Geometric mean & 30.8 & 30.9 & 30.6 & 31.2 & 33.1 & 32.6 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 1ns}.}
-\label{table:standard_path}
-\end{table}
-
-% TODO
-% "extra FPGA routing channel width" vs. wires across the interposer
-%\subsection{Intra-FPGA Routing  width vs. wires across the interposer}
-
-%Can you take a crack at adding a section that tries to present the data as "extra FPGA routing channel width" vs.
-wires across the interposer?
-The idea would be to address reviewer 4's comments:  the graph will show we need some minimum number of wires across the interposer, and it will also directly show how many more wires we need to add to the FPGA to maintain routability.
-I think you could get this data simply by processing data you already have.
-%So: 
-%x-axis:  geomean wires crossing interposer
-%y-axis:  geomean channel width required
-
-Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability.
-This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 4-die system.
-When 108 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die.
-As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer.
-The increase is gradual as the interposer routing is reduced from 108 tracks per channel to 52 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 108 tracks per channel to 129 tracks per channel.
-As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate.
-At 49 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{243} tracks to successfully route the designs.
-Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (48\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{numberofcrossingwires.eps}
-\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 3 cuts and $1ns$ of \textit{delay increase}.}
-\label{fig:crossingwires}
-\end{figure}
-
-%   - Delay vs % wires cut & wire delay
-\subsection{Circuit Speed vs. Interposer Delay}
-
-To investigate the impact of the interposer delay (\textit{delay increase}), we kept the \textit{number of cuts} constant at $3$ while \textit{delay increase} was swept between $0$ and $1.5ns$.
-
-Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}.
-The penalty in critical path delay is significant, ranging between $5$ and $9$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay.
-Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA.
-The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 20\%, while a $1ns$ interposer delay increases critical path delay by approximately 35\% vs.
-a monolithic FPGA.
-Once again, the critical path increase shows little correlation to the \textit{\% wires cut}, however.
-Tables \ref{table:delay0}, \ref{table:delay500}, \ref{table:standard_path} and \ref{table:delay1500} show the critical path delay for each circuit for each of the different values of \textit{delay increase}.
-Once again the results for individual circuits closely parallel the overall averages.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{delays_crit_path.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 3 cuts and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
-\label{fig:delays_crit}
-\end{figure}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline  \hline
-bgm & 25.9 & 26.3 & 25.8 & 26.1 & 25.8 & 26.7 \\ \hline
-LU32PEEng & 113 & 113 & 117 & 118 & 113 & 114 \\ \hline
-LU8PEEng & 115 & 114 & 116 & 115 & 118 & 115 \\ \hline
-mcml & 77.2 & 78.7 & 79.5 & 80.5 & 77.3 & 76.7 \\ \hline
-mkDelayWorker32B & 7.8 & 7.6 & 7.5 & 7.7 & 7.5 & 7.6 \\ \hline
-stereovision0 & 4.5 & 4.3 & 4.5 & 4.5 & 4.5 & 4.5 \\ \hline
-stereovision1 & 6.0 & 6.2 & 5.9 & 6.3 & 6.2 & 6.2 \\ \hline
-stereovision2 & 18.0 & 17.1 & 17.1 & 17.5 & 17.6 & 17.7 \\ \hline
-Geometric mean & 23.7 & 23.5 & 23.6 & 23.9 & 23.6 & 23.7 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 0ns}.}
-\label{table:delay0}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 31.6 & 30.3 & 29.2 & 29.2 & 29.9 & 31.6 \\ \hline
-LU32PEEng & 119 & 114 & 115 & 114 & 115 & 115 \\ \hline
-LU8PEEng & 118 & 116 & 119 & 117 & 114 & 115 \\ \hline
-mcml & 82.6 & 82.7 & 79.6 & 81.9 & 84.5 & 82.6 \\ \hline
-mkDelayWorker32B & 9.8 & 10.6 & 10.4 & 9.6 & 10.2 & 9.1 \\ \hline
-stereovision0 & 5.7 & 5.3 & 4.8 & 4.9 & 4.9 & 4.8 \\ \hline
-stereovision1 & 8.0 & 7.5 & 7.4 & 8.2 & 7.9 & 8.0 \\ \hline
-stereovision2 & 20.1 & 19.7 & 20.3 & 19.9 & 20.2 & 21.6 \\ \hline
-Geometric mean & 27.5 & 26.9 & 26.4 & 26.5 & 26.7 & 26.8 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 0.5ns}.}
-\label{table:delay500}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 47.2 & 52.4 & 43.1 & 42.7 & 44.8 & 48.3 \\ \hline
-LU32PEEng & 115 & 112 & 117 & 114 & 116 & 113 \\ \hline
-LU8PEEng & 161 & 159 & 122 & 146 & 118 & 129 \\ \hline
-mcml & 94.0 & 86.0 & 93.9 & 88.7 & 77.3 & 87.5 \\ \hline
-mkDelayWorker32B & 15.3 & 13.8 & 14.3 & 15.3 & 17.9 & 15.4 \\ \hline
-stereovision0 & 7.5 & 9.4 & 8.9 & 7.8 & 9.6 & 8.3 \\ \hline
-stereovision1 & 14.6 & 12.0 & 13.1 & 12.7 & 14.6 & 14.8 \\ \hline
-stereovision2 & 29.9 & 30.3 & 28.7 & 30.3 & 30.1 & 29.8 \\ \hline
-Geometric mean & 37.7 & 37.3 & 35.9 & 36.1 & 37.0 & 36.9 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 1.5ns}.}
-\label{table:delay1500}
-\end{table}
+Using the best CAD settings found in Section~\ref{cadSection} and the best CAD flow described in Section~\ref{sec:JaveedSection}, we analyze the impact of the key architecture parameters: \textit{\% wires cut}, \textit{delay increase}, \textit{number of cuts}, and \textit{interposer-routing interface}.
+We run all the experiments using the setup explained in Section~\ref{sec:CADeffect}.
 
@@
 
-Tables \ref{table:1cut_minW}, \ref{table:1cut_path}, \ref{table:2cut_minW}, \ref{table:2cut_path}, \ref{table:standard_minW} and \ref{table:standard_path} show the minimum channel width and critical path delay for each circuit for each of the different scenarios.
-The critical path results for individual circuits again follow the overall averages well.
-Some of the minimum channel widths for individual circuits show anomalous trends as the number of cutlines increase, however.
-For example, \textit{mcml} often has a smaller minimum channel width with 2 cuts than with 1 cut.
-This may be due to a circuit structure that leads to easier division into thirds than halves, but also likely indicates room for further CAD optimization.
-
 \begin{figure}[!htbp]
@@
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 114 & 114 & 122 & 126 & 142 & 156 \\ \hline
-LU32PEEng & 170 & 176 & 182 & 206 & 318 & 396 \\ \hline
-LU8PEEng & 110 & 114 & 116 & 138 & 164 & 236 \\ \hline
-mcml & 108 & 114 & 106 & 124 & 150 & 238 \\ \hline
-mkDelayWorker32B & 74 & 78 & 80 & 88 & 98 & 138 \\ \hline
-stereovision0 & 62 & 60 & 62 & 76 & 132 & 146 \\ \hline
-stereovision1 & 102 & 104 & 106 & 108 & 144 & 186 \\ \hline
-stereovision2 & 160 & 158 & 164 & 200 & 264 & 432 \\ \hline
-Geometric mean & 107 & 109 & 111 & 126 & 165 & 221 \\ \hline
-\end{tabular}
-\caption{Minimum channel width when \textit{number of cuts = 1} and \textit{increased delay = 1ns}.}
-\label{table:1cut_minW}
-\end{table}
+%   - Channel width vs % wires cut
+\subsection{Interposer Wiring Capacity (\% wires cut)}
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 26.5 & 32.1 & 28.3 & 30.2 & 27.5 & 30.0 \\ \hline
-LU32PEEng & 113 & 113 & 113 & 115 & 114 & 111 \\ \hline
-LU8PEEng & 117 & 118 & 116 & 114 & 114 & 116 \\ \hline
-mcml & 82.5 & 82.0 & 83.0 & 84.5 & 83.2 & 77.9 \\ \hline
-mkDelayWorker32B & 11.2 & 11.4 & 11.2 & 9.9 & 10.6 & 11.5 \\ \hline
-stereovision0 & 4.5 & 5.1 & 5.0 & 4.9 & 4.9 & 5.4 \\ \hline
-stereovision1 & 7.4 & 8.2 & 8.4 & 7.2 & 7.2 & 7.9 \\ \hline
-stereovision2 & 21.3 & 23.1 & 20.9 & 22.3 & 21.6 & 20.9 \\ \hline
-Geometric mean & 26.3 & 28.1 & 27.2 & 26.8 & 26.5 & 27.3 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 1} and \textit{increased delay = 1ns}.}
-\label{table:1cut_path}
-\end{table}
+To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the other parameters constant.
+We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{increased delay} $= 1ns$, \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 114 & 116 & 130 & 140 & 172 & 272 \\ \hline
-LU32PEEng & 170 & 180 & 186 & 198 & 264 & 426 \\ \hline
-LU8PEEng & 110 & 114 & 116 & 124 & 158 & 246 \\ \hline
-mcml & 108 & 110 & 126 & 138 & 114 & 150 \\ \hline
-mkDelayWorker32B & 76 & 80 & 82 & 88 & 100 & 132 \\ \hline
-stereovision0 & 62 & 58 & 60 & 88 & 114 & 166 \\ \hline
-stereovision1 & 104 & 108 & 116 & 136 & 184 & 296 \\ \hline
-stereovision2 & 160 & 156 & 162 & 188 & 246 & 362 \\ \hline
-Geometric mean & 108 & 109 & 116 & 132 & 160 & 237 \\ \hline
-\end{tabular}
-\caption{Minimum channel width when \textit{number of cuts = 2} and \textit{increased delay = 1ns}.}
-\label{table:2cut_minW}
-\end{table}
+Figure~\ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}.
+It can be noted from this graph that the minimum channel width is almost steady up to 60\% of wires cut, indicating that the placement engine is able to avoid saturating the interposer routing until that point.
+When more than 60\% of the wires are cut however (i.e.
+the interposer provides less than 40\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 31.0 & 30.3 & 28.3 & 36.1 & 35.9 & 37.5 \\ \hline
-LU32PEEng & 113 & 113 & 119 & 113 & 117 & 112 \\ \hline
-LU8PEEng & 118 & 118 & 117 & 116 & 118 & 114 \\ \hline
-mcml & 82.9 & 86.3 & 81.3 & 80.9 & 82.4 & 82.8 \\ \hline
-mkDelayWorker32B & 12.3 & 12.1 & 9.8 & 11.1 & 13.5 & 11.3 \\ \hline
-stereovision0 & 6.4 & 5.7 & 6.3 & 5.3 & 5.3 & 5.9 \\ \hline
-stereovision1 & 9.3 & 10.4 & 9.5 & 9.6 & 10.2 & 9.9 \\ \hline
-stereovision2 & 22.9 & 24.2 & 24.0 & 21.6 & 24.2 & 25.5 \\ \hline
-Geometric mean & 29.5 & 29.7 & 28.6 & 28.8 & 30.4 & 30.1 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 2} and \textit{increased delay = 1ns}.}
-\label{table:2cut_path}
-\end{table}
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{minChW.eps}
+\caption{Minimum channel width vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
+\label{fig:standard_minW}
+\end{figure}
+
+The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
+Note, however, that this same critical path delay is achieved at a much higher channel width.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{standard_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
+\label{fig:standard_crit}
+\end{figure}
+
+% TODO
+% "extra FPGA routing channel width" vs. wires across the interposer
+%\subsection{Intra-FPGA Routing  width vs. wires across the interposer}
+
+%Can you take a crack at adding a section that tries to present the data as "extra FPGA routing channel width" vs.
+wires across the interposer?
+The idea would be to address reviewer 4's comments:  the graph will show we need some minimum number of wires across the interposer, and it will also directly show how many more wires we need to add to the FPGA to maintain routability.
+I think you could get this data simply by processing data you already have.
+%So: 
+%x-axis:  geomean wires crossing interposer
+%y-axis:  geomean channel width required
+
+Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability.
+This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 2-die system.
+When 110 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die.
+As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer.
+The increase is gradual as the interposer routing is reduced from 110 tracks per channel to 34 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 110 tracks per channel to 114 tracks per channel.
+As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate.
+At 18 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{188} tracks to successfully route the designs.
+Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (40\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{numberofcrossingwires_new.eps}
+\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 2 dice and $1ns$ of \textit{delay increase}.}
+\label{fig:crossingwires}
+\end{figure}
+
+%   - Delay vs % wires cut & wire delay
+\subsection{Circuit Speed vs. Interposer Delay}
+
+To investigate the impact of the interposer delay (\textit{delay increase}), we keep all parameters constant and we sweep the \textit{delay increase} from 0 to 1.5 \textit{ns}.
+We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
+
+Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}.
+The penalty in critical path delay is significant, ranging between $1$ and $3.7$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay.
+Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA.
+The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 4\%, while a $1ns$ interposer delay increases critical path delay by approximately 12\% vs.
+a monolithic FPGA.
+But once again, the critical path delay shows little correlation to the \textit{\% wires cut}.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
+\label{fig:delays_crit}
+\end{figure}
+
+%   - routing-interposer interface
+\subsection{Routing-Interposer Interface}
+As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces.
+These interfaces provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.
+
+To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters as well as different combinations of them.
+Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{arch_experiments.eps}
+\caption{Exploring more flexible interposer-based architectures by modifying routing-interposer interfaces.}
+\label{fig:all_arch_experiments}
+\end{figure}
+
+Experiments show that using bidirectional interposer wires instead of unidirectional wires (column \textit{b} in Figure~\ref{fig:all_arch_experiments}) improves minimum channel width by 1\% and critical path delay by 0.5\%.
+Therefore, while it is beneficial to use bidirectional interposer wires, it does not dramatically improve performance.
+We believe that is because a well placed and routed design that fills the chip completely with little spare room has approximately equal number of signals crossing the interposer in each direction (i.e.
+number of signals transferred from die 1 to die 2 is close to the number of signals transferred from die 2 to die 1).
+
+Adding multiplexers in front of interposer wires allows dangling routing wires to connect to interposer wires (\textit{``fanin transfer"}) and adding extra load on the output of the interposer wires (\textit{``fanout transfer"}) allows the interposer wire to drive dangling wires on the other side of the cut.
+Therefore these modifications provide higher flexibility to better utilize interposer wires.
+Experiments show that enabling these two options alone without using bidirectional interposer wires (column \textit{c} in Figure~\ref{fig:all_arch_experiments}) reduces minimum channel width by 3\% and reduces critical path delay by 1\%.
+
+Moreover, better performance can be achieved if either of these two options is used together with bidirectional interposer wires.
+Enabling \textit{fanout transfer} with bidirectional interposer wires (column \textit{d} in Figure~\ref{fig:all_arch_experiments}) reduces the minimum channel width by 9\% and critical path delay by 0.7\%.
+Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%.
+This signifies that the flexibility to drive an interposer wire is more important for routability compared to the flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
+
+Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay.
+These models show that interposer-based FPGAs can extensively benefit from hardware that provides a flexible routing-interposer interface.
+
+The \textit{fanin transfer} and \textit{fanout transfer} options ensure that no vertical routing wires are dangling at the cutline, hence, each vertical routing wire either has exactly \textit{one} interposer wire driving it or it drives a mux of exactly \textit{one} interposer wire.
+As a final experiment, we try to add additional fanins and fanouts to interposer wires, so that each vertical routing wire can \textit{potentially} drive multiple interposer wires, or be driven by one of many interposer wires.
+We found that adding extra fanouts to interposer wires does not help performance or area, but adding additional fanins to interposer wire muxes helps reduce the minimum channel width while maintaining the same critical path delay.
+As shown in Figure~\ref{fig:additional_fanins}, adding 4 additional inputs to interposer wire muxes reduces the minimum channel width by 1\% and adding 8 additional inputs reduces minimum channel width by 1.5\%.
+After that point the performance is no longer limited by the ability to get on an interposer wire and hence we do not see any gains for adding more than 8 additional inputs.
+Notice that column \textit{0} of Figure~\ref{fig:additional_fanins} is the same as column \textit{f} of Figure~\ref{fig:all_arch_experiments}.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{additional_fanins.eps}
+\caption{Adding up to 8 extra fanins to interposer wire muxes helps reduce minimum channel width.}
+\label{fig:additional_fanins}
+\end{figure}
 
@@
 \label{conclusionSection}
-
 We have extended VPR to model and optimize for interposer-based multi-FPGA systems.

### 2014-11-24 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Updating according to Vaughn's second review. Almost all sections by Ehsan are done by this update.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4509 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \label{sec:CADeffect}
-For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the ``flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}).
+For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the ``flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}) \cite{luu2014vtr}.
 The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices.
@@
 
-We believe the key to the good performance of the new placer cost functions is that it produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
+We believe the key to the good performance of the new placer cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
 Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
-Note that bounding box (a) and (b) both cross the cutline, but (b) is penalized less by (\ref{eq:cost2}) because bounding box (b) is mostly on the lower side of the cutline; it is more likely that later smaller placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
+Note that bounding box (a) and (b) both cross the cutline, but (b) is penalized less by (\ref{eq:cost2}) because bounding box (b) is mostly on the lower side of the cutline; it is more likely that later placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
 Cost function (\ref{eq:cost2}) penalize bounding box (a) more than (b) and (b) more than (c) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
@@
 \centering
-\includegraphics[width=\linewidth]{bb.eps}
+\includegraphics[width=\linewidth]{bb_new.eps}
 \caption{Illustration of three different scenarios for the wiring cost.
@@
 \label{sec:effectiveness_of_router_enhancements}
-The router optimizations described in Section~\ref{sec:router_optimizations} extensively improve the runtime of the router lookahead algorithm.
-Figure~\ref{fig:lookahead_runtime} shows geomean of the router runtime for the 8 circuits for different \textit{\% wires cut} values.
+The router optimizations described in Section~\ref{sec:router_optimizations} greatly reduce the runtime of the router.
+Figure~\ref{fig:lookahead_runtime} shows the geomean of the router runtime for the 8 largest VTR circuits for different \textit{\% wires cut} values.
 On average, the router runs 31 times faster and finds identical minimum channel widths for all circuits.
@@
 \includegraphics[width=\linewidth]{router_opt.eps}
-\caption{Router runtime improvement using updated lookahead function}
+\caption{Router runtime with original and updated lookahead function.}
 \label{fig:lookahead_runtime}
@@
 Using the new router lookahead function increases the critical path delay by 1.3\% on average as shown in Figure~\ref{fig:lookahead_fmax}.
+We consider this small quality loss acceptable and use the enhanced router lookahead in all our experiments due to its greatly reduced runtime.
 
@@
 \includegraphics[width=\linewidth]{router_opt_fmax.eps}
-\caption{The new router lookahead increases the critical path delay.}
+\caption{Impact of new vs. original router lookahead on the critical path delay.}
 \label{fig:lookahead_fmax}
 \end{figure}
+
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+\section{Adding Partitioning to the CAD Flow}
+\label{sec:CADflow}
+TODO: Javeed
 
@@
 
-Using the best CAD settings found in Section~\ref{cadSection} and the best CAD flow described in Section~\ref{sec:JaveedSection}, we analyze the impact of the key architecture parameters: \textit{\% wires cut}, \textit{delay increase}, \textit{number of cuts}, and \textit{interposer-routing interface}.
-We run all the experiments using the setup explained in Section~\ref{sec:CADeffect}.
+Using the best CAD settings found in Section~\ref{cadSection} and the best CAD flow described in Section~\ref{sec:CADflow}, we analyze the impact of the key architecture parameters: \textit{\% wires cut}, \textit{delay increase}, \textit{number of cuts}, and \textit{interposer-routing interface}.
+We run all the experiments using the setup detailed in Section~\ref{sec:CADeffect}.
+
+%   - routing-interposer interface
+\subsection{Routing-Interposer Interface}
+As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces.
+These interfaces provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.
+
+To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters in different combinations.
+Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments; all results are normalized to the least flexible routing-interposer interface.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{arch_experiments2.eps}
+\caption{Impact of routing-interposer interface on minimum channel width and circuit speed.}
+\label{fig:all_arch_experiments}
+\end{figure}
+
+Figure~\ref{fig:all_arch_experiments} shows that using bidirectional interposer wires instead of unidirectional wires (column \textit{b} in Figure~\ref{fig:all_arch_experiments}) improves minimum channel width by 1\% and critical path delay by 0.5\%.
+Therefore, while it is beneficial to use bidirectional interposer wires, in isolation they do not dramatically improve performance.
+
+Adding multiplexers in front of interposer wires allows dangling routing wires to connect to interposer wires (\textit{``fanin transfer"}) and adding extra load on the output of the interposer wires (\textit{``fanout transfer"}) allows the interposer wire to drive dangling wires on the other side of the cut.
+Therefore these modifications provide higher flexibility to better utilize interposer wires.
+Figure~\ref{fig:all_arch_experiments} shows that enabling these two options alone without using bidirectional interposer wires (column \textit{c} in Figure~\ref{fig:all_arch_experiments}) reduces minimum channel width by 3\% and reduces critical path delay by 1\%.
+
+Moreover, better performance can be achieved if either of these two options is used together with bidirectional interposer wires.
+Enabling \textit{fanout transfer} with bidirectional interposer wires (column \textit{d} in Figure~\ref{fig:all_arch_experiments}) reduces the minimum channel width by 9\% and critical path delay by 0.7\%.
+Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%.
+This signifies that the flexibility to drive an interposer wire is more important for routability compared to the flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
+
+Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay.
+These models show that interposer-based FPGAs can extensively benefit from hardware that provides a flexible routing-interposer interface.
+It is important both to make the interposer wires easier to access with additional routing switches \textit{and} to make them bidirectional to accommodate asymmetric routing demand.
+
+The \textit{fanin transfer} and \textit{fanout transfer} options ensure that no vertical routing wires are dangling at the cutline; hence, each vertical routing wire either has exactly \textit{one} interposer wire driving it or it drives a mux feeding exactly \textit{one} interposer wire.
+As a final experiment, we try to add additional fanins and fanouts to interposer wires, so that each vertical routing wire can \textit{potentially} drive multiple interposer wires, or be driven by one of many interposer wires.
+We found that adding these extra fanouts to interposer wires does not help performance or area, but adding additional fanins to interposer wire muxes helps reduce the minimum channel width while maintaining the same critical path delay.
+As shown in Figure~\ref{fig:additional_fanins}, adding 4 additional inputs to interposer wire muxes reduces the minimum channel width by 1\% and adding 8 additional inputs reduces minimum channel width by 1.5\%.
+After that point the performance is no longer limited by the ability to get on an interposer wire and hence we do not see any gains for adding more than 8 additional inputs.
+Notice that column \textit{0} of Figure~\ref{fig:additional_fanins} is the same as column \textit{f} of Figure~\ref{fig:all_arch_experiments}.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{additional_fanins.eps}
+\caption{Adding up to 8 extra fanins to interposer wire muxes helps reduce minimum channel width.}
+\label{fig:additional_fanins}
+\end{figure}
+
+Clearly using bidirectional interposer wires with fanin transfer and fanout transfer is very beneficial, so we use these parameters in all future experiments.
+Adding additional fanin beyond transferring the fanin of cut wires is only a small further benefit so we use additional fanin=0 in subsequent sections.
+Adding these additional multiplexers will have a very small impact on die size as 10,000 muxes per cut would be typical in a large FPGA~\cite{xilinx7series} which is a very small fraction of the several million routing multiplexers in a large FPGA die.
+
+%   - Channel width vs % wires cut
+\subsection{Interposer Wiring Capacity (\% wires cut)}
+
+To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the other parameters constant.
+We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{increased delay} $= 1ns$, \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
+
+Figure~\ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}.
+It can be noted from this graph that the minimum channel width is essentially constant up to 60\% of wires cut, indicating that the CAD flow is able to avoid saturating the interposer routing until that point.
+When more than 80\% of the wires are cut however (i.e.
+the interposer provides less than 20\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{minChW.eps}
+\caption{Minimum channel width vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
+\label{fig:standard_minW}
+\end{figure}
+
+The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
+Note, however, that this same critical path delay is achieved at a much higher channel width.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{standard_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
+\label{fig:standard_crit}
+\end{figure}
+
+Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability.
+This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 2-die system.
+When 110 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die.
+As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer.
+The increase is gradual as the interposer routing is reduced from 110 tracks per channel to 34 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 110 tracks per channel to 114 tracks per channel.
+As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate.
+At 18 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{188} tracks to successfully route the designs.
+Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (20\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=\linewidth]{numberofcrossingwires_new.eps}
+\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 2 dice and $1ns$ of \textit{delay increase}.}
+\label{fig:crossingwires}
+\end{figure}
+
+%   - Delay vs % wires cut & wire delay
+\subsection{Circuit Speed vs. Interposer Delay}
+
+\begin{figure}[!b]
+\centering
+\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
+\label{fig:delays_crit}
+\end{figure}
+
+To investigate the impact of the interposer delay (\textit{delay increase}), we keep all other interposer parameters constant and sweep the \textit{delay increase} from 0 to 1.5 \textit{ns}.
+We use \textit{number of cuts} $= 1$ (i.e.
+two dice), \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
+
+Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}.
+The penalty in critical path delay is significant, ranging between $1$ and $3.7$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay.
+Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA.
+The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 4\%, while a $1ns$ interposer delay increases critical path delay by approximately 12\% vs.
+a monolithic FPGA.
+But once again, the critical path delay shows little correlation to the \textit{\% wires cut}.
 
@@
 \subsection{Impact of Number of Dice}
-
+\label{num_dice_impact}
+TODO: Javeed\\
 To examine the impact of the number of die used to construct an interposer-based FPGA, we varied the number of cuts from 1 to 3 (which varies the number of die from 2 to 4).
@@
 
-%   - Channel width vs % wires cut
-\subsection{Interposer Wiring Capacity (\% wires cut)}
+%   - Impact of aspect ratio
+\subsection{Impact of Aspect Ratio}
+\label{aspect_ratio_impact}
+TODO: Javeed
 
-To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the other parameters constant.
-We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{increased delay} $= 1ns$, \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
-
-Figure~\ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}.
-It can be noted from this graph that the minimum channel width is almost steady up to 60\% of wires cut, indicating that the placement engine is able to avoid saturating the interposer routing until that point.
-When more than 60\% of the wires are cut however (i.e.
-the interposer provides less than 40\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{minChW.eps}
-\caption{Minimum channel width vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
-\label{fig:standard_minW}
-\end{figure}
-
-The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
-Note, however, that this same critical path delay is achieved at a much higher channel width.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{standard_crit_path_new.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
-\label{fig:standard_crit}
-\end{figure}
-
-% TODO
-% "extra FPGA routing channel width" vs. wires across the interposer
-%\subsection{Intra-FPGA Routing  width vs. wires across the interposer}
-
-%Can you take a crack at adding a section that tries to present the data as "extra FPGA routing channel width" vs.
-wires across the interposer?
-The idea would be to address reviewer 4's comments:  the graph will show we need some minimum number of wires across the interposer, and it will also directly show how many more wires we need to add to the FPGA to maintain routability.
-I think you could get this data simply by processing data you already have.
-%So: 
-%x-axis:  geomean wires crossing interposer
-%y-axis:  geomean channel width required
-
-Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability.
-This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 2-die system.
-When 110 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die.
-As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer.
-The increase is gradual as the interposer routing is reduced from 110 tracks per channel to 34 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 110 tracks per channel to 114 tracks per channel.
-As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate.
-At 18 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{188} tracks to successfully route the designs.
-Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (40\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{numberofcrossingwires_new.eps}
-\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 2 dice and $1ns$ of \textit{delay increase}.}
-\label{fig:crossingwires}
-\end{figure}
-
-%   - Delay vs % wires cut & wire delay
-\subsection{Circuit Speed vs. Interposer Delay}
-
-To investigate the impact of the interposer delay (\textit{delay increase}), we keep all parameters constant and we sweep the \textit{delay increase} from 0 to 1.5 \textit{ns}.
-We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
-
-Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}.
-The penalty in critical path delay is significant, ranging between $1$ and $3.7$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay.
-Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA.
-The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 4\%, while a $1ns$ interposer delay increases critical path delay by approximately 12\% vs.
-a monolithic FPGA.
-But once again, the critical path delay shows little correlation to the \textit{\% wires cut}.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
-\label{fig:delays_crit}
-\end{figure}
-
-%   - routing-interposer interface
-\subsection{Routing-Interposer Interface}
-As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces.
-These interfaces provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.
-
-To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters as well as different combinations of them.
-Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\linewidth]{arch_experiments.eps}
-\caption{Exploring more flexible interposer-based architectures by modifying routing-interposer interfaces.}
-\label{fig:all_arch_experiments}
-\end{figure}
-
-Experiments show that using bidirectional interposer wires instead of unidirectional wires (column \textit{b} in Figure~\ref{fig:all_arch_experiments}) improves minimum channel width by 1\% and critical path delay by 0.5\%.
-Therefore, while it is beneficial to use bidirectional interposer wires, it does not dramatically improve performance.
-We believe that is because a well placed and routed design that fills the chip completely with little spare room has approximately equal number of signals crossing the interposer in each direction (i.e.
-number of signals transferred from die 1 to die 2 is close to the number of signals transferred from die 2 to die 1).
-
-Adding multiplexers in front of interposer wires allows dangling routing wires to connect to interposer wires (\textit{``fanin transfer"}) and adding extra load on the output of the interposer wires (\textit{``fanout transfer"}) allows the interposer wire to drive dangling wires on the other side of the cut.
-Therefore these modifications provide higher flexibility to better utilize interposer wires.
-Experiments show that enabling these two options alone without using bidirectional interposer wires (column \textit{c} in Figure~\ref{fig:all_arch_experiments}) reduces minimum channel width by 3\% and reduces critical path delay by 1\%.
-
-Moreover, better performance can be achieved if either of these two options is used together with bidirectional interposer wires.
-Enabling \textit{fanout transfer} with bidirectional interposer wires (column \textit{d} in Figure~\ref{fig:all_arch_experiments}) reduces the minimum channel width by 9\% and critical path delay by 0.7\%.
-Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%.
-This signifies that the flexibility to drive an interposer wire is more important for routability compared to the flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
-
-Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay.
-These models show that interposer-based FPGAs can extensively benefit from hardware that provides a flexible routing-interposer interface.
-
-The \textit{fanin transfer} and \textit{fanout transfer} options ensure that no vertical routing wires are dangling at the cutline, hence, each vertical routing wire either has exactly \textit{one} interposer wire driving it or it drives a mux of exactly \textit{one} interposer wire.
-As a final experiment, we try to add additional fanins and fanouts to interposer wires, so that each vertical routing wire can \textit{potentially} drive multiple interposer wires, or be driven by one of many interposer wires.
-We found that adding extra fanouts to interposer wires does not help performance or area, but adding additional fanins to interposer wire muxes helps reduce the minimum channel width while maintaining the same critical path delay.
-As shown in Figure~\ref{fig:additional_fanins}, adding 4 additional inputs to interposer wire muxes reduces the minimum channel width by 1\% and adding 8 additional inputs reduces minimum channel width by 1.5\%.
-After that point the performance is no longer limited by the ability to get on an interposer wire and hence we do not see any gains for adding more than 8 additional inputs.
-Notice that column \textit{0} of Figure~\ref{fig:additional_fanins} is the same as column \textit{f} of Figure~\ref{fig:all_arch_experiments}.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\linewidth]{additional_fanins.eps}
-\caption{Adding up to 8 extra fanins to interposer wire muxes helps reduce minimum channel width.}
-\label{fig:additional_fanins}
-\end{figure}
-
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{Conclusions and Future Work}

### 2014-11-30 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
updated figure caption
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4510 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
updated based on Vaughn's comments on Draft 2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4515 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces.
-These interfaces provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.
+Some interface choices provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.
 
 To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters in different combinations.
-Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments; all results are normalized to the least flexible routing-interposer interface.
+Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments; all results are normalized to the least flexible routing-interposer interface: one in which 80\% of wires end at the interposer and the other 20\% cross the interposer with the same directionality and routing switches they would have in a conventional FPGA.
 
@@
 \centering
-\includegraphics[width=\linewidth]{arch_experiments2.eps}
-\caption{Impact of routing-interposer interface on minimum channel width and circuit speed.}
+\includegraphics[width=\linewidth]{arch_experiments3.eps}
+\caption{Impact of routing-interposer interface flexibility on minimum channel width and circuit speed.}
 \label{fig:all_arch_experiments}
@@
 Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%.
-This signifies that the flexibility to drive an interposer wire is more important for routability compared to the flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
+This signifies that added flexibility to drive an interposer wire is more important for routability compared to adding flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
 
 Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay.
-These models show that interposer-based FPGAs can extensively benefit from hardware that provides a flexible routing-interposer interface.
+These models show that interposer-based FPGAs significantly benefit from hardware that provides a flexible routing-interposer interface.
 It is important both to make the interposer wires easier to access with additional routing switches \textit{and} to make them bidirectional to accommodate asymmetric routing demand.
@@
 When more than 80\% of the wires are cut however (i.e.
-the interposer provides less than 20\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.
+the interposer provides less than 20\% of the usual within-die routing capacity), the minimum channel width grows rapidly, indicating that the interposer routing bandwidth has become a limiting factor.
 
@@
 The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
-Note, however, that this same critical path delay is achieved at a much higher channel width.
+Note, however, that this same critical path delay is achieved at a much higher channel width when \textit{\% wires cut} is greater than 80\%.
 

### 2014-12-07 doc/journal/fpgapaperfinal/fpga2014paper.tex
updating files to the latest version before changing template from ACM to IEEETran
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4520 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 %   - Define parameters
-To properly model a silicon interposer FPGA, the popular FPGA exploration toolset, Verilog-to-Routing (VTR)~\cite{vtr2012}, was used.
-The logic synthesis portions of the flow (ODIN II and  ABC) were left untouched, while the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) was modified to model and optimize for interposer-based FPGAs.
-The modifications were made in such a way that they require no changes to any of the input files, so one can experiment with interposer-based FPGAs with any existing benchmark circuits and any existing VPR-format FPGA architecture description simply by specifying appropriate command-line parameters.
+To properly model a silicon interposer FPGA, we use version 7.0 of the popular FPGA exploration toolset, Verilog-to-Routing (VTR)~\cite{luu2014vtr}.
+The logic synthesis portions of the flow (ODIN II and  ABC) are left untouched, while we modify the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) to model and optimize for interposer-based FPGAs.
+The modifications are made in such a way that they require no changes to any of the input files, so one can experiment with interposer-based FPGAs with any existing benchmark circuits and any existing VPR-format FPGA architecture description simply by specifying appropriate command-line parameters.
 
-Three parameters were added to VPR: \textit{\% wires cut}, \textit{increased delay} and \textit{number of cuts}.
-These three parameters describe the interposer portion of the 2.5D FPGA as detailed below.
+We add the following parameters to VPR: \textit{number of cuts}, \textit{\% wires cut}, \textit{increased delay}, and \textit{interposer-routing interface}.
+These parameters describe the interposer portion of the 2.5D FPGA as detailed below.
+
+%	- number of cuts
+\subsection{Number of cuts}
+\textit{Number of cuts} describes how many cuts are made to the interposer-based FPGA versus a monolithic die.
+If \textit{number of cuts} equals 1 then there are 2 dice, and so on.
+We investigate values of 1 and 3 (2 and 4 dice).
+The Virtex 7 family has members with \textit{number of cuts} $=$ 2 and 3.
+Figure \ref{fig:fpga} shows a sample architecture with one cutline.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{fpga.eps}
+\caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
+\label{fig:fpga}
+\end{figure}
 
@@
 
-%	- number of cuts
-\subsection{Number of cuts}
-\textit{Number of cuts} describes how many cuts were made to the interposer-based FPGA versus a monolithic die.
-If \textit{number of cuts} equals 1 then there are 2 dies, and so on.
-We investigate values of this parameter between 1 and 3 (between 2 and 4 dies), reflecting commercial practice: the Virtex 7 family has members with \textit{number of cuts} $=$ 2 and 3.
-Figure \ref{fig:fpga} shows a sample architecture with one cutline.
+%   - Interposer-Routing interface
+\subsection{Interposer-Routing interface}
+\label{sec:interposer-routing-interface}
+In a simple architecture, the vertical routing wires that can not connect to an interposer wire are dangling.
+Since interposer wires are scarce, we want to be able to study architectures in which the interposer wires are more accessible.
+To do so, we have enhanced the router to support the following interposer-routing interface options.
+
+\begin{enumerate}[label=(\alph*)]
+\item Fan-in Transfer for Interposer Wires\\
+For an architecture with \textit{\% wires cut} $=0$, all vertical routing wires have a corresponding interposer wire that connects them to the adjacent die.
+Practical interposers, however, can only accommodate a fraction of all vertical routing wires.
+When \textit{fanin transfer} option is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
+This allows for additional flexibility in routing by ensuring that \textit{every} vertical routing wire could \textit{potentially} drive an interposer wire through a multiplexer.
+As an example, if \textit{\% wires cut} $=75\%$, 3 of every 4 interposer wires are cut.
+Therefore, we place a 4-to-1 multiplexer at the input of the one interposer wire that is not cut, and we connect the 3 vertical routing wires that do not have a corresponding interposer wire to the multiplexer.
+
+\item Fan-out Transfer for Interposer Wires\\
+When \textit{fanout transfer} option is \textit{on}, we allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
+
+\item Bidirectional Interposer Wires\\
+Dominant routing architectures in modern FPGAs use unidirectional wires~\cite{unidirectional, lewis2005stratix}.
+Such wires can only be driven at one point, normally their start point.
+When this option is \textit{on}, we allow interposer wires to be driven from one end or the other by adding tri-state buffers at both ends.
+\end{enumerate}
 
@@
 \centering
-\includegraphics[width=\linewidth]{fpga.eps}
-\caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
-\label{fig:fpga}
+\includegraphics[width=\linewidth]{architecture_params.eps}
+\caption{For a channel width of 4 and 75\% wires cut, only 1 interposer wire exists (in green) and 3 interposer wires are cut (in the red).
+(a)~Fan-in Transfer for Interposer Wires allows red routing wires in die 1  to drive a mux that allows them to cross the interposer layer.
+(b)~Fan-out Transfer for Interposer Wires allows the output of the interposer wire to drive wires in other tracks in die 2.}
+\label{fig:architecture_params}
 \end{figure}
+
+These options are independent and can be combined to achieve more flexible routing.
+Enabling any of these options requires adding specific circuitry (such as multiplexers) on the dice which causes very small area and delay increase that we neglect in our experiments.
 
@@
 \subsection{Implementation}
-To model an interposer-based FPGA in VPR the Routing Resource Graph (\textit{rrgraph}) must be modified.
-The rrgraph is the data structure that defines all the available routing wires and switches in the FPGA, as well the delay of each.
+To model an interposer-based multi-FPGA system in the CAD tool, we use the existing VPR infrastructure for monolithic FPGAs and make several modifications to it.
+
+First, The presence of multiple dice in an interposer-based FPGA is modeled by creating horizontal cuts in the FPGA.
+The \textit{number of cuts} is specified by the user, and we use that to determine the y-location of the cuts.
+The cutlines ``snap to the grid" such that the height of all dice are equal and no cutline cuts through a block.
+For instance, a RAM block of height 4 must remain a unit and can not be sliced by a cutline.
+
+Second, knowing the location of the cutlines, the placer uses the \textit{(x,y)} coordinate of any pin or block to determine the die to which it belongs.
+The placer uses this information as well as \textit{interposer delay} to better decide where to place each netlist block in order to minimize the cost of inter-die signal transfer as described in Section \ref{cad_enh_placer_subsection}.
+
+Third, the Routing Resource Graph (\textit{rrgraph}) must be modified.
+The rrgraph is the data structure that defines all the available routing wires (\textit{rrnodes}) and switches in the FPGA, as well as the delay of each~\cite{betz1999architecture}.
 Given a suitable rrgraph, the VPR router can implement circuits in the desired FPGA, and the VPR timing analyzer can estimate their delay.
+In the experiments below, we use an architecture that has only unidirectional wires.
 
-The presence of multiple dice in an interposer-based FPGA was modeled by creating horizontal cuts in the FPGA, which are equally spaced vertically.
-Every wire that crosses one of these cuts in fact passes through the interposer.
-In the experiments below, we use an architecture that has only unidirectional wires~\cite{unidirectional}, as these are the dominant routing architecture in modern FPGAs.
-Such wires can only be driven at their beginning.
-To control the wiring capacity between dies \textit{\% wires cut} wires of each channel segment crossing a cutline have their connections to wires and block inputs on the die opposite to their starting point removed, and the wires which aren't cut have \textit{increased delay} added to their connections which cross the cutline.
-Note that the combination of a unidirectional routing architecture and a silicon interposer results in some blocks near the cutline having reduced routing connectivity as some of the routing wires driving their inputs are disconnected from the (single) wire driver, when that driver is on the other side of the cutline and the wire is one of the cut wires.
+Previous work by Pereira et al \cite{hahn2014cad} modeled the interposer indirectly: they did not introduce new rrnodes in the rrgraph to represent the interposer wires.
+Rather, they found the rrnodes that cross a cutline.
+Then they either removed all outgoing edges of these rrnodes (if the routing node was part of the \textit{\% wires cut}), or, they increased the switch delay of the rrnode fan-outs (if the wire was not cut) to account for the additional interposer delay.
 
-Figure \ref{fig:implementation} illustrates the approach used.
+While this is a reasonable method to model such architectures, it makes it difficult to experiment with different routing-interposer interfaces such as placing input muxes for interposer wires.
+Therefore, we modify the rrgraph differently to more accurately reflect the interposer layer directly:
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{implementation.eps}
-\caption{Illustration of the approach used to simulate crossing the interposer.
-The red dashed line indicates the interposer, the black arrows indicate the direction of the wires.
-Blue arrows indicate normal connections from the wires to the CLBs.
-The red arrows indicate removed connections and green arrows indicate connections with increased delay.}
-\label{fig:implementation}
+\begin{figure}[!t]
+\subfigure[]{\includegraphics[width=0.24\linewidth]{a.eps}}
+\subfigure[]{\includegraphics[width=0.24\linewidth]{b.eps}}
+\subfigure[]{\includegraphics[width=0.24\linewidth]{a_prime.eps}}
+\subfigure[]{\includegraphics[width=0.24\linewidth]{b_prime.eps}}
+\caption{(a) Regular routing wire connectivity.
+(b) Replacing the long wire with two wires and connecting them via an interposer wire.
+(c) rrgraph for regular connectivity.
+(d) rrgraph after modifications.}
+\label{fig:rrgraph_ops}
 \end{figure}
 
-For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the "flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}).
-The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices.
-It consists of logic clusters with 10 fracturable 6-LUTs per block (N=10, k=6), and also includes 32kb RAM blocks and DSP blocks configurable to perform 9x9, 18x18 or 36x36 multipliers.
-The delay values in the architecture are taken from 40 nm circuit simulations and 40 nm commercial FPGAs.
-It uses unidirectional routing, with all wire segments having length L = 4.
+\begin{figure*}[!thpb]
+\subfigure[]{\includegraphics[width=0.6\textwidth]{gui_1.eps}}
+\subfigure[]{\includegraphics[width=0.4\textwidth]{gui_2.eps}}
+\caption{VPR Graphical User Interface. (a) Full chip view. (b) Connectivity at the cutline. Due to the limited signal capacity of the interposer, some vertical routing wires are left dangling.}
+\label{fig:vpr_interposer_gui}
+\end{figure*}
 
+For every vertical routing wire in the channel, we add a new routing node (``\textit{interposer node}") into the rrgraph.
+Then, we determine rrnodes (e.g.
+vertical routing wires) that cross a cutline in the rrgraph, and we replace them with 2 rrnodes: one that is above the cutline and one that is below the cutline.
+We properly transfer fan-ins and fan-outs of the original node onto the two newly created nodes and we connect the two nodes using an interposer node.
+Finally, we remove \textit{all} edges from \textit{\%wires cut} percent of interposer nodes and increase the switch delay of output edges of the remaining interposer nodes by \textit{increased delay}.
+Figure~\ref{fig:rrgraph_ops} shows an example of these operations.
+
+As illustrated in Figure~\ref{fig:vpr_interposer_gui}, no vertical routing wires cross a cutline boundary and a fraction of them connect to other dies via interposer wires.
+
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{CAD Enhancements}
@@
 
-Once the routing resource graph was modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
-Placement, however, is crucial in mitigating the impact of the reduced wiring and increased delay when crossing between dice; a good placement should minimize the number of signals crossing between dice, particularly time-critical ones.
-We investigated several enhancements to VPR's placement cost function to improve result quality.
+\subsection{Placer optimization}
+\label{cad_enh_placer_subsection}
+Placement is crucial in mitigating the impact of the reduced wiring and increased delay when crossing between dice; a good placement should minimize the number of signals crossing between dice, particularly time-critical ones.
+VPR uses 2 different costs as the metrics for its placer algorithm: the timing cost and the bounding box (wiring) cost.
+We modified the placer's bounding box and timing costs to account for the increased latency to cross the interposer and for the reduced wire capacity close to the cutlines, as the wire availability becomes more sparse.
 
-VPR uses 2 different costs as the metrics for its placer algorithm: the timing cost and the bounding box (wiring) cost.
+%   - Delay
+\subsubsection{Placer timing cost}
 The usual VPR timing cost is a (criticality-weighted) summation of the estimated delay (given the current placement) of every connection required by the circuit~\cite{timing2000}:
@@
 \end{equation}
+where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
 
-where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
+This cost function assumes that the FPGA is homogeneous and consequently the delay between 2 points $(x_1,y_1)$ and $(x_2,y_2)$ only depends on $(\Delta x,\Delta y)$.
+This is obviously not true for interposer-based FPGAs, as the cutlines make them heterogeneous in the \textit{y} direction.
+To solve this problem and improve the quality of the results, we add an extra term to the delay function:
+
+\begin{equation} \label{eq:timing_cost}
+\begin{split}
+delay(i,j) =& delay(\Delta x_{ij}, \Delta y_{ij}) + \\
+&times\_crossed(i,j) \times delay\_increase
+\end{split}
+\end{equation}
+
+where $times\_crossed(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
+
+%   - Wiring / cut cost
+\subsubsection{Placer wiring cost}
 The bounding box cost estimates the amount of wiring required for a net, based on the number of pins and size of the net's bounding box. VPR's original formulation is~\cite{betz1997vpr}:
@@
 Finally, $q(n)$ is a function obtained from \cite{icann} which models the fact that bounding boxes underpredict the required routing for high fanout nets.
-$q(n)$ slowly increases with the fanout of net \textit{n}, rising from 1 for nets with 3 or fewer terminals to 2.79 for nets with 50 terminals.
+$q(n)$ slowly increases with the fanout of net \textit{n}.
 
-We modified both the placer's bounding box and timing costs to account for the increased latency to cross the interposer and for the reduced wire capacity close to the cutlines, as the wire availability becomes more sparse.
-
-%   - Delay
-\subsection{Placer timing cost}
-The standard placer timing cost in VPR assumes that the FPGA is homogeneous and consequently the delay between 2 points $(x_1,y_1)$ and $(x_2,y_2)$ only depends on $(\Delta x,\Delta y)$.
-This is obviously not true for interposer-based FPGAs, as the cutlines make them heterogeneous in the \textit{y} direction.
-
-To solve this problem and improve the quality of the results, an extra term was added to the delay function. The delay function becomes:
-
-\begin{equation} \label{eq:timing_cost}
-\begin{split}
-delay(i,j) =& delay(\Delta x_{ij}, \Delta y_{ij}) + \\
-&times\_crossed(i,j) \times delay\_increase
-\end{split}
-\end{equation}
-
-where $times\_crossed(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
-
-%   - Wiring / cut cost
-\subsection{Placer wiring cost}
 VPR's wiring cost also considers the FPGA to be homogeneous, and uses only the number of nets, the size of the net's bounding box and the average number of wires per channel to calculate the cost.
-Thus, to account for the reduced connectivity near the cutlines an extra cost term, \textit{cut\_cost}, was created.
+Thus, to account for the reduced connectivity near the cutlines we create an extra cost term, \textit{cut\_cost}.
 This new cost is added to (\ref{eq:bbcost}) to create the total wiring cost.
@@
 
-We tested several different \textit{cut\_cost} formulations, as well as different weighting (C values) for each. For all of the formulations the variable $C'$ was defined as:
-
-\begin{equation} \label{eq:cprime}
-\begin{split}
-C' = \frac{C \times ratio\_wires\_cut}{avg\_chany\_W(n)}
-\end{split}
-\end{equation}
-
-where $ratio\_wires\_cut$ is the ratio of wires cut at the cutline.
-This formulation of $C'$ ensures that when we choose a \textit{C} value of 1, the new term cost term is of roughly the same magnitude as $wire\_cost_{orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
-
-The first \textit{cut\_cost} term we tested was:
-
-\begin{equation} \label{eq:cost0}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times times\_crossed(n)
-\end{equation}
-
-which penalizes the net according to the number of cutlines the bounding box spans; this directly penalizes each signal crossing between dice.
-Surprisingly this extra term did not result in significant improvements to the quality of results and in some cases even made them worse.
-We believe this is due to the discontinuous, non-smooth, nature of this cost function -- it does not guide placement optimization well as a bounding box shrinks toward a cutline.
-Placement changes that make bounding boxes "almost not cross" a cutline are not given any gain; only a sudden change in the placement that moves a bounding box entirely to one side or the other of the cutline yields a cost reduction.
-This insight led us to the other proposed functions, which give gradual gains as bounding boxes change in their size and come closer to avoiding a cutline crossing.
-
-The other tested terms were:
-
-%add C' x times_crossed(i)
-
-\begin{equation} \label{eq:cost1}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbWidth(n) \times times\_crossed(n)
-\end{equation}
+In~\cite{hahn2014cad}, Pereira et al tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
+They found that the following \textit{cut\_cost} results in the best quality of results in terms of minimum routable channel width and critical path delay.
 
@@
 
-\begin{equation} \label{eq:cost5}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbHeight(n)
-\end{equation}
+\textit{bbHeight(n)} is the height of the bounding box of the net \textit{n} and $ratio\_wires\_cut$ is the ratio of wires cut at the cutline.
+This formulation of $C'$ ensures that the new cost term is of roughly the same magnitude as $wire\_cost_{orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
 
-where \textit{bbWidth(n)} and \textit{bbHeight(n)} are the width and height of the bounding box of the net \textit{n}, respectively, and \textit{minDist(n)} is the minimum distance from the top or bottom of the bounding box to a cutline.
+This cost function performs better than several others as shown in \cite{hahn2014cad}, and we believe the key to the good performance of this cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
+Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
+Note that bounding box (a) and (b) both cross the cutline, but (b) is penalized less by (\ref{eq:cost2}) because bounding box (b) is mostly on the lower side of the cutline; it is more likely that later placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
+Cost function (\ref{eq:cost2}) penalize bounding box (a) more than (b) and (b) more than (c) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
 
-%   - Effectiveness / results
-\subsection{Effectiveness of the enhancements}
-\label{sec:CADeffect}
-
-We used VPR 7.0, with the enhancements we detailed above, and the architecture file \textit{k6\_frac\_N10\_mem32K\_40nm} in the experiments below.
-All experiments targeted the smallest FPGA (with number of rows equal to number of columns) which could accomodate a benchmark circuit; this represents a very full FPGA with little white space left, and hence presents a difficult case to an interposer-based FPGA as no die can be left mostly empty.
-
-Figures \ref{fig:minW_constant_sweep} and \ref{fig:areadelay_constant_sweep} show the efficacy of the five placer wiring cost modifications as their weight, \textit{C} varies.
-Each point in the graphs is the geometric mean of the results with \textit{\% wires cut} = 60 and 80, and the results for 60\% and 80\% are themselves geometric means of the results of 6 circuits from the VTR benchmark suite~\cite{vtr2012}:  \textit{stereovision0}, \textit{stereovision1}, \textit{mkDelayWorker32B}, \textit{mkSMAdapter4B}, \textit{or1200} and \textit{blob\_merge}.
-These values for \textit{\% wires cut} were chosen because they in the likely range of wires which can cross the interposer without exceeding the microbump capacity, as described in Section~\ref{virtex7section}.
-The \textit{area-delay product} is the product of the \textit{minimum channel width} and the \textit{critical path delay}.
-To obtain the \textit{critical path delay} the circuits were run with a low stress routing with a channel width, $W = 1.3 \times minW$, where $minW$ is the minimum channel width for which the circuit is still routable.
-
-\begin{figure}[!htbp]
+\begin{figure}[!h]
 \centering
-\includegraphics[width=\linewidth]{minW_constant_sweep_log.eps}
-\caption{Minimum channel width vs. weighting for different \textit{cut\_cost} terms.}
-\label{fig:minW_constant_sweep}
-\end{figure}
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{areadelay_constant_sweep_log.eps}
-\caption{Area-delay product vs. weighting for different \textit{cut\_cost} terms.}
-\label{fig:areadelay_constant_sweep}
-\end{figure}
-
-\begin{table}[htbp]
-\begin{tabular}{|r|r|r|r|r|}
-\hline
-\multicolumn{1}{|l|}{Term} & \multicolumn{1}{l|}{Best C} & \multicolumn{1}{l|}{minW} & \multicolumn{1}{l|}{crit\_path(ns)} & \multicolumn{1}{l|}{Area-delay} \\ \hline \hline
-None & - & 124.636 & 9.844 & 1226.955 \\ \hline
-\ref{eq:cost0} & 0.3 & 122.929 & 9.413 & 1157.09 \\ \hline
-\ref{eq:cost1} & 5 & 114.444 & 9.565 & 1094.68 \\ \hline
-\ref{eq:cost2} & 1 & 107.303 & 9.285 & 996.26 \\ \hline
-\ref{eq:cost4} & 1 & 113.381 & 9.280 & 1052.16 \\ \hline
-\ref{eq:cost5} & 5 & 112.703 & 9.841 & 1109.11 \\ \hline
-\end{tabular}
-\caption{Best weighting and performance for each \textit{cut\_cost} term.}
-\label{table:constant_sweep}
-\end{table}
-
-The best \textit{C} value and the resulting performance for each term is summarized in Table~\ref{table:constant_sweep}.
-Note that cost term (\ref{eq:cost2}) with weighting $C = 1.0$ has the best performance; this is the configuration used in the rest of our experiments.
-
-We believe the key to the good performance of this cost function is that it produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
-Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
-Note that bounding box (a) and (b) both cross the cutline and hence are penalized equally by (\ref{eq:cost0}), while bounding box (c) does not cross a cutline and hence is not penalized.
-However, bounding box (b) is mostly on the lower side of the cutline; it is more likely that later smaller placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
-Cost function terms (\ref{eq:cost2}) and (\ref{eq:cost4}) will both penalize bounding box (a) more than bounding box (b) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
-Figure \ref{fig:areadelay_constant_sweep} shows that these are the two best performing \textit{cut\_cost} terms.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{bb.eps}
+\includegraphics[width=\linewidth]{bb_new.eps}
 \caption{Illustration of three different scenarios for the wiring cost.
@@
 
-Figure \ref{fig:comparison_minW} shows the minimum channel width for different values of \textit{\% wires cut} when there are no optimizations and when \textit{cut\_const} (\ref{eq:cost2}) is used.
-% comment about impact vs % wires cut
-Figure \ref{fig:comparison_bars} compares result quality with both enhanced timing and wiring cost functions.
-These experiments were run with term (\ref{eq:cost2}), and are the geometric means of the runs with \textit{\% wires cut = 60, 70, 80}.
+\subsection{Router optimization}
+\label{sec:router_optimizations}
+Once the Routing Resource Graph is modified as detailed in Section \ref{archSection}, the VPR router adapted automatically to the interposer architecture.
+The router starts at a net source node and expands by discovering adjacent routing nodes until it reaches the destination node.
+At any point, the router knows the cost of the path taken to the current point and it uses a \textit{lookahead cost function} to estimate the resource and delay cost from the current node to the destination.
+Using this lookahead cost, the router can properly sort routing alternatives.
+The lookahead cost function in VPR is not aware of the interposer, and therefore it takes a long time to converge to a routing decision.
+To make the lookahead aware of the interposer, we modify the lookahead cost function by adding terms that account for extra \textit{interposer delay} and extra \textit{interposer nodes} that may exist on the route from the current node to the destination node.
+
+The VPR router uses the following to calculate a cost for a routing node \textit{r} on the route from \textit{i} to \textit{j}:
+\begin{equation} \label{eq:routing_cost}
+\begin{split}
+Cost(r) = Criticality(i,j) \times Delay_{Elmore}(r, topology) + \\ [1-Criticality(i,j)] \times Congestion(r)
+\end{split}
+\end{equation}
+
+Given this cost function, the router's lookahead algorithm determines the total cost of the route from \textit{i} to \textit{j} when routing node \textit{r} is used:
+
+\begin{equation} \label{eq:lookahead_cost}
+\begin{split}
+Total\_Cost(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectedCost(r,j)
+\end{split}
+\end{equation}
+
+where \textit{RT(i,r)} is the set of all routing nodes used so far from \textit{i} to \textit{r} and $\alpha$ is a constant number that determines the aggressiveness of the directed search.
+$\alpha$ is 1.2 by default in VPR.
+VPR's original $ExpectedCost(n,j)$ function is~\cite{betz1999architecture}:
+
+\begin{equation} \label{eq:expected_routing_cost}
+\begin{split}
+ExpectedCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(r,j);
+\end{split}
+\end{equation}
+
+$T_{delay}$ is an estimate of the delay that the router predicts it will see on the route from \textit{r} to \textit{j}.
+Since (\ref{eq:expected_routing_cost}) assumes a homogeneous FPGA, $T_{delay}$ does not include the additional delay some routes experience by going through the interposer.
+Thus, for interposer-based architectures, we replace $T_{delay}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{delay}$ to include the delay of expected interposer hops from \textit{r} to \textit{j}.
+\begin{equation} \label{eq:tdelay}
+\begin{split}
+T^\prime_{delay} = T_{delay} + T_{interposer} \times ExpectedInterposerHops
+\end{split}
+\end{equation}
+
+In a similar way, we modify the $ExpectedCongestionCost(r,j)$ in (\ref{eq:expected_routing_cost}) to account for the new interposer routing nodes.
+
+The interposer-aware router lookahead is able to make routing decisions more quickly given the large delay difference of a path that crosses the interposer as opposed to a path that doesn't.
+
+%   - Effectiveness / results
+\subsection{Effectiveness of the enhancements}
+\label{sec:CADeffect}
+For all experiments in this paper, the remaining architecture parameters of the FPGA are taken from the ``flagship" architecture of the VTR project (\textit{k6\_frac\_N10\_mem32K\_40nm.xml}) \cite{luu2014vtr}.
+The parameters of this architecture are in line with both current commercial FPGAs and academic research into best practices.
+It consists of logic clusters with 10 fracturable 6-LUTs per block (N=10, k=6), and also includes 32kb RAM blocks and DSP blocks configurable to perform 9x9, 18x18 or 36x36 multipliers.
+The delay values in the architecture are taken from 40 nm circuit simulations and 40 nm commercial FPGAs.
+It uses unidirectional routing, with all wire segments having length L = 4.
+
+We used VPR 7.0, with the enhancements we detailed above, and the architecture file \textit{k6\_frac\_N10\_mem32K\_40nm} in the experiments below.
+All experiments targeted the smallest FPGA (with number of rows equal to number of columns) which could accommodate a benchmark circuit; this represents a very full FPGA with little white space left, and hence presents a difficult case to an interposer-based FPGA as no die can be left mostly empty.
+
+We have used the eight largest circuits from the VTR benchmark~\cite{vtr2012}, namely: \textit{bgm}, \textit{LU8PEEng}, \textit{LU32PEEng}, \textit{mcml}, \textit{mkDelayWorker32B}, \textit{stereovision0}, \textit{stereovision1} and \textit{stereovision2}.
+The size of these circuits ranges from $9,100$ to $153,000$ primitives (LUTs, FFs, etc.), with an average size of $52,600$ primitives.
+All results are the geometric mean over all circuits for a given interposer architecture.
+
+To obtain the \textit{critical path delay} the circuits were run with a low stress routing with a channel width, $W = 1.3 \times minW$, where $minW$ is the minimum channel width for which the circuit is still routable.
+
+\subsubsection{Effectiveness of placer optimizations}
+\label{sec:effectiveness_of_placer_enhancements}
+Enabling the enhancements described in Section~\ref{cad_enh_placer_subsection} improves the results ....
+haven't finished running these yet.
+So, <TODO EHSAN>: fill this section with actual results.
+For now, let's use a graph from Andre as a placeholder!
+The graphs show that combining the timing with the bounding box optimizations led to an average of 20\% improvement on the area-delay product and the most significant contribution was made by incorporation of \textit{cut\_cost} into the wire cost function.
 
@@
 
-\begin{figure}[!htbp]
+\subsubsection{Effectiveness of router optimizations}
+\label{sec:effectiveness_of_router_enhancements}
+The router optimizations described in Section~\ref{sec:router_optimizations} greatly reduce the runtime of the router.
+Figure~\ref{fig:lookahead_runtime} shows the geomean of the router runtime for the 8 largest VTR circuits for different \textit{\% wires cut} values.
+On average, the router runs 31 times faster and finds identical minimum channel widths for all circuits.
+
+\begin{figure}[!hb]
 \centering
-\includegraphics[width=\linewidth]{comparison_bars_relative.eps}
-\caption{Performance of different placer optimizations, relative to the original VPR results. The area-delay product is calculated as the product $minW \times critical path delay.$}
-\label{fig:comparison_bars}
+\includegraphics[width=\linewidth]{router_opt.eps}
+\caption{Router runtime with original and updated lookahead function.}
+\label{fig:lookahead_runtime}
 \end{figure}
 
-The graphs show that combining the timing with the bounding box optimizations led to an average of 20\% improvement on the area-delay product and the most significant contribution was made by incorporation of \textit{cut\_cost} into the wire cost function.
+The new router lookahead changes the priority of the routes discovered by the router.
+Using the new router lookahead function increases the critical path delay by 1.3\% on average as shown in Figure~\ref{fig:lookahead_fmax}.
+We consider this small quality loss acceptable and use the enhanced router lookahead in all our experiments due to its greatly reduced runtime.
 
-%----------------------------------------------------
+\begin{figure}[!hb]
+\centering
+\includegraphics[width=\linewidth]{router_opt_fmax.eps}
+\caption{Impact of new vs. original router lookahead on the critical path delay.}
+\label{fig:lookahead_fmax}
+\end{figure}
+
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+\section{Adding Partitioning to the CAD Flow}
+\label{sec:CADflow}
+TODO: Javeed
+
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{Architecture results}
@@
 
-Using the best CAD settings found in Section~\ref{cadSection} we analyzed the impact of three key architecture parameters:  \textit{\% wires cut}, \textit{delay increase} and \textit{number of cuts}.
-All the experiments were run with the eight largest circuits from the VTR benchmark~\cite{vtr2012}, namely: \textit{bgm}, \textit{LU8PEEng}, \textit{LU32PEEng}, \textit{mcml}, \textit{mkDelayWorker32B}, \textit{stereovision0}, \textit{stereovision1} and \textit{stereovision2}.
-The size of these circuits ranges from $9,100$ to $153,000$ primitives (LUTs, FFs, etc.), with an average size of $52,600$ primitives.
-All results are the geometric mean over all circuits for a given interposer architecture.
+Using the best CAD settings found in Section~\ref{cadSection} and the best CAD flow described in Section~\ref{sec:CADflow}, we analyze the impact of the key architecture parameters: \textit{\% wires cut}, \textit{delay increase}, \textit{number of cuts}, and \textit{interposer-routing interface}.
+We run all the experiments using the setup detailed in Section~\ref{sec:CADeffect}.
+
+%   - routing-interposer interface
+\subsection{Routing-Interposer Interface}
+As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces.
+Some interface choices provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.
+
+To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters in different combinations.
+Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments; all results are normalized to the least flexible routing-interposer interface: one in which 80\% of wires end at the interposer and the other 20\% cross the interposer with the same directionality and routing switches they would have in a conventional FPGA.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{arch_experiments3.eps}
+\caption{Impact of routing-interposer interface flexibility on minimum channel width and circuit speed.}
+\label{fig:all_arch_experiments}
+\end{figure}
+
+Figure~\ref{fig:all_arch_experiments} shows that using bidirectional interposer wires instead of unidirectional wires (column \textit{b} in Figure~\ref{fig:all_arch_experiments}) improves minimum channel width by 1\% and critical path delay by 0.5\%.
+Therefore, while it is beneficial to use bidirectional interposer wires, in isolation they do not dramatically improve performance.
+
+Adding multiplexers in front of interposer wires allows dangling routing wires to connect to interposer wires (\textit{``fanin transfer"}) and adding extra load on the output of the interposer wires (\textit{``fanout transfer"}) allows the interposer wire to drive dangling wires on the other side of the cut.
+Therefore these modifications provide higher flexibility to better utilize interposer wires.
+Figure~\ref{fig:all_arch_experiments} shows that enabling these two options alone without using bidirectional interposer wires (column \textit{c} in Figure~\ref{fig:all_arch_experiments}) reduces minimum channel width by 3\% and reduces critical path delay by 1\%.
+
+Moreover, better performance can be achieved if either of these two options is used together with bidirectional interposer wires.
+Enabling \textit{fanout transfer} with bidirectional interposer wires (column \textit{d} in Figure~\ref{fig:all_arch_experiments}) reduces the minimum channel width by 9\% and critical path delay by 0.7\%.
+Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%.
+This signifies that added flexibility to drive an interposer wire is more important for routability compared to adding flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
+
+Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay.
+These models show that interposer-based FPGAs significantly benefit from hardware that provides a flexible routing-interposer interface.
+It is important both to make the interposer wires easier to access with additional routing switches \textit{and} to make them bidirectional to accommodate asymmetric routing demand.
+
+The \textit{fanin transfer} and \textit{fanout transfer} options ensure that no vertical routing wires are dangling at the cutline; hence, each vertical routing wire either has exactly \textit{one} interposer wire driving it or it drives a mux feeding exactly \textit{one} interposer wire.
+As a final experiment, we try to add additional fanins and fanouts to interposer wires, so that each vertical routing wire can \textit{potentially} drive multiple interposer wires, or be driven by one of many interposer wires.
+We found that adding these extra fanouts to interposer wires does not help performance or area, but adding additional fanins to interposer wire muxes helps reduce the minimum channel width while maintaining the same critical path delay.
+As shown in Figure~\ref{fig:additional_fanins}, adding 4 additional inputs to interposer wire muxes reduces the minimum channel width by 1\% and adding 8 additional inputs reduces minimum channel width by 1.5\%.
+After that point the performance is no longer limited by the ability to get on an interposer wire and hence we do not see any gains for adding more than 8 additional inputs.
+Notice that column \textit{0} of Figure~\ref{fig:additional_fanins} is the same as column \textit{f} of Figure~\ref{fig:all_arch_experiments}.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{additional_fanins.eps}
+\caption{Adding up to 8 extra fanins to interposer wire muxes helps reduce minimum channel width.}
+\label{fig:additional_fanins}
+\end{figure}
+
+Clearly using bidirectional interposer wires with fanin transfer and fanout transfer is very beneficial, so we use these parameters in all future experiments.
+Adding additional fanin beyond transferring the fanin of cut wires is only a small further benefit so we use additional fanin=0 in subsequent sections.
+Adding these additional multiplexers will have a very small impact on die size as 10,000 muxes per cut would be typical in a large FPGA~\cite{xilinx7series} which is a very small fraction of the several million routing multiplexers in a large FPGA die.
 
@@
 
-To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the number of cuts and the increased delay constant.
-The values used for these parameters were \textit{number of cuts} $= 3$ (corresponding to four dice) and \textit{increased delay} $= 1ns$, as these values are similar to those of the \textit{XC7V2000T} device.
+To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the other parameters constant.
+We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{increased delay} $= 1ns$, \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
 
-Figure \ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}.
-It can be noted from this graph that the minimum channel width increases slowly up to 60\% of wires cut, indicating that the placement engine is able to avoid saturating the interposer routing until that point.
-When more than 60\% of the wires are cut however (i.e.
-the interposer provides less than 40\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.
-The minW value at 60\% is 20\% greater than with no wires cut, while at 70\% it is 52\% above and at 80\% it is 125\% larger than the minW with no cut wires.
-
-The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 60\% of the wires cut is essentially the same as at 0\% wires cut.
-At 80\% of the wires cut the critical path delay does rise somewhat, by 6\%; this is due to two factors: the placement is being modified to improve routability at the expense of timing, and some circuitous routes are occuring due to saturated interconnect across the interposer.
-%Andre, what is the critical path delay with no interposer (i.e. 0% cut and 0 delay)?
-
-Tables \ref{table:standard_minW} and \ref{table:standard_path} show the minimum channel width and critical path delay for each circuit.
-The trends for individual circuits follow those of the averages quite closely.
+Figure~\ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}.
+It can be noted from this graph that the minimum channel width is essentially constant up to 60\% of wires cut, indicating that the CAD flow is able to avoid saturating the interposer routing until that point.
+When more than 80\% of the wires are cut however (i.e.
+the interposer provides less than 20\% of the usual within-die routing capacity), the minimum channel width grows rapidly, indicating that the interposer routing bandwidth has become a limiting factor.
 
@@
 \centering
-\includegraphics[width=\linewidth]{standard_minW.eps}
-\caption{Minimum channel width vs. \textit{\% wires cut} for 3 cuts and $1ns$ of \textit{delay increase}.}
+\includegraphics[width=\linewidth]{minChW.eps}
+\caption{Minimum channel width vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
 \label{fig:standard_minW}
@@
 
+The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
+Note, however, that this same critical path delay is achieved at a much higher channel width when \textit{\% wires cut} is greater than 80\%.
+
 \begin{figure}[!htbp]
 \centering
-\includegraphics[width=\linewidth]{standard_crit_path.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 3 cuts and $1ns$ of \textit{delay increase}.}
+\includegraphics[width=\linewidth]{standard_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
 \label{fig:standard_crit}
@@
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 114 & 128 & 124 & 132 & 152 & 226 \\ \hline
-LU32PEEng & 172 & 178 & 182 & 218 & 304 & 426 \\ \hline
-LU8PEEng & 110 & 116 & 118 & 128 & 158 & 236 \\ \hline
-mcml & 108 & 124 & 128 & 116 & 152 & 218 \\ \hline
-mkDelayWorker32B & 76 & 80 & 82 & 88 & 100 & 132 \\ \hline
-stereovision0 & 62 & 60 & 70 & 88 & 110 & 166 \\ \hline
-stereovision1 & 104 & 108 & 116 & 120 & 178 & 312 \\ \hline
-stereovision2 & 162 & 164 & 168 & 194 & 244 & 360 \\ \hline
-Geometric mean & 108 & 114 & 118 & 129 & 164 & 243 \\ \hline
-\end{tabular}
-\caption{Minimum channel width when \textit{number of cuts = 3} and \textit{increased delay = 1ns}.}
-\label{table:standard_minW}
-\end{table}
+Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability.
+This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 2-die system.
+When 110 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die.
+As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer.
+The increase is gradual as the interposer routing is reduced from 110 tracks per channel to 34 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 110 tracks per channel to 114 tracks per channel.
+As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate.
+At 18 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{188} tracks to successfully route the designs.
+Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (20\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 35.1 & 33.6 & 32.2 & 32.9 & 37.8 & 40.3 \\ \hline
-LU32PEEng & 115 & 114 & 114 & 114 & 115 & 113 \\ \hline
-LU8PEEng & 118 & 117 & 119 & 116 & 117 & 124 \\ \hline
-mcml & 82.3 & 82.9 & 84.0 & 83.4 & 88.3 & 85.7 \\ \hline
-mkDelayWorker32B & 11.0 & 12.3 & 12.0 & 12.3 & 14.4 & 12.8 \\ \hline
-stereovision0 & 6.9 & 7.8 & 7.0 & 7.4 & 7.6 & 7.0 \\ \hline
-stereovision1 & 11.6 & 9.2 & 10.4 & 11.0 & 11.6 & 11.8 \\ \hline
-stereovision2 & 23.3 & 25.4 & 23.9 & 24.9 & 25.6 & 24.8 \\ \hline
-Geometric mean & 30.8 & 30.9 & 30.6 & 31.2 & 33.1 & 32.6 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 1ns}.}
-\label{table:standard_path}
-\end{table}
-
-% TODO
-% "extra FPGA routing channel width" vs. wires across the interposer
-%\subsection{Intra-FPGA Routing  width vs. wires across the interposer}
-
-%Can you take a crack at adding a section that tries to present the data as "extra FPGA routing channel width" vs.
-wires across the interposer?
-The idea would be to address reviewer 4's comments:  the graph will show we need some minimum number of wires across the interposer, and it will also directly show how many more wires we need to add to the FPGA to maintain routability.
-I think you could get this data simply by processing data you already have.
-%So: 
-%x-axis:  geomean wires crossing interposer
-%y-axis:  geomean channel width required
-
-Figure \ref{fig:crossingwires} provides an alternative way to visualize the relationship between interposer routing supply and routability.
-This figure shows how the geometric average minimum channel width required within the FPGA dice varies as the geometric average of the \emph{absolute} number of wires crossing the interposer in each channel varies, again for 4-die system.
-When 108 tracks cross the interposer, the interposer channels have the same capacity as the vertical routing channels within each FPGA die.
-As fewer wires cross the interposer, the channel width required within the FPGA dice increases to compensate for the routing difficulty in crossing the interposer.
-The increase is gradual as the interposer routing is reduced from 108 tracks per channel to 52 tracks per channel; over this range the routing per channel required in the FPGA dice increases from 108 tracks per channel to 129 tracks per channel.
-As the routing crossing the interposer is further reduced however, it becomes very difficult to increase the within-die routing sufficiently to compensate.
-At 49 tracks crossing the interposer channels, for example, the within-die routing must have a channel width of \emph{243} tracks to successfully route the designs.
-Clearly the CAD tools have the ability to trade-off interposer routing for within-die routing over a reasonable range but below a certain level (48\% of the original within-die minimum channel width in our experiments) routability becomes almost solely limited by the wiring crossing the interposer and further reduction in interposer routing is not productive.
-
-\begin{figure}[!htbp]
+\begin{figure}[!t]
 \centering
-\includegraphics[width=\linewidth]{numberofcrossingwires.eps}
-\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 3 cuts and $1ns$ of \textit{delay increase}.}
+\includegraphics[width=\linewidth]{numberofcrossingwires_new.eps}
+\caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 2 dice and $1ns$ of \textit{delay increase}.}
 \label{fig:crossingwires}
@@
 
-To investigate the impact of the interposer delay (\textit{delay increase}), we kept the \textit{number of cuts} constant at $3$ while \textit{delay increase} was swept between $0$ and $1.5ns$.
-
-Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}.
-The penalty in critical path delay is significant, ranging between $5$ and $9$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay.
-Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA.
-The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 20\%, while a $1ns$ interposer delay increases critical path delay by approximately 35\% vs.
-a monolithic FPGA.
-Once again, the critical path increase shows little correlation to the \textit{\% wires cut}, however.
-Tables \ref{table:delay0}, \ref{table:delay500}, \ref{table:standard_path} and \ref{table:delay1500} show the critical path delay for each circuit for each of the different values of \textit{delay increase}.
-Once again the results for individual circuits closely parallel the overall averages.
-
-\begin{figure}[!htbp]
+\begin{figure}[!b]
 \centering
-\includegraphics[width=\linewidth]{delays_crit_path.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 3 cuts and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
+\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
 \label{fig:delays_crit}
@@
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline  \hline
-bgm & 25.9 & 26.3 & 25.8 & 26.1 & 25.8 & 26.7 \\ \hline
-LU32PEEng & 113 & 113 & 117 & 118 & 113 & 114 \\ \hline
-LU8PEEng & 115 & 114 & 116 & 115 & 118 & 115 \\ \hline
-mcml & 77.2 & 78.7 & 79.5 & 80.5 & 77.3 & 76.7 \\ \hline
-mkDelayWorker32B & 7.8 & 7.6 & 7.5 & 7.7 & 7.5 & 7.6 \\ \hline
-stereovision0 & 4.5 & 4.3 & 4.5 & 4.5 & 4.5 & 4.5 \\ \hline
-stereovision1 & 6.0 & 6.2 & 5.9 & 6.3 & 6.2 & 6.2 \\ \hline
-stereovision2 & 18.0 & 17.1 & 17.1 & 17.5 & 17.6 & 17.7 \\ \hline
-Geometric mean & 23.7 & 23.5 & 23.6 & 23.9 & 23.6 & 23.7 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 0ns}.}
-\label{table:delay0}
-\end{table}
+To investigate the impact of the interposer delay (\textit{delay increase}), we keep all other interposer parameters constant and sweep the \textit{delay increase} from 0 to 1.5 \textit{ns}.
+We use \textit{number of cuts} $= 1$ (i.e.
+two dice), \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 31.6 & 30.3 & 29.2 & 29.2 & 29.9 & 31.6 \\ \hline
-LU32PEEng & 119 & 114 & 115 & 114 & 115 & 115 \\ \hline
-LU8PEEng & 118 & 116 & 119 & 117 & 114 & 115 \\ \hline
-mcml & 82.6 & 82.7 & 79.6 & 81.9 & 84.5 & 82.6 \\ \hline
-mkDelayWorker32B & 9.8 & 10.6 & 10.4 & 9.6 & 10.2 & 9.1 \\ \hline
-stereovision0 & 5.7 & 5.3 & 4.8 & 4.9 & 4.9 & 4.8 \\ \hline
-stereovision1 & 8.0 & 7.5 & 7.4 & 8.2 & 7.9 & 8.0 \\ \hline
-stereovision2 & 20.1 & 19.7 & 20.3 & 19.9 & 20.2 & 21.6 \\ \hline
-Geometric mean & 27.5 & 26.9 & 26.4 & 26.5 & 26.7 & 26.8 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 0.5ns}.}
-\label{table:delay500}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 47.2 & 52.4 & 43.1 & 42.7 & 44.8 & 48.3 \\ \hline
-LU32PEEng & 115 & 112 & 117 & 114 & 116 & 113 \\ \hline
-LU8PEEng & 161 & 159 & 122 & 146 & 118 & 129 \\ \hline
-mcml & 94.0 & 86.0 & 93.9 & 88.7 & 77.3 & 87.5 \\ \hline
-mkDelayWorker32B & 15.3 & 13.8 & 14.3 & 15.3 & 17.9 & 15.4 \\ \hline
-stereovision0 & 7.5 & 9.4 & 8.9 & 7.8 & 9.6 & 8.3 \\ \hline
-stereovision1 & 14.6 & 12.0 & 13.1 & 12.7 & 14.6 & 14.8 \\ \hline
-stereovision2 & 29.9 & 30.3 & 28.7 & 30.3 & 30.1 & 29.8 \\ \hline
-Geometric mean & 37.7 & 37.3 & 35.9 & 36.1 & 37.0 & 36.9 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 3} and \textit{increased delay = 1.5ns}.}
-\label{table:delay1500}
-\end{table}
+Figure \ref{fig:delays_crit} shows critical path delay versus \textit{\% wires cut} for 4 different values of \textit{delay increase}.
+The penalty in critical path delay is significant, ranging between $1$ and $3.7$ times the interposer \textit{delay increase}, when compared to the case where the interposer adds no delay.
+Note that the 0\% wires cut with a 0 ns \textit{delay increase} in Figure \ref{fig:delays_crit} corresponds to a traditional monolithic FPGA.
+The speed of an interposer-based FPGA is strongly correlated to \textit{delay increase}: a $0.5ns$ interposer delay increases the critical path delay by 4\%, while a $1ns$ interposer delay increases critical path delay by approximately 12\% vs.
+a monolithic FPGA.
+But once again, the critical path delay shows little correlation to the \textit{\% wires cut}.
 
@@
 With \textit{number of cuts} $= 3$, the critical path delay is typically 7 ns higher than that of a monolithic FPGA.
-
-Tables \ref{table:1cut_minW}, \ref{table:1cut_path}, \ref{table:2cut_minW}, \ref{table:2cut_path}, \ref{table:standard_minW} and \ref{table:standard_path} show the minimum channel width and critical path delay for each circuit for each of the different scenarios.
-The critical path results for individual circuits again follow the overall averages well.
-Some of the minimum channel widths for individual circuits show anomalous trends as the number of cutlines increase, however.
-For example, \textit{mcml} often has a smaller minimum channel width with 2 cuts than with 1 cut.
-This may be due to a circuit structure that leads to easier division into thirds than halves, but also likely indicates room for further CAD optimization.
 
@@
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 114 & 114 & 122 & 126 & 142 & 156 \\ \hline
-LU32PEEng & 170 & 176 & 182 & 206 & 318 & 396 \\ \hline
-LU8PEEng & 110 & 114 & 116 & 138 & 164 & 236 \\ \hline
-mcml & 108 & 114 & 106 & 124 & 150 & 238 \\ \hline
-mkDelayWorker32B & 74 & 78 & 80 & 88 & 98 & 138 \\ \hline
-stereovision0 & 62 & 60 & 62 & 76 & 132 & 146 \\ \hline
-stereovision1 & 102 & 104 & 106 & 108 & 144 & 186 \\ \hline
-stereovision2 & 160 & 158 & 164 & 200 & 264 & 432 \\ \hline
-Geometric mean & 107 & 109 & 111 & 126 & 165 & 221 \\ \hline
-\end{tabular}
-\caption{Minimum channel width when \textit{number of cuts = 1} and \textit{increased delay = 1ns}.}
-\label{table:1cut_minW}
-\end{table}
+%   - Impact of aspect ratio
+\subsection{Impact of Aspect Ratio}
+\label{aspect_ratio_impact}
+TODO: Javeed
 
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 26.5 & 32.1 & 28.3 & 30.2 & 27.5 & 30.0 \\ \hline
-LU32PEEng & 113 & 113 & 113 & 115 & 114 & 111 \\ \hline
-LU8PEEng & 117 & 118 & 116 & 114 & 114 & 116 \\ \hline
-mcml & 82.5 & 82.0 & 83.0 & 84.5 & 83.2 & 77.9 \\ \hline
-mkDelayWorker32B & 11.2 & 11.4 & 11.2 & 9.9 & 10.6 & 11.5 \\ \hline
-stereovision0 & 4.5 & 5.1 & 5.0 & 4.9 & 4.9 & 5.4 \\ \hline
-stereovision1 & 7.4 & 8.2 & 8.4 & 7.2 & 7.2 & 7.9 \\ \hline
-stereovision2 & 21.3 & 23.1 & 20.9 & 22.3 & 21.6 & 20.9 \\ \hline
-Geometric mean & 26.3 & 28.1 & 27.2 & 26.8 & 26.5 & 27.3 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 1} and \textit{increased delay = 1ns}.}
-\label{table:1cut_path}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 114 & 116 & 130 & 140 & 172 & 272 \\ \hline
-LU32PEEng & 170 & 180 & 186 & 198 & 264 & 426 \\ \hline
-LU8PEEng & 110 & 114 & 116 & 124 & 158 & 246 \\ \hline
-mcml & 108 & 110 & 126 & 138 & 114 & 150 \\ \hline
-mkDelayWorker32B & 76 & 80 & 82 & 88 & 100 & 132 \\ \hline
-stereovision0 & 62 & 58 & 60 & 88 & 114 & 166 \\ \hline
-stereovision1 & 104 & 108 & 116 & 136 & 184 & 296 \\ \hline
-stereovision2 & 160 & 156 & 162 & 188 & 246 & 362 \\ \hline
-Geometric mean & 108 & 109 & 116 & 132 & 160 & 237 \\ \hline
-\end{tabular}
-\caption{Minimum channel width when \textit{number of cuts = 2} and \textit{increased delay = 1ns}.}
-\label{table:2cut_minW}
-\end{table}
-
-\begin{table}[!htbp]
-\begin{tabular}{|l|r|r|r|r|r|r|}
-\hline
-\% wires cut & 0 & 20 & 40 & 60 & 70 & 80 \\ \hline \hline
-bgm & 31.0 & 30.3 & 28.3 & 36.1 & 35.9 & 37.5 \\ \hline
-LU32PEEng & 113 & 113 & 119 & 113 & 117 & 112 \\ \hline
-LU8PEEng & 118 & 118 & 117 & 116 & 118 & 114 \\ \hline
-mcml & 82.9 & 86.3 & 81.3 & 80.9 & 82.4 & 82.8 \\ \hline
-mkDelayWorker32B & 12.3 & 12.1 & 9.8 & 11.1 & 13.5 & 11.3 \\ \hline
-stereovision0 & 6.4 & 5.7 & 6.3 & 5.3 & 5.3 & 5.9 \\ \hline
-stereovision1 & 9.3 & 10.4 & 9.5 & 9.6 & 10.2 & 9.9 \\ \hline
-stereovision2 & 22.9 & 24.2 & 24.0 & 21.6 & 24.2 & 25.5 \\ \hline
-Geometric mean & 29.5 & 29.7 & 28.6 & 28.8 & 30.4 & 30.1 \\ \hline
-\end{tabular}
-\caption{Critical path delay (in \textit{ns}) when \textit{number of cuts = 2} and \textit{increased delay = 1ns}.}
-\label{table:2cut_path}
-\end{table}
-
+%---------------------------------------------------------------
+%---------------------------------------------------------------
+%---------------------------------------------------------------
 \section{Conclusions and Future Work}
 \label{conclusionSection}
-
 We have extended VPR to model and optimize for interposer-based multi-FPGA systems.

### 2014-12-07 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
testing my svn access
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4519 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
changing paper template from ACM Conference to IEEETran. this also requires changing some of the Latex function calls. Also adding the section on placer optimization results.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4521 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
-% THIS IS SIGPROC-SP.TEX - VERSION 3.1
-% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
-% APRIL 2009
+%% bare_jrnl.tex
+%% V1.4
+%% 2012/12/27
+%% by Michael Shell
+%% see http://www.michaelshell.org/
+%% for current contact information.
+%%
+%% This is a skeleton file demonstrating the use of IEEEtran.cls
+%% (requires IEEEtran.cls version 1.8 or later) with an IEEE journal paper.
+%%
+%% Support sites:
+%% http://www.michaelshell.org/tex/ieeetran/
+%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
+%% and
+%% http://www.ieee.org/
+
+\documentclass[journal]{IEEEtran}
 %
-% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
-% LaTeX2e document class file for Conference Proceedings submissions.
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) Page numbering
-% ---------------------------------------------------------------------------------------------------------------
-% It is an example which *does* use the .bib file (from which the .bbl file
-% is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission,
-% you need to 'insert'  your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% Questions regarding SIGS should be sent to
-% Adrienne Griscti ---> griscti@acm.org
-%
-% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
-% Gerald Murray ---> murray@hq.acm.org
-%
-% For tracking purposes - this is V3.1SP - APRIL 2009
+% If IEEEtran.cls has not been installed into the LaTeX system files,
+% manually specify the path to it like:
+% \documentclass[journal]{../sty/IEEEtran}
 
-\documentclass{sig-alternate-2013}
 \usepackage{graphicx}
 \usepackage{subfigure}
-
 \usepackage{epstopdf}
 \usepackage{enumitem}
-%\linespread{0.94}
+\usepackage{cite}
+\usepackage[cmex10]{amsmath}
 
-\newfont{\mycrnotice}{ptmr8t at 7pt}
-\newfont{\myconfname}{ptmri8t at 7pt}
-\let\crnotice\mycrnotice%
-\let\confname\myconfname%
-
-\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
-Copyrights for components of this work owned by others than the author(s) must be honored.
-Abstracting with credit is permitted.
-To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
-Request permissions from permissions@acm.org.}
-\conferenceinfo{FPGA'14,}{February 26--28, 2014, Monterey, CA, USA. \\
-{\mycrnotice{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}}}
-\copyrightetc{ACM \the\acmcopyr}
-\crdata{978-1-4503-2671-1/14/02\ ...\$15.00.\\
-http://dx.doi.org/10.1145/2554688.2554776}
-
-\clubpenalty=10000
-\widowpenalty = 10000
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
 
 \begin{document}
-
-% --- Author Metadata here ---
-%\conferenceinfo{FPGA}{'14 Monterey, California USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
-
+%
+% paper title
+% can use linebreaks \\ within to get better formatting as desired
+% Do not put math or special symbols in the title.
 \title{CAD and Routing Architecture for Interposer-Based Multi-FPGA Systems}
 %
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
 %
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
+% author names and IEEE memberships
+% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
+% a structure at a ~ so this keeps an author's name from being broken across
+% two lines.
+% use \thanks{} to gain access to the first footnote area
+% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
+% was not built to handle multiple paragraphs
 %
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
 
-\numberofauthors{2}
-%
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Andre Hahn Pereira\\
-	   \affaddr{Computer and Digital Systems Engineering Department - Escola Polit\'{e}cnica}\\
-       \affaddr{University of S\~{a}o Paulo}\\
-       \affaddr{S\~{a}o Paulo, SP}\\
-       \affaddr{Brazil}\\
-       \email{andre.hahn@usp.br}
-% 2nd. author
-\alignauthor
-Vaughn Betz\\
-       \affaddr{Department of Electrical and Computer Engineering}\\
-       \affaddr{University of Toronto}\\
-       \affaddr{Toronto, ON}\\
-       \affaddr{Canada}\\
-       \email{vaughn@eecg.utoronto.ca}
-}
-\date{18 December 2013}
+		% 1st author
+		Ehsan~Nasiri,
+		% 2nd author
+		Javeed~Shaikh,
+		% 3rd author
+		Andre~Hahn~Pereira,
+		% 4th author
+        and~Vaughn~Betz,~\IEEEmembership{Member,~IEEE,}%
+\thanks{E. Nasiri is with Altera Corporation, Toronto Technology Center, Toronto, Ontario, Canada (email: enasiri@altera.com)}% 
+\thanks{J. Shaikh is with Google, Mountain View, California, USA (email: jvds@google.com)}%
+\thanks{A. H. Pereira is with the Computer and Digital Systems Engineering Department - Escola Polit\'{e}cnica, University of S\~{a}o Paulo, S\~{a}o Paulo, SP, Brazil (email: andre.hahn@usp.br)}%
+\thanks{V. Betz is with the Department of Electrical and Computer Engineering, University of Toronto, Toronto, Ontario, Canada (email: vaughn@eecg.utoronto.ca)}%
+\thanks{Manuscript received January 19, 2014.}}
 
+% make the title area
 \maketitle
+
+% As a general rule, do not put math, special symbols or citations in the abstract or keywords.
 \begin{abstract}
-%
 Interposer-based multi-FPGA systems are composed of multiple FPGA dice connected through a silicon interposer.
@@
 
-\category{B.7.1}{Integrated Circuits}{Types and Design Style}[VLSI]
-\category{B.7.2}{Integrated Circuits}{Design Aids}[Placement and Routing]
-\terms{Algorithms, Design}
-\keywords{FPGA; Silicon interposer; 2.5D ICs}
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+Algorithms, Design, FPGA, Silicon interposer, 2.5D ICs. 
+\end{IEEEkeywords}
 
@@
 %---------------------------------------------------------------
-\section{Architecture models}
+\section{Architecture Models}
 \label{archSection}
@@
 %	- number of cuts
-\subsection{Number of cuts}
+\subsection{Number of Cuts}
 \textit{Number of cuts} describes how many cuts are made to the interposer-based FPGA versus a monolithic die.
@@
 %   - % wires cut
-\subsection{\% wires cut}
+\subsection{\% Wires Cut}
 This variable describes and models the reduced connectivity between different dice by specifying the fraction of routing wires that are removed at the border between dice.
@@
 %   - Added delay
-\subsection{Increased delay}
+\subsection{Increased Delay}
 Interposer wires are longer and wider than on-die wires and have microbumps on each end.
@@
 %   - Interposer-Routing interface
-\subsection{Interposer-Routing interface}
+\subsection{Interposer-Routing Interface}
 \label{sec:interposer-routing-interface}
@@
 
-\begin{enumerate}[label=(\alph*)]
+\begin{enumerate}
 \item Fan-in Transfer for Interposer Wires\\
@@
 
-\subsection{Placer optimization}
+\subsection{Placer Optimization}
 \label{cad_enh_placer_subsection}
@@
 %   - Delay
-\subsubsection{Placer timing cost}
+\subsubsection{Placer Timing Cost}
 The usual VPR timing cost is a (criticality-weighted) summation of the estimated delay (given the current placement) of every connection required by the circuit~\cite{timing2000}:
 
-\begin{equation} \label{eq:timing_eq_full}
-\begin{split}
-Timing\_Cost = \sum_{\forall i, j \subset circuit} delay(\Delta x_{ij},\Delta y_{ij}) \times \\ 
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:timing_eq_full}
+Timing\_Cost=\sum_{\forall i, j \subset circuit} delay(\Delta x_{ij},\Delta y_{ij}) \times \\ 
 criticality(i,j)
-\end{split}
-\end{equation}
+\end{multline}
+\setlength{\arraycolsep}{5pt}
+
 where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
@@
 
-\begin{equation} \label{eq:timing_cost}
-\begin{split}
-delay(i,j) =& delay(\Delta x_{ij}, \Delta y_{ij}) + \\
-&times\_crossed(i,j) \times delay\_increase
-\end{split}
-\end{equation}
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:timing_cost}
+delay(i,j) = delay(\Delta x_{ij}, \Delta y_{ij}) + \\
+times\_crossed(i,j) \times delay\_increase
+\end{multline}
+\setlength{\arraycolsep}{5pt}
 
@@
 %   - Wiring / cut cost
-\subsubsection{Placer wiring cost}
+\subsubsection{Placer Wiring Cost}
 The bounding box cost estimates the amount of wiring required for a net, based on the number of pins and size of the net's bounding box. VPR's original formulation is~\cite{betz1997vpr}:
 
-\begin{equation} \label{eq:bbcost}
-\begin{split}
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:bbcost}
 wiring\_cost_{orig} = \sum_{n=1}^{N_{nets}} q(n) \times [\frac{bb_x(n)}{avg\_chanx\_W(n)} + \\
 \frac{bb_y(n)}{avg\_chany\_W(n)}]
-\end{split}
-\end{equation}
+\end{multline}
+\setlength{\arraycolsep}{5pt}
 
@@
 
-\begin{equation} \label{eq:total_wiring}
+\setlength{\arraycolsep}{0.0em}
+\begin{equation}
+\label{eq:total_wiring}
 wiring\_cost = wiring\_cost_{orig} + cut\_cost
 \end{equation}
+\setlength{\arraycolsep}{5pt}
 
@@
 
-\begin{equation} \label{eq:cost2}
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:cost2}
 cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbHeight(i) \times times\_crossed(n)
-\end{equation}
+\end{multline}
 
 where 
-\begin{equation} \label{eq:cprime}
-\begin{split}
+\setlength{\arraycolsep}{0.0em}
+\begin{equation}
+\label{eq:cprime}
 C' = \frac{ratio\_wires\_cut}{avg\_chany\_W(n)}
-\end{split}
 \end{equation}
+\setlength{\arraycolsep}{5pt}
 
@@
 
-\subsection{Router optimization}
+This cost function performs better than several others as shown in \cite{hahn2014cad}, and we believe the key to the good performance of this cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
+Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
+Note that bounding box (a) and (b) both cross the cutline, but (b) is penalized less by (\ref{eq:cost2}) because bounding box (b) is mostly on the lower side of the cutline; it is more likely that later placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
+Cost function (\ref{eq:cost2}) penalizes bounding box (a) more than (b) and (b) more than (c) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{bb_new.eps}
+\caption{Illustration of three different scenarios for the wiring cost.
+The dashed green box shows a case where \textit{(a)} it crosses the interposer, the dotted black box shows a case where \textit{(b)} it barely crosses the cutline, and the solid blue one shows a case where \textit{(c)} the bounding box does not span the cutline.}
+\label{fig:bb_illustration}
+\end{figure}
+
+\subsection{Router Optimization}
 \label{sec:router_optimizations}
@@
 The VPR router uses the following to calculate a cost for a routing node \textit{r} on the route from \textit{i} to \textit{j}:
-\begin{equation} \label{eq:routing_cost}
-\begin{split}
-Cost(r) = Criticality(i,j) \times Delay_{Elmore}(r, topology) + \\ [1-Criticality(i,j)] \times Congestion(r)
-\end{split}
-\end{equation}
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:routing_cost}
+Cost(r) = Criticality(i,j) \times Delay_{Elmore}(r, topology) \\ + [1-Criticality(i,j)] \times Congestion(r)
+\end{multline}
+\setlength{\arraycolsep}{5pt}
 
@@
 
-\begin{equation} \label{eq:lookahead_cost}
-\begin{split}
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:lookahead_cost}
 Total\_Cost(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectedCost(r,j)
-\end{split}
-\end{equation}
+\end{multline}
+\setlength{\arraycolsep}{5pt}
 
@@
 
-\begin{equation} \label{eq:expected_routing_cost}
-\begin{split}
-ExpectedCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(r,j);
-\end{split}
-\end{equation}
+\setlength{\arraycolsep}{0.0em}
+\begin{multline}
+\label{eq:expected_routing_cost}
+ExpectedCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(r,j)
+\end{multline}
+\setlength{\arraycolsep}{5pt}
 
@@
 Thus, for interposer-based architectures, we replace $T_{delay}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{delay}$ to include the delay of expected interposer hops from \textit{r} to \textit{j}.
-\begin{equation} \label{eq:tdelay}
-\begin{split}
+
+\begin{multline}
+\label{eq:tdelay}
 T^\prime_{delay} = T_{delay} + T_{interposer} \times ExpectedInterposerHops
-\end{split}
-\end{equation}
+\end{multline}
+\setlength{\arraycolsep}{5pt}
 
@@
 %   - Effectiveness / results
-\subsection{Effectiveness of the enhancements}
+\subsection{Effectiveness of the Enhancements}
 \label{sec:CADeffect}
@@
 
-\subsubsection{Effectiveness of placer optimizations}
+\subsubsection{Effectiveness of Placer Optimizations}
 \label{sec:effectiveness_of_placer_enhancements}
-Enabling the enhancements described in Section~\ref{cad_enh_placer_subsection} improves the results ....
-haven't finished running these yet.
-So, <TODO EHSAN>: fill this section with actual results.
-For now, let's use a graph from Andre as a placeholder!
-The graphs show that combining the timing with the bounding box optimizations led to an average of 20\% improvement on the area-delay product and the most significant contribution was made by incorporation of \textit{cut\_cost} into the wire cost function.
-
-We believe the key to the good performance of the new placer cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
-Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
-Note that bounding box (a) and (b) both cross the cutline, but (b) is penalized less by (\ref{eq:cost2}) because bounding box (b) is mostly on the lower side of the cutline; it is more likely that later placement changes will result in the bounding box moving entirely below the cutline, reducing interposer wiring demand.
-Cost function (\ref{eq:cost2}) penalize bounding box (a) more than (b) and (b) more than (c) to guide placement to gradually move bounding boxes to one side or the other of a cutline.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\linewidth]{bb_new.eps}
-\caption{Illustration of three different scenarios for the wiring cost.
-The dashed green box shows a case where \textit{(a)} it crosses the interposer, the dotted black box shows a case where \textit{(b)} it barely crosses the cutline, and the solid blue one shows a case where \textit{(c)} the bounding box does not span the cutline.}
-\label{fig:bb_illustration}
-\end{figure}
+Enabling the enhancements described in Section~\ref{cad_enh_placer_subsection} improves the minimum routable channel width and circuit speed.
+As shown in Figure~\ref{fig:comparison_minW}, for architectures with low \textit{\% wires cut} (high interposer capacity), the placer optimizations are only marginally beneficial compared to the original VPR placer since the interposer wires are not scarce; but as more wires are cut and we reach the realistic interposer capacity of commercial FPGAs, the placer optimizations provide higher and higher reduction in minimum channel width.
+When 80\% of wires are cut, the placer optimizations result in 11.3\% reduction in minimum channel width.
 
@@
 \centering
-\includegraphics[width=\linewidth]{comparison_minW.eps}
-\caption{Minimum channel width vs. \textit{\% wires cut}.}
+\includegraphics[width=\linewidth]{placer_opt_minChW.eps}
+\caption{Impact of placer optimizations on minimum channel width.}
 \label{fig:comparison_minW}
@@
 
-\subsubsection{Effectiveness of router optimizations}
+The optimizations also help the placer to place related entities on the same die to reduce the number of inter-die signal crossings and hence increase circuit speed.
+At 80\% wires cut, the placer optimizations result in 8.4\% higher circuit speed.
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{placer_opt_fmax.eps}
+\caption{Impact of placer optimizations on circuit speed.}
+\label{fig:comparison_minW}
+\end{figure}
+
+\subsubsection{Effectiveness of Router Optimizations}
 \label{sec:effectiveness_of_router_enhancements}
@@
 %---------------------------------------------------------------
-\section{Architecture results}
+\section{Architecture Results}
 \label{resultsSection}
@@
 
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
+% use section* for acknowledgement
+\section*{Acknowledgment}
+This work was supported by a Ci\^{e}ncia sem Fronteiras scholarship from CNPq - Brazil and the NSERC/Altera Industrial Research Chair in Programmable Silicon.
+Some of our computations were performed on the GPC supercomputer at the SciNet HPC Consortium.
+SciNet is funded by: the Canada Foundation for Innovation under the auspices of Compute Canada; the Government of Ontario; Ontario Research Fund - Research Excellence; and the University of Toronto.
 
-This work was supported by a Ci\^{e}ncia sem Fronteiras scholarship from CNPq - Brazil and the NSERC/Altera Industrial Research Chair in Programmable Silicon.
-%
-\bibliographystyle{abbrv}
-\bibliography{fpga2014paper}
+\bibliographystyle{IEEEtran}
+% argument is your BibTeX string definitions and bibliography database(s)
+\bibliography{IEEEabrv,./fpga2014paper}
 
-\balancecolumns
-% That's all folks!
+% that's all folks
 \end{document}

### 2014-12-08 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
fixing the graph that had a duplicate label as an earlier graph
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4522 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Adding incomplete draft of partitioning section.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4523 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
+\usepackage[cmex10]{amsmath}
 \usepackage{graphicx}
+\usepackage{mathtools}
 \usepackage{subfigure}
@@
 \usepackage{cite}
-\usepackage[cmex10]{amsmath}
+
+\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
 
@@
 \caption{Impact of placer optimizations on circuit speed.}
-\label{fig:comparison_minW}
+\label{fig:comparison_fmax}
 \end{figure}
@@
 %---------------------------------------------------------------
-\section{Adding Partitioning to the CAD Flow}
-\label{sec:CADflow}
-TODO: Javeed
+\input{partitioning.tex}
 
@@
 \label{num_dice_impact}
-TODO: Javeed\\
+TODO: Javeed
 To examine the impact of the number of die used to construct an interposer-based FPGA, we varied the number of cuts from 1 to 3 (which varies the number of die from 2 to 4).

### 2014-12-08 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Adding incomplete draft of partitioning section.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4523 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\section{Adding Partitioning to the CAD Flow}
\label{sec:CADflow}
TODO: Javeed

Given that the number of signals that can cross the interposer between FPGA dice is considerably more limited than within an FPGA, using a partitioner to divide the circuit into one partition per die is a promising CAD flow.

\subsection{CAD Flow Variations}
The overall CAD flow, including the partitioning steps, is shown in Figure \ref{fig:partitioning_cad_flow}.
\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{partitioning_cad_flow.eps}
\caption{Possible ways to add partitioning to the VPR CAD flow.}
\label{fig:partitioning_cad_flow}
\end{figure}

\subsubsection{Partitioner Stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. In this flow the partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new legality constraint.

The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can to pass the partition information to the placement stage as well. The post-packing placement constraints can be derived from the primitive block partitioning, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:

\begin{enumerate}
  \item Partitioning the primitive netlist to guide the packer, leaving the placer unconstrained.
  \item Partitioning the primitive netlist to guide the packer, and constraining the placer based on constraints derived from partitioning, as described above.
  \item Leaving the packer unconstrained, and partitioning the function block (clustered) netlist to guide the placer.
  \item Leaving both the packer and the placer unconstrained. This is the base flow that does not use partitioning, as described in Section \ref{cadSection}.
\end{enumerate}

\subsubsection{Partitioning Tool}
Two well-known freely available graph partitioning tools are Metis [] and hMetis []. Metis operates on graphs, while hMetis operates on hypergraphs -- a hypergraph is a generalization of a graph, where each edge can connect any number of vertices. It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins. Based on that criterion alone, hMetis is the most appropriate choice of partitioner. However, there are other issues to consider for the circuit netlist partitioning use case.

Graph partitioning algorithms generate partitions of a circuit while satisfying ``balance constraints''. The balance constraints exist to drive the algorithm away from the trivial partitioning solution where all vertices are in the same partition, leaving all other partitions empty. An example of a basic balance constraint is one which requires an equal number of vertices in each partition. Generally such a perfect balance overly restricts the solution, so some fractional ``unbalance'' is allowed; in this work we use $0.05$ ($5\%$). For two partitions, $P1$ and $P2$, this translates to $[ref Metis]$:

\begin{equation}\label{eq:balance_constraint}
\frac{|P1|}{|P1 \cup P2|} \in [0.5 - unbalance, 0.5 + unbalance]
\end{equation}

To derive an appropriate balance constraint for our use case, consider that FPGA circuit netlists may contain different block types such as LUTs, registers, RAM slices, and multipliers, among others. The basic balance constraint of (\ref{eq:balance_constraint}), applied to the two partition case with only LUTs and multipliers would ensure that the total block count (sum of the LUT count + multiplier count) in each partition is roughly balanced.

In our architecture, each FPGA die is identical, and therefore each die has the same number of resources as every other die. However, the number of available instances of each block type is not equal -- there are usually more LUTs than multipliers, for example. The problem with the simple balance constraint in the case of multiple block types is that the capacity of each partition is not taken into account on a per-block-type basis. It is therefore possible to generate partitions that cannot be realized because the number of resources of each type is not balanced across partitions, even though the total number of resources is balanced. For example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which would not be physically realizable because partition 1 may not have enough LUTs to implement the desired partitioning (and the LUTs in partition 2 would be completely unused.) A technique is to enforce a balance constraint on each type of block, $i$:

\begin{equation}
\forall i, \abs*{ \frac{| P1_i |}{|P1_i \cup P2_i |} - 0.5} < unbalance
\end{equation}
where $P1_i$ denotes the set of all blocks of type $i$ in partition 1.

We refer to this type of per-block set of constraints as heterogeneous balance constraints. Note that there are complex legality constraints governing which LUTs and FFs can be packed into legal logic blocks, as well as which multipliers can be packed into legal DSP blocks [ref AAPack]. As the partitioner is not aware of these constraints, flow 3 above (which runs the partitioner after function block packing) will be able to more precisely balance resource use across partitions.

The latest public version of hMetis does not support heterogeneous balance constraints, thourgh Metis does. Metis was used in our CAD flow. However, as discussed earlier, the circuit netlist is most naturally represented as a hypergraph, and so it needs to be transformed to a graph before Metis can process it.

\subsubsection{Hypergraph to Graph Transformation}
We transformed the circuit netlist hypergraph to a graph in the following way:
\begin{enumerate}
\item For each hyperedge of the circuit netlist hypergraph, we generated a graph (the \emph{per-net subgraph}).
\item We combined all of the per-net subgraphs, summing the edge weights for edges appearing in more than one per-net subgraph, to generate the total netlist graph.
\end{enumerate}

We explored several ways to generate the per-net subgraphs. The two parameters in this operation are the graph topology and the edge weight scheme. We considered two subgraph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.

\begin{figure}[!htbp]
\centering
\subfigure[]{\includegraphics[width=0.45\linewidth]{star.eps}}
\subfigure[]{\includegraphics[width=0.45\linewidth]{clique.eps}}
\caption{(a) Star and (b) clique graph topologies.}
\label{fig:star_clique}
\end{figure}

We assigned the same edge weight to every edge of the generated subgraph. Edge weights were computed based on the number of vertices in the hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$. Note that the vertices in the generated total netlist graph are the same as the vertices in the original circuit netlist hypergraph.

Figure \ref{fig:hyperedge_to_graph} shows how a netlist of three blocks and two nets is transformed into its component subgraphs.

\begin{figure}[!htbp]
\centering
\subfigure[]{\includegraphics[width=0.45\linewidth]{hg2g_star.pdf}}
\subfigure[]{\includegraphics[width=0.45\linewidth]{hg2g_clique.pdf}}
\caption{The hypergraph to graph transformation with (a) the star topology, and (b) the clique topology.}
\label{fig:hyperedge_to_graph}
\end{figure}

The parameters of graph topology and edge weight were evaluated by generating the total primitive netlist graph and running it through the Metis partitioner to generate two partitions. The results of the different schemes were ranked based on the hyperedge cutsize, which is the number of hyperedges crossing the cutline, as illustrated in Figure ??. The hyperedge cutsize is a proxy for the routability of the circuit, as the cutsize is the number of nets that cross the interposer boundary. To compute the hyperedge cutsize, we applied the partitioning result (which assigns each vertex to a partition number) to the original untransformed hypergraph, and counted the number of hyperedges whose vertices were spread across different partitions. We computed the hyperedge cutsize for every combination of graph topology and graph edge weight scheme, across several circuits. The results are shown in Figure \ref{fig:graph_topology_cutsize}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{graph_topology_cutsize.eps}
\caption{Hypergraph cut size achieved by Metis with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
\label{fig:graph_topology_cutsize}
\end{figure}

The performance of the clique topology is significantly worse than that of the star topology, regardless of edge weight scheme. An intuitive explanation for this result is that with the clique topology, the partitioner does not have knowledge of which vertex is the source of a net and which vertices are sinks. In contrast, the star topology clearly differentiates sources from sinks and this appears to give the partitioner an “anchor point” that pulls all hyperedge fanouts toward the source. Additionally, if $n$ is the average number of vertices in a hyperedge, the clique topology generates $O(n^2)$ edges while the star topology generates only $O(n)$ edges, so the star topology saves both memory and runtime.

For both topologies, the $1/n$ edge weight scheme gives the smallest cut size on average. This can be explained intuitively in terms of the total edge weight over all edges of the graph. The star $1$ (constant) edge weight scheme assigns a total weight of $n$ to each net, which heavily weights high-fanout nets. In contrast, the star $1/n$ scheme assigns the same total weight of $1$ to each net. The star $1/n^2$ scheme penalizes high-fanout nets relative to lower-fanout nets. Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ scheme achieves this.

To validate the choice of hyperedge cutsize as a proxy for circuit routability, we ran the partitioning CAD flow for each graph topology and edge weight scheme, across several circuits. We use an unbalance of $5\%$, split the clustered netlist into two partitions (constraining only the placer), cut $80\%$ of the wires crossing the interposer, and impose a $1ns$ delay penalty for wires crossing the interposer. To accommodate unbalance in the placement engine, we increase the size of the grid of complex blocks that make up the FPGA device. Relative to the minimum device size required for placement (without partitioning constraints), we add $10\%$ to the complex block grid width and $10\%$ to the grid height. Figure ?? shows the geometric mean of the minimum channel width required for a successful route, for each of the graph topology and edge weight combinations.

### 2014-12-09 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
More updates to the partitioning sections of the paper.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4524 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \label{num_dice_impact}
-TODO: Javeed
-To examine the impact of the number of die used to construct an interposer-based FPGA, we varied the number of cuts from 1 to 3 (which varies the number of die from 2 to 4).
-In these experiments, the \textit{delay increase} was kept constant at $1ns$.
-
-As shown in Figure \ref{fig:cuts_minW}, the number of cuts does have an impact on the minimum channel width, but not a very large or constant one.
-At 80\% of wires cut the experiments with 2 and 3 cuts had the minimum channel width only 8\% and 10\%, respectively, greater than the scenario with only 1 cut.
-At lower values of \textit{\% wires cut} the difference was even smaller.
-
-Figure \ref{fig:cuts_crit} shows that the number of dice in an interposer-based FPGA impacts circuit speed significantly, as the critical path delay rises significantly as systems have more cuts.
-Recall that a monolithic FPGA has a geometric average critical path delay of $23.7ns$ for our test architecture.
-With \textit{number of cuts} $= 1$ the critical path delay increases by 3 to 4 ns, while with \textit{number of cuts} $= 2$ the critical path delay increases by 5 to 6 ns.
-With \textit{number of cuts} $= 3$, the critical path delay is typically 7 ns higher than that of a monolithic FPGA.
+To examine the impact of the number of dice used to construct an interposer-based FPGA, we measured the minimum channel width for an FPGA device with four dice, varying the fraction of wires cut.
+We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section (TODO JAVEED -- spell them out?).
+The results, shown in Figure \ref{fig:wires_cut_4part}, show the same trend as Figure \ref{fig:wires_cut}.
 
@@
 \centering
-\includegraphics[width=\linewidth]{cuts_minW.eps}
-\caption{Minimum channel width vs. \textit{\% wires cut} for 1, 2 and 3 cuts and $1ns$ of \textit{delay increase}.}
-\label{fig:cuts_minW}
+\includegraphics[width=\linewidth]{wires_cut_4part.eps}
+\caption{Impact of partitioning-based CAD flow on routability for a range of architectures with 4 dice. TODO JAVEED -- LU32 90\% result is missing}
+\label{fig:wires_cut_4part}
 \end{figure}
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{cuts_crit_path.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 1, 2 and 3 cuts and $1ns$ of \textit{delay increase}.}
-\label{fig:cuts_crit}
-\end{figure}
+TODO JAVEED -- show critical path delay? (it's uninteresting)
 
@@
 TODO: Javeed
+We investigated the implications of the device aspect ratio on routability for the four-partition case.
+In particular, we compared a device with a square interposer (four rectangular dice) to a device with a rectangular interposer (four square dice.) We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section.
+The results are shown in Table \ref{table:squares}.
+
+\begin{table}[!htbp]
+\centering
+\begin{tabular}{|l|r|r|}
+\hline
+ & Square Interposer & Square Dice \\ \hline \hline
+bgm & 102 & 306 \\ \hline
+stereovision0 & 86 & 256 \\ \hline
+stereovision1 & 166 & 256 \\ \hline
+LU8PEEng & 156 & 126 \\ \hline
+mkDelayWorker32B & 96 & 194 \\ \hline
+mcml & 196 & 496 \\ \hline
+LU32PEEng & 256 & 264 \\ \hline
+geomean & 140.7481367 & 251.805728 \\ \hline
+\end{tabular}
+\caption{Minimum channel width for square interposer and square dice configurations.}
+\label{table:squares}
+\end{table}
+
+The square interposer configuration performs significantly better than the square dice configuration.
+
+TODO JAVEED -- Vaughn commented that we should talk about which is easier to manufacture.
+I think the square dice device would actually be easier to manufacture, because the more complex dice will fit well on a round wafer.
+Isn't it fine for the interposer to be non-square?
 

### 2014-12-09 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
More updates to the partitioning sections of the paper.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4524 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-\subsubsection{Partitioner Stage}
+\subsubsection{Partitioner Stage}\label{sec:partitioner_stage}
 As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow.
@@
 \begin{enumerate}
+  \item Leaving both the packer and the placer unconstrained. This is the base flow that does not use partitioning, as described in Section \ref{cadSection}.
   \item Partitioning the primitive netlist to guide the packer, leaving the placer unconstrained.
+  \item Leaving the packer unconstrained, and partitioning the function block (clustered) netlist to guide the placer.
   \item Partitioning the primitive netlist to guide the packer, and constraining the placer based on constraints derived from partitioning, as described above.
-  \item Leaving the packer unconstrained, and partitioning the function block (clustered) netlist to guide the placer.
-  \item Leaving both the packer and the placer unconstrained. This is the base flow that does not use partitioning, as described in Section \ref{cadSection}.
 \end{enumerate}
@@
 
-\subsubsection{Hypergraph to Graph Transformation}
+\subsubsection{Hypergraph to Graph Transformation}\label{sec:hypergraph_to_graph}
 We transformed the circuit netlist hypergraph to a graph in the following way:
@@
 Relative to the minimum device size required for placement (without partitioning constraints), we add $10\%$ to the complex block grid width and $10\%$ to the grid height.
-Figure ??
-shows the geometric mean of the minimum channel width required for a successful route, for each of the graph topology and edge weight combinations.
+Figure \ref{fig:graph_topology_mcw} shows the geometric mean of the minimum channel width required for a successful route, for each of the graph topology and edge weight combinations.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{graph_topology_mcw.eps}
+\caption{Minimum channel width achieved by VPR with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
+\label{fig:graph_topology_mcw}
+\end{figure}
+
+The star topology with 1/n edge weights achieves the best minimum channel width, confirming that hyperedge cutsize is a good proxy for routability.
+Consequently, we use this scheme in all future results in this paper.
+Unlike routability, we found that the post-routing critical path delay was not strongly impacted by the hypergraph conversion scheme used.
+
+\subsubsection{Partitioner Stage Results}\label{sec:partitioner_stage_results}
+We compare the performance of the four CAD flow variations described in Section \ref{sec:partitioner_stage} on the 8 largest VTR benchmarks.
+We again use an unbalance of $5\%$ and split the clustered netlist into two partitions (constraining the packer and/or the placer, as required by the flow under test).
+The interposer parameters (\% wires cut = $80\%$ and delay increase = 1 ns) and bloat factor ($10\%$ in each dimension) are the same as in Section \ref{sec:hypergraph_to_graph}.
+Figure \ref{fig:flows_mcw} shows the minimum routable channel width for each partitioning CAD flow variation.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{vpr_flows_mcw.eps}
+\caption{Minimum channel width achieved by VPR with the different partitioning flows.}
+\label{fig:flows_mcw}
+\end{figure}
+
+On average, the best minimum channel width was obtained when constraining both packing and placement.
+However, constraining packing alone performed significantly better than constraining placement alone.
+This shows that the VPR packer strongly benefits from partitioning constraints when targeting an interposer-based architecture, likely because it forbids the packing of primitives that are not naturally related into one function block.
+Interestingly, other recent work targeting conventional FPGAs has shown that using a partitioner to guide packing benefits routability [Wenyi Feng, et al, “Rent’s Rule Based Packing for Routability Optimization,” FPGA 2014 ].
+On the other hand, The VPR placement algorithm with the enhancements described in Section 4 sees only a modest routability benefit from the partitioning constraints.
+
+For the same set of circuits, we computed the critical path delay at a channel width equal to 1.3x the minimum. The results are shown in Figure \ref{fig:flows_crit_path}.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{vpr_flows_crit_path.eps}
+\caption{Critical path delay achieved by VPR with the different partitioning flows.}
+\label{fig:flows_crit_path}
+\end{figure}
+
+On average, the critical path delay did not vary significantly between the different partitioning CAD flows, but it appears that partitioning before packing may slightly improve circuit speed, while enforcing these partitioning constraints during placement may slightly reduce circuit speed.
+
+Considering both routability and delay, the best CAD flow constrained both packing and placement with the partitioning result and this is the flow we use in the remainder of this paper.
+
+\subsubsection{\% wires cut}
+To see how well the CAD flow works for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits.
+As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of [] for FPGAs where more than half the wires are cut at interposer boundaries.
+It is slightly worse than the traditionally FPGA CAD flow even for an architecture with 0\% of the wires cut, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{wires_cut.eps}
+\caption{Impact of partitioning-based CAD flow on routability for a range of architectures.}
+\label{fig:wires_cut}
+\end{figure}
+
+\subsubsection{Bloating of the Clustered Netlist due to Partitioning Constraints}
+The additional constraints on the packer imposed by partitioning may result in a less dense packing.
+Figure \ref{fig:packing_bloat} compares the number of complex blocks in the partitioning-constrained and unconstrained clustered netlists.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{packing_bloat.eps}
+\caption{Packing bloat.}
+\label{fig:packing_bloat}
+\end{figure}
+
+On average, the packing-constrained netlist contained $3.2\%$ more complex blocks than the unconstrained netlist.
+The LU8PEEng and LU32PEEng circuits were the most affected by this packing bloat.
+In those circuits, we found that the packing bloat was due almost entirely to poor packing of memory blocks; an interesting area for future work would be to augment the partitioner to understand that some RAM primitives are best kept together in one partition.

### 2014-12-09 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Changed title.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4525 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Edits to effectiveness of placer enhancements.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4526 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 % Do not put math or special symbols in the title.
-\title{CAD and Routing Architecture for Interposer-Based Multi-FPGA Systems}
+\title{Making Multiple Dice Work as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs}
 %
@@
 \thanks{V. Betz is with the Department of Electrical and Computer Engineering, University of Toronto, Toronto, Ontario, Canada (email: vaughn@eecg.utoronto.ca)}%
-\thanks{Manuscript received January 19, 2014.}}
+\thanks{Manuscript received December 19, 2014.}}
 
@@
 \label{sec:effectiveness_of_placer_enhancements}
-Enabling the enhancements described in Section~\ref{cad_enh_placer_subsection} improves the minimum routable channel width and circuit speed.
-As shown in Figure~\ref{fig:comparison_minW}, for architectures with low \textit{\% wires cut} (high interposer capacity), the placer optimizations are only marginally beneficial compared to the original VPR placer since the interposer wires are not scarce; but as more wires are cut and we reach the realistic interposer capacity of commercial FPGAs, the placer optimizations provide higher and higher reduction in minimum channel width.
+Enabling the enhancements described in Section~\ref{cad_enh_placer_subsection} improves both the minimum routable channel width and circuit speed.
+As shown in Figure~\ref{fig:comparison_minW}, for architectures with low \textit{\% wires cut} (high interposer capacity), the placer optimizations have only a marginal benefit on routability compared to the original VPR placer since the interposer wires are not scarce, but as more wires are cut and we reach the realistic interposer capacity of commercial FPGAs, the placer optimizations provide higher and higher reductions in minimum channel width.
 When 80\% of wires are cut, the placer optimizations result in 11.3\% reduction in minimum channel width.
@@
 
-The optimizations also help the placer to place related entities on the same die to reduce the number of inter-die signal crossings and hence increase circuit speed.
-At 80\% wires cut, the placer optimizations result in 8.4\% higher circuit speed.
+The optimizations also guide the placer to keep related entities on the same die to reduce the number of inter-die signal crossings and hence increase circuit speed.
+This speed improvement is fairly consistent across a wide range of \% wires cut, averaging 12.5\% higher circuit speed.
 \begin{figure}[!htbp]

### 2014-12-09 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Adding square/tall FPGA figures.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4528 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 We investigated the implications of the device aspect ratio on routability for the four-partition case.
-In particular, we compared a device with a square interposer (four rectangular dice) to a device with a rectangular interposer (four square dice.) We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section.
+In particular, we compared a device with a square interposer (four rectangular dice) to a device with a rectangular interposer (four square dice), as shown in \ref{fig:square_tall}.
+We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section.
 The results are shown in Table \ref{table:squares}.
+
+\begin{figure}[!htbp]
+\centering
+\subfigure[]{\includegraphics[height = 5cm]{square_fpga.pdf}}
+\subfigure[]{\includegraphics[height = 5cm]{tall_fpga.pdf}}
+\caption{Devices with (a) a square interposer and (b) square dice.}
+\label{fig:square_tall}
+\end{figure}
 

### 2014-12-09 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4529 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Re-adding aspect ratio figure
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4530 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Vaughn and Javeed's meeting updates.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4531 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 % Do not put math or special symbols in the title.
-\title{Making Multiple Dice Work as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs}
+\title{Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs}
 %
@@
 \subsection{Interposer Wiring Capacity (\% wires cut)}
+\label{sec:resultsWiringCapacity}
 
@@
 \label{num_dice_impact}
-To examine the impact of the number of dice used to construct an interposer-based FPGA, we measured the minimum channel width for an FPGA device with four dice, varying the fraction of wires cut.
-We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section (TODO JAVEED -- spell them out?).
-The results, shown in Figure \ref{fig:wires_cut_4part}, show the same trend as Figure \ref{fig:wires_cut}.
+To examine the impact of the number of dice used to construct an interposer-based FPGA, we compare the minimum channel width for FPGA systems composed of four dice and two dice, respectively, while varying the fraction of wires cut at the interposer boundaries.
+We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same interposer routing architecture parameters as Section \ref{sec:resultsWiringCapacity}.
+As Figure \ref{fig:wires_cut_4part} shows, four-die systems perform almost as well in routability as two-die systems until the \textit{\% wires cut} exceeds 80\%, at which point routability is worse for the four-die system.
+TODO JAVEED -- add extra line to graph, make sure same circuits for all points.
 
@@
 \includegraphics[width=\linewidth]{wires_cut_4part.eps}
-\caption{Impact of partitioning-based CAD flow on routability for a range of architectures with 4 dice. TODO JAVEED -- LU32 90\% result is missing}
+\caption{Impact of number of dice on routability. TODO JAVEED -- LU32 90\% result is missing}
 \label{fig:wires_cut_4part}
@@
 
-TODO JAVEED -- show critical path delay? (it's uninteresting)
+Systems with more dice are slower, on average the critical path delay is xx\% higher for a four-die system than a two-die system.
+This speed difference is not significantly affected by the \textit{\% wires cut}.
+ TODO JAVEED -- figure out critical path delay %.
 
@@
 \label{aspect_ratio_impact}
-TODO: Javeed
-We investigated the implications of the device aspect ratio on routability for the four-partition case.
-In particular, we compared a device with a square interposer (four rectangular dice) to a device with a rectangular interposer (four square dice), as shown in \ref{fig:square_tall}.
-We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section.
-The results are shown in Table \ref{table:squares}.
+It is easier to manufacture silicon dice that have an aspect ratio near 1 (i.e.
+square).
+Commercial FPGAs are laid out with identical columns, and hence combine multiple identical FPGA dice in \textit{one dimension} on an interposer \cite{xilinxTSV}.
+As Figure \ref{fig:square_tall} shows, we can make \textit{either} the FPGA dice square, or the interposer square, but not both.
+Table \ref{table:squares} compares the routability of these two options for the four-die case: a square interposer (four rectangular dice) or a rectangular interposer (four square dice.) We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same routing architecture parameters as Section \ref{sec:resultsWiringCapacity}.
 
@@
 LU32PEEng & 256 & 264 \\ \hline
-geomean & 140.7481367 & 251.805728 \\ \hline
+geomean & 141 & 252 \\ \hline
 \end{tabular}
@@
 The square interposer configuration performs significantly better than the square dice configuration.
-
-TODO JAVEED -- Vaughn commented that we should talk about which is easier to manufacture.
-I think the square dice device would actually be easier to manufacture, because the more complex dice will fit well on a round wafer.
-Isn't it fine for the interposer to be non-square?
+A square interposer results in cutlines that are closer together, and hence more logic blocks are near a cutline.
+On the other hand, a square interposer (composed of rectangular FPGA dice) has more vertical routing channels, giving it a higher absolute number of wires between FPGA dice.
+The second effect is the more important one, leading to much better routability for a square interposer.
 

### 2014-12-14 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Added metis and hmetis references
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4527 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Updated my figures to look more like Ehsan's.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4537 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \subsubsection{Partitioning Tool}
-Two well-known freely available graph partitioning tools are Metis [] and hMetis [].
+Two well-known freely available graph partitioning tools are Metis~\cite{karypis1998multilevelmetis} and hMetis~\cite{karypis1999multilevelhmetis}.
 Metis operates on graphs, while hMetis operates on hypergraphs -- a hypergraph is a generalization of a graph, where each edge can connect any number of vertices.
@@
 \centering
-\includegraphics[width=\linewidth]{graph_topology_cutsize.eps}
+%\includegraphics[width=\linewidth]{graph_topology_cutsize.eps}
+%\input{graph_topology_cutsize.tex}
+\input{graph_topology_cutsize}
 \caption{Hypergraph cut size achieved by Metis with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
@@
 \centering
-\includegraphics[width=\linewidth]{graph_topology_mcw.eps}
+\input{graph_topology_mcw}
+%\includegraphics[width=\linewidth]{graph_topology_mcw.eps}
 \caption{Minimum channel width achieved by VPR with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
@@
 \centering
-\includegraphics[width=\linewidth]{vpr_flows_mcw.eps}
+%\includegraphics[width=\linewidth]{vpr_flows_mcw.eps}
+\input{vpr_flows_mcw}
 \caption{Minimum channel width achieved by VPR with the different partitioning flows.}
@@
 \centering
-\includegraphics[width=\linewidth]{vpr_flows_crit_path.eps}
+\input{vpr_flows_crit_path}
 \caption{Critical path delay achieved by VPR with the different partitioning flows.}
@@
 \centering
-\includegraphics[width=\linewidth]{wires_cut.eps}
+\input{wires_cut}
 \caption{Impact of partitioning-based CAD flow on routability for a range of architectures.}
@@
 \centering
-\includegraphics[width=\linewidth]{packing_bloat.eps}
-\caption{Packing bloat.}
+\input{packing_bloat}
+\caption{The number of blocks in the clustered netlist with and without partitioning constraints. This is a measure of packing bloat.}
 \label{fig:packing_bloat}

### 2014-12-15 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Fixed a typo.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4533 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
updating all my graphs to be blakc&white friendly.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4535 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Updated my figures to look more like Ehsan's.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4537 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Updated square vs tall interposer figures to have the same area, compressed cad flow flowchart.
git-svn-id: https://vtr-verilog-t

@@
 \begin{abstract}
-Interposer-based multi-FPGA systems are composed of multiple FPGA dice connected through a silicon interposer.
-Such devices allow larger FPGA systems to be built than one monolithic die can accomodate and are now commercially available.
-An open question, however, is how efficient such systems are compared to a monolithic FPGA, as the number of signals passing between dice is reduced and the signal delay between dice is increased in an interposer system vs.
-a monolithic FPGA.
-
-We create a new version of VPR to investigate the architecture of such systems, and show that by modifying the placement cost function to minimize the number of signals that must cross between dice we can reduce routing demand by 18\% and delay by 2\%.
-We also show that the signal count between dice and the signal delay between dice are key architecture parameters for interposer-based FPGA systems.
-We find that if an interposer supplies (between dice) 60\% of the routing capacity that the normal (within-die) FPGA routing channels supply, there is little impact on the routability of circuits.
-Smaller routing capacities in the interposer do impact routability however: minimum channel width increases by 20\% and 50\% when an interposer supplies only 40\% and 30\% of the within-die routing, respectively.
-The interposer also impacts delay, increasing circuit delay by 34\% on average for a 1 ns interposer signal delay and a four-die system.
+Multiple FPGA dice can be connected through a silicon interposer to form a larger FPGA system than one monolithic die can accomodate.
+Such systems are now commercially available but many open questions remain concerning their architecture and efficiency, as the number of signals passing between dice is reduced and the signal delay between dice is increased compared to a monolithic FPGA.
+We create a new version of VPR to target such systems and investigate modifications to placement and routing and the incorporation of partitioning into the flow to best exploit interposer-based FPGAs.
+Our best CAD flow can reduce the routing demand for interposer FPGAs with realistic amounts of connectivity between dice by 47\% and improve the circuit speed by 13\% on average.
+Modifications to the FPGA architecture to add routing flexibility when crossing the interposer are very beneficial and can improve routability by a further 11\%.
+With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 40\% of the routing capacity that the normal (within-die) FPGA routing channels supply there is negligible impact on the routability of circuits.
+Smaller interposer routing capacities do impact routability however: minimum channel width increases by 10\% and 70\% when an interposer supplies only 20\% and 10\% of the within-die routing, respectively.
+The interposer also impacts delay, increasing circuit delay by 11\% on average for a 1 ns interposer signal delay and a two-die system.
 Reducing the interposer delay has a greater benefit in improving circuit speed than does reducing the number of dice in the system.
@@
 %   - Describe multi-chip
-Interposer-based multi-FPGA systems are composed of multiple FPGA dice, which are connected through a silicon interposer.
-The interposer is fabricated with an older technology than that used for the dies, and links between dice include both a micro-bump on each FPGA die and a metal wire on the interposer~\cite{xilinxTSVperformance}.
-This results in a reduced connectivity between dice and increased delay for connections between dice, as compared to the total routing connectivity and the delay one can achieve within a single die.
-In this paper we present an architecture study of silicon interposer-based FPGAs that analyzes how the reduced connectivity and increased delay impact their performance.
+Interposer-based FPGA systems are composed of multiple FPGA dice connected through a silicon interposer.
+The interposer is fabricated with an older technology than that used for the dice, and links between dice include both a micro-bump on each FPGA die and a metal wire on the interposer~\cite{xilinxTSVperformance}.
+This results in a reduced connectivity between dice and increased delay for connections between dice compared to the total routing connectivity and the delay one can achieve within a single die.
+In this work we present an architecture study of silicon interposer-based FPGAs that analyzes how the reduced connectivity and increased delay impact their performance.
 
 %   - Motivate: yield, bigger FPGA, no previous study
-The silicon interposer FPGA technology is interesting because it enables the creation of large FPGAs composed of small dies and also of very large FPGAs, with higher logic capacity than one can achieve with a single die.
-Such FPGA systems are sometimes called 2.5D FPGAs, since they make use of vertical stacking of dies on an interposer to enable higher integration levels.
+Silicon interposer FPGA technology is interesting because it enables the creation of large FPGAs composed of small dice and also of very large FPGAs, with higher logic capacity than one can achieve with a single die.
+Such interposer systems are sometimes called 2.5D devices, since they make use of vertical stacking of dice on an interposer to enable higher integration levels~\cite{3dhandbook}.
 
 % yield rate
-Being able to make large FPGAs with multiple smaller dies is particularly interesting at the beginning of a new manufacturing process, when defect densities are high.
+Making large FPGAs from multiple smaller dice is particularly interesting at the beginning of a new manufacturing process, when defect densities are high.
 In such a case, good-die yield drops dramatically as the die size increases, and this drastically impacts the availability of large FPGAs early in the process lifetime.
@@
 \cite{3dfpga1995} proposed the creation of 3D FPGAs by stacking 2D FPGAs and connecting them with solder bumps.
-Lin et al also studied 3D FPGAs, but using a different approach~\cite{3dfpga}, proposing multiple active layers, with different FPGA functions in each.
-While promising, both approaches have manufacturing challenges: \cite{3dfpga1995} requires a higher density of through-silicon vias than can currently be manufactured and the multiple active layers required by \cite{3dfpga} are not yet widely available.
-In contrast, silicon interposers linked to dies via microbumps are now manufacturable.
+Lin et al also studied 3D FPGAs, but using a different approach~\cite{3dfpga} that utilized multiple active layers with different FPGA functions (logic, routing, and configuration) in each.
+While promising, both these approaches have manufacturing challenges: \cite{3dfpga1995} requires a higher density of through-silicon vias than can currently be manufactured and the multiple active layers required by \cite{3dfpga} are not yet widely available.
+In contrast, silicon interposers linked to dice via microbumps are now manufacturable.
 
@@
 \begin{itemize}
-%     * Study on routability vs wires cut
-	\item A detailed study of how the multi-FPGA system routability is impacted by the amount of connectivity between dice provided by the interposer.
-%     * Timing vs wires cut + delay
-	\item An Analysis of the impact of the interposer on timing.
-%     * CAD flow optimizations for 2.5D FPGAs
-	\item Modification of the VPR CAD tool to model and target 2.5D silicon interposer-based FPGAs, as well as CAD changes to optimize for 2.5D FPGAs.
-%     * Analysis of existing Virtex-7 FPGAs
-	\item Analysis of the commercially available Xilinx Virtex-7 silicon interposer FPGAs~\cite{xilinxWP}\cite{xilinx7series}.
-We measure the connectivity reduction and delay increase between dies caused by the interposer on such FPGAs in order to see where these devices lie within the architcture space we explore.
+	\item Modifications of the VPR CAD tool to model and target 2.5D silicon interposer-based FPGAs, as well as CAD changes to optimize for 2.5D FPGAs.
+	\item An investigation of the efficacy of incorporating partitioning into various stages of the CAD flow.
+	\item A detailed study of how the multi-FPGA system routability is impacted by the amount of between-die connectivity provided by the interposer, and by the switching flexibility at the FPGA die - interposer boundary.
+	\item An analysis of the interposer's impact on timing.
 \end{itemize}
@@
 Section \ref{archSection} describes the changes made to the FPGA architecture description and to VPR to target 2.5D FPGAs.
-Section \ref{cadSection} describes optimizations made to the VPR placement algorithm to improve quality on such FPGAs and Section \ref{resultsSection} presents architectural results for 2.5D FPGAs.
+Section \ref{cadSection} describes optimizations made to the VPR placement and routing algorithms to improve quality on such FPGAs and Section~\ref{sec:CADflow} describes how partitioning can be used to further enhance the CAD flow.
+Section \ref{resultsSection} presents architectural results for 2.5D FPGAs.
+
+An earlier and less detailed version of this work appeared in \cite{interposer2014}.
+We extend this earier work by enhancing the CAD tools, incorporating partitioning into the flows and investigating a broader range of architecture questions.
 
@@
 
-% 1 - FPGAs "More than Moore" - want(?) to combine chips more tightly / with more connectivity than on a board
-As mentioned in the previous section, interposer-based FPGAs allow the creation of chips larger than a single die, making a "More than Moore" improvement on the size and number of logic elements possible, and with chips combined far more tightly and with more connectivity than if they were on separate boards connected through conventional means.
+Interposer-based FPGAs allow the creation of chips larger than a single die, making a "More than Moore" improvement on the size and number of logic cells possible, and with chips combined far more tightly and with more connectivity than if they were on separate boards connected through conventional means.
 
-% 2 - a) - Can sell bigger chips
-%        - Can we make bigger? -> 28nm 1.95m LE vs 979k (V7) / 952k (SV)
-The improvement in the number of logic elements of 2.5D FPGAs over conventional ones is very significant.
-Xilinx's largest interposer-based FPGA, the Virtex-7 XC7V2000T, has 4 dies (which Xilinx calls Super Logic Regions) and 1.954 million logic elements~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die (the XC7VX980T), has 979k logic elements and Altera's largest FPGA, the Stratix V 5SGXBB, has 952k logic elements~\cite{stratixV}.
+The improvement in the capacity of 2.5D FPGAs over conventional ones is very significant.
+Xilinx's largest 28 nm interposer-based FPGA, the Virtex-7 XC7V2000T, contains 4 FPGA dice (which Xilinx calls Super Logic Regions) and 1.954 million logic cells~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die (the XC7VX980T), has 979k logic cells and Altera's largest 28 nm FPGA, the Stratix V 5SGXBB, has 952k logic elements~\cite{stratixV}.
 Even though all these FPGAs use a 28 nm process, silicon interposer technology allows the creation of FPGAs with twice the resources possible on even an extremely large single die.
+Xilinx's recently announced Virtex UltraScale line of 20 nm FPGAs makes even more aggressive use of silicon interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.
 
-%     b) - Yield results
-%               -> Ref paper / use lecture slides
-%               -> Defect density of 1/cm^2 is typical at early days of new process
-%               -> FPGA early adopters of new process
 Another major advantage of interposer-based FPGAs comes at the beginning of a new manufacturing process, when the defect density is high~\cite{xilinxTSVperformance}.
@@
 
-To illustrate this impact, consider a new process in which the defect density is 1/$cm^2$, which is a reasonable value early in the process lifecycle, and the die area is $6cm^2$, which roughly matches the size of the largest member of a high-end FPGA family such as Virtex 7.
+To illustrate this impact consider a new process in which the defect density is 1/$cm^2$, which is a reasonable value early in the process lifecycle, and the die area is $6cm^2$, which roughly matches the size of the largest member of a high-end FPGA family such as Virtex 7.
 Using the Poisson Yield Model~\cite{yieldmodel}, the yield is only $0.25\%$ of die.
-If instead the chip is composed of four $1.5cm^2$ dies, the yield is $22\%$.
-This means that a 12 inch silicon wafer with $730cm^2$ of area would produce on average $0.3$ working $6cm^2$ dies, while the same wafer would produce on average $107$ working $1.5cm^2$ dies.
-Therefore, as a $6cm^2$ chip would be composed of four $1.5cm^2$ dies, the wafer would yield $26.75$ systems on average, as the "assembly yield" of placing these four die on an interposer is very high~\cite{xilinxTSV}.
+If instead the chip is composed of four $1.5cm^2$ dice, the yield is $22\%$.
+This means that a 12 inch silicon wafer with $730cm^2$ of area would produce on average $0.3$ working $6cm^2$ dice, while the same wafer would produce on average $107$ working $1.5cm^2$ dice.
+Therefore, as a $6cm^2$ chip would be composed of four $1.5cm^2$ dice, the wafer would yield $26.75$ systems on average, as the "assembly yield" of placing these four die on an interposer is very high~\cite{xilinxTSV}.
 Hence the number of interposer-based FPGAs created from the same silicon wafer would be almost $100\times$ greater than a monolithic FPGA of the same size.
@@
 
-The Xilinx 2.5D FPGAs from the Virtex-7 family are currently the only commercially available silicon interposer-based FPGAs.
+The Xilinx 2.5D FPGAs from the Virtex-7 and Virtex Ultrascale families are currently the only commercially available silicon interposer-based FPGAs~\cite{xilinxWP}.
 As described in Section~\ref{archSection} we are studying the impact of several key interposer parameters on the performance of the multi-FPGA system, including (i) the percentage of the wiring normally present between rows of the FPGA which are cut when crossing between dice, and (ii) the extra delay (vs.
@@
 
-The 28nm dies are connected to the 65nm silicon interposer through microbumps with a 45$\mu$m pitch.
+The 28nm dice are connected to the 65nm silicon interposer through microbumps with a 45$\mu$m pitch.
 Hence the area occupied by microbumps at one edge of one die is 13440 $\times$ (45$\mu$m)\textsuperscript{2} $=$ 27mm\textsuperscript{2}.
@@
 
-Overall, these interposer-based FPGAs have increased delay and reduced connectivity between dies, with approximately $23\%$ of the usual number of vertical wires crossing between dies and approximately $1ns$ of increased delay to cross the interposer.
+Overall, these interposer-based FPGAs have increased delay and reduced connectivity between dice, with approximately $23\%$ of the usual number of vertical wires crossing between dice and approximately $1ns$ of increased delay to cross the interposer.
 
@@
 To properly model a silicon interposer FPGA, we use version 7.0 of the popular FPGA exploration toolset, Verilog-to-Routing (VTR)~\cite{luu2014vtr}.
-The logic synthesis portions of the flow (ODIN II and  ABC) are left untouched, while we modify the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) to model and optimize for interposer-based FPGAs.
+The logic synthesis portions of the flow (ODIN II and  ABC) are used as-is, while we modify the placement and routing portion of the flow (VPR~\cite{betz1997vpr}) to model and optimize for interposer-based FPGAs.
 The modifications are made in such a way that they require no changes to any of the input files, so one can experiment with interposer-based FPGAs with any existing benchmark circuits and any existing VPR-format FPGA architecture description simply by specifying appropriate command-line parameters.
@@
 \centering
-\includegraphics[width=\linewidth]{fpga.eps}
+\includegraphics[width=0.65\linewidth]{fpga.eps}
 \caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
@@
 \label{sec:interposer-routing-interface}
-In a simple architecture, the vertical routing wires that can not connect to an interposer wire are dangling.
+In a simple architecture, the vertical routing wires at the edge of an FPGA die that can not connect to an interposer wire are dangling.
 Since interposer wires are scarce, we want to be able to study architectures in which the interposer wires are more accessible.
@@
 
-As illustrated in Figure~\ref{fig:vpr_interposer_gui}, no vertical routing wires cross a cutline boundary and a fraction of them connect to other dies via interposer wires.
+As illustrated in Figure~\ref{fig:vpr_interposer_gui}, no vertical routing wires cross a cutline boundary and a fraction of them connect to other dice via interposer wires.
 
@@
 \centering
-\includegraphics[width=\linewidth]{bb_new.eps}
+\includegraphics[width=0.75\linewidth]{bb_new.eps}
 \caption{Illustration of three different scenarios for the wiring cost.
@@
 \centering
-\includegraphics[width=\linewidth]{arch_experiments3.eps}
+\includegraphics[width=\linewidth]{arch_experiments4.eps}
 \caption{Impact of routing-interposer interface flexibility on minimum channel width and circuit speed.}
@@
 \centering
-\includegraphics[width=\linewidth]{wires_cut_4part.eps}
+\input{wires_cut_4part}
 \caption{Impact of number of dice on routability. TODO JAVEED -- LU32 90\% result is missing}
@@
 
-Systems with more dice are slower, on average the critical path delay is xx\% higher for a four-die system than a two-die system.
+Systems with more dice are slower; on average the critical path delay is xx\% higher for a four-die system than a two-die system.
 This speed difference is not significantly affected by the \textit{\% wires cut}.
@@
 \subfigure[]{\includegraphics[height = 5cm]{square_fpga.pdf}}
-\subfigure[]{\includegraphics[height = 5cm]{tall_fpga.pdf}}
+\subfigure[]{\includegraphics[height = 9.335cm]{tall_fpga.pdf}}
 \caption{Devices with (a) a square interposer and (b) square dice.}
@@
 %---------------------------------------------------------------
-\section{Conclusions and Future Work}
+\section{Conclusion}
 \label{conclusionSection}
@@
 While such systems are now a commercial reality, we do not know of any prior study of their key architectural parameters.
-We found that by modifying VPR's placement cost function we could improve the routability (reduce minW) by 18\%, while simultaneously improving speed by 2\%.
-Interestingly, we found that a smooth cost function that less precisely models the interposing wiring demand outperformed direct monitoring of the number of required interposer connections during annealing.
+We found that by modifying VPR's placement cost function we can improve routability (reduce minW), while simultaneously improving speed.
+Modifications to the VPR router lookahead function are crucial to efficiently routing interposer-based FPGAs, yielding a 31$\times$ speed-up.
+Incorporating partitioning into the CAD flow is very beneficial.
+Our best partitioning flow converts nets (hyperedges) to a star graph model, uses Metis to partition the resulting circuit graph, and forces VPR's packer to respect the Metis partitioning to FPGA dice.
+Taken together, these CAD changes reduce the signal wires that must cross the interposer by an average of 47\% while simultaneously improving circuit speed by 13\% for realistic interposer architectures.
 
-We defined three key architecture parameters for interposer-based FPGAs, and used this extended VPR to analyze their impact on minimum channel width and critical path delay.
-We found that the \textit{minimum channel width} increases steeply after more than 60\% of the within-die wires are cut at the interposer, and when 80\% of wires are cut the \textit{minimum channel width} is more than double that required by a monolithic FPGA.
-The Xilinx \textit{XC7V2000T} FPGA has 77\% of the normal wiring cut at the interposer, indicating that commercial interposer-based FPGAs will require high-quality CAD optimization to maintain good routability.
+We defined four key architecture parameters for interposer-based FPGAs, and used this extended VPR to analyze their impact on minimum channel width and critical path delay.
+We find that adding a moderate amount of switching flexibility at the FPGA - interposer boundary has negligible area cost but reduces the required FPGA minimum channel width by 13\%.
+We also find that the multi-die system has good routability so long as the interposer provides sufficient routing bandwidth, and this routing bandwidth can be much less than one would find in a monolithic FPGA.
+Specifically, if the interposer provides 40\% or more of the channel width required by the monolithic FPGA (an average of 45 wires per channel across our benchmark suite) the routability of the interposer-based FPGA is equivalent to that of a monolithic one.
+As the interposer signal bandwidth becomes more constrained, there is some impact to routability, but even an interposer providing only 20\% of the monolithic FPGA routing channel bandwidth (22 wires in our experiments) only needs a 10\% increase in the within-FPGA-die routing channel width to compensate.
+
 The \textit{critical path delay} is not strongly influenced by the \textit{\% wires cut} but is strongly influenced by the interposer delay and the number of cuts.
-Our results show that the critical path usually crosses the interposer cutline more than once, as the \textit{critical path delay} is increased by more than the interposer \textit{delay increase}.
-Increasing the number of dice in an interposer-based FPGA does not significantly impact the \textit{minimum channel width}, but does lead to a larger \textit{critical path delay}.
+Our results show that the critical path often crosses the interposer cutline more than once, as the \textit{critical path delay} is increased by more than the interposer \textit{delay increase}; on average we find that a 1 ns interposer delay causes an 11\% delay increase in a two-die system vs.
+a monolithic FPGA .
+Increasing the number of dice in an interposer-based FPGA does not greatly impact the \textit{minimum channel width}, but does lead to a larger \textit{critical path delay}.
+We also found that the aspect ratio of the interposer-based FPGA is an important architectural parameter; for the best routability one should use a square interposer and rectangular FPGA dice.
 
-There are many interesting CAD and architecture questions for interposer-based FPGAs which we plan to investigate.
-Currently the wires that cross the interposer are accessed with the same switch structure as other wires in the FPGA.
-Since interposer wires are more scarce, possibly these wires should have larger multiplexers feeding them, which would make them easier to use; this may help routability of the system.
-Such changes will have to be carefully considered however, to make sure the FPGA can still be laid out as an array of regular tiles.
-Alternative CAD flows to improve the interposer routability are also possible.
-For example, instead of following the synthesize, place and route CAD flow of a conventional FPGA, we could add a partitioning step before placement which would divide the circuit into one partition per die.
-Such a flow may improve routability, as a partitioner's main focus is minimizing the number of cut signals.
-Enhanced partitioners will be required, as current partitioners such as hMetis~\cite{hMetis} cannot model the heterogeneous balance constraint (i.e.
-use no more than the available device logic, RAM, DSP and I/O blocks in any die) present in FPGAs.
-Additionally, by dividing the placement problem into two pieces -- partitioning and within-die placement -- we may increase the critical path delay as we can no longer globally optimize the placement of timing critical paths in one unified placement step.
-Nonetheless, this is a viable alternative approach and a comparison to the approach taken in this paper would be very interesting.
+Overall we find the results for interposer-based FPGAs to be very positive.
+While an interposer can only provide a fraction of the signal count between dice that one would have across cutline within a die, the system still has excellent routability and behaves similarly to a monolithic FPGA so long as the interposer signal count is not pushed to very low values; approximately 15\% of the within-die routing in our experiments.
+We have also found that modifications to the FPGA CAD flow and routing architecture to exploit interposer-based FPGAs have been very productive.
+As this is a new area we believe more routability and timing enhancements are likely with further research, and this would further enhance the utility of interposer-based FPGAs.
 

### 2014-12-15 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4540 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Overall we find the results for interposer-based FPGAs to be very positive.
-While an interposer can only provide a fraction of the signal count between dice that one would have across cutline within a die, the system still has excellent routability and behaves similarly to a monolithic FPGA so long as the interposer signal count is not pushed to very low values; approximately 15\% of the within-die routing in our experiments.
+While an interposer can only provide a fraction of the signal count between dice that one would have across cutline within a die, the system still has excellent routability and behaves similarly to a monolithic FPGA so long as the interposer signal count is not pushed to very low values; below 20\% of the within-die routing in our experiments.
 We have also found that modifications to the FPGA CAD flow and routing architecture to exploit interposer-based FPGAs have been very productive.

### 2014-12-16 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Updated square vs tall interposer figures to have the same area, compressed cad flow flowchart.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4538 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Adding reference to Metis manual, removing done TODOs
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4543 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Adding a hyperedge cutline figure.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4544 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \centering
-\includegraphics[width=\linewidth]{partitioning_cad_flow.eps}
+\includegraphics[height=3.5in]{partitioning_cad_flow.eps}
 \caption{Possible ways to add partitioning to the VPR CAD flow.}
@@
 Generally such a perfect balance overly restricts the solution, so some fractional ``unbalance'' is allowed; in this work we use $0.05$ ($5\%$).
-For two partitions, $P1$ and $P2$, this translates to $[ref Metis]$:
+For two partitions, $P1$ and $P2$, this translates to~\cite{karypismanual}
 
@@
 The parameters of graph topology and edge weight were evaluated by generating the total primitive netlist graph and running it through the Metis partitioner to generate two partitions.
-The results of the different schemes were ranked based on the hyperedge cutsize, which is the number of hyperedges crossing the cutline, as illustrated in Figure ??.
+The results of the different schemes were ranked based on the hyperedge cutsize, which is the number of hyperedges crossing the cutline, as illustrated in Figure \ref{fig:hyperedge_cutline}.
 The hyperedge cutsize is a proxy for the routability of the circuit, as the cutsize is the number of nets that cross the interposer boundary.
@@
 The results are shown in Figure \ref{fig:graph_topology_cutsize}.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{hyperedge_cutline.pdf}
+\caption{A graph with six vertices and three hyperedges partitioned into two halves. As there are two hyperedges crossing the cutline, the hyperedge cutsize is 2.}
+\label{fig:hyperedge_cutline}
+\end{figure}
 

### 2014-12-16 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Replaced oldfpaper with a citation
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4546 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \label{sec:CADflow}
-TODO: Javeed
-
 Given that the number of signals that can cross the interposer between FPGA dice is considerably more limited than within an FPGA, using a partitioner to divide the circuit into one partition per die is a promising CAD flow.

### 2014-12-16 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Updating reference for packing constraints
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4547 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 We refer to this type of per-block set of constraints as heterogeneous balance constraints.
-Note that there are complex legality constraints governing which LUTs and FFs can be packed into legal logic blocks, as well as which multipliers can be packed into legal DSP blocks [ref AAPack].
+Note that there are complex legality constraints governing which LUTs and FFs can be packed into legal logic blocks, as well as which multipliers can be packed into legal DSP blocks~\cite{luu2014vtr}.
 As the partitioner is not aware of these constraints, flow 3 above (which runs the partitioner after function block packing) will be able to more precisely balance resource use across partitions.

### 2014-12-16 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Adding critical path graph for 4 partitions.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4541 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Adding reference to Metis manual, removing done TODOs
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4543 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Updating reference for packing constraints
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4547 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Adding double line to geomean in table
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/s

@@
 \usepackage{cite}
+\usepackage{hhline}
 
@@
 As Figure \ref{fig:wires_cut_4part} shows, four-die systems perform almost as well in routability as two-die systems until the \textit{\% wires cut} exceeds 80\%, at which point routability is worse for the four-die system.
-TODO JAVEED -- add extra line to graph, make sure same circuits for all points.
 
@@
 This speed difference is not significantly affected by the \textit{\% wires cut}.
- TODO JAVEED -- figure out critical path delay %.
+\begin{figure}[!htbp]
+\centering
+\input{crit_path_4part}
+\caption{Impact of number of dice on critical path delay. TODO JAVEED -- LU32 90\% result is missing}
+\label{fig:crit_path_4part}
+\end{figure}
 
@@
 \centering
-\subfigure[]{\includegraphics[height = 5cm]{square_fpga.pdf}}
-\subfigure[]{\includegraphics[height = 9.335cm]{tall_fpga.pdf}}
+\subfigure[]{\includegraphics[height = 3cm]{square_fpga.pdf}}
+\subfigure[]{\includegraphics[height = 4.6675cm]{tall_fpga.pdf}}
 \caption{Devices with (a) a square interposer and (b) square dice.}
@@
 mcml & 196 & 496 \\ \hline
-LU32PEEng & 256 & 264 \\ \hline
+LU32PEEng & 256 & 264 \\ \hhline{|=|=|=|}
 geomean & 141 & 252 \\ \hline

### 2014-12-16 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Adding feng rent's rule reference
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4549 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \end{equation}
-where $P1_i$ denotes the set of all blocks of type $i$ in partition 1.
+where $P1_i$ is the set of all blocks of type $i$ in partition 1.
 
@@
 This shows that the VPR packer strongly benefits from partitioning constraints when targeting an interposer-based architecture, likely because it forbids the packing of primitives that are not naturally related into one function block.
-Interestingly, other recent work targeting conventional FPGAs has shown that using a partitioner to guide packing benefits routability [Wenyi Feng, et al, “Rent’s Rule Based Packing for Routability Optimization,” FPGA 2014 ].
+Interestingly, other recent work targeting conventional FPGAs has shown that using a partitioner to guide packing benefits routability\cite{feng2014rent}.
 On the other hand, The VPR placement algorithm with the enhancements described in Section 4 sees only a modest routability benefit from the partitioning constraints.

### 2014-12-23 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Did some wordsmithing
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4557 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks.
-In this flow the partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block.
-We updated the VPR clustering algorithm to respect this new legality constraint.
-
+The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block.
+We updated the VPR clustering algorithm to respect this new constraint.
 The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist.
@@
 The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines.
-These two partitioning flow options can be combined -- if partitioning is applied before packing, one can to pass the partition information to the placement stage as well.
-The post-packing placement constraints can be derived from the primitive block partitioning, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block.
+These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage.
+The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block.
 There are, therefore, four possible partitioning flows:
@@
 It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins.
-Based on that criterion alone, hMetis is the most appropriate choice of partitioner.
-However, there are other issues to consider for the circuit netlist partitioning use case.
+Based on that criterion alone, hMetis would be the most appropriate choice of partitioner.
+However, there are other important criteria to consider for our use case.
 
@@
 
-In our architecture, each FPGA die is identical, and therefore each die has the same number of resources as every other die.
-However, the number of available instances of each block type is not equal -- there are usually more LUTs than multipliers, for example.
-The problem with the simple balance constraint in the case of multiple block types is that the capacity of each partition is not taken into account on a per-block-type basis.
+The problem with (\ref{eq:balance_constraint}) in the case of multiple block types is that the capacity of each partition is not taken into account on a per-block-type basis.
 It is therefore possible to generate partitions that cannot be realized because the number of resources of each type is not balanced across partitions, even though the total number of resources is balanced.
-For example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which would not be physically realizable because partition 1 may not have enough LUTs to implement the desired partitioning (and the LUTs in partition 2 would be completely unused.) A technique is to enforce a balance constraint on each type of block, $i$:
+For example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which would not be physically realizable because partition 1 may not have enough LUTs to implement the desired partitioning (and no LUTs in partition 2 would be used.) One solution is to enforce a balance constraint on each block type, $i$:
 
@@
 
-The latest public version of hMetis does not support heterogeneous balance constraints, thourgh Metis does.
+The latest public version of hMetis does not support heterogeneous balance constraints, though Metis does.
 Metis was used in our CAD flow.
@@
 We considered two subgraph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.
+We assigned the same edge weight to every edge of the generated subgraph.
+Edge weights were computed based on the number of vertices in the hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.
 
@@
 
-We assigned the same edge weight to every edge of the generated subgraph.
-Edge weights were computed based on the number of vertices in the hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.
-Note that the vertices in the generated total netlist graph are the same as the vertices in the original circuit netlist hypergraph.
-
-Figure \ref{fig:hyperedge_to_graph} shows how a netlist of three blocks and two nets is transformed into its component subgraphs.
+Figure \ref{fig:hyperedge_to_graph} shows how a netlist of three blocks and two nets is transformed into its component subgraphs. TODO JAVEED: Maybe just show star 1/n from start to finish.
 
@@
 %\includegraphics[width=\linewidth]{graph_topology_mcw.eps}
-\caption{Minimum channel width achieved by VPR with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
+\caption{Minimum channel width achieved by VPR with different hyperedge models and weightings. n is the number of vertices on a hyperedge. TODO JAVEED: Which circuits?}
 \label{fig:graph_topology_mcw}

### 2014-12-24 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
More wordsmithing
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4558 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
   \item Leaving the packer unconstrained, and partitioning the function block (clustered) netlist to guide the placer.
-  \item Partitioning the primitive netlist to guide the packer, and constraining the placer based on constraints derived from partitioning, as described above.
+  \item Partitioning the primitive netlist to guide the packer, and, as described above, passing these constraints to the placer.
 \end{enumerate}
@@
 It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins.
-Based on that criterion alone, hMetis would be the most appropriate choice of partitioner.
-However, there are other important criteria to consider for our use case.
+Owing to the natural hypergraph representation, hMetis would be the most appropriate choice of partitioner.
+However, there are other issues to consider for our use case.
 
-Graph partitioning algorithms generate partitions of a circuit while satisfying ``balance constraints''.
+Graph partitioning algorithms assign vertices to partitions while satisfying ``balance constraints''.
 The balance constraints exist to drive the algorithm away from the trivial partitioning solution where all vertices are in the same partition, leaving all other partitions empty.
@@
 To derive an appropriate balance constraint for our use case, consider that FPGA circuit netlists may contain different block types such as LUTs, registers, RAM slices, and multipliers, among others.
-The basic balance constraint of (\ref{eq:balance_constraint}), applied to the two partition case with only LUTs and multipliers would ensure that the total block count (sum of the LUT count + multiplier count) in each partition is roughly balanced.
+The basic balance constraint of (\ref{eq:balance_constraint}), applied to a netlist with only LUTs and multipliers, would ensure that the total block count (sum of the LUT count + multiplier count) in each partition is roughly balanced.
+The problem with (\ref{eq:balance_constraint}) in the case of multiple block types is that the capacity, per block type, of each partition is not taken into account.
+It is therefore possible to generate partitions that cannot be realized because the number of resources of each type is not balanced across partitions, even though the total number of resources is balanced.
+For example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which may not be realizable given the per-block-type capacity of each partition.
 
-The problem with (\ref{eq:balance_constraint}) in the case of multiple block types is that the capacity of each partition is not taken into account on a per-block-type basis.
-It is therefore possible to generate partitions that cannot be realized because the number of resources of each type is not balanced across partitions, even though the total number of resources is balanced.
-For example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which would not be physically realizable because partition 1 may not have enough LUTs to implement the desired partitioning (and no LUTs in partition 2 would be used.) One solution is to enforce a balance constraint on each block type, $i$:
+One solution is to enforce a balance constraint on each block type, $i$:
 
@@
 where $P1_i$ is the set of all blocks of type $i$ in partition 1.
-
 We refer to this type of per-block set of constraints as heterogeneous balance constraints.
@@
 
-The latest public version of hMetis does not support heterogeneous balance constraints, though Metis does.
+The current version of hMetis does not support heterogeneous balance constraints, though Metis does.
 Metis was used in our CAD flow.
-However, as discussed earlier, the circuit netlist is most naturally represented as a hypergraph, and so it needs to be transformed to a graph before Metis can process it.
+However, as the circuit netlist is naturally represented as a hypergraph, it needs to be transformed to a graph before Metis can process it.
 
@@
 The two parameters in this operation are the graph topology and the edge weight scheme.
-We considered two subgraph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.
+We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.
 We assigned the same edge weight to every edge of the generated subgraph.
-Edge weights were computed based on the number of vertices in the hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.
+Edge weights were computed based on the number of vertices in the originating hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.
+We use \emph{hyperedge model} to refer to the combination of graph topology and edge weight scheme.
 
@@
 
-The parameters of graph topology and edge weight were evaluated by generating the total primitive netlist graph and running it through the Metis partitioner to generate two partitions.
-The results of the different schemes were ranked based on the hyperedge cutsize, which is the number of hyperedges crossing the cutline, as illustrated in Figure \ref{fig:hyperedge_cutline}.
-The hyperedge cutsize is a proxy for the routability of the circuit, as the cutsize is the number of nets that cross the interposer boundary.
-To compute the hyperedge cutsize, we applied the partitioning result (which assigns each vertex to a partition number) to the original untransformed hypergraph, and counted the number of hyperedges whose vertices were spread across different partitions.
-We computed the hyperedge cutsize for every combination of graph topology and graph edge weight scheme, across several circuits.
-The results are shown in Figure \ref{fig:graph_topology_cutsize}.
+The number of hyperedges crossing the cutline is referred to as the \emph{hyperedge cutsize}, as illustrated in Figure \ref{fig:hyperedge_cutline}.
+The hyperedge cutsize is a measure of the routability of the circuit, as it equals the number of nets that cross the interposer boundary.
+Using the Metis partitioner to generate graph partitions (which assigns each vertex to a partition number) and applying the partitioning result to the original untransformed hypergraph, we computed the hyperedge cutsize for every hyperedge model, across several circuits.
+We used the results, shown in Figure \ref{fig:graph_topology_cutsize}, to select the ``best'' hyperedge model.
 
@@
 \includegraphics[width=\linewidth]{hyperedge_cutline.pdf}
-\caption{A graph with six vertices and three hyperedges partitioned into two halves. As there are two hyperedges crossing the cutline, the hyperedge cutsize is 2.}
+\caption{A graph with six vertices, three hyperedges, and two partitions. As there are two hyperedges crossing the cutline, the hyperedge cutsize is 2.}
 \label{fig:hyperedge_cutline}
@@
 \input{graph_topology_cutsize}
-\caption{Hypergraph cut size achieved by Metis with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
+\caption{Hypergraph cut size achieved by Metis with different hyperedge models and weightings. $n$ is the number of vertices on a hyperedge.}
 \label{fig:graph_topology_cutsize}
@@
 The star $1$ (constant) edge weight scheme assigns a total weight of $n$ to each net, which heavily weights high-fanout nets.
-In contrast, the star $1/n$ scheme assigns the same total weight of $1$ to each net.
-The star $1/n^2$ scheme penalizes high-fanout nets relative to lower-fanout nets.
-Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ scheme achieves this.
+In contrast, the star $1/n$ model assigns the same total weight of $1$ to each net.
+The star $1/n^2$ model penalizes high-fanout nets relative to lower-fanout nets.
+Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ model achieves this.
 
-To validate the choice of hyperedge cutsize as a proxy for circuit routability, we ran the partitioning CAD flow for each graph topology and edge weight scheme, across several circuits.
+To validate the choice of hyperedge cutsize as a proxy for circuit routability, we ran the partitioning CAD flow for each hyperedge model, across several circuits.
 We use an unbalance of $5\%$, split the clustered netlist into two partitions (constraining only the placer), cut $80\%$ of the wires crossing the interposer, and impose a $1ns$ delay penalty for wires crossing the interposer.
@@
 Relative to the minimum device size required for placement (without partitioning constraints), we add $10\%$ to the complex block grid width and $10\%$ to the grid height.
-Figure \ref{fig:graph_topology_mcw} shows the geometric mean of the minimum channel width required for a successful route, for each of the graph topology and edge weight combinations.
+Figure \ref{fig:graph_topology_mcw} shows the geometric mean of the minimum channel width required for a successful route, for each hyperedge model.
 
@@
 The star topology with 1/n edge weights achieves the best minimum channel width, confirming that hyperedge cutsize is a good proxy for routability.
-Consequently, we use this scheme in all future results in this paper.
-Unlike routability, we found that the post-routing critical path delay was not strongly impacted by the hypergraph conversion scheme used.
+Consequently, we use this model in all future results in this paper.
+Unlike routability, we found that the post-routing critical path delay was not strongly impacted by the hypergraph model used.
 
@@
 \input{vpr_flows_mcw}
-\caption{Minimum channel width achieved by VPR with the different partitioning flows.}
+\caption{Minimum channel width achieved by VPR for each partitioning flow.}
 \label{fig:flows_mcw}
@@
 \input{vpr_flows_crit_path}
-\caption{Critical path delay achieved by VPR with the different partitioning flows.}
+\caption{Critical path delay achieved by VPR for each partitioning flow.}
 \label{fig:flows_crit_path}
@@
 
-Considering both routability and delay, the best CAD flow constrained both packing and placement with the partitioning result and this is the flow we use in the remainder of this paper.
+Considering both routability and delay, the best CAD flow used the partitioning result to constrain both packing and placement, and we use that flow in the remainder of this paper.
 
@@
 To see how well the CAD flow works for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits.
-As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of [] for FPGAs where more than half the wires are cut at interposer boundaries.
-It is slightly worse than the traditionally FPGA CAD flow even for an architecture with 0\% of the wires cut, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.
+As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of~\cite{interposer2014} for FPGAs where more than half the wires are cut at interposer boundaries.
+It is slightly worse than the traditional FPGA CAD flow even for an architecture with 0\% of the wires cut, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.
 

### 2014-12-24 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Made some changes to the hyperedge cutsize results section
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4559 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-The number of hyperedges crossing the cutline is referred to as the \emph{hyperedge cutsize}, as illustrated in Figure \ref{fig:hyperedge_cutline}.
-The hyperedge cutsize is a measure of the routability of the circuit, as it equals the number of nets that cross the interposer boundary.
-Using the Metis partitioner to generate graph partitions (which assigns each vertex to a partition number) and applying the partitioning result to the original untransformed hypergraph, we computed the hyperedge cutsize for every hyperedge model, across several circuits.
-We used the results, shown in Figure \ref{fig:graph_topology_cutsize}, to select the ``best'' hyperedge model.
+To select the best hyperedge model, we used the Metis partitioner to generate graph partitions (which assign each vertex to a partition number) and applied the partitioning result to the original untransformed hypergraph.
+To rank the different models, we computed the \emph{hyperedge cutsize} for every hyperedge model, across several circuits.
+The hyperedge cutsize is the number of hyperedges crossing the cutline, as illustrated in Figure \ref{fig:hyperedge_cutline}.
+As it equals the number of nets that cross the interposer boundary, it is a measure of circuit routability.
+The results are shown in Figure \ref{fig:graph_topology_cutsize}.
 

### 2014-12-24 vpr/SRC/base/OptionTokens.c
Print timing analysis runtime slack_definition is now a command-line option Backend graphics change: converted create_button to variadic template function to allow additional arguments to be passed to button callbacks Fixed all warnings in Linux and Windows builds (disabling some Windows warnings), and treat warnings as errors in the future Define CLOCKS_PER_SEC as CLK_PER_SEC if CLOCKS_PER_SEC is not defined Fixed SDC info printout bug and file descriptor leak Switched from "Multi-threaded" to "Multi-threaded Debug" libraries for Visual Studio debug build to avoid exiting immediately on asser

@@ -18,6 +18,7 @@ struct s_TokenPair OptionBaseTokenList[] = {
 		{ "gen_postsynthesis_netlist", OT_GENERATE_POST_SYNTHESIS_NETLIST }, 
 		{ "timing_analysis", OT_TIMING_ANALYSIS }, 
 		{ "timing_analyze_only_with_net_delay", OT_TIMING_ANALYZE_ONLY_WITH_NET_DELAY },
+        { "slack_definition", OT_SLACK_DEFINITION },
 		{ "init_t", OT_INIT_T }, 
 		{ "alpha_t", OT_ALPHA_T }, 
 		{ "exit_t", OT_EXIT_T }, 

### 2014-12-24 vpr/SRC/base/OptionTokens.h

@@ -17,6 +17,7 @@ enum e_OptionBaseToken {
 	OT_CREATE_ECHO_FILE,
 	OT_TIMING_ANALYSIS,
 	OT_TIMING_ANALYZE_ONLY_WITH_NET_DELAY,
+    OT_SLACK_DEFINITION,
 	OT_GENERATE_POST_SYNTHESIS_NETLIST,
 	OT_INIT_T,
 	OT_ALPHA_T,

### 2014-12-24 vpr/SRC/draw/create_button.h

@@ -1,0 +1,167 @@
+/* Note: everything in this file used to be part of graphics.c, 
+   but was moved here because the template function create_button 
+   needs to be defined in its own header file. */
+
+#define BUTTON_TEXT_LEN	100
+#include <string.h>
+#include <functional>
+#include <vector>
+
+#ifdef X11
+
+/*********************************************
+* X-Windows Specific Preprocessor Directives *
+*********************************************/
+
+#include <X11/Xlib.h>
+#include <X11/Xutil.h>
+#include <X11/Xos.h>
+#include <X11/Xatom.h>
+#include <X11/Xft/Xft.h>
+
+/* Uncomment the line below if your X11 header files don't define XPointer */
+/* typedef char *XPointer;                                                 */
+
+#endif /* X11 Preprocessor Directives */
+
+/* Indicates if this button displays text, a polygon or is just a separator.
+*/
+typedef enum {
+    BUTTON_TEXT = 0,
+    BUTTON_POLY,
+    BUTTON_SEPARATOR
+} t_button_type;
+
+/* Structure used to define the buttons on the right hand side of the main window (menu region).
+* width, height: button size, in pixels.
+* xleft, ytop: coordinates, in pixels, of the top-left corner of the button relative to its
+*    containing (menu) window.
+* fcn: a callback function that is called when the button is pressed. This function takes one
+*    argument, a function pointer to the routine that can draw the graphics area (user routine).
+* win, hwnd:  X11 and Win32 data pointer to the window, respectively.
+* button_type: indicates if this button displays text, a polygon or is just a separator.
+* text: the text to display if this is a text button.
+* poly: the polygon (up to 3 points right now) to display if this is a polygon button
+* is_pressed: has the button been pressed, and is currently executing its callback?
+* is_enabled: can you press this button right now? Visually will look "pushed in" when
+*      not enabled, and won't respond to clicks.
+*/
+
+struct t_button {
+    int width;
+    int height;
+    int xleft;
+    int ytop;
+    std::function<void(void(*drawscreen)(void))> fcn;
+#ifdef X11
+    Window win;
+    XftDraw* draw;
+#else
+// TODO this is hacky - should actually include windows.h 
+// to get access to HWND, but doing so would cause a namespace collision
[…]

### 2014-12-24 vpr/SRC/draw/draw.h

@@ -5,7 +5,7 @@
 #include "graphics.h"
 
 void update_screen(int priority, char *msg, enum pic_type pic_on_screen_val,
-		boolean crit_path_button_enabled);
+		boolean crit_path_button_enabled, const t_timing_inf &timing_inf);
 
 void alloc_draw_structs(void);
 

### 2014-12-24 vpr/SRC/draw/graphics.c

@@ -2,15 +2,21 @@
 * Easygl Version 3.0                                                       *
 * Written by Vaughn Betz at the University of Toronto, Department of       *
 * Electrical and Computer Engineering, with additions by Paul Leventis     *
-* and William Chow of Altera, Guy Lemieux of the University of Brish       *
-* Columbia, Long Yu (Mike) Wang of the University of Toronto, and          *
-* Matthew J.P. Walker of the University of Toronto                         *
+* and William Chow of Altera, Guy Lemieux of the University of British     *
+* Columbia, and Long Yu (Mike) Wang, Matthew J.P. Walker and Michael       *
+* Wainberg of the University of Toronto.                                   *
 * All rights reserved by U of T, etc.                                      *
 *                                                                          *
 * You may freely use this graphics interface for non-commercial purposes   *
 * as long as you leave the author info above in it.                        *
 *                                                                          *
 * Revision History:                                                        *
+*                                                                          *
+* V3.0.1 December 2014 (Michael Wainberg)                                  *
+* - converted create_button to variadic template (moved to                 *
+*   create_button.h) to allow additional arguments to be passed to button  *
+*   callbacks                                                              *
+* - converted button array to std::vector, simplifying button bookkeeping  *
 *                                                                          *
 * V3.0 May 2014 - June 2014 (Matthew J.P. Walker)                          *
 * - continued integration with c++                                         *
@@ -197,7 +203,6 @@ close_graphics() to release all drawing structures and close the graphics.*/
 #include <math.h>
 #include <stdlib.h>
 #include <stdio.h>
-#include <string.h>
 #include <string>
 #include <iostream>
 #include <algorithm>
@@ -210,6 +215,7 @@ using namespace std;
 
[…]

### 2014-12-24 vpr/SRC/power/power_lowlevel.c

@@ -399,7 +399,9 @@ static float power_calc_leakage_st_pass_transistor(float size, float v_ds) {
 	}
 	power_low = i_ds * g_power_tech->Vdd;
 
-	if (!over_range) {
+    if (over_range) {
+        return power_low;
+    } else {
 		power_find_nmos_leakage(nmos_high, &lower, &upper, v_ds);
 		if (lower->v_ds == v_ds || !upper) {
 			i_ds = lower->i_ds;
@@ -409,11 +411,7 @@ static float power_calc_leakage_st_pass_transistor(float size, float v_ds) {
 			i_ds = (1 - perc_upper) * lower->i_ds + perc_upper * upper->i_ds;
 		}
 		power_high = i_ds * g_power_tech->Vdd;
-	}
 
-	if (over_range) {
-		return power_low;
-	} else {
 		float perc_upper = (size - nmos_low->nmos_size)
 				/ (nmos_high->nmos_size - nmos_low->nmos_size);
 		return power_high * perc_upper + power_low * (1 - perc_upper);
@@ -643,7 +641,9 @@ float power_calc_mux_v_out(int num_inputs, float transistor_size, float v_in,
 	}
 	v_out_low = in_prob_avg * v_out_max + (1 - in_prob_avg) * v_out_min;
 
-	if (!over_range) {
+    if (over_range) {
+        return v_out_low;
+    } else {
 		mux_volt_inf_high = &mux_nmos_inf_upper->mux_voltage_inf[num_inputs];
 		power_find_mux_volt_inf(&lower, &upper, mux_volt_inf_high, v_in);
 		if (lower->v_in == v_in || !upper) {
@@ -658,11 +658,7 @@ float power_calc_mux_v_out(int num_inputs, float transistor_size, float v_in,
 					+ perc_upper * upper->v_out_max;
 		}
 		v_out_high = in_prob_avg * v_out_max + (1 - in_prob_avg) * v_out_min;
-	}
 
-	if (over_range) {
-		return v_out_low;
-	} else {
 		float perc_upper =
 				(transistor_size - mux_nmos_inf_lower->nmos_size)
 						/ (mux_nmos_inf_upper->nmos_size

### 2014-12-24 vpr/SRC/util/token.c

@@ -183,15 +183,10 @@ void my_atof_2D(INOUTP float **matrix, INP int max_i, INP int max_j,
 bool check_my_atof_2D(INP int max_i, INP int max_j,
 		INP char *instring, OUTP int * num_entries){
 
-	int i, j, entry_count;
-	char* cur;
-	bool in_str;
-
 	/* Check if max_i * max_j matches number of entries in instring */
-	cur = instring;
-	i = j = 0;
-	in_str = FALSE;
-	entry_count = 0;
+	char *cur = instring;
+	bool in_str = FALSE;
+	int entry_count = 0;
 
 	/* First count number of entries in instring */
 	while (*cur != '\0'){

### 2014-12-29 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Updating table caption
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4552 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Fix typos
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4566 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 		% 4th author
-        and~Vaughn~Betz,~\IEEEmembership{Member,~IEEE,}%
-\thanks{E. Nasiri is with Altera Corporation, Toronto Technology Center, Toronto, Ontario, Canada (email: enasiri@altera.com)}% 
-\thanks{J. Shaikh is with Google, Mountain View, California, USA (email: jvds@google.com)}%
-\thanks{A. H. Pereira is with the Computer and Digital Systems Engineering Department - Escola Polit\'{e}cnica, University of S\~{a}o Paulo, S\~{a}o Paulo, SP, Brazil (email: andre.hahn@usp.br)}%
-\thanks{V. Betz is with the Department of Electrical and Computer Engineering, University of Toronto, Toronto, Ontario, Canada (email: vaughn@eecg.utoronto.ca)}%
-\thanks{Manuscript received December 19, 2014.}}
+        and~Vaughn~Betz,~\IEEEmembership{Member,~IEEE}%
+\thanks{E. Nasiri is with Altera, Toronto, ON, Canada (email: enasiri@altera.com)}% 
+\thanks{J. Shaikh is with Google, Mountain View, CA, USA (email: jvds@google.com)}%
+\thanks{A. H. Pereira is with the Computer and Digital Systems Engineering Dept., University of S\~{a}o Paulo, SP, Brazil (email: andre.hahn@usp.br)}%
+\thanks{V. Betz is with the Dept. of Electrical and Computer Engineering, University of Toronto, ON, Canada (email: vaughn@eecg.utoronto.ca)}%
+\thanks{Manuscript received January 6, 2015.}}
 
@@
 \begin{abstract}
-Multiple FPGA dice can be connected through a silicon interposer to form a larger FPGA system than one monolithic die can accomodate.
-Such systems are now commercially available but many open questions remain concerning their architecture and efficiency, as the number of signals passing between dice is reduced and the signal delay between dice is increased compared to a monolithic FPGA.
-We create a new version of VPR to target such systems and investigate modifications to placement and routing and the incorporation of partitioning into the flow to best exploit interposer-based FPGAs.
-Our best CAD flow can reduce the routing demand for interposer FPGAs with realistic amounts of connectivity between dice by 47\% and improve the circuit speed by 13\% on average.
-Modifications to the FPGA architecture to add routing flexibility when crossing the interposer are very beneficial and can improve routability by a further 11\%.
-With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 40\% of the routing capacity that the normal (within-die) FPGA routing channels supply there is negligible impact on the routability of circuits.
+Multiple FPGA dice can be connected through a silicon interposer to form a larger system than one monolithic die can accommodate.
+Such systems are now commercially available but many open questions remain concerning their key architecture parameters and efficiency, as the signal count between dice is reduced and the signal delay between dice is increased compared to a monolithic FPGA.
+We create a new version of VPR to target interposer-based FPGAs and investigate modifications to placement and routing and the incorporation of partitioning into the flow to improve results.
+Our best CAD flow reduces the routing demand for interposer FPGAs with realistic amounts of connectivity between dice by 47\% and improves the circuit speed by 13\% on average.
+Architecture modifications to add routing flexibility when crossing the interposer are very beneficial and improve routability by a further 11\%.
+With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 40\% of the routing capacity that the normal (within-die) FPGA routing channels supply there is negligible impact on circuit routability.
 Smaller interposer routing capacities do impact routability however: minimum channel width increases by 10\% and 70\% when an interposer supplies only 20\% and 10\% of the within-die routing, respectively.
 The interposer also impacts delay, increasing circuit delay by 11\% on average for a 1 ns interposer signal delay and a two-die system.
-Reducing the interposer delay has a greater benefit in improving circuit speed than does reducing the number of dice in the system.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Algorithms, Design, FPGA, Silicon interposer, 2.5D ICs. 
+Algorithms, FPGA, Silicon interposer, 2.5D ICs. 
 \end{IEEEkeywords}
@@
 Interposer-based FPGA systems are composed of multiple FPGA dice connected through a silicon interposer.
-The interposer is fabricated with an older technology than that used for the dice, and links between dice include both a micro-bump on each FPGA die and a metal wire on the interposer~\cite{xilinxTSVperformance}.
-This results in a reduced connectivity between dice and increased delay for connections between dice compared to the total routing connectivity and the delay one can achieve within a single die.
+The interposer uses an older fabrication technology than the FPGA dice, and links between dice include both a micro-bump on each FPGA die and a metal wire on the interposer~\cite{xilinxTSVperformance}.
+This results in reduced connectivity and increased delay between dice compared to the routing bandwidth and delay across a cutline within a single die.
+Such interposer systems are sometimes called 2.5D devices, since they make use of vertical stacking of dice on an interposer to enable higher integration levels~\cite{3dhandbook}.
 In this work we present an architecture study of silicon interposer-based FPGAs that analyzes how the reduced connectivity and increased delay impact their performance.
 
-%   - Motivate: yield, bigger FPGA, no previous study
-Silicon interposer FPGA technology is interesting because it enables the creation of large FPGAs composed of small dice and also of very large FPGAs, with higher logic capacity than one can achieve with a single die.
-Such interposer systems are sometimes called 2.5D devices, since they make use of vertical stacking of dice on an interposer to enable higher integration levels~\cite{3dhandbook}.
-
-% yield rate
+Silicon interposers enable the creation of large FPGAs from smaller dice, which can both improve yield and fabricate FPGAs that are too large to be manufacturable on a single die.
 Making large FPGAs from multiple smaller dice is particularly interesting at the beginning of a new manufacturing process, when defect densities are high.
-In such a case, good-die yield drops dramatically as the die size increases, and this drastically impacts the availability of large FPGAs early in the process lifetime.
+In this case, good-die yield drops dramatically as the die size increases, drastically impacting the availability of large FPGAs early in the process lifetime.
 
@@
 \begin{itemize}
-	\item Modifications of the VPR CAD tool to model and target 2.5D silicon interposer-based FPGAs, as well as CAD changes to optimize for 2.5D FPGAs.
+	\item VPR CAD tool modifications to model and optimize for 2.5D silicon interposer-based FPGAs.
 	\item An investigation of the efficacy of incorporating partitioning into various stages of the CAD flow.
-	\item A detailed study of how the multi-FPGA system routability is impacted by the amount of between-die connectivity provided by the interposer, and by the switching flexibility at the FPGA die - interposer boundary.
+	\item A study of how multi-FPGA system routability is impacted by the amount of between-die connectivity and by the switching flexibility at the FPGA die - interposer boundary.
 	\item An analysis of the interposer's impact on timing.
@@
 Section \ref{siliconSection} gives more information about silicon interposer technology.
-Section \ref{archSection} describes the changes made to the FPGA architecture description and to VPR to target 2.5D FPGAs.
-Section \ref{cadSection} describes optimizations made to the VPR placement and routing algorithms to improve quality on such FPGAs and Section~\ref{sec:CADflow} describes how partitioning can be used to further enhance the CAD flow.
+Section \ref{archSection} describes the changes made to the FPGA architecture description and VPR to target 2.5D FPGAs.
+Section \ref{cadSection} describes optimizations to the VPR placement and routing algorithms for interposer FPGAs and Section~\ref{sec:CADflow} describes how partitioning can further enhance the CAD flow.
 Section \ref{resultsSection} presents architectural results for 2.5D FPGAs.
@@
 An earlier and less detailed version of this work appeared in \cite{interposer2014}.
-We extend this earier work by enhancing the CAD tools, incorporating partitioning into the flows and investigating a broader range of architecture questions.
+We greatly extend this earlier work by enhancing the CAD tools, incorporating partitioning into the flows and investigating a broader range of architecture questions.
 
@@
 
-Interposer-based FPGAs allow the creation of chips larger than a single die, making a "More than Moore" improvement on the size and number of logic cells possible, and with chips combined far more tightly and with more connectivity than if they were on separate boards connected through conventional means.
+Interposer-based FPGAs allow systems larger than a single die, making a "More than Moore" improvement on the size and number of logic cells possible, and with chips combined with far more connectivity and less delay than if they were in separate packages connected through a printed circuit board.
 
 The improvement in the capacity of 2.5D FPGAs over conventional ones is very significant.
-Xilinx's largest 28 nm interposer-based FPGA, the Virtex-7 XC7V2000T, contains 4 FPGA dice (which Xilinx calls Super Logic Regions) and 1.954 million logic cells~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die (the XC7VX980T), has 979k logic cells and Altera's largest 28 nm FPGA, the Stratix V 5SGXBB, has 952k logic elements~\cite{stratixV}.
+Xilinx's largest 28 nm interposer-based FPGA contains 4 FPGA dice and 1.954 million logic cells~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die has 979k logic cells and Altera's largest 28 nm FPGA has 952k logic elements~\cite{stratixV}.
 Even though all these FPGAs use a 28 nm process, silicon interposer technology allows the creation of FPGAs with twice the resources possible on even an extremely large single die.
-Xilinx's recently announced Virtex UltraScale line of 20 nm FPGAs makes even more aggressive use of silicon interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.
+Xilinx's recently announced Virtex UltraScale line of 20 nm FPGAs makes even more aggressive use of interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.
 
 Another major advantage of interposer-based FPGAs comes at the beginning of a new manufacturing process, when the defect density is high~\cite{xilinxTSVperformance}.
-Bigger dice suffer a much reduced yield compared to smaller dice, and this greatly affects the supply and cost of top-of-the-line FPGAs to early adopters.
-
-To illustrate this impact consider a new process in which the defect density is 1/$cm^2$, which is a reasonable value early in the process lifecycle, and the die area is $6cm^2$, which roughly matches the size of the largest member of a high-end FPGA family such as Virtex 7.
-Using the Poisson Yield Model~\cite{yieldmodel}, the yield is only $0.25\%$ of die.
+To illustrate this impact consider a new process in which the defect density is 1/$cm^2$ (a reasonable value early in the process lifecycle) and the die area is $6cm^2$, which roughly matches the size of the largest member of a high-end FPGA family such as Virtex 7.
+Using the Poisson Yield Model~\cite{yieldmodel}, the yield is only $0.25\%$ of dice.
 If instead the chip is composed of four $1.5cm^2$ dice, the yield is $22\%$.
@@
 Therefore, as a $6cm^2$ chip would be composed of four $1.5cm^2$ dice, the wafer would yield $26.75$ systems on average, as the "assembly yield" of placing these four die on an interposer is very high~\cite{xilinxTSV}.
-Hence the number of interposer-based FPGAs created from the same silicon wafer would be almost $100\times$ greater than a monolithic FPGA of the same size.
+Hence the number of interposer-based FPGAs created from the same silicon wafer would be almost $100\times$ greater than a monolithic FPGA of the same size, greatly impacting not only cost but also the availability of such a large FPGA.
 
-When the process matures and the defect density decreases this advantage drops significantly.
+When the process matures and the defect density decreases this yield advantage drops.
 Consider a mature process with defect density of $0.1/cm^2$.
@@
 Hence the number of single die FPGAs created from a $730cm^2$ silicon wafer would be $66.9$ and the number of interposer-based FPGAs created would be $104.6$.
-While the interposer-based FPGA still has a yield advantage it might not lead to a major cost advantage, particularly when the cost of the interposer and assembling the die to it are included.
+While the interposer-based FPGA still has a yield advantage it might not lead to a major cost advantage when the cost of the interposer and assembling the die to it are included.
 For the large, state-of-the-art FPGAs that are built early in a process cycle and heavily used for prototyping, however, there is clearly a compelling cost advantage to an interposer-based solution.
@@
 The Xilinx 2.5D FPGAs from the Virtex-7 and Virtex Ultrascale families are currently the only commercially available silicon interposer-based FPGAs~\cite{xilinxWP}.
-As described in Section~\ref{archSection} we are studying the impact of several key interposer parameters on the performance of the multi-FPGA system, including (i) the percentage of the wiring normally present between rows of the FPGA which are cut when crossing between dice, and (ii) the extra delay (vs.
+As described in Section~\ref{archSection} we are studying the impact of several key interposer parameters on the performance of the multi-die system, including (i) the percentage of the wiring normally present between rows of the FPGA which is cut when crossing between dice, and (ii) the extra delay (vs.
 a normal connection between adjacent rows) added when one must traverse the interposer.
@@
 There are also 40 clock wires crossing the interposer.
-The average number of wires per vertical channel of this FPGA is 210 and there are approximately 280 vertical channels on the FPGA, resulting in approximately 58800 vertical wires crossing a horizontal cutline within a die.
+The average vertical routing channel in this FPGA contains 210 wires and there are approximately 280 vertical channels, resulting in approximately 58800 vertical wires crossing a horizontal cutline within a die.
 Hence the number of wires which cross the interposer is about 23\% of the total number of within-die vertical wires.
 
-The 28nm dice are connected to the 65nm silicon interposer through microbumps with a 45$\mu$m pitch.
+The 28 nm dice are connected to the 65 nm silicon interposer through microbumps with a 45$\mu$m pitch.
 Hence the area occupied by microbumps at one edge of one die is 13440 $\times$ (45$\mu$m)\textsuperscript{2} $=$ 27mm\textsuperscript{2}.
@@
 in \cite{xilinxTSV}, the bumps have to be spread out near the edge and need to go as far as $2.25mm$ away from the edge of the die.
-This greater distance from the border increaes the length of the inter-die connections, and along with the presence of the micro-bumps and their capacitance, leads to an increased delay for these crossing wires vs.
+This greater distance from the border increase the length of the inter-die connections, and along with the presence of the micro-bumps and their capacitance, leads to an increased delay for these crossing wires vs.
 that of a typical on-die routing wire.
@@
 
-Overall, these interposer-based FPGAs have increased delay and reduced connectivity between dice, with approximately $23\%$ of the usual number of vertical wires crossing between dice and approximately $1ns$ of increased delay to cross the interposer.
+Overall, these interposer FPGAs have increased delay and reduced connectivity between dice, with approximately $23\%$ of the usual number of vertical wires between dice and approximately $1ns$ of increased delay to cross the interposer.
 
@@
 
-Previous work by Pereira et al \cite{hahn2014cad} modeled the interposer indirectly: they did not introduce new rrnodes in the rrgraph to represent the interposer wires.
+Previous work by Pereira et al \cite{interposer2014} modeled the interposer indirectly: they did not introduce new rrnodes in the rrgraph to represent the interposer wires.
 Rather, they found the rrnodes that cross a cutline.
@@
 
-In~\cite{hahn2014cad}, Pereira et al tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
+In~\cite{interposer2014}, Pereira et al tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
 They found that the following \textit{cut\_cost} results in the best quality of results in terms of minimum routable channel width and critical path delay.
@@
 
-This cost function performs better than several others as shown in \cite{hahn2014cad}, and we believe the key to the good performance of this cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
+This cost function performs better than several others as shown in \cite{interposer2014}, and we believe the key to the good performance of this cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
 Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
@@
 \centering
+\caption{Square interposer and square dice minimum channel widths}
 \begin{tabular}{|l|r|r|}
@@
 \end{tabular}
-\caption{Minimum channel width for square interposer and square dice configurations.}
 \label{table:squares}
@@
 \label{conclusionSection}
-We have extended VPR to model and optimize for interposer-based multi-FPGA systems.
-While such systems are now a commercial reality, we do not know of any prior study of their key architectural parameters.
-We found that by modifying VPR's placement cost function we can improve routability (reduce minW), while simultaneously improving speed.
-Modifications to the VPR router lookahead function are crucial to efficiently routing interposer-based FPGAs, yielding a 31$\times$ speed-up.
+We have extended VPR to target interposer-based multi-FPGA systems.
+We found that by modifying VPR's placement cost function we can improve routability, while simultaneously improving speed.
+Modifications to the VPR router lookahead function are crucial to keep compile time reasonable and yield a 31$\times$ speed-up.
 Incorporating partitioning into the CAD flow is very beneficial.
@@
 We defined four key architecture parameters for interposer-based FPGAs, and used this extended VPR to analyze their impact on minimum channel width and critical path delay.
-We find that adding a moderate amount of switching flexibility at the FPGA - interposer boundary has negligible area cost but reduces the required FPGA minimum channel width by 13\%.
+We find that adding a moderate amount of switching flexibility at the FPGA - interposer boundary has negligible area cost but reduces the required minimum channel width by 13\%.
 We also find that the multi-die system has good routability so long as the interposer provides sufficient routing bandwidth, and this routing bandwidth can be much less than one would find in a monolithic FPGA.
 Specifically, if the interposer provides 40\% or more of the channel width required by the monolithic FPGA (an average of 45 wires per channel across our benchmark suite) the routability of the interposer-based FPGA is equivalent to that of a monolithic one.
-As the interposer signal bandwidth becomes more constrained, there is some impact to routability, but even an interposer providing only 20\% of the monolithic FPGA routing channel bandwidth (22 wires in our experiments) only needs a 10\% increase in the within-FPGA-die routing channel width to compensate.
+As the interposer signal bandwidth becomes more constrained, routability is impacted but even an interposer providing only 20\% of the monolithic FPGA routing channel bandwidth (22 wires in our experiments) only needs a 10\% increase in the within-FPGA-die routing channel width to compensate.
 
 The \textit{critical path delay} is not strongly influenced by the \textit{\% wires cut} but is strongly influenced by the interposer delay and the number of cuts.
-Our results show that the critical path often crosses the interposer cutline more than once, as the \textit{critical path delay} is increased by more than the interposer \textit{delay increase}; on average we find that a 1 ns interposer delay causes an 11\% delay increase in a two-die system vs.
-a monolithic FPGA .
-Increasing the number of dice in an interposer-based FPGA does not greatly impact the \textit{minimum channel width}, but does lead to a larger \textit{critical path delay}.
-We also found that the aspect ratio of the interposer-based FPGA is an important architectural parameter; for the best routability one should use a square interposer and rectangular FPGA dice.
+On average we find that a 1 ns interposer delay causes an 11\% delay increase in a two-die system vs.
+a monolithic FPGA.
+Increasing the number of dice in an interposer-based FPGA only moderately impacts the \textit{minimum channel width}, but does lead to a larger \textit{critical path delay}.
+The aspect ratio of the interposer-based FPGA is an important architectural parameter; a square interposer containing rectangular FPGA dice has the best routability.
 
 Overall we find the results for interposer-based FPGAs to be very positive.
-While an interposer can only provide a fraction of the signal count between dice that one would have across cutline within a die, the system still has excellent routability and behaves similarly to a monolithic FPGA so long as the interposer signal count is not pushed to very low values; below 20\% of the within-die routing in our experiments.
-We have also found that modifications to the FPGA CAD flow and routing architecture to exploit interposer-based FPGAs have been very productive.
-As this is a new area we believe more routability and timing enhancements are likely with further research, and this would further enhance the utility of interposer-based FPGAs.
+While an interposer can only provide a fraction of the signal count between dice that one would have across cutline within a die, the system still has excellent routability so long as the interposer signal count is not pushed to very low values; below 20\% of the within-die routing in our experiments.
+We have also found that modifications to the CAD flow and routing architecture to exploit interposer-based FPGAs have been very productive.
+As this is a new area we believe more routability and timing enhancements are likely with further research; for example, a timing-driven partitioner may reduce the delay impact of the interposer.
 
@@
 \section*{Acknowledgment}
-This work was supported by a Ci\^{e}ncia sem Fronteiras scholarship from CNPq - Brazil and the NSERC/Altera Industrial Research Chair in Programmable Silicon.
-Some of our computations were performed on the GPC supercomputer at the SciNet HPC Consortium.
-SciNet is funded by: the Canada Foundation for Innovation under the auspices of Compute Canada; the Government of Ontario; Ontario Research Fund - Research Excellence; and the University of Toronto.
+This work was supported by Brazil's Ci\^{e}ncia sem Fronteiras scholarship and the NSERC/Altera Industrial Research Chair in Programmable Silicon.
+Some computations were performed on the GPC supercomputer at the SciNet HPC Consortium.
 

### 2014-12-29 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Trivial edits
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4568 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Architecture modifications to add routing flexibility when crossing the interposer are very beneficial and improve routability by a further 11\%.
-With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 40\% of the routing capacity that the normal (within-die) FPGA routing channels supply there is negligible impact on circuit routability.
+With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 40\% of the routing capacity that the normal (within-die) FPGA routing channels supply, there is negligible impact on circuit routability.
 Smaller interposer routing capacities do impact routability however: minimum channel width increases by 10\% and 70\% when an interposer supplies only 20\% and 10\% of the within-die routing, respectively.
@@
 
-First, The presence of multiple dice in an interposer-based FPGA is modeled by creating horizontal cuts in the FPGA.
+First, the presence of multiple dice in an interposer-based FPGA is modeled by creating horizontal cuts in the FPGA.
 The \textit{number of cuts} is specified by the user, and we use that to determine the y-location of the cuts.

### 2014-12-30 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Ehsan's final revisions. removed a couple of figures and wordsmith the rest of my sections. I should re-read the placer optimizations section to make sure the formulas are consistent.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4569 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Practical interposers, however, can only accommodate a fraction of all vertical routing wires.
-When \textit{fanin transfer} option is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect some adjacent ``dangling" vertical routing wires to the multiplexer.
-This allows for additional flexibility in routing by ensuring that \textit{every} vertical routing wire could \textit{potentially} drive an interposer wire through a multiplexer.
+When \textit{fanin transfer} option is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect adjacent ``dangling" vertical routing wires to the multiplexer.
+This enables additional flexibility in routing by ensuring that \textit{every} vertical routing wire could \textit{potentially} drive an interposer wire through a multiplexer.
 As an example, if \textit{\% wires cut} $=75\%$, 3 of every 4 interposer wires are cut.
-Therefore, we place a 4-to-1 multiplexer at the input of the one interposer wire that is not cut, and we connect the 3 vertical routing wires that do not have a corresponding interposer wire to the multiplexer.
+Therefore, we place a 4-to-1 mux at the input of the one interposer wire that is not cut, and we connect the 3 vertical routing wires that do not have a corresponding interposer wire to the multiplexer.
 
 \item Fan-out Transfer for Interposer Wires\\
-When \textit{fanout transfer} option is \textit{on}, we allow the interposer wire output to drive not only its own corresponding vertical wire at the other side of the cut, but also the other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
+When \textit{fanout transfer} option is \textit{on}, we allow an interposer wire to drive not only its own corresponding vertical wire at the other side of the cut, but also other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
 
 \item Bidirectional Interposer Wires\\
-Dominant routing architectures in modern FPGAs use unidirectional wires~\cite{unidirectional, lewis2005stratix}.
-Such wires can only be driven at one point, normally their start point.
+Dominant routing architectures in modern FPGAs use unidirectional wires~\cite{unidirectional, lewis2005stratix} which can only be driven at one point, normally their start point.
 When this option is \textit{on}, we allow interposer wires to be driven from one end or the other by adding tri-state buffers at both ends.
@@
 vertical routing wires) that cross a cutline in the rrgraph, and we replace them with 2 rrnodes: one that is above the cutline and one that is below the cutline.
-We properly transfer fan-ins and fan-outs of the original node onto the two newly created nodes and we connect the two nodes using an interposer node.
+We properly transfer fan-ins and fan-outs of the original node onto the two newly created nodes and connect the two nodes using an interposer node.
 Finally, we remove \textit{all} edges from \textit{\%wires cut} percent of interposer nodes and increase the switch delay of output edges of the remaining interposer nodes by \textit{increased delay}.
@@
 \label{eq:timing_eq_full}
-Timing\_Cost=\sum_{\forall i, j \subset circuit} delay(\Delta x_{ij},\Delta y_{ij}) \times \\ 
-criticality(i,j)
+Cost_{T}=\sum_{\forall i, j \subset circuit} delay(\Delta x_{ij},\Delta y_{ij}) \times 
+Criticality(i,j)
 \end{multline}
@@
 delay(i,j) = delay(\Delta x_{ij}, \Delta y_{ij}) + \\
-times\_crossed(i,j) \times delay\_increase
+N_{crossed}(i,j) \times delay\_increase
 \end{multline}
@@
 
-where $times\_crossed(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
+where $N_{crossed}(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
 
@@
 \label{eq:bbcost}
-wiring\_cost_{orig} = \sum_{n=1}^{N_{nets}} q(n) \times [\frac{bb_x(n)}{avg\_chanx\_W(n)} + \\
-\frac{bb_y(n)}{avg\_chany\_W(n)}]
+Cost_{W\_orig} = \sum_{n=1}^{N_{nets}} q(n) \times [\frac{bb_x(n)}{W_{chanx}} + \frac{bb_y(n)}{W_{chany}}]
 \end{multline}
@@
 where $bb_x(n)$ and $bb_y(n)$ are the dimensions of the bounding box of net \textit{n} in the x and y directions, respectively.
-$avg\_chanx\_W$ and $avg\_chany\_W$ are the average x-directed and y-directed channel widths over this bounding box.
+$W_{chanx}$ and $W_{chany}$ are the average x-directed and y-directed channel widths over this bounding box.
 Finally, $q(n)$ is a function obtained from \cite{icann} which models the fact that bounding boxes underpredict the required routing for high fanout nets.
@@
 VPR's wiring cost also considers the FPGA to be homogeneous, and uses only the number of nets, the size of the net's bounding box and the average number of wires per channel to calculate the cost.
-Thus, to account for the reduced connectivity near the cutlines we create an extra cost term, \textit{cut\_cost}.
+Thus, to account for the reduced connectivity near the cutlines we create an extra cost term, $Cost_{cut}$.
 This new cost is added to (\ref{eq:bbcost}) to create the total wiring cost.
@@
 \label{eq:total_wiring}
-wiring\_cost = wiring\_cost_{orig} + cut\_cost
+Cost_{W} = Cost_{W\_orig} + Cost_{cut}
 \end{equation}
@@
 
-In~\cite{interposer2014}, Pereira et al tested several different \textit{cut\_cost} formulations, as well as different weighting for each.
-They found that the following \textit{cut\_cost} results in the best quality of results in terms of minimum routable channel width and critical path delay.
+In~\cite{interposer2014}, Pereira et al tested several different $Cost_{cut}$ formulations, and different weighting for each.
+They found that the following $Cost_{cut}$ results in the best quality of results in terms of minimum routable channel width and critical path delay.
 
@@
 \label{eq:cost2}
-cut\_cost = \sum_{n=1}^{N_{nets}} C' \times bbHeight(i) \times times\_crossed(n)
+Cost_{cut} = \sum_{n=1}^{N_{nets}} C' \times bbHeight(i) \times N_{crossed}(n)
 \end{multline}
@@
 \label{eq:lookahead_cost}
-Total\_Cost(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectedCost(r,j)
+Cost_{Total}(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectedCost(r,j)
 \end{multline}
@@
 
-In a similar way, we modify the $ExpectedCongestionCost(r,j)$ in (\ref{eq:expected_routing_cost}) to account for the new interposer routing nodes.
+We modify $ExpectedCongestionCost(r,j)$ in (\ref{eq:expected_routing_cost}) in a similar way to account for the new interposer routing nodes.
 
-The interposer-aware router lookahead is able to make routing decisions more quickly given the large delay difference of a path that crosses the interposer as opposed to a path that doesn't.
+The interposer-aware router lookahead makes routing decisions more quickly given the large delay difference of a path that crosses the interposer as opposed to a path that doesn't.
 
@@
 The new router lookahead changes the priority of the routes discovered by the router.
-Using the new router lookahead function increases the critical path delay by 1.3\% on average as shown in Figure~\ref{fig:lookahead_fmax}.
+Using the new router lookahead function increases the critical path delay by 1.3\% on average.
 We consider this small quality loss acceptable and use the enhanced router lookahead in all our experiments due to its greatly reduced runtime.
-
-\begin{figure}[!hb]
-\centering
-\includegraphics[width=\linewidth]{router_opt_fmax.eps}
-\caption{Impact of new vs. original router lookahead on the critical path delay.}
-\label{fig:lookahead_fmax}
-\end{figure}
 
@@
 To analyze the impact of the number of cut wires we ran experiments varying only this parameter while leaving the other parameters constant.
-We use \textit{number of cuts} $= 1$ (corresponding to two dice), \textit{increased delay} $= 1ns$, \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
+We use \textit{number of cuts} $= 1$ (i.e.
+two dice), \textit{increased delay} $= 1ns$, \textit{bidirectional interposer wires = on}, \textit{fanin transfer = on}, and \textit{fanout transfer = on}.
 
@@
 \label{fig:standard_minW}
-\end{figure}
-
-The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
-Note, however, that this same critical path delay is achieved at a much higher channel width when \textit{\% wires cut} is greater than 80\%.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{standard_crit_path_new.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
-\label{fig:standard_crit}
 \end{figure}
@@
 
+The critical path delay, depicted in Figure~\ref{fig:delays_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut.
+Note, however, that this same critical path delay is achieved at a much higher channel width when \textit{\% wires cut} is greater than 80\%.
+
 %   - Delay vs % wires cut & wire delay
 \subsection{Circuit Speed vs. Interposer Delay}
-
-\begin{figure}[!b]
-\centering
-\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
-\label{fig:delays_crit}
-\end{figure}
-
 To investigate the impact of the interposer delay (\textit{delay increase}), we keep all other interposer parameters constant and sweep the \textit{delay increase} from 0 to 1.5 \textit{ns}.
@@
 But once again, the critical path delay shows little correlation to the \textit{\% wires cut}.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
+\label{fig:delays_crit}
+\end{figure}
 

### 2014-12-30 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Ehsan: move graph to the right section
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4570 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \subsection{Circuit Speed vs. Interposer Delay}
+
+\begin{figure}[!b]
+\centering
+\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
+\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
+\label{fig:delays_crit}
+\end{figure}
+
 To investigate the impact of the interposer delay (\textit{delay increase}), we keep all other interposer parameters constant and sweep the \textit{delay increase} from 0 to 1.5 \textit{ns}.
@@
 But once again, the critical path delay shows little correlation to the \textit{\% wires cut}.
-
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=\linewidth]{delays_crit_path_new.eps}
-\caption{Critical path delay vs. \textit{\% wires cut} for 2 dice and $0.0$, $0.5$, $1.0$ and $1.5ns$ of \textit{delay increase}.}
-\label{fig:delays_crit}
-\end{figure}
 

### 2014-12-31 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Updated 4 cut results to only go up to 85% wires cut, since we had trouble at 90%
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4571 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
fixing the text to be consistent with the terms in the equations. Ehsan: done with my edits. what's left to do is just proofreading the whole paper.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4572 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-\subsection{Virtex-7 Interposer-based FPGAs}
+\subsection{Virtex Interposer-based FPGAs}
 \label{virtex7section}
@@
 \label{eq:cprime}
-C' = \frac{ratio\_wires\_cut}{avg\_chany\_W(n)}
+C' = \frac{ratio\_wires\_cut}{W_{chany}}
 \end{equation}
@@
 \textit{bbHeight(n)} is the height of the bounding box of the net \textit{n} and $ratio\_wires\_cut$ is the ratio of wires cut at the cutline.
-This formulation of $C'$ ensures that the new cost term is of roughly the same magnitude as $wire\_cost_{orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
+This formulation of $C'$ ensures that the new cost term is of roughly the same magnitude as $Cost_{W\_orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
 
-This cost function performs better than several others as shown in \cite{interposer2014}, and we believe the key to the good performance of this cost functions is that \textit{cut\_cost} produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
+This cost function performs better than several others as shown in \cite{interposer2014}, and we believe the key to the good performance of this cost functions is that $Cost_{cut}$ produces gradual gains as bounding boxes crossing the cutline shrink to be closer and closer to being captured entirely on one side of the cutline.
 Consider for example the 3 bounding boxes shown in Figure \ref{fig:bb_illustration}.
@@
 The optimizations also guide the placer to keep related entities on the same die to reduce the number of inter-die signal crossings and hence increase circuit speed.
-This speed improvement is fairly consistent across a wide range of \% wires cut, averaging 12.5\% higher circuit speed.
+This speed improvement is fairly consistent across a wide range of \%~wires~cut, averaging 12.5\% higher circuit speed.
 \begin{figure}[!htbp]
@@
 \input{wires_cut_4part}
-\caption{Impact of number of dice on routability. TODO JAVEED -- LU32 90\% result is missing}
+\caption{Impact of number of dice on routability.}
 \label{fig:wires_cut_4part}
@@
 \input{crit_path_4part}
-\caption{Impact of number of dice on critical path delay. TODO JAVEED -- LU32 90\% result is missing}
+\caption{Impact of number of dice on critical path delay.}
 \label{fig:crit_path_4part}

### 2014-12-31 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
resolved some todos
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4573 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-Figure \ref{fig:hyperedge_to_graph} shows how a netlist of three blocks and two nets is transformed into its component subgraphs. TODO JAVEED: Maybe just show star 1/n from start to finish.
+Figure \ref{fig:hyperedge_to_graph} shows how a netlist of three blocks and two nets is transformed into its component subgraphs.
 
@@
 To select the best hyperedge model, we used the Metis partitioner to generate graph partitions (which assign each vertex to a partition number) and applied the partitioning result to the original untransformed hypergraph.
-To rank the different models, we computed the \emph{hyperedge cutsize} for every hyperedge model, across several circuits.
+To rank the different models, we computed the \emph{hyperedge cutsize} for every hyperedge model, across the same circuits as in Section \ref{sec:CADeffect}.
 The hyperedge cutsize is the number of hyperedges crossing the cutline, as illustrated in Figure \ref{fig:hyperedge_cutline}.
@@
 %\includegraphics[width=\linewidth]{graph_topology_mcw.eps}
-\caption{Minimum channel width achieved by VPR with different hyperedge models and weightings. n is the number of vertices on a hyperedge. TODO JAVEED: Which circuits?}
+\caption{Minimum channel width achieved by VPR with different hyperedge models and weightings. n is the number of vertices on a hyperedge.}
 \label{fig:graph_topology_mcw}

### 2015-01-06 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
resolved some todos
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4573 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
fix typo in formula (5)
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4574 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \label{eq:cost2}
-Cost_{cut} = \sum_{n=1}^{N_{nets}} C' \times bbHeight(i) \times N_{crossed}(n)
+Cost_{cut} = \sum_{n=1}^{N_{nets}} C' \times bbHeight(n) \times N_{crossed}(n)
 \end{multline}
@@
 
-Systems with more dice are slower; on average the critical path delay is xx\% higher for a four-die system than a two-die system.
+Systems with more dice are slower; on average the critical path delay is 40\% higher for a four-die system than a two-die system.
 This speed difference is not significantly affected by the \textit{\% wires cut}.

### 2015-01-07 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Fixing hardcoded section number
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4576 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Interestingly, other recent work targeting conventional FPGAs has shown that using a partitioner to guide packing benefits routability\cite{feng2014rent}.
-On the other hand, The VPR placement algorithm with the enhancements described in Section 4 sees only a modest routability benefit from the partitioning constraints.
+On the other hand, The VPR placement algorithm with the enhancements described in Section \ref{cad_enh_placer_subsection} sees only a modest routability benefit from the partitioning constraints.
 

### 2015-01-07 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Use roman numerals where lists can be confused with subsections
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4577 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Making a note on the absence of the 90% wires cut result from old paper
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4578 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Updated text for missing 90% wires cut data
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4579 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-\begin{enumerate}
+\begin{enumerate}[label=(\roman*)]
   \item Leaving both the packer and the placer unconstrained. This is the base flow that does not use partitioning, as described in Section \ref{cadSection}.
@@
 We transformed the circuit netlist hypergraph to a graph in the following way:
-\begin{enumerate}
+\begin{enumerate}[label=(\roman*)]
 \item For each hyperedge of the circuit netlist hypergraph, we generated a graph (the \emph{per-net subgraph}).
@@
 \input{wires_cut}
-\caption{Impact of partitioning-based CAD flow on routability for a range of architectures.}
+\caption{Impact of partitioning-based CAD flow on routability for a range of architectures.\protect\footnotemark}
 \label{fig:wires_cut}
 \end{figure}
-
+\footnotetext{The best flow from~\cite{interposer2014} does not provide data at 90\% wires cut.}
 \subsubsection{Bloating of the Clustered Netlist due to Partitioning Constraints}

### 2015-03-04 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Updated hierarchy in section V
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4587 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-\subsection{CAD Flow Variations}
 The overall CAD flow, including the partitioning steps, is shown in Figure \ref{fig:partitioning_cad_flow}.
@@
 
-\subsubsection{Partitioner Stage}\label{sec:partitioner_stage}
+\subsection{Partitioner Stage}\label{sec:partitioner_stage}
 As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow.
@@
 
-\subsubsection{Partitioning Tool}
+\subsection{Partitioning Tool}
 Two well-known freely available graph partitioning tools are Metis~\cite{karypis1998multilevelmetis} and hMetis~\cite{karypis1999multilevelhmetis}.
@@
 
-\subsubsection{Hypergraph to Graph Transformation}\label{sec:hypergraph_to_graph}
+\subsection{Hypergraph to Graph Transformation}\label{sec:hypergraph_to_graph}
 We transformed the circuit netlist hypergraph to a graph in the following way:
@@
 
-\subsubsection{Partitioner Stage Results}\label{sec:partitioner_stage_results}
+\subsection{Partitioner Stage Results}\label{sec:partitioner_stage_results}
 We compare the performance of the four CAD flow variations described in Section \ref{sec:partitioner_stage} on the 8 largest VTR benchmarks.
@@
 
-\subsubsection{\% wires cut}
+\subsection{\% wires cut}
 To see how well the CAD flow works for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits.
@@
 \footnotetext{The best flow from~\cite{interposer2014} does not provide data at 90\% wires cut.}
-\subsubsection{Bloating of the Clustered Netlist due to Partitioning Constraints}
+\subsection{Bloating of the Clustered Netlist due to Partitioning Constraints}
 The additional constraints on the packer imposed by partitioning may result in a less dense packing.

### 2015-03-04 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Reworded issues to issue, re: metis vs hmetis
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4588 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Commented on lack of optimality of random partition assignment
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4589 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Commented on origin of edge weight schemes
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4590 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Each partition is then assigned (randomly) to a separate FPGA die.
+Our use of random assignment is not optimal, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice.
 The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines.
@@
 Owing to the natural hypergraph representation, hMetis would be the most appropriate choice of partitioner.
-However, there are other issues to consider for our use case.
+However, another important issue is the resource balancing ability of each tool.
 
@@
 Edge weights were computed based on the number of vertices in the originating hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.
+Our intuition is that assigning lower edge weight to high-fanout nets may be beneficial, and we selected the edge weights accordingly.
 We use \emph{hyperedge model} to refer to the combination of graph topology and edge weight scheme.

### 2015-03-04 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
changing axis label from "MinChW" to "Minimum Channel Width"
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4580 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
TVLSI requires at least 5 Index terms. We had 4. I added one more: CAD.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4586 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Commented on interposer crossings as contributing to path delay
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4591 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \begin{IEEEkeywords}
-Algorithms, FPGA, Silicon interposer, 2.5D ICs. 
+Algorithms, FPGA, Silicon interposer, 2.5D ICs, CAD. 
 \end{IEEEkeywords}
@@
 \centering
-\includegraphics[width=\linewidth]{placer_opt_minChW.eps}
+\includegraphics[width=\linewidth]{placer_opt_minChW_2.eps}
 \caption{Impact of placer optimizations on minimum channel width.}
@@
 \centering
-\includegraphics[width=\linewidth]{additional_fanins.eps}
+\includegraphics[width=\linewidth]{additional_fanins_2.eps}
 \caption{Adding up to 8 extra fanins to interposer wire muxes helps reduce minimum channel width.}
@@
 \centering
-\includegraphics[width=\linewidth]{minChW.eps}
+\includegraphics[width=\linewidth]{minChW_2.eps}
 \caption{Minimum channel width vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}
@@
 \centering
-\includegraphics[width=\linewidth]{numberofcrossingwires_new.eps}
+\includegraphics[width=\linewidth]{numberofcrossingwires_new_2.eps}
 \caption{Geometric mean of required intra-die minimum channel width vs. geometric mean of the number of wires crossing the interposer for 2 dice and $1ns$ of \textit{delay increase}.}
@@
 
-Systems with more dice are slower; on average the critical path delay is 40\% higher for a four-die system than a two-die system.
+As shown in \ref{fig:crit_path_4part}, systems with more dice are slower; on average the critical path delay is 40\% higher for a four-die system than a two-die system.
 This speed difference is not significantly affected by the \textit{\% wires cut}.
+This suggests that the main cause of the slowdown is due to interposer crossings, rather than wire congestion.
 \begin{figure}[!htbp]

### 2015-03-08 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Final revisions by Ehsan based on Google reviewers' comments
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4592 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-where $N_{crossed}(i,j)$ is the number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
+where $N_{crossed}(i,j)$ is the minimum number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
 
@@
 Adding additional fanin beyond transferring the fanin of cut wires is only a small further benefit so we use additional fanin=0 in subsequent sections.
-Adding these additional multiplexers will have a very small impact on die size as 10,000 muxes per cut would be typical in a large FPGA~\cite{xilinx7series} which is a very small fraction of the several million routing multiplexers in a large FPGA die.
+Adding these additional multiplexers and making the interposer drivers tri-statable causes less than a 0.1\% die area increase as 10,000 muxes per cut would be typical in a large FPGA~\cite{xilinx7series} which is a very small fraction of the several million routing multiplexers in a large FPGA die.
 
@@
 \centering
-\includegraphics[width=\linewidth]{minChW_2.eps}
+\includegraphics[width=\linewidth]{MinChW_3.eps}
 \caption{Minimum channel width vs. \textit{\% wires cut} for 2 dice and $1ns$ of \textit{delay increase}.}

### 2015-03-08 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Adding "more than moore" reference.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4593 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-Interposer-based FPGAs allow systems larger than a single die, making a "More than Moore" improvement on the size and number of logic cells possible, and with chips combined with far more connectivity and less delay than if they were in separate packages connected through a printed circuit board.
+Interposer-based FPGAs allow systems larger than a single die, making a "More than Moore" improvement \cite{arden2010morethan} on the size and number of logic cells possible, and with chips combined with far more connectivity and less delay than if they were in separate packages connected through a printed circuit board.
 

### 2015-04-14 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Two minor updates by Vaughn (Virtex Ultrascale is shipping and our partition assignment is suboptimal only for >2 dice) plus added cover letter.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4595 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Even though all these FPGAs use a 28 nm process, silicon interposer technology allows the creation of FPGAs with twice the resources possible on even an extremely large single die.
-Xilinx's recently announced Virtex UltraScale line of 20 nm FPGAs makes even more aggressive use of interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.
+Xilinx's latest (Virtex UltraScale, 20 nm) line of FPGAs makes even more aggressive use of interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.
 

### 2015-04-14 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Two minor updates by Vaughn (Virtex Ultrascale is shipping and our partition assignment is suboptimal only for >2 dice) plus added cover letter.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4595 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 Each partition is then assigned (randomly) to a separate FPGA die.
-Our use of random assignment is not optimal, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice.
+Our use of random assignment is sub-optimal for systems with more than two dice, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice.
 The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines.

### 2015-04-14 doc/journal/fpgapaperfinal/FPGAPaper/difference_from_prior.tex
Shortened abstract and added differences from prior conference paper section.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4596 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\section*{Differences from Previous Work}
This work has not been previously published. It builds upon an earlier conference publication ``CAD and Routing Architecture for Interposer-Based Multi-FPGA Systems,'' A. Hahn Pereira and V. Betz, \textit{ACM Int. Symposium on FPGAs}, 2014, pp. 75 - 84, but greatly improves on and extends this earlier work.  Approximately 75\% of the material in this paper is entirely new methods and new results that are not in the conference paper; the remaining 25\% of the material (chiefly the Introduction, Background, some of the Architecture Models and the Placer Optimization subsection of CAD Enhancements) have been updated since the conference paper but cover the same basic material as the conference paper. In detail, new material includes:

\begin{itemize}
\item Partitioning is now included in the CAD flow. We investigate multiple ways to formulate the partitioning problem for interposer-based FPGAs and compare the results to CAD flows without partitioning. All of Section V of the paper (4 pages) is completely new material vs. the conference paper.

\item Section III, Architecture Models has been expanded vs. the conference paper, and the method of modeling the architecture in the CAD tools is new.  Section III.D is completely new and Section III.E uses a different method than the conference paper and has been rewritten and expanded.

\item Section IV, CAD enhancements contains mostly new material vs. the conference paper. The router enhancements section (IV.B) is entirely new, while the placement enhancement section (IV.A) is shortened and the effectiveness of the enhancements section (IV.C) is new.

\item All of the architecture results have been re-run with the new CAD flow and architectural enhancements proposed in this paper, so all the results in Section VI are different and markedly better than those of the conference paper; this entire section has been re-written. Sections VI.A and VI.E investigate entirely new architecture questions proposed in this paper; there is no analog to these sections in the earlier conference paper.

\item The Introduction, Background and Conclusions (Section I, II and VII) have been updated to highlight the new results of this paper and to describe recent developments in commercial interposer-based FPGAs at 20 nm.
\end {itemize}

### 2015-04-14 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Shortened abstract and added differences from prior conference paper section.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4596 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \begin{abstract}
-Multiple FPGA dice can be connected through a silicon interposer to form a larger system than one monolithic die can accommodate.
-Such systems are now commercially available but many open questions remain concerning their key architecture parameters and efficiency, as the signal count between dice is reduced and the signal delay between dice is increased compared to a monolithic FPGA.
-We create a new version of VPR to target interposer-based FPGAs and investigate modifications to placement and routing and the incorporation of partitioning into the flow to improve results.
-Our best CAD flow reduces the routing demand for interposer FPGAs with realistic amounts of connectivity between dice by 47\% and improves the circuit speed by 13\% on average.
+Large FPGA systems with multiple dice connected by a silicon interposer are now commercially available.
+However, many questions remain concerning their key architecture parameters and efficiency, as the signal count between dice is reduced and the delay between dice is increased compared to a monolithic FPGA.
+We modify VPR to target interposer-based FPGAs and investigate placement and routing changes and incorporating partitioning into the flow to improve results.
+Our best CAD flow reduces the routing demand for interposer FPGAs with realistic connectivity between dice by 47\% and improves the circuit speed by 13\% on average.
 Architecture modifications to add routing flexibility when crossing the interposer are very beneficial and improve routability by a further 11\%.
-With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 40\% of the routing capacity that the normal (within-die) FPGA routing channels supply, there is negligible impact on circuit routability.
-Smaller interposer routing capacities do impact routability however: minimum channel width increases by 10\% and 70\% when an interposer supplies only 20\% and 10\% of the within-die routing, respectively.
+With these CAD and architecture enhancements, we find that if an interposer supplies (between dice) 20\% of the routing capacity that the normal (within-die) FPGA routing channels supply, there is only a modest impact on circuit routability.
+Smaller interposer routing capacities do impact routability however: minimum channel width increases by 70\% when an interposer supplies only 10\% of the within-die routing.
 The interposer also impacts delay, increasing circuit delay by 11\% on average for a 1 ns interposer signal delay and a two-die system.
@@
 
+\input{difference_from_prior.tex}
+
 % that's all folks

### 2015-04-20 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Adding biographies with photos.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4597 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
+% ----- BIOGRAPHY ------------------
+% \vspace{-2in}
+\vfill
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{enasiri}}]{Ehsan Nasiri} received the Bachelor of Applied Science in Computer Engineering from the University of Toronto in 2012, and the M.Eng.
+degree in ECE from the University of Toronto in 2015.
+He has been a Software Engineer at Altera since 2012.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{javeed}}]{Javeed Shaikh} received the BEng degree in Electrical Engineering from McGill University in 2011 and the MEng degree in ECE from the University of Toronto in 2014.
+He has been a Software Engineer at Google since 2013.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{apereira}}]{Andre Hahn Pereira} is currently pursuing his B.Sc degree in ECE at the University of S\~{a}o Paulo.
+He was an exchange student at the University of Toronto from September 2012 until August 2013.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{vbetz}}]{Vaughn Betz}(S'88--M'91) received the B.Sc degree in EE from the University of Manitoba in 1991, the MS degree in ECE from the University of Illinois at Urbana-Champaign in 1993, and the PhD degree in ECE from the University of Toronto in 1998.
+
+Dr.
+Betz was a co-founder of Right Track CAD in 1998 and VP of Engineering until its acquisition by Altera in 2000.
+He was at Altera from 2000 to 2011, ultimately as Senior Director of Software Engineering.
+He joined the University of Toronto as an Associate Professor in 2011, where his research covers FPGA architecture, CAD and FPGA-based computation.
+\end{IEEEbiography}
+
+% ----------------------------------
+
 \input{difference_from_prior.tex}

### 2015-04-26 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Cut some less important text and modified figure sizes to fit in 14 pages.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4601 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \centering
-\includegraphics[width=0.65\linewidth]{fpga.eps}
+\includegraphics[width=0.6\linewidth]{fpga.eps}
 \caption{A sample two-die / 1-cutline architecture containing both logic blocks and RAM blocks.}
@@
 \centering
-\includegraphics[width=\linewidth]{architecture_params.eps}
+\includegraphics[width=0.75\linewidth]{architecture_params.eps}
 \caption{For a channel width of 4 and 75\% wires cut, only 1 interposer wire exists (in green) and 3 interposer wires are cut (in the red).
@@
 While this is a reasonable method to model such architectures, it makes it difficult to experiment with different routing-interposer interfaces such as placing input muxes for interposer wires.
-Therefore, we modify the rrgraph differently to more accurately reflect the interposer layer directly:
+Therefore, we modify the rrgraph differently to more accurately reflect the interposer layer directly.
 
 \begin{figure}[!t]
-\subfigure[]{\includegraphics[width=0.24\linewidth]{a.eps}}
-\subfigure[]{\includegraphics[width=0.24\linewidth]{b.eps}}
-\subfigure[]{\includegraphics[width=0.24\linewidth]{a_prime.eps}}
-\subfigure[]{\includegraphics[width=0.24\linewidth]{b_prime.eps}}
+\subfigure[]{\includegraphics[width=0.17\linewidth]{a.eps}}
+\subfigure[]{\includegraphics[width=0.17\linewidth]{b.eps}}
+\subfigure[]{\includegraphics[width=0.17\linewidth]{a_prime.eps}}
+\subfigure[]{\includegraphics[width=0.17\linewidth]{b_prime.eps}}
 \caption{(a) Regular routing wire connectivity.
@@
 \begin{figure*}[!thpb]
-\subfigure[]{\includegraphics[width=0.6\textwidth]{gui_1.eps}}
+\centering
+\subfigure[]{\includegraphics[width=0.5\textwidth]{gui_1.eps}}
 \subfigure[]{\includegraphics[width=0.4\textwidth]{gui_2.eps}}
@@
 Commercial FPGAs are laid out with identical columns, and hence combine multiple identical FPGA dice in \textit{one dimension} on an interposer \cite{xilinxTSV}.
-As Figure \ref{fig:square_tall} shows, we can make \textit{either} the FPGA dice square, or the interposer square, but not both.
-Table \ref{table:squares} compares the routability of these two options for the four-die case: a square interposer (four rectangular dice) or a rectangular interposer (four square dice.) We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same routing architecture parameters as Section \ref{sec:resultsWiringCapacity}.
-
-\begin{figure}[!htbp]
-\centering
-\subfigure[]{\includegraphics[height = 3cm]{square_fpga.pdf}}
-\subfigure[]{\includegraphics[height = 4.6675cm]{tall_fpga.pdf}}
-\caption{Devices with (a) a square interposer and (b) square dice.}
-\label{fig:square_tall}
-\end{figure}
+We can make \textit{either} the FPGA dice square, or the interposer square, but not both.
+Table \ref{table:squares} compares the routability of these two options for the four-die case.
+We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same routing architecture parameters as Section \ref{sec:resultsWiringCapacity}.
 
@@
 
-The square interposer configuration performs significantly better than the square dice configuration.
-A square interposer results in cutlines that are closer together, and hence more logic blocks are near a cutline.
-On the other hand, a square interposer (composed of rectangular FPGA dice) has more vertical routing channels, giving it a higher absolute number of wires between FPGA dice.
-The second effect is the more important one, leading to much better routability for a square interposer.
+Compared to square dice, a square interposer (composed of rectangular FPGA dice) results in cutlines that are closer together, and hence more logic blocks are near a cutline.
+On the other hand, a square interposer has more vertical routing channels, giving it a higher absolute number of wires between FPGA dice.
+The second effect is the more important one, leading to much better routability for a square interposer than for square dice.
 
@@
 Overall we find the results for interposer-based FPGAs to be very positive.
-While an interposer can only provide a fraction of the signal count between dice that one would have across cutline within a die, the system still has excellent routability so long as the interposer signal count is not pushed to very low values; below 20\% of the within-die routing in our experiments.
+While an interposer can only provide a fraction of the signal count between dice that one would have within a die, the system still has excellent routability so long as the interposer signal count is not pushed to very low values; below 20\% of the within-die routing in our experiments.
 We have also found that modifications to the CAD flow and routing architecture to exploit interposer-based FPGAs have been very productive.
-As this is a new area we believe more routability and timing enhancements are likely with further research; for example, a timing-driven partitioner may reduce the delay impact of the interposer.
+As this is a new area, we believe more routability and timing enhancements can be achieved with further research; for example, a timing-driven partitioner may reduce the delay impact of the interposer.
 

### 2015-04-26 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Cut some less important text and modified figure sizes to fit in 14 pages.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4601 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

--- 2015-03-04 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
@@
 The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist.
-Each partition is then assigned (randomly) to a separate FPGA die.
-Our use of random assignment is not optimal, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice.
+Each partition is then assigned (randomly, though this is not optimal) to a separate FPGA die.
 The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines.
@@
 To derive an appropriate balance constraint for our use case, consider that FPGA circuit netlists may contain different block types such as LUTs, registers, RAM slices, and multipliers, among others.
-The basic balance constraint of (\ref{eq:balance_constraint}), applied to a netlist with only LUTs and multipliers, would ensure that the total block count (sum of the LUT count + multiplier count) in each partition is roughly balanced.
+The basic balance constraint of (\ref{eq:balance_constraint}), applied to a netlist with different block types, would ensure that the total block count (over all blocks, regardless of type) in each partition is roughly balanced.
 The problem with (\ref{eq:balance_constraint}) in the case of multiple block types is that the capacity, per block type, of each partition is not taken into account.
-It is therefore possible to generate partitions that cannot be realized because the number of resources of each type is not balanced across partitions, even though the total number of resources is balanced.
-For example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which may not be realizable given the per-block-type capacity of each partition.
+It is therefore possible to generate partitions that cannot be realized -- for example, there is nothing preventing the partitioner from moving all LUTs to partition 1 and all multipliers to partition 2, which may not be realizable given the per-block-type capacity of each partition.
 
@@
 
-We explored several ways to generate the per-net subgraphs.
-The two parameters in this operation are the graph topology and the edge weight scheme.
+We explored several ways to generate the per-net subgraphs, varying the graph topology and the edge weight scheme.
 We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.
@@
 Our intuition is that assigning lower edge weight to high-fanout nets may be beneficial, and we selected the edge weights accordingly.
-We use \emph{hyperedge model} to refer to the combination of graph topology and edge weight scheme.
+We refer to the combination of graph topology and edge weight scheme as the \emph{hyperedge model}.
 
@@
 \centering
-\subfigure[]{\includegraphics[width=0.45\linewidth]{star.eps}}
-\subfigure[]{\includegraphics[width=0.45\linewidth]{clique.eps}}
+\subfigure[]{\includegraphics[width=0.35\linewidth]{star.eps}}
+\subfigure[]{\includegraphics[width=0.35\linewidth]{clique.eps}}
 \caption{(a) Star and (b) clique graph topologies.}
@@
 \centering
-\subfigure[]{\includegraphics[width=0.45\linewidth]{hg2g_star.pdf}}
-\subfigure[]{\includegraphics[width=0.45\linewidth]{hg2g_clique.pdf}}
+\subfigure[]{\includegraphics[width=0.4\linewidth]{hg2g_star.pdf}}
+\subfigure[]{\includegraphics[width=0.4\linewidth]{hg2g_clique.pdf}}
 \caption{The hypergraph to graph transformation with (a) the star topology, and (b) the clique topology.}
@@
 
-To select the best hyperedge model, we used the Metis partitioner to generate graph partitions (which assign each vertex to a partition number) and applied the partitioning result to the original untransformed hypergraph.
-To rank the different models, we computed the \emph{hyperedge cutsize} for every hyperedge model, across the same circuits as in Section \ref{sec:CADeffect}.
-The hyperedge cutsize is the number of hyperedges crossing the cutline, as illustrated in Figure \ref{fig:hyperedge_cutline}.
-As it equals the number of nets that cross the interposer boundary, it is a measure of circuit routability.
+To select the best hyperedge model, we used the Metis partitioner to generate graph partitions and applied the partitioning result to the original untransformed hypergraph.
+We used the \emph{hyperedge cutsize} to rank the different hyperedge models, across the same circuits as in Section \ref{sec:CADeffect}.
+The hyperedge cutsize is the number of hyperedges (nets) crossing the cutline, as illustrated in Figure \ref{fig:hyperedge_cutline}, and it measures circuit routability.
 The results are shown in Figure \ref{fig:graph_topology_cutsize}.
@@
 \centering
-\includegraphics[width=\linewidth]{hyperedge_cutline.pdf}
+\includegraphics[width=0.75\linewidth]{hyperedge_cutline.pdf}
 \caption{A graph with six vertices, three hyperedges, and two partitions. As there are two hyperedges crossing the cutline, the hyperedge cutsize is 2.}
@@
 \subsection{\% wires cut}
-To see how well the CAD flow works for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits.
+To evaluate the CAD flow performance for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits.
 As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of~\cite{interposer2014} for FPGAs where more than half the wires are cut at interposer boundaries.

### 2015-05-07 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Shortened a few equation variables and fixed a few confusing sentences.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4610 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \label{eq:lookahead_cost}
-Cost_{Total}(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectedCost(r,j)
+Cost_{Total}(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectCost(r,j)
 \end{multline}
@@
 $\alpha$ is 1.2 by default in VPR.
-VPR's original $ExpectedCost(n,j)$ function is~\cite{betz1999architecture}:
+VPR's original $ExpectCost(n,j)$ function is~\cite{betz1999architecture}:
 
@@
 \label{eq:expected_routing_cost}
-ExpectedCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectedCongestionCost(r,j)
+ExpectCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectCongestionCost(r,j)
 \end{multline}
@@
 \label{eq:tdelay}
-T^\prime_{delay} = T_{delay} + T_{interposer} \times ExpectedInterposerHops
+T^\prime_{delay} = T_{delay} + T_{interposer} \times ExpectInterposerHops
 \end{multline}
@@
 
-We modify $ExpectedCongestionCost(r,j)$ in (\ref{eq:expected_routing_cost}) in a similar way to account for the new interposer routing nodes.
+We modify $ExpectCongestionCost(r,j)$ in (\ref{eq:expected_routing_cost}) in a similar way to account for the new interposer routing nodes.
 
@@
 Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%.
-This signifies that added flexibility to drive an interposer wire is more important for routability compared to adding flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.
+This signifies that added flexibility to drive an interposer wire is more important for routability than added flexibility allowing an interposer wire to drive multiple vertical routing wires on the other side of the cut.
 

### 2015-05-07 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Various changes to shorten a few things and make room for the square vs. rectangular interposer picture.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4611 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \begin{IEEEkeywords}
-Algorithms, FPGA, Silicon interposer, 2.5D ICs, CAD. 
+FPGA, Silicon interposer, 2.5D ICs, CAD. 
 \end{IEEEkeywords}
@@
 Another major advantage of interposer-based FPGAs comes at the beginning of a new manufacturing process, when the defect density is high~\cite{xilinxTSVperformance}.
-To illustrate this impact consider a new process in which the defect density is 1/$cm^2$ (a reasonable value early in the process lifecycle) and the die area is $6cm^2$, which roughly matches the size of the largest member of a high-end FPGA family such as Virtex 7.
+To illustrate this impact consider a new process in which the defect density is 1/$cm^2$ and the die area is $6cm^2$, roughly matching the size of the largest member of a high-end FPGA family such as Virtex 7.
 Using the Poisson Yield Model~\cite{yieldmodel}, the yield is only $0.25\%$ of dice.
@@
 Hence the number of single die FPGAs created from a $730cm^2$ silicon wafer would be $66.9$ and the number of interposer-based FPGAs created would be $104.6$.
-While the interposer-based FPGA still has a yield advantage it might not lead to a major cost advantage when the cost of the interposer and assembling the die to it are included.
+While the interposer-based FPGA still has a yield advantage it might not lead to a major cost advantage when the cost of the interposer and die assembly costs are included.
 For the large, state-of-the-art FPGAs that are built early in a process cycle and heavily used for prototyping, however, there is clearly a compelling cost advantage to an interposer-based solution.
@@
 The \textit{XC7V2000T} is composed of four identical dice arranged such that the vertical routing crosses between the dice.
-Each horizontal edge of each die has 280 groups of 48 length-12 wires crossing the interposer, which sums to a total of 13440 wires between dice.
+Each horizontal edge of each die has 280 groups of 48 length-12 wires crossing the interposer, for a total of 13440 wires between dice.
 There are also 40 clock wires crossing the interposer.
@@
 in \cite{xilinxTSV}, the bumps have to be spread out near the edge and need to go as far as $2.25mm$ away from the edge of the die.
-This greater distance from the border increase the length of the inter-die connections, and along with the presence of the micro-bumps and their capacitance, leads to an increased delay for these crossing wires vs.
+This greater distance from the border increases the length of the inter-die connections, and along with the presence of the micro-bumps and their capacitance, leads to an increased delay for these crossing wires vs.
 that of a typical on-die routing wire.
@@
 \centering
-\subfigure[]{\includegraphics[width=0.5\textwidth]{gui_1.eps}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{gui_1.eps}}
 \subfigure[]{\includegraphics[width=0.4\textwidth]{gui_2.eps}}
@@
 \label{eq:lookahead_cost}
-Cost_{Total}(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectCost(r,j)
+Cost_{Tot}(r) = \sum_{\forall l \in RT(i,r)} Cost(l) + \alpha \times ExpectCost(r,j)
 \end{multline}
@@
 \label{eq:expected_routing_cost}
-ExpectCost(r,j) = Criticality_{fac} \times T_{delay} + \\(1 - Criticality_{fac}) \times ExpectCongestionCost(r,j)
+ExpectCost(r,j) = Criticality_{fac} \times T_{del} + \\(1 - Criticality_{fac}) \times ExpectCongestionCost(r,j)
 \end{multline}
@@
 
-$T_{delay}$ is an estimate of the delay that the router predicts it will see on the route from \textit{r} to \textit{j}.
-Since (\ref{eq:expected_routing_cost}) assumes a homogeneous FPGA, $T_{delay}$ does not include the additional delay some routes experience by going through the interposer.
-Thus, for interposer-based architectures, we replace $T_{delay}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{delay}$ to include the delay of expected interposer hops from \textit{r} to \textit{j}.
+$T_{del}$ is an estimate of the delay that the router predicts it will see on the route from \textit{r} to \textit{j}.
+Since (\ref{eq:expected_routing_cost}) assumes a homogeneous FPGA, $T_{del}$ does not include the additional delay some routes experience by going through the interposer.
+Thus, for interposer-based architectures, we replace $T_{del}$ in (\ref{eq:expected_routing_cost}) with $T^\prime_{del}$ to include the delay of expected interposer hops from \textit{r} to \textit{j}.
 
@@
 \label{eq:tdelay}
-T^\prime_{delay} = T_{delay} + T_{interposer} \times ExpectInterposerHops
+T^\prime_{del} = T_{del} + T_{interpose} \times ExpectInterposeHops
 \end{multline}
@@
 Commercial FPGAs are laid out with identical columns, and hence combine multiple identical FPGA dice in \textit{one dimension} on an interposer \cite{xilinxTSV}.
-We can make \textit{either} the FPGA dice square, or the interposer square, but not both.
+As Figure~\ref{fig:square_tall} shows, we can make \textit{either} the FPGA dice square, or the interposer square, but not both.
 Table \ref{table:squares} compares the routability of these two options for the four-die case.
 We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same routing architecture parameters as Section \ref{sec:resultsWiringCapacity}.
+
+\begin{figure}[!htbp]
+\centering
+\subfigure[]{\includegraphics[height = 2.3cm]{square_fpga.pdf}}
+\subfigure[]{\includegraphics[height = 3.6cm]{tall_fpga.pdf}}
+\caption{Devices with (a) a square interposer and (b) square dice.}
+\label{fig:square_tall}
+\end{figure}
 
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{vbetz}}]{Vaughn Betz}(S'88--M'91) received the B.Sc degree in EE from the University of Manitoba in 1991, the MS degree in ECE from the University of Illinois at Urbana-Champaign in 1993, and the PhD degree in ECE from the University of Toronto in 1998.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{vbetz}}]{Vaughn Betz}(S'88--M'91) received the B.Sc(EE) from the University of Manitoba in 1991, the MS(ECE) from the University of Illinois at Urbana-Champaign in 1993, and the PhD(ECE) from the University of Toronto in 1998.
 
 Dr.
-Betz was a co-founder of Right Track CAD in 1998 and VP of Engineering until its acquisition by Altera in 2000.
+Betz co-founded Right Track CAD, which was acquired by Altera in 2000.
 He was at Altera from 2000 to 2011, ultimately as Senior Director of Software Engineering.
-He joined the University of Toronto as an Associate Professor in 2011, where his research covers FPGA architecture, CAD and FPGA-based computation.
+He is now an Associate Professor at the University of Toronto and researchs FPGA architecture, CAD and FPGA-based computation.
 \end{IEEEbiography}

### 2015-05-07 src/libtatum/base/sta_util.cpp
Add support for dumping tags.

@@ -141,7 +141,7 @@ void print_levelization(const TimingGraph& tg) {
 void print_timing_tags_histogram(const TimingGraph& tg, SerialTimingAnalyzer& analyzer, int nbuckets) {
     const int int_width = 8;
     const int flt_width = 2;
-     
+
     cout << "Node Arrival Tag Count Histogram:" << endl;
     std::map<int,int> arr_tag_cnts;
     for(NodeId i = 0; i < tg.num_nodes(); i++) {
@@ -149,7 +149,7 @@ void print_timing_tags_histogram(const TimingGraph& tg, SerialTimingAnalyzer& an
     }
 
     auto totaler = [](int total, const std::map<int,int>::value_type& kv) {
-        return total + kv.second; 
+        return total + kv.second;
     };
 
     int total_arr_tags = std::accumulate(arr_tag_cnts.begin(), arr_tag_cnts.end(), 0, totaler);
@@ -167,3 +167,19 @@ void print_timing_tags_histogram(const TimingGraph& tg, SerialTimingAnalyzer& an
         cout << "\t" << kv.first << " Tags: " << std::setw(int_width) << kv.second << " (" << std::setw(flt_width) << std::fixed << (float) kv.second / total_req_tags << ")" << endl;
     }
 }
+
+void print_timing_tags(const TimingGraph& tg, SerialTimingAnalyzer& analyzer) {
+    cout << std::scientific;
+    for(NodeId i = 0; i < tg.num_nodes(); i++) {
+        for(const TimingTag& tag : analyzer.arrival_tags(i)) {
+            cout << "Node " << i << ": clk: " << tag.clock_domain() << " Arr: " << tag.time().value() << endl;
+        }
+    }
+    for(NodeId i = 0; i < tg.num_nodes(); i++) {
+        for(const TimingTag& tag : analyzer.required_tags(i)) {
+            cout << "Node " << i << ": clk: " << tag.clock_domain() << " Req: " << tag.time().value() << endl;
+        }
+    }
+}
+
+

### 2015-05-07 src/libtatum/base/sta_util.hpp

@@ -16,3 +16,5 @@ void print_timing_graph(const TimingGraph& tg);
 void print_levelization(const TimingGraph& tg);
 
 void print_timing_tags_histogram(const TimingGraph& tg, SerialTimingAnalyzer& analyzer, int nbuckets);
+
+void print_timing_tags(const TimingGraph& tg, SerialTimingAnalyzer& analyzer);

### 2015-05-07 src/tatum_sta/main.cpp

@@ -199,6 +199,7 @@ int main(int argc, char** argv) {
 
         //Tag stats
         print_timing_tags_histogram(timing_graph, serial_analyzer, 10);
+        print_timing_tags(timing_graph, serial_analyzer);
     }
 
 /*

### 2015-06-24 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Added response_to_reviewers.txt
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4647 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

Associate Editor
Comments to the Author:
All the reviewers think this paper tackles a problem of practical interests, and the proposed solution/results are competent. Some major concerns raised by the reviewer is that the authors need to clearly clarify the details of hypergraph partitioning. Other detailed suggestions on the writing, implementations, and etc, please refer to the reviewers' comments.

Reviewer's Comments

Reviewer: 1

SPECIFIC FEEDBACK TO AUTHORS
This paper describes a method to model an interposer-based FPGA, CAD enhancements for an interposer-based architecture, and a detailed architecture study. The paper is solid architecture work, and presents many detailed and useful results.  That being said, the novelty is a bit low --  the techniques to modify the CAD flow are fairly straightforward, and the results are not all that surprising.   Still, I think the research question is important enough that readers will find useful data here.

The authors have made significant changes to their previously published conference paper including the addition of a partitioning-based flow; in my opinion, this meets the 30% new material requirement.

The paper presents significant space talking about how to transform a hyper graph to a graph using various models.  This has been well studied in the past.  It would be useful to present the proposed techniques in the context of this previous work.  As a start, there are numerous references within reference [22] that should be considered.  Doing so would help the reader understand the novelty of the hypergraph->graph methodology and results.

The partitioning flow does not appear to be timing driven.  Given that, I wonder how meaningful the results in Section VI-C (Circuit speed vs. interposer delay) are.  These results show that the critical path delay is strongly affected by the delay_increase, but any reasonable timing-driven CAD flow would try to keep the critical path on one side of the partition to minimize the number of critical path segments crossing a partition.  Can you comment on this?  Can you at least provide data regarding how many cut crossings are typical on the critical paths?

A key result that would help the reader interpret the results would be a table that shows, for each benchmark circuit, the number of signals that cross the partition (independent of the place and route).   It is important to break this down by benchmark circuit.  If one was to build such a device, it would be important to consider the range of cut requirements rather than just the average.  This is especially important considering the results of Figure 23 which suggests that increasing the channel width is not an efficient way to make up for a lack of cut capacity.

In Figure 7, it was not clear to me why, for 30% of the wires cut, the placer optimizations do worse than the placer without optimizations.  Is this just experimental noise?

Did you adjust the range windows in VPR to account for the new architecture?

Section III-E contains a lot of information about the detailed implementation of the model within VPR.  While this is useful to someone who wants to replicate the results, if it becomes necessary to cut content to address reviewer comments, some of this detail could be omitted without losing much from the paper (the fact that an extra RR_Graph node is created for each crossing is likely not as interesting to most readers, compared to the other experimental results).

I wasn’t clear on how useful the results in VI-E were.   The experiments in the rest of the paper assumed a square FPGA in VPR (before cutting) so more like Figure 27(a). The alternative in Figure 27b intuitively seems like a bad idea: (a) it is known that highly rectangle FPGAs (more columns than rows) lead to higher channel width, and (b) the number of interposers is lower, which from the rest of the paper we know is a bad idea.  So the fact that Figure 27b gives worse results should not be surprising.  If there is something else to conclude from this table, perhaps it could be highlighted.

Reviewer: 2

SPECIFIC FEEDBACK TO AUTHORS
This is a well-written paper. I think, this will allow researchers to perform research on multi-die devices.

I have few simple comments:

1. In section IV A (Placer Optimization), you have assumed that all the FPGA dies are homogeneous. Can you explain a bit about how would you handle heterogeneous dies? For example, there can be 4 dies in a chip and 1 of them is DSP-heavy, another one is Block-RAM heavy and other two are "regular" with lots of different types of logics (some CLB, some DSP, some Block-RAM, some IO etc)

2. Once initial placement is done, can you think of performing some physically knowledgeable optimization techniques? For example, if there are lot of connections between different dies, can you perform some minimization operation on that?

3. In section V B (partitioning Tool), you assumed 0.05 (5%) in line 30. What is the rationale behind choosing that number? What happens if we choose some other number like 10% or 15%?

4. In line 50 of Section V B, are you giving any special consideration to DSP or Block-RAM etc? Or are those treated the same way you treat simple LUTs and Flops?

5. In your work, maximum how many partitions are allowed? Was there any situation where partitioned could not partition the design?

6. In page 13, section E (Impact of Aspect Ratio), can you explain some other different styles?
Like

  die-1      die-2
  die-3      die-4
  die-5      die-6
  die-7      die-8

7. What is the frequency loss between monolithic design and multi-die design?

8. Any data regarding the runtime comparison of this tool between for a monolithic design and a similar multi-die design? In the FPGA-industry, runtime is a very important issue

9. Any analyze on the power consumption for multi-die designs? Any comparison of power consumption with respect to that of a monolithic design?

In spite of all these feedbacks, I like your paper. Good job

Reviewer: 3

SPECIFIC FEEDBACK TO AUTHORS
Compared to the conference version of the submission, this is a more thorough study of the performance of interposer-based FPGAs when key parameters are varied. To carry out the study, it adapted and twisted the VPR flow for conventional homogeneous FPGAs a bit. Though it may not be the ideal CAD flow for interposer-based FPGAs, some valuable insights can still be gained from the experiments.

In section IV.A, you should mention that VPR's placer is simulated annealing-based. In particular, the reason you cited in the last paragraph of IV.A is based on the assumption that you are using an iterative improvement-based placer (and not, say, an analytical placer).

For the discussion about hypergraph to graph transformation in Sec V.C, there were many papers discussing how to set the edge weights in the 90's, please cite them (for example, see the survey paper "Recent directions in netlist partioning: a survey"). In particular, setting the edge weight to 1/(n-1) instead of 1/n is a popular choice since it can model a 2-pin net exactly. So, it makes sense to report the results for 1/(n-1). It looks like your star model picks one node of a net and connect it to other nodes of the net, but the more popular star model in the physical design literature would add a dummy node for each net that connects to all nodes of the net. So please explicitly define your star model in words. And from p.9, it looks like that you also want to distinguish the source of a net from the sinks of a net in the star model, if so, please state that in the definition of your star model.

p.9, middle of left column said "the star topology clearly differentiates sources from sinks and this appears to give the partitioner an anchor point that pulls all hyperedge fanouts toward the source", you should mention that Metis is an iterative improvement-based partitioner which is related to your reasoning.

p.9, left column, 2nd last paragraph, your star 1 edge weight scheme actually assigns a total weight of n-1 (not n) to each net. Similarly, your star 1/n model assigns a total weight of (n-1)/n (not 1) to each net.

Some writing issues:

- Sec II.A, 1st paragraph, read "As described in Section III". It is a bit odd since the readers have not read section III at that point.

- The text never refers to Fig.3.

- There should not be indentation following eqts. (1), (2), (3), (5), (6).

- Sec. V.E., "It is slightly worse ... even for ..." -> "But it is slightly worse ... for ..."

Reviewer: 4

SPECIFIC FEEDBACK TO AUTHORS
This paper describes a set of extensions to the VPR tool to accommodate multi-FPGA platforms that use silicon interposers. The paper includes both architectural enhancements and modifications to the VPR CAD flow.

The main challenge is that silicon interposers have increased delay and reduced connectivity between dice. Rather than cutting wires at the border of each FPGA, the authors propose to add multiplexers at the interface (Fig. 3). This entails some minor modifications to VPR’s routing resource graph (RRG) at the cut-points. In terms of CAD support, the primary contribution of the paper are new cost functions for the placer and router, and a pre-placement partitioning phase using METIS.

I find the paper to be competent and reasonable, but at the same time, a collection of very small incremental ideas, and lacking in terms of a significant contribution to the field. (Other than a VPR update, which is always welcome because VPR is widely used in FPGA architecture and CAD research).

==

I do have two fairly significant technical quibbles with the paper; although they are somehow interdependent, they can each be taken at face value.

(1) The case for METIS over hMETIS is weak

Both METIS and hMETIS are well known tools for graph and hypergraph partitioning, respectively. Since circuits are known to be hypergraphs, the natural inclination would be to use hMETIS; however, the authors argue that METIS is more appropriate because hMETIS does not support heterogeneous balance constraints. Therefore, the authors take great pains to convert the hypergraph to a graph.

This is not satisfying to me at all. If the correct solution is to use hypergraph partitioning with heterogeneous balance constraints, then the authors should make appropriate modifications to hMETIS to produce the appropriate partitioning tool. It is unclear how much relevant information is lost by transforming the hypergraph to a graph, and what the impact on relative CAD metrics (W_min, f_max, etc) would be.

(2) I disagree with the case for balance constraints

I do not understand why a good partioning solution is approximately equal among the different dice in an interposer-based FPGA, because doing so intuitively seems as though it would make maximal use of interposer wires, which are both slow and constrained (in terms of general availability).

As a generic example, suppose that I have a two dice, each with 100 CLBs, and the circuit that I am trying to partition has 101 CLBs. Do I really want a 50/51 partition using a large number of interposer wires? I don’t think that I do. If my partition was 100/1 instead, I would probably use just a handful of interposer wires, and the likelihood that the interposer wire is critical (i.e., it is on the f_max path) is fairly low.

Now, I realize that my example is toy, and I also know that you don’t want 100% LUT/CLB utilization (referring to the title of DeHon’s FPGA 1999 paper). So, 100/1 is probably going to be unroutable on the densely used side, by 90/11 or 80/21 may work out well, and would use far fewer interposer wires than 50/51.

In summary, I question the motivation to impose balance constraints. It seems like an unbalanced partition (within the parameters that provide reasonable routability) is the right way to go. Taking points (1) and (2) together (and assuming that I am correct—which I may not be), if balance constraints are unneeded, then the case for METIS over hMETIS falls apart, so perhaps hMETIS (with an appropriate objective) is perhaps a better way to go.

==

At the very end of Section V.F, the authors briefly discuss packing bloat, which “was due almost entirely to poor packing of memory blocks: an interesting area for future work would be to augment the partitioner to understand that some RAM primitives are best kept together in one partition.”

Since I’ve already argued that the paper is somewhat lightweight in terms of technical contribution, perhaps doing a proper investigation of this issue and adding the necessary algorithmic support could bring greater novelty to this paper, rather than leaving it open to future work.

==

In Section VI.E, there is a short study on the impact of FPGA die aspect ratio on the minimum channel width (W_min). I’d like for the authors to clarify this issue a bit, because of my understanding of the way VPR’s routers work.

Firstly, the modifications to the router cost functions described earlier in the paper were for VPR’s timing driven router. To the best of my knowledge, the timing-driven router is very good at reducing critical path delay, but not particular good at minimizing W_min.

VPR also has a routability-driven router which is very good at minimizing W_min, but not particularly effective at reducing critical path delay.

So, to produce Table I, in principle, the routability-driven router should have been chosen; however, if so, then the cost function used by the routability-driven router needs to be modified as well in order to model the interposer and interface at the cut lines.

This leads to a few interdependent questions:
(1) Which router was used?
(2) If the timing-driven router was used, is it justifiable to report W_min values? Should the routability-driven router (enhanced as needed) have been used instead?
(3) If the routability-driven router was used, how were the cost functions enhanced?

==

Minor points:

Section I, third paragraph
- Don’t use [3] as a word. Either say “Ref. [3]” or “Alexander et al. [3]”
- et al should be “et al.” (there is a period after al, but not after et)

Regarding the placer timing cost described in Section IV.A.1), does the delay term include the multiplexers or just the bumps and the wires through the interposer?

Regarding Equation (3), can you give a better explanation of the W_chanx and W_chany terms? I didn’t completely understand them from the short 1-sentence description.

Below Equation (6), I don’t understand the ratio_of_wires_cut metric, especially vis-a-vis Figure 6.

In Figure 6, the green box and label (a) are nearly impossible to see given the light blue background with white spacing. Please switch to a darker shade of green.

In Equation (1), does the ExpectInterposeHops include the delay of the muxes?

Fig. 23 is somehow redundant, as it is simply a different way to display data that was reported earlier.

### 2015-06-24 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Reviewer 1 actions.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4648 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
-Associate Editor
-Comments to the Author:
-All the reviewers think this paper tackles a problem of practical interests, and the proposed solution/results are competent.
-Some major concerns raised by the reviewer is that the authors need to clearly clarify the details of hypergraph partitioning.
-Other detailed suggestions on the writing, implementations, and etc, please refer to the reviewers' comments.
-
-Reviewer's Comments
-
 Reviewer: 1
-
-SPECIFIC FEEDBACK TO AUTHORS
-This paper describes a method to model an interposer-based FPGA, CAD enhancements for an interposer-based architecture, and a detailed architecture study.
-The paper is solid architecture work, and presents many detailed and useful results.
-That being said, the novelty is a bit low --  the techniques to modify the CAD flow are fairly straightforward, and the results are not all that surprising.
-Still, I think the research question is important enough that readers will find useful data here.
-
-The authors have made significant changes to their previously published conference paper including the addition of a partitioning-based flow; in my opinion, this meets the 30% new material requirement.
 
@@
 Doing so would help the reader understand the novelty of the hypergraph->graph methodology and results.
+
+[Javeed to add survey paper reference and discussion on why we use a non-dummy node star topology]
 
@@
 
+[Javeed add table on number of cut crossings on critical paths vs.
+total connections on critical paths.
+Also some discussion on timing-driven partitioning -- our star model should help timing, and a timing-driven partitioner might help further but will the challenge of estimating critical paths accurately is hard.
+Could reference Mike Hutton paper on APEX 20k timing-driven partitioning.
+
+Ref is:  Hutton, M., Adibsamii, K. and Leaver, A. “Adaptive Delay Estimation for Partitioning-Driven PLD Placement”, IEEE Trans. VLSI, 11:1, pp. 60-63, 2003.
+
+My summary of the paper is: 
+"A commercial recursive partitioning placement algorithm for the Altera Apex 20K family is described in [92].
+Apex has a hierarchical routing architecture, making it well suited to partitioning-based placement.
+Recursive partitioning is conducted along the natural cut-lines formed by the various hierarchy levels of the routing architecture, as shown in Figure 13.
+Notice that the sequence of partitions in this algorithm is significantly different than that of ALTOR, showing the large impact an FPGA’s routing architecture has on placement algorithms.
+This algorithm is made timing-driven by weighting connections with low slack highly during each partitioning phase to encourage partitioning solutions in which such connections can be routed using only fast, lower-hierarchy-level routing.
+To improve the prediction of the critical path, the delay estimate for each connection is a function both of the known number of hierarchy boundaries the net must traverse due to partitionings at the higher levels of the routing hierarchy, and statistical estimates of how many hierarchy boundaries the connection will cross at future partitioning steps.
+"]
+
 A key result that would help the reader interpret the results would be a table that shows, for each benchmark circuit, the number of signals that cross the partition (independent of the place and route).
@@
 
+[Javeed to get table and add to paper if feasible.  Try to add in same table as critical path one.  Best flow, 2 partitions, cut net count vs. benchmark and cut net per channel vs. benchmark.]
+
 In Figure 7, it was not clear to me why, for 30% of the wires cut, the placer optimizations do worse than the placer without optimizations.  Is this just experimental noise?
+
+[ This is due to experimental noise; the placer optimizations simply appear to be unnecessary for a small % wires cut so we get small random variations of the same magnitude as those due to changing the placement random seed.]
 
@@
 
+[ No, we did not adjust the range windows.
+As the range window initially covers the entire FPGA and gradually shrinks to a box that allows moves only to adjacent blocks of the compatible type, VPR's natural range limit still permits full flexibility in moving across the interposers, even at the end of the anneal.
+We agree it is a possible direction for future work to see if altering the move generation functions could save CPU time.
+]
+
 Section III-E contains a lot of information about the detailed implementation of the model within VPR.
 While this is useful to someone who wants to replicate the results, if it becomes necessary to cut content to address reviewer comments, some of this detail could be omitted without losing much from the paper (the fact that an extra RR_Graph node is created for each crossing is likely not as interesting to most readers, compared to the other experimental results).
+
+[ Ehsan: after we see how long new material makes the paper Ehsan could shorten this section.]
 
@@
 If there is something else to conclude from this table, perhaps it could be highlighted.
+
+[ Ehsan: could shorten if we run out of space. ]
 

### 2015-06-24 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Review 2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4649 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-SPECIFIC FEEDBACK TO AUTHORS
-This is a well-written paper. I think, this will allow researchers to perform research on multi-die devices.
-
-I have few simple comments:
-
 1.
@@
 
+[ The device floorplan could be modified in VPR and the placement and routing engine would automatically adapt to generate legal solutions and optimize within those constraints; this would not require any code changes beyond the architecture file description.
+The partitioner would need a modified balance criterion where it targeted not resource-balanced partitions, but instead forbade partitions that deviated too much from the resources available in each portion (die) of the system.
+This would require source code changes to metis.]
+
 2.
@@
 
+[ We rely on VPR's annealer to perform this minimization.
+Many random moves are generated, including those of blocks connected by nets that span the interposer.
+Our modified cost functions make it visible to VPR that moves which reduce cut size, or produce a placement that has a higher potential to later be perturbed to reduce cut size, are preferable.
+]
+
 3. In section V B (partitioning Tool), you assumed 0.05 (5%) in line 30. What is the rationale behind choosing that number? What happens if we choose some other number like 10% or 15%?
+
+[ We assumed a 5% balance constraint to give the partitioner some flexibility to minimize cut size, but we are forcing the partitions to be within 5% of balanced so that we can implement designs in fairly full FPGAs (e.g.
+90% logic utilization).
+If we allowed a very loose balance constraint of 10%, then a circuit could be split into 60% and 40% portions, and the interposer system would need to provide enough circuitry in one partition to support 60% of the design logic.
+This implies the chip would need 120% of the logic (or DSP, or RAM if they are the limiting resource) required by the circuit.
+This would limit logic utilization to 83% of the chip (1/1.2), but many FPGA designs go beyond that level of utilization.]
 

### 2015-06-24 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Reviewer 2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4650 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-[ We assumed a 5% balance constraint to give the partitioner some flexibility to minimize cut size, but we are forcing the partitions to be within 5% of balanced so that we can implement designs in fairly full FPGAs (e.g.
+[ We assumed a 5% balance constraint to give the partitioner some flexibility to minimize cut size, while still forcing the partitions to be within 5% of balanced so that we can implement designs in fairly full FPGAs (e.g.
 90% logic utilization).
-If we allowed a very loose balance constraint of 10%, then a circuit could be split into 60% and 40% portions, and the interposer system would need to provide enough circuitry in one partition to support 60% of the design logic.
-This implies the chip would need 120% of the logic (or DSP, or RAM if they are the limiting resource) required by the circuit.
-This would limit logic utilization to 83% of the chip (1/1.2), but many FPGA designs go beyond that level of utilization.]
+If we allowed a more relaxed balance constraint of 10%, then a circuit could be split into 60% and 40% portions, and a 2-die interposer system would need to provide enough circuitry in one partition to support 60% of the design logic.
+This implies the interposer system would need 120% (2 x 60%) of the logic (or DSP, or RAM if they are the limiting resource) required by the circuit.
+This would limit logic utilization to 83% of the system (1/1.2), but many FPGA designs go beyond that level of utilization.]
+
+[ Javeed: try a balance constraint that reflects the real limits in each die. ]
 
@@
 
+[ DSP and block RAM primitives are treated the same way as LUTs and FFs by the partitioner but it does enforce a balance criterion on each resource. ]
+
 5. In your work, maximum how many partitions are allowed? Was there any situation where partitioned could not partition the design?
+
+[ We experiment with 2-die and 4-die systems, which require 2 and 4 partitions respectively. Partitioning was successful on all designs.]
 
@@
 
+[ Two dimensional arrays of dice are interesting, but we have not studied such systems and they will require some further changes to VPR's architecture generation phase and model of the system.
+Hence we consider this interesting future work, but beyond the scope of this paper.]
+
 7. What is the frequency loss between monolithic design and multi-die design?
+
+[ This is discussed in Section VI.C.
+The frequency loss is highly dependent on the interposer delay, but mostly independent of the % wires cut.
+For a 2-die system, a 0.5 ns interposer delay reduces frequency by 4% vs.
+a monolithic FPGA, while a 1 ns interposer delay reduces frequency by 11%.]
 
@@
 
+[ Ehsan: to run 3 experiments.  conventional VPR, modified VPR on 2-die system with 0% wires cut and 0 ns delay, and modified VPR on 2-die system with 80% wires cut and 1 ns delay.]
+  Javeed:  Send CPU time for 2 partition and 4 partition case.
+
 9. Any analyze on the power consumption for multi-die designs? Any comparison of power consumption with respect to that of a monolithic design?
 
+[ This is interesting future work, but we have not investigated the power impact of an interposer.
+Our intuition is that the power increase will be small as relatively few signals cross the interposer compared to the within-die signals throughout the system, but we have not quantiied this effect.
+]
+
 In spite of all these feedbacks, I like your paper. Good job
+
+[ Thank you for your comments!]
 

### 2015-06-24 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Reviewer 3
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4651 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-SPECIFIC FEEDBACK TO AUTHORS
-Compared to the conference version of the submission, this is a more thorough study of the performance of interposer-based FPGAs when key parameters are varied.
-To carry out the study, it adapted and twisted the VPR flow for conventional homogeneous FPGAs a bit.
-Though it may not be the ideal CAD flow for interposer-based FPGAs, some valuable insights can still be gained from the experiments.
-
 In section IV.A, you should mention that VPR's placer is simulated annealing-based.
 In particular, the reason you cited in the last paragraph of IV.A is based on the assumption that you are using an iterative improvement-based placer (and not, say, an analytical placer).
+
+[Ehsan: add mention of annealing].
 
@@
 
+[Javeed: make explicit and correct text if it says we make all hyperedges the same weight with 1/n (they will vary from 1/2 to 1 in total hyperedge weight).]
+
 p.9, middle of left column said "the star topology clearly differentiates sources from sinks and this appears to give the partitioner an anchor point that pulls all hyperedge fanouts toward the source", you should mention that Metis is an iterative improvement-based partitioner which is related to your reasoning.
+
+[ Javeed: add.]
 
@@
 Similarly, your star 1/n model assigns a total weight of (n-1)/n (not 1) to each net.
+
+[Javeed: fix.]
 
@@
 
+[Ehsan: take out.]
+
 - The text never refers to Fig.3.
+[Ehsan: add reference]
 
 - There should not be indentation following eqts. (1), (2), (3), (5), (6).
+[Ehsan: fix. Shouldn't start a new paragraph after.]
 
 - Sec. V.E., "It is slightly worse ... even for ..." -> "But it is slightly worse ... for ..."
+[Ehsan: fix.]
 

### 2015-06-24 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Reviewer 4
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4652 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
+[Javeed: mention hMetis is not open source (emailed person), give results on hmetis vs.
+metis cut size (with and without balance), and mention the star model should help timing, as timing is a connection / edge, phenomenon, not a net / hyperedge phenomenon.]
+
 (2) I disagree with the case for balance constraints
@@
 
+[Vaughn: write up utilization argument.
+ Javeed: do the tuned balance / limit ].
+
 ==
@@
 Since I’ve already argued that the paper is somewhat lightweight in terms of technical contribution, perhaps doing a proper investigation of this issue and adding the necessary algorithmic support could bring greater novelty to this paper, rather than leaving it open to future work.
+
+[Vaughn: write beyond the scope].
 
@@
 
+[Vaughn: answer - always use timing driven].
+
 ==
@@
 
+[Vaughn: fix]
+
 Regarding the placer timing cost described in Section IV.A.1), does the delay term include the multiplexers or just the bumps and the wires through the interposer?
+[Ehsan: modify text.]
 
 Regarding Equation (3), can you give a better explanation of the W_chanx and W_chany terms? I didn’t completely understand them from the short 1-sentence description.
+[Ehsan: change.]
 
 Below Equation (6), I don’t understand the ratio_of_wires_cut metric, especially vis-a-vis Figure 6.
+[Use % wires cut instead].
 
 In Figure 6, the green box and label (a) are nearly impossible to see given the light blue background with white spacing. Please switch to a darker shade of green.
+[Ehsan: fix].
 
 In Equation (1), does the ExpectInterposeHops include the delay of the muxes?
+[Ehsan: explain / fix].
 
 Fig. 23 is somehow redundant, as it is simply a different way to display data that was reported earlier.
+[Vaughn: write up defense of figure].

### 2015-06-24 vtr_flow/scripts/benchtracker/flask_cors/__init__.py
Sync - benchtracker now includes flask cors for buildbot machine (since it can't install it)
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/trunk@4653 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@ -1,0 +1,15 @@
+# -*- coding: utf-8 -*-
+"""
+    flask_cors
+    ~~~~
+    Flask-CORS is a simple extension to Flask allowing you to support cross
+    origin resource sharing (CORS) using a simple decorator.
+
+    :copyright: (c) 2014 by Cory Dolphin.
+    :license: MIT, see LICENSE for more details.
+"""
+from .decorator import cross_origin
+from .extension import CORS
+from .version import __version__
+
+__all__ = ['CORS', 'cross_origin']

### 2015-06-24 vtr_flow/scripts/benchtracker/flask_cors/core.py

@@ -1,0 +1,372 @@
+# -*- coding: utf-8 -*-
+"""
+    flask_cors
+    ~~~~
+    Flask-CORS is a simple extension to Flask allowing you to support cross
+    origin resource sharing (CORS) using a simple decorator.
+
+    :copyright: (c) 2014 by Cory Dolphin.
+    :license: MIT, see LICENSE for more details.
+"""
+import re
+import logging
+import collections
+from datetime import timedelta
+from six import string_types
+from flask import request, current_app
+try:
+    from flask import _app_ctx_stack as stack
+except ImportError:
+    from flask import _request_ctx_stack as stack
+from werkzeug.datastructures import MultiDict
+
+# Compatibility with old Pythons!
+if not hasattr(logging, 'NullHandler'):
+    class NullHandler(logging.Handler):
+        def emit(self, record):
+            pass
+    logging.NullHandler = NullHandler
+
+# Response Headers
+ACL_ORIGIN = 'Access-Control-Allow-Origin'
+ACL_METHODS = 'Access-Control-Allow-Methods'
+ACL_ALLOW_HEADERS = 'Access-Control-Allow-Headers'
+ACL_EXPOSE_HEADERS = 'Access-Control-Expose-Headers'
+ACL_CREDENTIALS = 'Access-Control-Allow-Credentials'
+ACL_MAX_AGE = 'Access-Control-Max-Age'
+
+# Request Header
+ACL_REQUEST_METHOD = 'Access-Control-Request-Method'
+ACL_REQUEST_HEADERS = 'Access-Control-Request-Headers'
+
+ALL_METHODS = ['GET', 'HEAD', 'POST', 'OPTIONS', 'PUT', 'PATCH', 'DELETE']
+CONFIG_OPTIONS = ['CORS_ORIGINS', 'CORS_METHODS', 'CORS_ALLOW_HEADERS',
+                  'CORS_EXPOSE_HEADERS', 'CORS_SUPPORTS_CREDENTIALS',
+                  'CORS_MAX_AGE', 'CORS_SEND_WILDCARD',
+                  'CORS_AUTOMATIC_OPTIONS', 'CORS_VARY_HEADER',
+                  'CORS_RESOURCES', 'CORS_INTERCEPT_EXCEPTIONS']
+
+# Attribute added to request object by decorator to indicate that CORS
+# was evaluated, in case the decorator and extension are both applied
+# to a view.
[…]

### 2015-06-24 vtr_flow/scripts/benchtracker/flask_cors/decorator.py

@@ -1,0 +1,127 @@
+# -*- coding: utf-8 -*-
+"""
+    flask_cors
+    ~~~~
+    Flask-CORS is a simple extension to Flask allowing you to support cross
+    origin resource sharing (CORS) using a simple decorator.
+
+    :copyright: (c) 2014 by Cory Dolphin.
+    :license: MIT, see LICENSE for more details.
+"""
+from functools import update_wrapper
+from flask import make_response, request, current_app
+from .core import *
+
+
+def cross_origin(*args, **kwargs):
+    '''
+    This function is the decorator which is used to wrap a Flask route with.
+    In the simplest case, simply use the default parameters to allow all
+    origins in what is the most permissive configuration. If this method
+    modifies state or performs authentication which may be brute-forced, you
+    should add some degree of protection, such as Cross Site Forgery
+    Request protection.
+
+
+    :param origins: The origin, or list of origins to allow requests from.
+        The origin(s) may be regular expressions, case-sensitive strings,
+        or else an asterisk
+
+        Default : '*'
+    :type origins: list, string or regex
+
+    :param methods: The method or list of methods which the allowed origins
+        are allowed to access for non-simple requests.
+
+        Default : [GET, HEAD, POST, OPTIONS, PUT, PATCH, DELETE]
+    :type methods: list or string
+
+    :param expose_headers: The header or list which are safe to expose to the
+        API of a CORS API specification.
+
+        Default : None
+    :type expose_headers: list or string
+
+    :param allow_headers: The header or list of header field names which can be
+        used when this resource is accessed by allowed origins. The header(s)
+        may be regular expressions, case-sensitive strings, or else an asterisk.
+
+        Default : '*', allow all headers
+    :type allow_headers: list, string or regex
+
+    :param supports_credentials: Allows users to make authenticated requests.
+        If true, injects the `Access-Control-Allow-Credentials` header in
+        responses. This allows cookies to be submitted across domains.
+
+        :note: This option cannot be used in conjuction with a '*' origin
+
+        Default : False
+    :type supports_credentials: bool
+
+    :param max_age: The maximum time for which this CORS request maybe cached.
+        This value is set as the `Access-Control-Max-Age` header.
+
+        Default : None
+    :type max_age: timedelta, integer, string or None
+
+    :param send_wildcard: If True, and the origins parameter is `*`, a
+        wildcard `Access-Control-Allow-Origin` header is sent, rather than
+        the request's `Origin` header.
+
[…]

### 2015-06-24 vtr_flow/scripts/benchtracker/flask_cors/extension.py

@@ -1,0 +1,116 @@
+# -*- coding: utf-8 -*-
+"""
+    flask_cors
+    ~~~~
+    Flask-CORS is a simple extension to Flask allowing you to support cross
+    origin resource sharing (CORS) using a simple decorator.
+
+    :copyright: (c) 2014 by Cory Dolphin.
+    :license: MIT, see LICENSE for more details.
+"""
+from flask import request
+from .core import *
+
+
+class CORS(object):
+    '''
+        Initializes Cross Origin Resource sharing for the application. The
+        arguments are identical to :py:func:`cross_origin`, with the
+        addition of a `resources` parameter. The resources parameter
+        defines a series of regular expressions for resource paths to match
+        and optionally, the associated options
+        to be applied to the particular resource. These options are
+        identical to the arguments to :py:func:`cross_origin`.
+
+        The settings for CORS are determined in the following order:
+            Resource level settings (e.g when passed as a dictionary)
+            Keyword argument settings
+            App level configuration settings (e.g. CORS_*)
+            Default settings
+
+        Note: as it is possible for multiple regular expressions to match a
+        resource path, the regular expressions are first sorted by length,
+        from longest to shortest, in order to attempt to match the most
+        specific regular expression. This allows the definition of a
+        number of specific resource options, with a wildcard fallback
+        for all other resources.
+
+        :param resources: the series of regular expression and (optionally)
+        associated CORS options to be applied to the given resource path.
+
+        If the argument is a dictionary, it is expected to be of the form:
+        regexp : dict_of_options
+
+        If the argument is a list, it is expected to be a list of regular
+        expressions, for which the app-wide configured options are applied.
+
+        If the argument is a string, it is expected to be a regular
+        expression for which the app-wide configured options are applied.
+
+        Default :'*'
+
+        :type resources: dict, iterable or string
+
+    '''
+
+    def __init__(self, app=None, **kwargs):
+        self._options = kwargs
+        if app is not None:
+            self.init_app(app, **kwargs)
+
+    def init_app(self, app, **kwargs):
+        # The resources and options may be specified in the App Config, the CORS constructor
+        # or the kwargs to the call to init_app.
+        options = get_cors_options(app, self._options, kwargs)
+
+        # Flatten our resources into a list of the form
+        # (pattern_or_regexp, dictionary_of_options)
+        resources = parse_resources(options.get('resources'))
+
+        # Compute the options for each resource by combining the options from
+        # the app's configuration, the constructor, the kwargs to init_app, and
[…]

### 2015-06-24 vtr_flow/scripts/benchtracker/server_db.py

@@ -17,7 +17,7 @@ except ImportError:
     parentdir = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
     os.sys.path.insert(0, parentdir)
 
-    from flask.ext.cors import CORS
+    from flask_cors import CORS
 
 
 

### 2015-06-24 vtr_flow/scripts/benchtracker/static/plotter.js

@@ -494,6 +494,15 @@ function simple_plot(params, series, overlay_list, xNM, t, titleMode) {
     // ..............
     // setup axis
     var x = null;
+    if (range[t]['x'][0] == range[t]['x'][1]) {
+        range[t]['x'][0] -= 0.5*range[t]['x'][0];
+        range[t]['x'][1] += 0.5*range[t]['x'][1];
+    }
+    if (range[t]['y'][0] == range[t]['y'][1]) {
+        range[t]['y'][0] -= 0.5*range[t]['y'][0];
+        range[t]['y'][1] += 0.5*range[t]['y'][1];
+    }
+
     if (xNM.values.length == 0) {
         x = d3.scale.linear()
               .domain([range[t]['x'][0] - 0.1*(range[t]['x'][1]-range[t]['x'][0]), range[t]['x'][1] + 0.1*(range[t]['x'][1]-range[t]['x'][0])])
@@ -512,11 +521,6 @@ function simple_plot(params, series, overlay_list, xNM, t, titleMode) {
         .scale(x)
         .orient("bottom")
         .tickSize(-height);
-    /*
-    if (xNM.values.length != 0) {
-        xAxis.tickValues([100,120,140,160]);
-    }
-    */
     var yAxis = d3.svg.axis()
         .scale(y)
         .orient("left")
@@ -541,7 +545,7 @@ function simple_plot(params, series, overlay_list, xNM, t, titleMode) {
     if (xNM.values.length == 0) {
         svg.call(zoom);
     }
-    var canvas = svg.append("rect")
+    var canvas = svg.append("rect").attr('fill', '#f0f0f0')
         .attr("width", width)
         .attr("height", height);
 
@@ -569,13 +573,13 @@ function simple_plot(params, series, overlay_list, xNM, t, titleMode) {
                     .x(function(d) {return x(d['x']);})
                     .y(function(d) {return y(d['y']);});
     var clip = svg.append('svg:clipPath').attr('id', 'clip')
-                  .append('svg:rect').attr('x', 0).attr('y', 0)
+                  .append('svg:rect').attr('fill', '#f0f0f0').attr('x', 0).attr('y', 0)
                   .attr('width', width).attr('height', height);
     var color = d3.scale.category10();
     
     for (var i in lineInfo){
         var svgg = svg.append('g').attr('clip-path', 'url(#clip)');
-        svgg.append('svg:path')
+        svgg.append('svg:path').attr('fill', 'none')
             .datum(lineInfo[i]['values'])
[…]

### 2015-07-01 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Ehsan: Making updates based on reviewer's comments. Done with almost all of edits that were assigned to me (see response_to_reviewers.txt). (i just have to summarize a section in case we run out of space). I also need to get runtime measurements to present to one of the reviewers, but that's not going to be added to the paper.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4670 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-[Ehsan: add mention of annealing].
+[Ehsan: add mention of annealing][DONE]
 
@@
 
-[Ehsan: take out.]
+[Ehsan: take out.][DONE]
 
 - The text never refers to Fig.3.
-[Ehsan: add reference]
+[Ehsan: add reference][DONE]
 
 - There should not be indentation following eqts. (1), (2), (3), (5), (6).
-[Ehsan: fix. Shouldn't start a new paragraph after.]
+[Ehsan: fix. Shouldn't start a new paragraph after.][DONE]
 
 - Sec. V.E., "It is slightly worse ... even for ..." -> "But it is slightly worse ... for ..."
-[Ehsan: fix.]
+[Ehsan: fix.][DONE]
 
@@
 Regarding the placer timing cost described in Section IV.A.1), does the delay term include the multiplexers or just the bumps and the wires through the interposer?
-[Ehsan: modify text.]
+[Ehsan: modify text.][DONE]
 
 Regarding Equation (3), can you give a better explanation of the W_chanx and W_chany terms? I didn’t completely understand them from the short 1-sentence description.
-[Ehsan: change.]
+[Ehsan: change.][DONE]
 
 Below Equation (6), I don’t understand the ratio_of_wires_cut metric, especially vis-a-vis Figure 6.
-[Use % wires cut instead].
+[Use % wires cut instead][DONE]
 
 In Figure 6, the green box and label (a) are nearly impossible to see given the light blue background with white spacing. Please switch to a darker shade of green.
-[Ehsan: fix].
+[Ehsan: fix][DONE]
 
 In Equation (1), does the ExpectInterposeHops include the delay of the muxes?
-[Ehsan: explain / fix].
+[Ehsan: explain / fix][DONE: ExpectInterposeHops is not the delay term.
+It's the number of expected interposer hops which is multiplied by the total delay of each hop.
+I changed the variable name to: NumExpectedInterpHops]
 

### 2015-07-01 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Ehsan: Making updates based on reviewer's comments. Done with almost all of edits that were assigned to me (see response_to_reviewers.txt). (i just have to summarize a section in case we run out of space). I also need to get runtime measurements to present to one of the reviewers, but that's not going to be added to the paper.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4670 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 \usepackage{hhline}
+\usepackage{color,soul}
 
@@
 The Xilinx 2.5D FPGAs from the Virtex-7 and Virtex Ultrascale families are currently the only commercially available silicon interposer-based FPGAs~\cite{xilinxWP}.
-As described in Section~\ref{archSection} we are studying the impact of several key interposer parameters on the performance of the multi-die system, including (i) the percentage of the wiring normally present between rows of the FPGA which is cut when crossing between dice, and (ii) the extra delay (vs.
+\hl{We are studying} the impact of several key interposer parameters on the performance of the multi-die system, including (i) the percentage of the wiring normally present between rows of the FPGA which is cut when crossing between dice, and (ii) the extra delay (vs.
 a normal connection between adjacent rows) added when one must traverse the interposer.
@@
 Practical interposers, however, can only accommodate a fraction of all vertical routing wires.
-When \textit{fanin transfer} option is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect adjacent ``dangling" vertical routing wires to the multiplexer.
+\hl{As shown in Figure}~\ref{fig:architecture_params}, when \textit{fanin transfer} option is \textit{on}, we place a multiplexer at the input of each available interposer wire and connect adjacent ``dangling" vertical routing wires to the multiplexer.
 This enables additional flexibility in routing by ensuring that \textit{every} vertical routing wire could \textit{potentially} drive an interposer wire through a multiplexer.
@@
 \item Fan-out Transfer for Interposer Wires\\
-When \textit{fanout transfer} option is \textit{on}, we allow an interposer wire to drive not only its own corresponding vertical wire at the other side of the cut, but also other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
+\hl{As shown in Figure}~\ref{fig:architecture_params}, when \textit{fanout transfer} option is \textit{on}, we allow an interposer wire to drive not only its own corresponding vertical wire at the other side of the cut, but also other ``dangling" vertical routing wires that are not driven by anything because their corresponding interposer wires are cut.
 
@@
 Placement is crucial in mitigating the impact of the reduced wiring and increased delay when crossing between dice; a good placement should minimize the number of signals crossing between dice, particularly time-critical ones.
-VPR uses 2 different costs as the metrics for its placer algorithm: the timing cost and the bounding box (wiring) cost.
+VPR uses 2 different costs as the metrics for its \hl{simulated annealing} placer algorithm: the timing cost and the bounding box (wiring) cost.
 We modified the placer's bounding box and timing costs to account for the increased latency to cross the interposer and for the reduced wire capacity close to the cutlines, as the wire availability becomes more sparse.
@@
 
-where \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
+\noindent{\hl{where}} \textit{ij} denotes a connection from block i to block j that exists in the circuit netlist.
 
@@
 
-where $N_{crossed}(i,j)$ is the minimum number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the timing penalty of crossing the interposer.
+\noindent{\hl{where}} $N_{crossed}(i,j)$ is the minimum number of times this path has to cross the interposer to go from $(x_i,y_i)$ to $(x_j,y_j)$ and $delay\_increase$ is the \hl{total timing penalty of crossing the interposer, including delays through muxes, bumps, and interposer wires.}
 
@@
 
-where $bb_x(n)$ and $bb_y(n)$ are the dimensions of the bounding box of net \textit{n} in the x and y directions, respectively.
-$W_{chanx}$ and $W_{chany}$ are the average x-directed and y-directed channel widths over this bounding box.
+\noindent{\hl{where}} $bb_x(n)$ and $bb_y(n)$ are the dimensions of the bounding box of net \textit{n} in the x and y directions, respectively.
+$W_{chanx}$ and $W_{chany}$ are the average \hl{horizontal and vertical channel widths} over this bounding box.
 Finally, $q(n)$ is a function obtained from \cite{icann} which models the fact that bounding boxes underpredict the required routing for high fanout nets.
@@
 \label{eq:cprime}
-C' = \frac{ratio\_wires\_cut}{W_{chany}}
+C' = \frac{\%~wires~cut}{W_{chany}}
 \end{equation}
@@
 
-\textit{bbHeight(n)} is the height of the bounding box of the net \textit{n} and $ratio\_wires\_cut$ is the ratio of wires cut at the cutline.
+\noindent{\hl{and \textit{bbHeight(n)}}} is the height of the bounding box of the net \textit{n}.
 This formulation of $C'$ ensures that the new cost term is of roughly the same magnitude as $Cost_{W\_orig}$ in (\ref{eq:bbcost}) and that it is weighted more heavily for interposer architectures in which the wiring between dice is more scarce.
@@
 \label{eq:tdelay}
-T^\prime_{del} = T_{del} + T_{interpose} \times ExpectInterposeHops
+T^\prime_{del} = T_{del} + T_{interposer} \times NumExpectedInterpHops
 \end{multline}

### 2015-07-01 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Shortened a few equation variables and fixed a few confusing sentences.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4610 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Ehsan: Making updates based on reviewer's comments. Done with almost all of edits that were assigned to me (see response_to_reviewers.txt). (i just have to summarize a section in case we run out of space). I also need to get runtime measurements to present to one of the reviewers, but that's not going to be added to the paper.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4

@@
 
-The current version of hMetis does not support heterogeneous balance constraints, though Metis does.
-Metis was used in our CAD flow.
+The current version of hMetis does not support heterogeneous balance constraints, though Metis does, so we use Metis in our CAD flow.
 However, as the circuit netlist is naturally represented as a hypergraph, it needs to be transformed to a graph before Metis can process it.
@@
 As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of~\cite{interposer2014} for FPGAs where more than half the wires are cut at interposer boundaries.
-It is slightly worse than the traditional FPGA CAD flow even for an architecture with 0\% of the wires cut, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.
+\hl{But it is slightly worse than the traditional FPGA CAD flow for an architecture with 0\% of the wires cut}, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.
 

### 2015-07-05 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Vaughn's updates: some text changes and responses to reviewer updates for all Vaughn's actions.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4671 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
+We would like to thank the reviewers for their detailed and constructive feedback. We believe we have addressed
+all the reviewer comments (or explained below why we did not implement a suggestion) and believe this
+has made a stronger paper. Significant text changes are highlighted in yellow in the paper, and 
+we summarize the changes made to address each comment or answer any detailed questions in the response
+below.
+
 Reviewer: 1
@@
 
-[ We assumed a 5% balance constraint to give the partitioner some flexibility to minimize cut size, while still forcing the partitions to be within 5% of balanced so that we can implement designs in fairly full FPGAs (e.g.
+[ We have added a discussion of why we choose a relatively tight balance constraint of 5% to Section V.B.  
+We assumed a 5% balance constraint to give the partitioner some flexibility to minimize cut size, while still forcing the partitions to be within 5% of balanced so that we can implement designs in fairly full FPGAs (e.g.
 90% logic utilization).
@@
 
-SPECIFIC FEEDBACK TO AUTHORS
-This paper describes a set of extensions to the VPR tool to accommodate multi-FPGA platforms that use silicon interposers.
-The paper includes both architectural enhancements and modifications to the VPR CAD flow.
-
-The main challenge is that silicon interposers have increased delay and reduced connectivity between dice.
-Rather than cutting wires at the border of each FPGA, the authors propose to add multiplexers at the interface (Fig.
-3).
-This entails some minor modifications to VPR’s routing resource graph (RRG) at the cut-points.
-In terms of CAD support, the primary contribution of the paper are new cost functions for the placer and router, and a pre-placement partitioning phase using METIS.
-
-I find the paper to be competent and reasonable, but at the same time, a collection of very small incremental ideas, and lacking in terms of a significant contribution to the field.
-(Other than a VPR update, which is always welcome because VPR is widely used in FPGA architecture and CAD research).
-
-==
-
-I do have two fairly significant technical quibbles with the paper; although they are somehow interdependent, they can each be taken at face value.
-
 (1) The case for METIS over hMETIS is weak
@@
 
-[Vaughn: write up utilization argument.
+[ We have added a discussion of why we use a balance constraint to Section V.B.
+The basic reason is that interposer-based FPGAs are intended to expand FPGA capacity beyond what is possible, or what yields at a reasonable/economic rate, on a single die.
+To make use of this increased capacity, however, one must place a significant fraction of the circuit in each die.
+Stated another way, we wish to examine the routability of circuits at typical FPGA customer utilization levels, which are usually in the 90% range of the limiting resource (usually logic, RAM, or multipliers).
+If one uses 90% of the resources in a two-die interposer system, for example, then a 0.05 unbalance constraint allows a partitioning with 45% of the design in one die and 55% in the other.
+0.55 * 0.9 = 0.5, meaning that the more full partition / die uses 50% of the two-dice systems resources, which is 100% of the resources available in that die.
+A larger unbalance between partitions would mean that a design that used 90% of the interposer system's resources would have a partition that is infeasible on one of the two dice.
+
+You are correct that if we targeted lower utilization rates for the interposer systems we could relax the balance constraint, and this might help routability.
+However, we do not feel this is the best point at which to evaluate interposer routability.
+Designers are accustomed to using most of an FPGA's logic, RAM, etc.
+capacity and will have the same expectation for interposer-based FPGAs.
+Commercial monolithic FPGAs are architected to have a high probability of successful routing at utilizations of 90% and even higher.
+Partially this is because the routing is cheaper and the function blocks more expensive than was assumed in DeHon's 1999 paper and partly this is because an FPGA is usually selected and integrated into a board before the FPGA design is complete.
+While designers can estimate logic, RAM and multiplier utlization with reasonable accuracy even before a design is complete, routability estimation is very hard, so it would be difficult to choose a chip before a design was 100% complete.]
+
  Javeed: do the tuned balance / limit ].
@@
 
-[Vaughn: write beyond the scope].
+[Thank you for your comment.
+We agree that this is important future work but feel it is beyond the scope of the current paper as we are already at the limit of 14 pages, and the best way to incorporate such "packing aware partitioning hints" is not clear.
+We have explored one method of making the partitioner packing-aware, which is to run partitioning after packing; however, this is an inferior CAD flow as the packer benefits more from the partitioning constraints than the partitioner benefits from the packing constraints.
+We have added text to Section V.F to higlight this.
+].
 
@@
 
-[Vaughn: answer - always use timing driven].
+[We used the VPR timing-driven router for all results in this paper, and believe it is the correct choice as it produces a better combination of routability and timing than VPR's routability-driven router, and its algorithm is more scalable to large circuits.
+
+  The timing-driven router in VPR optimizes for both timing and routability (timing is optimized as much as possible, but the cost of congestion grows to eventually swamp the timing terms in congested parts of the design if necessary to achieve routability).
+Hence its routability is good; in "Architecture and CAD for Deep Submicron FPGAs" by Betz, Marquardt and Rose, for example, the timing driven router was found to produce circuits with a xx% higher Wmin, but that were yy% faster.
+Hence it achieves a superior combination of routability and timing than the routability-driven router, although you are correct that the routability-driven router produces slightly lower absolute channel widths.
+A further problem with the routability-driven router in VPR is that its algorithm does not scale as well as the timing-driven router, and hence its runtime will be much higher and likely impractical on the larger circuits we use.
+
+For these reasons, the timing-driven router in VPR is the default and is normally used to find both the minimum channel width and the achievable timing of circuits; in the recent VTR 7.0 paper for example, only results for the timing-driven VPR router are given as the routability-driven router is not widely used.]
 
@@
 
-[Vaughn: fix]
+[Thanks you; we have fixed this text.]
 
@@
 
-Regarding Equation (3), can you give a better explanation of the W_chanx and W_chany terms? I didn’t completely understand them from the short 1-sentence description.
+Regarding Equation (3), can you give a better explanation of t  he W_chanx and W_chany terms? I didn’t completely understand them from the short 1-sentence description.
 [Ehsan: change.][DONE]
@@
 Fig. 23 is somehow redundant, as it is simply a different way to display data that was reported earlier.
-[Vaughn: write up defense of figure].
+[You are correct that Fig.
+23 is simply a different way to visualize the data of Fig.
+22.
+We prefer to retain the figure however, as it highlights that the interposer wiring capacity becomes the dominant factor in routability when its absolute signal count drops below a certain level.
+Other figures show the within-die channel width required when one cuts a certain percentage of the normal FPGA wiring across the interposer, while this figure presents the same data as within-die channel width required vs.
+the absolute (not relative) interposer wiring capacity.
+We believe this highlights the important fact that the interposer for this FPGA architecture requires about 20 wires per vertical channel at the absolute minimum.]

### 2015-07-05 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Vaughn's updates: some text changes and responses to reviewer updates for all Vaughn's actions.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4671 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-To derive an appropriate balance constraint for our use case, consider that FPGA circuit netlists may contain different block types such as LUTs, registers, RAM slices, and multipliers, among others.
+\hl{We chose an unbalance of 0.05 to allow the partitioner some flexibility, while still ensuring the partitions are sufficiently equal in size to make good use of each die.
+Highly unbalanced partitions would leave some dice nearly empty, defeating the key advantage of interposer-based FPGAs: larger capacity.} FPGA circuit netlists may contain different block types such as LUTs, registers, RAM slices, and multipliers, among others.
 The basic balance constraint of (\ref{eq:balance_constraint}), applied to a netlist with different block types, would ensure that the total block count (over all blocks, regardless of type) in each partition is roughly balanced.

### 2015-07-05 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Added in VPR routability vs. timing-driven router numbers.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4672 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
   The timing-driven router in VPR optimizes for both timing and routability (timing is optimized as much as possible, but the cost of congestion grows to eventually swamp the timing terms in congested parts of the design if necessary to achieve routability).
-Hence its routability is good; in "Architecture and CAD for Deep Submicron FPGAs" by Betz, Marquardt and Rose, for example, the timing driven router was found to produce circuits with a xx% higher Wmin, but that were yy% faster.
+Hence its routability is good; in "Architecture and CAD for Deep Submicron FPGAs" by Betz, Marquardt and Rose, for example, the timing driven router was found to produce circuits with a 6% higher Wmin, but that had an operating frequency 2.58x higher.
 Hence it achieves a superior combination of routability and timing than the routability-driven router, although you are correct that the routability-driven router produces slightly lower absolute channel widths.
 A further problem with the routability-driven router in VPR is that its algorithm does not scale as well as the timing-driven router, and hence its runtime will be much higher and likely impractical on the larger circuits we use.
+On the largest circuit for which results are presented in "Architecture and CAD for Deep-Submicron FPGAs" the VPR routability-driven router requires 14.6x more CPU time than the timing-driven router, and the CPU time gap increases with circuit size.
+Hence on the much larger circuits (>10x larger) we use the CPU time of the routability-driven router would make it very difficult to use.
 

### 2015-07-06 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Added reference on hypergraph to graph transformation. Added section for cut critical connections.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4673 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.
+\hl{Ref.\mbox{\cite{hypergraph_to_graph_survey}}
+considers the addition of ``dummy'' nodes to model a hypergraph as a graph, with the objective of obtaining equivalent edge cut weights after partitioning.
+In our star model, we preferred to designate the net source as the center of the star, rather than introduce a dummy node.
+This choice was motivated by a desire to optimize timing -- the star topology clearly differentiates sources from sinks, and we expect this to guide the partitioner to keep sources and sinks in the same partition.}
+
 We assigned the same edge weight to every edge of the generated subgraph.
@@
 
+\subsection{\hl{Interposer Wire Utilization}}
+\hl{Table \mbox{\ref{table:interposer_wire_utilization}} shows the number of connections on the critical path and how many of those critical connections cross the interposer boundary.
+It also shows the number of cut nets per vertical channel.
+This provides a lower bound on the number of wires required in an interposer channel.
+The data was obtained at 70\% wires cut.}
+\begin{table}[!htbp]
+\centering
+\caption{Interposer wire utilization}
+\begin{tabular}{|l|r|r||r|r|}
+\hline
+ & \multicolumn{2}{c||}{Critical Connections} & \multicolumn{2}{c|}{Cut Nets} \\
+ & Cut & Total & Total & Per Channel \\ \hline
+bgm & 1 & 373 & 622 & 8.89 \\ \hline
+stereovision0 & 3 & 57 & 985 & 25 \\ \hline
+stereovision1 & 1 & 21 & 876 & 21 \\ \hline
+LU8PEEng & 0 & 1661 & 1795 & 20 \\ \hline
+mkDelayWorker32B & 4 & 58 & 771 & 17 \\ \hline
+mcml & 5 & 1443 & 2891 & 28 \\ \hline
+LU32PEEng & 0 & 1664 & 1130 & 5 \\ \hline
+\end{tabular}
+\label{table:interposer_wire_utilization}
+\end{table}
+
+\hl{Though our choice of the star topology benefits timing, our partitioning flow is not timing-driven.
+We would expect a timing-driven partitioner to reduce the number of cut critical connections, but obtaining good critical path estimates at the partitioning stage is nontrivial\mbox{\cite{hutton}}.}
+
 \subsection{\% wires cut}

### 2015-07-06 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt
Minor revisions in response to reviewer comments.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4674 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 
-[Javeed to add survey paper reference and discussion on why we use a non-dummy node star topology]
+[Javeed to add survey paper reference and discussion on why we use a non-dummy node star topology][DONE]
 
@@
 To improve the prediction of the critical path, the delay estimate for each connection is a function both of the known number of hierarchy boundaries the net must traverse due to partitionings at the higher levels of the routing hierarchy, and statistical estimates of how many hierarchy boundaries the connection will cross at future partitioning steps.
-"]
+"] [DONE]
 
@@
 
-[Javeed to get table and add to paper if feasible.  Try to add in same table as critical path one.  Best flow, 2 partitions, cut net count vs. benchmark and cut net per channel vs. benchmark.]
+[Javeed to get table and add to paper if feasible.
+Try to add in same table as critical path one.
+Best flow, 2 partitions, cut net count vs.
+benchmark and cut net per channel vs.
+benchmark.] [DONE, but p&r dependent]
 
@@
 
-[ Javeed: try a balance constraint that reflects the real limits in each die. ]
+[ Javeed: try a balance constraint that reflects the real limits in each die. ] [ DONE? Need to respond.]
 
@@
 
-[Javeed: make explicit and correct text if it says we make all hyperedges the same weight with 1/n (they will vary from 1/2 to 1 in total hyperedge weight).]
+[Javeed: make explicit and correct text if it says we make all hyperedges the same weight with 1/n (they will vary from 1/2 to 1 in total hyperedge weight).] [DONE]
 
@@
 
-[ Javeed: add.]
+[ Javeed: add.] [DONE]
 
@@
 
-[Javeed: fix.]
+[Javeed: fix.] [DONE]
 
@@
 [Javeed: mention hMetis is not open source (emailed person), give results on hmetis vs.
-metis cut size (with and without balance), and mention the star model should help timing, as timing is a connection / edge, phenomenon, not a net / hyperedge phenomenon.]
+metis cut size (with and without balance), and mention the star model should help timing, as timing is a connection / edge, phenomenon, not a net / hyperedge phenomenon.] [DONE]
+The hMetis source code is not freely available.
+We contacted the authors of hMetis, who stated that an hMetis update is in development and will contain support for heterogeneous balance constraints.
+Our experiments showed that hMetis performed 30% better on average when compared to Metis on our benchmark circuits, when both were used with homogeneous balance constraints.
 
@@
 
- Javeed: do the tuned balance / limit ].
+ Javeed: do the tuned balance / limit ]. [DONE?]
 

### 2015-07-06 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Minor revisions in response to reviewer comments.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4674 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

@@
 However, as the circuit netlist is naturally represented as a hypergraph, it needs to be transformed to a graph before Metis can process it.
+\hl{Our experiments showed that hMetis performed 30\% better on average when compared to Metis on our benchmark circuits with homogeneous balance constraints.}
 
@@
 We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}.
-\hl{Ref.\mbox{\cite{hypergraph_to_graph_survey}}
-considers the addition of ``dummy'' nodes to model a hypergraph as a graph, with the objective of obtaining equivalent edge cut weights after partitioning.
+\hl{Ref.\mbox{\cite{hypergraph_to_graph_survey}} surveys several methods of transforming hypergraphs to graphs, and considers the addition of ``dummy'' nodes to the star model.
 In our star model, we preferred to designate the net source as the center of the star, rather than introduce a dummy node.
@@
 An intuitive explanation for this result is that with the clique topology, the partitioner does not have knowledge of which vertex is the source of a net and which vertices are sinks.
-In contrast, the star topology clearly differentiates sources from sinks and this appears to give the partitioner an “anchor point” that pulls all hyperedge fanouts toward the source.
+In contrast, the star topology clearly differentiates sources from sinks and this appears to give the \hl{iterative refinement based} partitioner an “anchor point” that pulls all hyperedge fanouts toward the source.
 Additionally, if $n$ is the average number of vertices in a hyperedge, the clique topology generates $O(n^2)$ edges while the star topology generates only $O(n)$ edges, so the star topology saves both memory and runtime.
 
-For both topologies, the $1/n$ edge weight scheme gives the smallest cut size on average.
+\hl{For both topologies, the $1/n$ edge weight scheme gives the smallest cut size on average.
 This can be explained intuitively in terms of the total edge weight over all edges of the graph.
-The star $1$ (constant) edge weight scheme assigns a total weight of $n$ to each net, which heavily weights high-fanout nets.
-In contrast, the star $1/n$ model assigns the same total weight of $1$ to each net.
+The star $1$ (constant) edge weight scheme assigns a total weight of $n - 1$ to each net, which heavily weights high-fanout nets.
+In contrast, the star $1/n$ model assigns a total weight of $1 - \frac{1}{n}$ to each net.
 The star $1/n^2$ model penalizes high-fanout nets relative to lower-fanout nets.
-Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ model achieves this.
+Since we seek to minimize the hyperedge cut, it makes intuitive sense to weight all hyperedges equally, and the star $1/n$ model best achieves this.}
 
@@
 
-The star topology with 1/n edge weights achieves the best minimum channel width, confirming that hyperedge cutsize is a good proxy for routability.
+The star topology with $1/n$ edge weights achieves the best minimum channel width, confirming that hyperedge cutsize is a good proxy for routability.
 Consequently, we use this model in all future results in this paper.

### 2019-06-18 doc/src/vtr/vtr_flow_fig.tex

\documentclass[tikz,border=10pt]{standalone}
\usepackage{tikz}
\usepackage{makecell}

\usetikzlibrary{arrows}
\usetikzlibrary{circuits.logic.US}
\usetikzlibrary{matrix}
\usetikzlibrary{positioning}
\usetikzlibrary{backgrounds}
\usetikzlibrary{decorations.pathreplacing, positioning, shapes.geometric, shapes.misc, shadows, fit}
\usetikzlibrary{shapes.geometric, shapes.symbols, arrows, shadows, fit, backgrounds, 3d, plotmarks, calc, intersections, matrix, patterns}

\tikzstyle{thin_arrow} = [draw,->,>=stealth]
\tikzstyle{arrow} = [thick,thin_arrow]
\tikzstyle{bidir_arrow} = [draw,thick,<->,>=stealth]
\tikzstyle{graph_node} = [draw, circle, inner sep=0.5pt, font=\footnotesize]
\tikzstyle{lb_color} =  [top color=cyan!5, bottom color=cyan!15]
\tikzstyle{lb} =  [rectangle, anchor=south west, draw, lb_color]
\tikzstyle{large_lb} = [lb, minimum width=1.5cm, minimum height=1.5cm]
\tikzstyle{ram} = [rectangle, anchor=south west, draw, top color=teal!10, bottom color=teal!20]
\tikzstyle{dsp} = [rectangle, anchor=south west, draw, top color=olive!30, bottom color=olive!40]

\tikzstyle{trapezoid} = [trapezium,
                    trapezium stretches=true,
                    trapezium left angle=70,
                    trapezium right angle=110,
                    ]

\tikzstyle{io} = [trapezoid,
                    minimum width=2.5cm,
                    minimum height=1cm,
                    text centered,
                    align=center,
                    draw=black,
                    drop shadow,
                    top color=blue!10,
                    bottom color=blue!30]

\tikzstyle{design} = [io]
\tikzstyle{metric} = [io,
                      minimum width=1cm]

\tikzstyle{device} = [rounded rectangle,
                    minimum width=3cm,
                    minimum height=1cm,
                    text centered,
                    align=center,
                    draw=black,
                    drop shadow,
                    top color=black!10!green!20,
                    bottom color=black!10!green!70]

\tikzstyle{process} = [rectangle,
                        minimum width=3cm,
                        minimum height=1cm,
                        text centered,
                        align=center,
                        draw=black,
                        drop shadow,
                        top color=orange!10,
                        bottom color=orange!35]

\tikzstyle{process_no_shadow} = [rectangle,
                        minimum width=3cm,
                        minimum height=1cm,
                        text centered,
                        align=center,
                        draw=black,
                        top color=orange!10,
                        bottom color=orange!35]

\begin{document}

\begin{tikzpicture}[node distance=1.0cm]
    \tikzstyle{vtr_callout} = [draw=black!40!green, fill=black!30!green, fill opacity=0.25, very thick, dash pattern={on 8pt off 4pt}, rounded corners=0.3cm]
    \tikzstyle{titan_callout} = [draw=black!20!red, fill=red, very thick, fill opacity=0.25, dash pattern={on 6pt off 2pt on 3pt off 2pt}, rounded corners=0.3cm]

    \pgfmathsetmacro{\vtroffset}{0.5cm}
    \pgfmathsetmacro{\titanoffset}{0.3cm}
    \pgfmathsetmacro{\metricoffset}{0.6cm}

    \newcommand{\arrowshift}{0.25cm}
    \newcommand{\arrowxshift}{0.07cm}

    \matrix [row sep=0.6cm, column sep=0.9cm] {
        \node (arch) [device] {FPGA \\ Architecture}; & & \node (hdl) [design] {HDL}; & \\
        & \node (odin_synth) [process] {Odin II}; & \node (quartus_map) [process] {Quartus \\ Map}; & \node (yosys_frontend) [] {Yosys}; \\
        & & \node (vqm2blif) [process] {VQM to BLIF}; & \node (abc_yosys) [process_no_shadow] {ABC}; \\
        & \node (abc_synth) [process] {ABC}; & & \\
        & & \node (techmapped_netlist) [design] {Tech. Mapped \\ Netlist}; & & \\
        & \node (vpr_pack) [process] {VPR \\ Pack}; & \node (packing) [design] {Packing}; & \node (other_pack) [process, dashed] {Other \\ Pack}; & \\
        & \node (vpr_place) [process] {VPR \\ Place}; & \node (placement) [design] {Placement}; & \node (other_place) [process, dashed] {Other \\ Place}; & \\
        & \node (vpr_route) [process] {VPR \\ Route}; & \node (routing) [design] {Routing}; & \node (other_route) [process, dashed] {Other \\ Route}; & \\
        & & & & \\
        & \node (vpr_analysis) [process] {VPR \\ Analysis}; & & \node (assembler) [process, dashed] {Bitstream Gen.}; & \\
    };
    \node(rr_graph) [device,left=0.25cm of arch] {RR-Graph};

    \begin{scope}[on background layer]
        \node (yosys) [process, fit=(yosys_frontend) (abc_yosys)] {};
    \end{scope}

    \node (implementation) [draw, densely dotted, inner sep=0.35cm, fit=(techmapped_netlist) (packing) (placement) (routing)] {};

    \node (bitstream) [design, below=0.9cm of assembler, dashed] {Bitstream};

    \node (timing) [metric, below=0.9cm of vpr_analysis] {Timing \\ Metrics};
    \node (area) [metric, left=0.1cm of timing] {Area \\ Metrics};
    \node (power) [metric, right=0.1cm of timing] {Power \\ Metrics};
    \node (post_impl_netlist) [design, left=0.8cm of area] {Post-Impl. \\ Netlist};

    \draw [arrow] (arch) |- (odin_synth);
    \draw [arrow] (arch) |- (vpr_pack.west);
    \draw [arrow] (arch) |- ([yshift=\arrowshift]vpr_place.west);
    \draw [arrow] (arch) |- ([yshift=\arrowshift]vpr_route.west);
    \draw [arrow] (arch) |- ([yshift=\arrowshift]vpr_analysis.west);
    \draw [arrow] (arch) |- ([yshift=\arrowshift]vpr_analysis.west);

    \draw [arrow] (hdl) -- (odin_synth.north);
    \draw [arrow] (hdl) -- (quartus_map);
    \draw [arrow] (hdl) -- (yosys.north);
    \draw[arrow] (odin_synth) -- (abc_synth);
    \draw[arrow] (quartus_map) -- (vqm2blif);
    \draw[arrow] (yosys) |- ([yshift=\arrowshift,xshift=\arrowxshift]techmapped_netlist.east);
    \draw[arrow] (vqm2blif) -- (abc_synth);
    \draw[arrow] (vqm2blif) -- (techmapped_netlist);
    \draw[arrow] (abc_synth) |- ([yshift=\arrowshift,xshift=\arrowxshift]techmapped_netlist.west);

    \draw[arrow] ([yshift=-\arrowshift,xshift=-\arrowxshift]techmapped_netlist.west) -| (vpr_pack);
    \draw[arrow, dashed] ([yshift=-\arrowshift,xshift=-\arrowxshift]techmapped_netlist.east) -| (other_pack);

    \draw[arrow] (vpr_pack) -- (vpr_place);
    \draw[arrow] (vpr_place) -- (vpr_route);

    \draw[arrow, dashed] (other_pack) -- (other_place);
    \draw[arrow, dashed] (other_place) -- (other_route);

    \draw[bidir_arrow] (vpr_pack) -- (packing);
    \draw[bidir_arrow] (vpr_place) -- (placement);
    \draw[bidir_arrow] (vpr_route) -- (routing);

    \draw[bidir_arrow, dashed] (other_pack) -- (packing);
    \draw[bidir_arrow, dashed] (other_place) -- (placement);
    \draw[bidir_arrow, dashed] (other_route) -- (routing);

    \draw[arrow] (vpr_route) -- (vpr_analysis);
    \draw[arrow] (implementation.south west) -- (vpr_analysis);
    \draw[arrow] (vpr_analysis) -- (area);
    \draw[arrow] (vpr_analysis) -- (timing);
    \draw[arrow] (vpr_analysis) -- (power);
    \draw[arrow] (vpr_analysis) -- (post_impl_netlist);

    \draw[arrow, dashed] (implementation.south east) -- (assembler);
    \draw[arrow, dashed] (assembler) -- (bitstream);

    \draw [arrow, dashed] (rr_graph) |- ([yshift=-\arrowshift]vpr_place.west);
    \draw [arrow, dashed] (rr_graph) |- ([yshift=-\arrowshift]vpr_route.west);
    \draw [arrow, dashed] (rr_graph) |- ([yshift=-\arrowshift]vpr_analysis.west);

    \begin{scope}[on background layer]
        %VTR Callout
        \draw[vtr_callout]
            ([xshift=-\vtroffset,yshift=\vtroffset]odin_synth.north west)
         -- ([xshift=-\vtroffset,yshift=-\vtroffset]vpr_analysis.south west)
         -- ([xshift=\vtroffset,yshift=-\vtroffset]vpr_analysis.south east)
         -- ([xshift=\vtroffset,yshift=-\vtroffset]vpr_pack.north east)
         -- ([xshift=-\vtroffset,yshift=-\vtroffset]techmapped_netlist.bottom left corner-|techmapped_netlist.bottom left corner)
         -- ([xshift=\vtroffset,yshift=-\vtroffset]techmapped_netlist.bottom right corner-|techmapped_netlist.top right corner)
         -- ([xshift=\vtroffset,yshift=\vtroffset]techmapped_netlist.top right corner)
         -- ([xshift=\vtroffset,yshift=\vtroffset]techmapped_netlist.top left corner-|abc_synth.south east)
         -- ([xshift=\vtroffset,yshift=\vtroffset]abc_synth.south east)
         -- ([xshift=\vtroffset,yshift=\vtroffset]odin_synth.north east)
         -- cycle;

        %Titan Callout
        \draw[titan_callout]
            ([xshift=-\titanoffset,yshift=\titanoffset]quartus_map.north west)
         -- ([xshift=-\titanoffset,yshift=\titanoffset]quartus_map.north west|-techmapped_netlist.top left corner)
         -- ([xshift=-\titanoffset,yshift=\titanoffset]vpr_pack.north west)
         -- ([xshift=-\titanoffset,yshift=-\titanoffset]vpr_analysis.south west)
         -- ([xshift=\titanoffset,yshift=-\titanoffset]vpr_analysis.south east)
         -- ([xshift=\titanoffset,yshift=\titanoffset]vpr_pack.north east)
         -- ([xshift=\titanoffset,yshift=-\titanoffset]techmapped_netlist.bottom right corner-|techmapped_netlist.top right corner)
         -- ([xshift=\titanoffset,yshift=\titanoffset]quartus_map.north east)
         -- cycle;
    \end{scope}

    %Legend
    \matrix (legend) [row sep=0.1cm, column sep=0.0cm, draw, right=0.50cm of yosys, outer sep=0.0cm, inner sep=0.1cm, nodes={anchor=center}] {
        \node [device, minimum width=1.5cm, minimum height=0.5cm] {}; & \node[] {\footnotesize \makecell{Device \\ Description}}; \\
        \node [design, minimum width=1.5cm, minimum height=0.5cm] {}; & \node[] {\footnotesize  \makecell{Design}}; \\
        \node [process, minimum width=1.5cm, minimum height=0.5cm] {}; & \node[] {\footnotesize \makecell{CAD Tool \\ Stage}}; \\
        \node [vtr_callout, minimum width=1.5cm, minimum height=0.5cm] {}; & \node[] {\footnotesize \makecell{VTR Flow}}; \\
        \node [titan_callout, minimum width=1.5cm, minimum height=0.5cm] {}; & \node[] {\footnotesize \makecell{Titan Flow}}; \\
        \node (optional) [rectangle, minimum width=1.0cm, minimum height=0.5cm] {}; & \node[] {\footnotesize \makecell{Standard \\Optional}}; \\
    };
    \draw [arrow] ([yshift=+0.20cm] optional.west) -- ([yshift=+0.20cm] optional.east);
    \draw [arrow, dashed] ([yshift=-0.15cm] optional.west) -- ([yshift=-0.15cm] optional.east);
\end{tikzpicture}
\end{document}

### 2019-06-27 libs/librtlnumber/src/include/internal_bits.hpp
librtlnumber - updated author list

@@ -1,8 +1,9 @@
 /* Authors: Aaron Graham (aaron.graham@unb.ca, aarongraham9@gmail.com),
- *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com) and
- *            Dr. Kenneth B. Kent (ken@unb.ca)
- *            for the Reconfigurable Computing Research Lab at the
- *             Univerity of New Brunswick in Fredericton, New Brunswick, Canada
+ *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com),
+ *            Alexandrea Demmings (alexandrea.demmings@unb.ca, lxdemmings@gmail.com) and
+ *             Dr. Kenneth B. Kent (ken@unb.ca)
+ *             for the Reconfigurable Computing Research Lab at the
+ *              Univerity of New Brunswick in Fredericton, New Brunswick, Canada
  */
 
 #ifndef INTERNAL_BITS_HPP

### 2019-06-27 libs/librtlnumber/src/include/rtl_utils.hpp

@@ -1,8 +1,9 @@
 /* Authors: Aaron Graham (aaron.graham@unb.ca, aarongraham9@gmail.com),
- *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com) and
- *            Dr. Kenneth B. Kent (ken@unb.ca)
- *            for the Reconfigurable Computing Research Lab at the
- *             Univerity of New Brunswick in Fredericton, New Brunswick, Canada
+ *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com),
+ *            Alexandrea Demmings (alexandrea.demmings@unb.ca, lxdemmings@gmail.com) and
+ *             Dr. Kenneth B. Kent (ken@unb.ca)
+ *             for the Reconfigurable Computing Research Lab at the
+ *              Univerity of New Brunswick in Fredericton, New Brunswick, Canada
  */
 
 #ifndef RTL_UTILS_H

### 2019-06-27 libs/librtlnumber/src/rtl_int.cpp

@@ -1,8 +1,9 @@
 /* Authors: Aaron Graham (aaron.graham@unb.ca, aarongraham9@gmail.com),
- *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com) and
- *            Dr. Kenneth B. Kent (ken@unb.ca)
- *            for the Reconfigurable Computing Research Lab at the
- *             Univerity of New Brunswick in Fredericton, New Brunswick, Canada
+ *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com),
+ *            Alexandrea Demmings (alexandrea.demmings@unb.ca, lxdemmings@gmail.com) and
+ *             Dr. Kenneth B. Kent (ken@unb.ca)
+ *             for the Reconfigurable Computing Research Lab at the
+ *              Univerity of New Brunswick in Fredericton, New Brunswick, Canada
  */
 
 #include <string>

### 2019-06-27 libs/librtlnumber/src/rtl_utils.cpp

@@ -1,8 +1,9 @@
 /* Authors: Aaron Graham (aaron.graham@unb.ca, aarongraham9@gmail.com),
- *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com) and
- *            Dr. Kenneth B. Kent (ken@unb.ca)
- *            for the Reconfigurable Computing Research Lab at the
- *             Univerity of New Brunswick in Fredericton, New Brunswick, Canada
+ *           Jean-Philippe Legault (jlegault@unb.ca, jeanphilippe.legault@gmail.com),
+ *            Alexandrea Demmings (alexandrea.demmings@unb.ca, lxdemmings@gmail.com) and
+ *             Dr. Kenneth B. Kent (ken@unb.ca)
+ *             for the Reconfigurable Computing Research Lab at the
+ *              Univerity of New Brunswick in Fredericton, New Brunswick, Canada
  */
 
 #include "rtl_utils.hpp"

### 2019-06-27 ODIN_II/SRC/ast_util.cpp
odin alloc library and fix memleak on argparse

@@ -75,11 +75,16 @@ void update_tree_tag(ast_node_t *node, int cases, int tagged);
 		high_level_id = -1;
 
 		ids type = NO_ID;
-		if(!strcmp(global_args.high_level_block,"if")){
+		if(global_args.high_level_block.value() == "if")
+		{
 			type = IF;
-		}else if(!strcmp(global_args.high_level_block,"always")){
+		}
+		else if(global_args.high_level_block.value() == "always")
+		{
 			type = ALWAYS;
-		}else if(!strcmp(global_args.high_level_block,"module")){
+		}
+		else if(global_args.high_level_block.value() == "module")
+		{
 			type = MODULE;
 		}
 
@@ -1086,6 +1091,9 @@ static void check_binary_operation(ast_node_t **node)
  *-------------------------------------------------------------------------------------------*/
 ast_node_t *fold_unary(ast_node_t *node)
 {
+	if(!node)
+		return NULL;
+		
 	operation_list op_id = node->types.operation.op;
 	ast_node_t *child_0 = node->children[0];
 

### 2019-06-27 ODIN_II/SRC/include/odin_memory.hpp

@@ -1,0 +1,138 @@
+#ifndef ODIN_MEMORY_H
+#define ODIN_MEMORY_H
+
+#include <cstdint>
+#include <cstddef>
+#include <limits>
+#include <cstring>
+#include <algorithm>
+
+#include "odin_error.h"
+
+#ifndef __GLIBC__
+#    include <stdlib.h>
+#else
+#    include <malloc.h>
+#endif
+
+namespace odin
+{
+    extern uintptr_t min_address;
+    extern uintptr_t max_address;
+    
+    char *strdup(const char *in);
+
+    template<typename T>
+    void free(T** ptr_ref)
+    {
+        oassert(ptr_ref != NULL
+            && "did not pass in a valid ref");
+
+        T *ptr = (*ptr_ref);
+        (*ptr_ref) = NULL;
+
+        if(ptr != NULL 
+        && (uintptr_t)ptr >= min_address 
+        && (uintptr_t)ptr <= max_address )
+        {
+            std::free(ptr);
+        }
+    }
+
+    template<typename T, typename S>
+    void *calloc(T _n_element, S _element_size)
+    {
+        void* ret = NULL;
+        size_t n_element = 0;
+        size_t element_size = 0;
+        size_t n_bytes = 0;
+
+        if(std::is_unsigned<T>() || _n_element > 0)
+        {
+            n_element = static_cast<size_t>(_n_element);
+        }
+
+        if(std::is_unsigned<S>() || _element_size > 0)
+        {
+            element_size = static_cast<size_t>(_element_size);
+        }
+
+        n_bytes = n_element*element_size;
+        if(n_bytes > 0)
+        {
+            ret = std::calloc(n_bytes, 1);
+
+            oassert(ret != NULL
+                && "odin::calloc failed, OOM?");
+
+        }
+
+        if(ret != NULL)
+        {
+            min_address = std::min(min_address, (uintptr_t)ret);
+            max_address = std::max(max_address, (uintptr_t)ret);
+        }
+
+        return (ret);
+    }
+    
+        // we use a template to check for valid size
+    template<typename T>
+    void *malloc(T _n_bytes)
+    {
+        return calloc(_n_bytes, 1);
+    }
+
+    template<typename T, typename S>
+    void realloc(T **ptr_ref, S _n_bytes)
+    {
+        oassert(ptr_ref != NULL
+            && "did not pass in a valid ref");
+
+        T *ptr = (*ptr_ref);
+        (*ptr_ref) = NULL;
+
[…]

### 2019-06-27 ODIN_II/SRC/include/simulate_blif.h

@@ -124,7 +124,6 @@ typedef struct sim_data_t_t
 	FILE *modelsim_out;
 	FILE *in  = NULL;
 	long num_vectors;
-	char *input_vector_file;
 
 	double total_time; // Includes I/O
 	double simulation_time; // Does not include I/O

### 2019-06-27 ODIN_II/SRC/netlist_create_from_ast.cpp

@@ -505,7 +505,7 @@ ast_node_t *find_top_module()
 	std::string desired_module("");
 	bool found_desired_module = false;
 
-	if ( global_args.top_level_module_name )
+	if ( global_args.top_level_module_name.provenance() == argparse::Provenance::SPECIFIED )
 	{
 		desired_module = global_args.top_level_module_name;
 		printf("Using Top Level Module: %s\n", desired_module.c_str());

### 2019-06-27 ODIN_II/SRC/odin_ii.cpp

@@ -116,7 +116,7 @@ static ODIN_ERROR_CODE synthesize_verilog()
 	 */
 
 	/* after the ast is made potentially do tagging for downstream links to verilog */
-	if (global_args.high_level_block)
+	if (global_args.high_level_block.provenance() == argparse::Provenance::SPECIFIED)
 		add_tag_data();
 
 	/**
@@ -177,11 +177,9 @@ static ODIN_ERROR_CODE synthesize_verilog()
 	 *	point for outputs.  This includes soft and hard mapping all structures to the
 		*	target format.  Some of these could be considred optimizations 
 		*/
-	char *output_file = global_args.output_file;
 	printf("Outputting the netlist to the specified output format\n");
-
 	
-	output_blif(output_file, verilog_netlist);
+	output_blif(global_args.output_file.value().c_str(), verilog_netlist);
 
 	global_param_table_sc = sc_free_string_cache(global_param_table_sc);
 	module_names_to_idx = sc_free_string_cache(module_names_to_idx);
@@ -190,7 +188,7 @@ static ODIN_ERROR_CODE synthesize_verilog()
 
 	elaboration_time = wall_time() - elaboration_time;
 
-	printf("Successful High-level synthesis by Odin\n\tBlif file available at %s\n\tRan in ",output_file);
+	printf("Successful High-level synthesis by Odin\n\tBlif file available at %s\n\tRan in ",global_args.output_file.value().c_str());
 	print_time(elaboration_time);
 	printf("\n");
 	printf("--------------------------------------------------------------------\n");
@@ -232,19 +230,19 @@ struct netlist_t_t *start_odin_ii(int argc,char **argv)
 	get_options(argc, argv);
 
 	/* read the confirguration file .. get options presets the config values just in case theyr'e not read in with config file */
-	if (global_args.config_file != NULL)
+	if (global_args.config_file.provenance() == argparse::Provenance::SPECIFIED)
 	{
 		printf("Reading Configuration file\n");
-		read_config_file(global_args.config_file);
+		read_config_file(global_args.config_file.value().c_str());
 	}
 
 	/* read the FPGA architecture file */
-	if (global_args.arch_file != NULL)
+	if (global_args.arch_file.provenance() == argparse::Provenance::SPECIFIED)
 	{
 		printf("Reading FPGA Architecture file\n");
 		try 
 		{
[…]

### 2019-06-27 ODIN_II/SRC/odin_memory.cpp

@@ -1,0 +1,26 @@
+#include "odin_memory.hpp"
+
+namespace odin
+{
+    uintptr_t min_address = std::numeric_limits<uintptr_t>::max();
+    uintptr_t max_address = std::numeric_limits<uintptr_t>::min();
+
+    char *strdup(const char *in)
+    {
+        size_t len = (in)? strlen(in): 0;
+        char *to_return = (char*)calloc(len+1, sizeof(char));
+
+        if(to_return)
+        {
+            if(in)
+            {
+                memcpy(to_return, in, len);
+            }
+
+            to_return[len] = '\0';
+        }
+        
+        return to_return;
+    }
+}
+

### 2019-06-27 ODIN_II/SRC/output_blif.cpp

@@ -56,25 +56,24 @@ void output_blif_pin_connect(nnode_t *node, FILE *out);
  * (function: output_blif)
  * 	The function that prints out the details for a blif formatted file
  *-------------------------------------------------------------------------*/
-void output_blif(char *file_name, netlist_t *netlist)
+void output_blif(const char *file_name, netlist_t *netlist)
 {
 	int i;
 	int count = 0;
 	short first_time_inputs = FALSE;
 	short first_time_outputs = FALSE;
 	FILE *out;
-	char *out_file;
 
 	/* open the file for output */
-	if (global_args.high_level_block != NULL)
+	if (global_args.high_level_block.provenance() == argparse::Provenance::SPECIFIED )
 	{
-		out_file = (char*)vtr::malloc(sizeof(char)*(1+strlen(file_name)+strlen(global_args.high_level_block)+6));
-		odin_sprintf(out_file, "%s_%s.blif", file_name, global_args.high_level_block.value());
-		out = fopen(out_file, "w");
+		std::string out_file = "";
+		out_file = out_file + file_name + "_" + global_args.high_level_block.value() + ".blif";
+		out = fopen(out_file.c_str(), "w+");
 	}
 	else
 	{
-		out = fopen(file_name, "w");
+		out = fopen(file_name, "w+");
 	}
 
 	if (out == NULL)
@@ -93,7 +92,7 @@ void output_blif(char *file_name, netlist_t *netlist)
 			first_time_inputs = TRUE;
 		}
 
-		if (global_args.high_level_block != NULL)
+		if (global_args.high_level_block.provenance() == argparse::Provenance::SPECIFIED)
 		{
 			if (strlen(netlist->top_input_nodes[i]->name) + count < 79)
 				count = count + fprintf(out, " %s^^%i-%i", netlist->top_input_nodes[i]->name, netlist->top_input_nodes[i]->related_ast_node->far_tag, netlist->top_input_nodes[i]->related_ast_node->high_number);
@@ -136,7 +135,7 @@ void output_blif(char *file_name, netlist_t *netlist)
 				first_time_outputs = TRUE;
 			}
 
-			if (global_args.high_level_block != NULL)
+			if ( global_args.high_level_block.provenance() == argparse::Provenance::SPECIFIED )
 			{
[…]

### 2019-06-27 ODIN_II/SRC/read_xml_config_file.cpp

@@ -50,7 +50,7 @@ void set_default_optimization_settings(config_t *config);
  *
  * See config_t in types.h to see the data structures used in this read.
  *-----------------------------------------------------------------------*/
-void read_config_file(char *file_name)
+void read_config_file(const char *file_name)
 {
 	pugi::xml_node config, next;
 
@@ -121,13 +121,13 @@ void read_outputs(pugi::xml_node a_node, config_t *config, const pugiutil::loc_d
 	child = get_single_child(a_node, "output_type", loc_data, OPTIONAL);
 	if (child != NULL)
 	{
-		config->output_type = vtr::strdup(child.child_value());
+		config->output_type = child.child_value();
 	}
 
 	child = get_single_child(a_node, "output_path_and_name", loc_data, OPTIONAL);
 	if (child != NULL)
 	{
-		global_args.output_file.set(vtr::strdup(child.child_value()), argparse::Provenance::SPECIFIED);
+		global_args.output_file.set(child.child_value(), argparse::Provenance::SPECIFIED);
 	}
 
 	child = get_single_child(a_node, "target", loc_data, OPTIONAL);
@@ -137,11 +137,11 @@ void read_outputs(pugi::xml_node a_node, config_t *config, const pugiutil::loc_d
 		if (child != NULL)
 		{
 			/* Two arch files specified? */
-			if (global_args.arch_file != NULL)
+			if (global_args.arch_file.value() != "")
 			{
 				error_message(ARG_ERROR, -1, -1, "%s", "Error: Arch file specified in config file AND command line\n");
 			}
-			global_args.arch_file.set(vtr::strdup(child.child_value()), argparse::Provenance::SPECIFIED);
+			global_args.arch_file.set(child.child_value(), argparse::Provenance::SPECIFIED);
 		}
 	}
 	return;
@@ -169,7 +169,7 @@ void read_debug_switches(pugi::xml_node a_node, config_t *config, const pugiutil
 	child = get_single_child(a_node, "debug_output_path", loc_data, OPTIONAL);
 	if (child != NULL)
 	{
-		config->debug_output_path = std::string(child.child_value());
+		config->debug_output_path = child.child_value();
 	}
 
 	child = get_single_child(a_node, "print_parse_tokens", loc_data, OPTIONAL);

### 2019-06-27 ODIN_II/SRC/simulate_blif.cpp

@@ -1,5 +1,4 @@
-/*	number_of_workers = std::max(1, global_args.parralelized_simulation.value());
-
+/**
 Permission is hereby granted, free of charge, to any person
 obtaining a copy of this software and associated documentation
 FILEs (the "Software"), to deal in the Software without
@@ -221,7 +220,7 @@ static void write_cycle_to_file(lines_t *l, FILE* file, int cycle);
 static void write_vector_to_modelsim_file(lines_t *l, FILE *modelsim_out, int cycle);
 static void write_cycle_to_modelsim_file(netlist_t *netlist, lines_t *l, FILE* modelsim_out, int cycle);
 
-static int verify_output_vectors(char* output_vector_file, int num_test_vectors);
+static int verify_output_vectors(const char* output_vector_file, int num_test_vectors);
 
 static void add_additional_items_to_lines(nnode_t *node, lines_t *l);
 
@@ -290,11 +289,10 @@ void simulate_netlist(netlist_t *netlist)
 
 	printf("\n");
 	// If a second output vector file was given via the -T option, verify that it matches.
-	char *output_vector_file = global_args.sim_vector_output_file;
-	if (output_vector_file)
+	if (global_args.sim_vector_output_file.provenance() == argparse::Provenance::SPECIFIED)
 	{
-		if (verify_output_vectors(output_vector_file, sim_data->num_vectors))
-			printf("Vector file \"%s\" matches output\n", output_vector_file);
+		if ( verify_output_vectors( global_args.sim_vector_output_file.value().c_str(), sim_data->num_vectors) )
+			printf("Vector file \"%s\" matches output\n", global_args.sim_vector_output_file.value().c_str());
 		else
 			error_message(SIMULATION_ERROR, 0, -1, "%s\n", "Vector files differ.");
 		printf("\n");
@@ -429,33 +427,33 @@ sim_data_t *init_simulation(netlist_t *netlist)
 	sim_data_t *sim_data = (sim_data_t *)vtr::malloc(sizeof(sim_data_t));
 
 	sim_data->netlist = netlist;
-	printf("Beginning simulation. Output_files located @: %s\n", ((char *)global_args.sim_directory)); 
[…]

### 2019-06-27 ODIN_II/SRC/include/config_t.h
make generic source file line reader (support for multiline, single line comment, escaped new line and truncates duplicate whitespaces)

@@ -15,7 +15,7 @@ struct config_t
 	bool output_ast_graphs; // switch that outputs ast graphs per node for use with GRaphViz tools
 	bool output_netlist_graphs; // switch that outputs netlist graphs per node for use with GraphViz tools
 	bool print_parse_tokens; // switch that controls whether or not each token is printed during parsing
-	bool output_preproc_source; // switch that outputs the pre-processed source
+	bool output_preproc_source; // TODO: unused
 	
 	int min_hard_multiplier; // threshold from hard to soft logic
 	int mult_padding; // setting how multipliers are padded to fit fixed size

### 2019-06-27 ODIN_II/SRC/include/verilog_preprocessor.h

@@ -1,91 +1,10 @@
 #ifndef verilog_preprocessor_h
 #define verilog_preprocessor_h
 
-#define DefaultSize 20
 #include <stdio.h>
 
-//#define BLOCK_EMPTY_DEFINES
-
-/* Structs */
-typedef struct veri_include
-{
-	char *path;
-	struct veri_include *included_from;
-	int	line;
-} veri_include;
-
-typedef struct veri_define
-{
-	char *symbol;
-	char *value;
-	int line;
-	veri_include *defined_in;
-} veri_define;
-
-struct veri_Includes
-{
-	veri_include **included_files;
-	int current_size;
-	int current_index;
-} ;
-
-struct veri_Defines
-{
-	veri_define **defined_constants;
-	int current_size;
-	int current_index;
-} ;
-
-/* Globals */
-extern struct veri_Includes veri_includes;
-extern struct veri_Defines veri_defines;
-
-/* Initalization and Cleanup */
 int init_veri_preproc();
 int cleanup_veri_preproc();
-void clean_veri_define(veri_define *current);
-void clean_veri_include(veri_include *current);
-
-/* Adding/Removing includes or defines */
-int add_veri_define(char *symbol, char *value, int line, veri_include *included_from);
-char* ret_veri_definedval(const char *symbol);
-int veri_is_defined(const char * symbol);
-veri_include* add_veri_include(const char *path, int line, veri_include *included_from);
-
-/* Preprocessor ------------------------------------------------------------- */
 FILE* veri_preproc(FILE *source);
-void veri_preproc_bootstraped(FILE *source, FILE *preproc_producer, veri_include *current_include);
-
-/* ------------------------------------------------------------------------- */
-
-
-/* Stack for tracking conditional branches --------------------------------- */
-typedef struct veri_flag_node
-{
-	int flag;
-	struct veri_flag_node *next;
-} veri_flag_node;
-
-typedef struct
-{
-	veri_flag_node *top;
-} veri_flag_stack;
-
-/* stack methods */
-int top(veri_flag_stack *stack);
-int pop(veri_flag_stack *stack);
-void push(veri_flag_stack *stack, int flag);
-
-/* ------------------------------------------------------------------------- */
-
-
-/* General Utility methods ------------------------------------------------- */
-bool is_whitespace(const char in);
-char *trim(char *input_str);
-char* trim(char *input_string, size_t n);
-char *get_line(char *line, long *size, FILE *source);
-
-/* ------------------------------------------------------------------------- */
-
 
 #endif

### 2019-06-27 ODIN_II/SRC/odin_ii.cpp

@@ -625,7 +625,7 @@ void set_default_config()
 	configuration.output_ast_graphs = 0;
 	configuration.output_netlist_graphs = 0;
 	configuration.print_parse_tokens = 0;
-	configuration.output_preproc_source = 1;
+	configuration.output_preproc_source = 0; // TODO: unused
 	configuration.debug_output_path = std::string(DEFAULT_OUTPUT);
 	configuration.arch_file = NULL;
 

### 2019-06-27 ODIN_II/SRC/parse_making_ast.cpp

@@ -77,34 +77,10 @@ short to_view_parse;
 /*
  * File-scope function declarations
  */
-void graphVizOutputPreproc(FILE *yyin);
 ast_node_t *newFunctionAssigning(ast_node_t *expression1, ast_node_t *expression2, int line_number);
 ast_node_t *newHardBlockInstance(char* module_ref_name, ast_node_t *module_named_instance, int line_number);
 ast_node_t *resolve_ports(ids top_type, ast_node_t *symbol_list);
 
-/*
- * Function implementations
- */
-void graphVizOutputPreproc(FILE *yyin)
-{
-	std::string file_out = 	configuration.debug_output_path 
-							+ "/" 
-							+ strip_path_and_ext(configuration.list_of_file_names[current_parse_file]).c_str() 
-							+ "_preproc.v";
-
-	FILE *fp = fopen(file_out.c_str(), "w");
-	oassert(fp);
-
-	char *line = NULL;
-
-	while ((line = get_line(line, NULL, yyin)) != NULL)
-	{
-		fprintf(fp, "%s", line);
-		vtr::free(line);
-	}
-	fclose(fp);
-	rewind(yyin);
-}
 
 static void assert_supported_file_extension(std::string input_file, int file_number)
 {
@@ -160,10 +136,6 @@ void parse_to_ast()
 		}
 
 		yyin = veri_preproc(yyin);
-
-		/* write out the pre-processed file */
-		if (configuration.output_preproc_source)
-			graphVizOutputPreproc(yyin);
 
 		/* reset the line count */
 		yylineno = 0;

### 2019-06-27 ODIN_II/SRC/read_xml_config_file.cpp

@@ -178,12 +178,6 @@ void read_debug_switches(pugi::xml_node a_node, config_t *config, const pugiutil
 		config->print_parse_tokens = atoi(child.child_value());
 	}
 
-	child = get_single_child(a_node, "output_preproc_source", loc_data, OPTIONAL);
-	if (child != NULL)
-	{
-		config->output_preproc_source = atoi(child.child_value());
-	}
-
 	return;
 }
 

### 2019-06-27 ODIN_II/SRC/simulate_blif.cpp

@@ -28,6 +28,7 @@ OTHER DEALINGS IN THE SOFTWARE.
 #include <cmath>
 #include "vtr_util.h"
 #include "vtr_memory.h"
+#include "odin_buffer.hpp"
 #include "odin_util.h"
 #include <string>
 #include <sstream>
@@ -320,7 +321,7 @@ void simulate_netlist(netlist_t *netlist)
 	int       progress_bar_position = -1;
 	const int progress_bar_length   = 50;
 	
-	int increment_vector_by = global_args.sim_num_test_vectors;;
+	int increment_vector_by = global_args.sim_num_test_vectors;
 	double min_coverage =0.0;
 	if(global_args.sim_min_coverage)
 	{
@@ -563,7 +564,7 @@ void simulate_steps_sequential(sim_data_t *sim_data,double min_coverage)
 	int       progress_bar_position = -1;
 	const int progress_bar_length   = 50;
 	
-	int increment_vector_by = global_args.sim_num_test_vectors;;
+	int increment_vector_by = global_args.sim_num_test_vectors;
 	
 
 	double current_coverage =0.0;
@@ -3185,48 +3186,6 @@ static void instantiate_memory(nnode_t *node, long data_width, long addr_width)
 	}
 }
 
-/*
- * Removes white space (except new lines) and comments from
- * the given mif file and returns the resulting temporary file.
- */
-static FILE *preprocess_mif_file(FILE *source)
-{
-	FILE *destination = tmpfile();
-	destination = freopen(NULL, "r+", destination);
-	rewind(source);
-
-	char line[BUFFER_MAX_SIZE];
-	int in_multiline_comment = FALSE;
-	while (fgets(line, BUFFER_MAX_SIZE, source))
-	{
-		unsigned int i;
-		for (i = 0; i < strlen(line); i++)
-		{
-			if (!in_multiline_comment)
-			{
-				// For a single line comment, skip the rest of the line.
-				if (line[i] == '-' && line[i+1] == '-')
-					break;
-				// Start of a multiline comment
-				else if (line[i] == '%')
-					in_multiline_comment = TRUE;
-				// Don't copy any white space over.
-				else if (line[i] != '\n' && line[i] != ' ' && line[i] != '\r' && line[i] != '\t' )
-					fputc(line[i], destination);
-			}
-			else
-			{
-				// If we're in a multi-line comment, search for the %
[…]

### 2019-06-27 ODIN_II/SRC/ast_elaborate.cpp
Odin: fix compile warning, fallthrough statement

@@ -314,11 +314,12 @@ void create_enode(ast_node_t *node)
 				case DIVIDE:
 					s->type.operation = '/';
 					s->priority = 1;
-				break;
+					break;
 
 				default:
-				break;
+					break;
 			}
+			break;
 		}
 
 		case IDENTIFIERS:
@@ -330,10 +331,9 @@ void create_enode(ast_node_t *node)
 		}
 
 		default:
+		{
 			break;
-		
-
-
+		}
 	}
 
 	p->next = s;

### 2019-06-27 ODIN_II/SRC/ast_elaborate.cpp
Odin: fix compile warning, fallthrough statement

@@ -277,11 +277,11 @@ void record_tree_info(ast_node_t *node)
  *-------------------------------------------------------------------------*/
 void create_enode(ast_node_t *node)
 {
-	enode *s;
-	s = (enode*)vtr::malloc(sizeof(enode));
+	enode *s = (enode*)vtr::malloc(sizeof(enode));
 	s->flag = -1;
 	s->priority = -1;
 	s->id = node->unique_count;
+	
 	switch (node->type)
 	{
 		case NUMBERS:
@@ -297,31 +297,36 @@ void create_enode(ast_node_t *node)
 			switch(node->types.operation.op)
 			{
 				case ADD:
+				{
 					s->type.operation = '+';
 					s->priority = 2;
-				break;
-
+					break;
+				}
 				case MINUS:
+				{
 					s->type.operation = '-';
 					s->priority = 2;
-				break;
-
+					break;
+				}
 				case MULTIPLY:
+				{
 					s->type.operation = '*';
 					s->priority = 1;
 					break;
-
+				}
 				case DIVIDE:
+				{
 					s->type.operation = '/';
 					s->priority = 1;
 					break;
-
+				}
 				default:
+				{
 					break;
+				}
 			}
 			break;
 		}
-
 		case IDENTIFIERS:
 		{
 			s->type.variable = node->types.identifier;
@@ -329,7 +334,6 @@ void create_enode(ast_node_t *node)
 			s->priority = 0;
 			break;
 		}
-
 		default:
 		{
 			break;

### 2019-06-27 ODIN_II/SRC/ast_util.cpp
null check for constant folding operation, and make unary_fold consistant with binary_fold

@@ -850,7 +850,7 @@ ast_node_t *resolve_node(STRING_CACHE *local_param_table_sc, char *module_name, 
 			break;
 
 			case UNARY_OPERATION:
-				newNode = fold_unary(node);
+				newNode = fold_unary(&node);
 				break;
 
 			case BINARY_OPERATION:
@@ -1089,13 +1089,13 @@ static void check_binary_operation(ast_node_t **node)
  * we currently make it as small as possible(+1) such that d'00001 becomes (in bin) 01
  * so it would become 10, is that what is defined?
  *-------------------------------------------------------------------------------------------*/
-ast_node_t *fold_unary(ast_node_t *node)
+ast_node_t *fold_unary(ast_node_t **node)
 {
-	if(!node)
+	if(!node || !(*node))
 		return NULL;
 		
-	operation_list op_id = node->types.operation.op;
-	ast_node_t *child_0 = node->children[0];
+	operation_list op_id = (*node)->types.operation.op;
+	ast_node_t *child_0 = (*node)->children[0];
 
 	if(node_is_constant(child_0))
 	{
@@ -1151,7 +1151,7 @@ ast_node_t *fold_unary(ast_node_t *node)
 
 			case CLOG2:
 				if(voperand_0.size() > ODIN_STD_BITWIDTH)
-					warning_message(PARSE_ERROR, child_0->line_number, child_0->file_number, "argument is %ld-bits but ODIN limit is %lu-bits \n",voperand_0.size(),ODIN_STD_BITWIDTH);
+					warning_message(PARSE_ERROR, (*node)->line_number, (*node)->file_number, "argument is %ld-bits but ODIN limit is %lu-bits \n",voperand_0.size(),ODIN_STD_BITWIDTH);
 
 				vresult = VNumber(clog2(voperand_0.get_value(), voperand_0.size()));
 				success = TRUE;
@@ -1163,14 +1163,14 @@ ast_node_t *fold_unary(ast_node_t *node)
 
 		if(success)
 		{
-			ast_node_t *new_num = create_tree_node_number(vresult, child_0->line_number, child_0->file_number);
+			ast_node_t *new_num = create_tree_node_number(vresult, (*node)->line_number, (*node)->file_number);
 			return new_num;		
 		}
 	}
 	else if (op_id == CLOG2)
 	{
 		/* $clog2() argument must be a constant expression */
[…]

### 2019-06-27 doc/src/vtr/vtr_task_fig.tex

\documentclass[tikz,border=10pt]{standalone}
\usepackage{tikz}
\usepackage{makecell}

\usetikzlibrary{arrows}
\usetikzlibrary{circuits.logic.US}
\usetikzlibrary{matrix}
\usetikzlibrary{positioning}
\usetikzlibrary{backgrounds}
\usetikzlibrary{decorations.pathreplacing, positioning, shapes.geometric, shapes.misc, shadows, fit}
\usetikzlibrary{shapes.geometric, shapes.symbols, arrows, shadows, fit, backgrounds, 3d, plotmarks, calc, intersections, matrix, patterns}

\tikzstyle{thin_arrow} = [draw,->,>=stealth]
\tikzstyle{arrow} = [thick,thin_arrow]
\begin{document}

\begin{tikzpicture}[]
    \tikzstyle{task} = [draw, rectangle, rounded corners=1.5mm, top color=red!30, bottom color=red!60, minimum width=1cm, minimum height=0.5cm]
    \tikzstyle{run} = [draw, rectangle, rounded corners=1.5mm, top color=blue!10, bottom color=blue!40, minimum width=1cm, minimum height=0.5cm]
    \tikzstyle{run_result} = [draw=blue!60, rectangle, rounded corners=1.5mm]
    \tikzstyle{symlink} = [draw, rectangle, rounded corners=1.5mm, top color=blue!5, bottom color=blue!10,]

    \newcommand{\levelspace}{0.75cm}
    \newcommand{\nodespace}{0.5cm}
    \node[task](task) {\textless task\_name\textgreater };

    \node[task, below=\levelspace of task, xshift=-2cm](config) {config};
    \node[run, right=\nodespace of config](run001) {run001};
    \node[run, right=\nodespace of run001](run002) {run002};
    \node[run, right=\nodespace of run002](run003) {run003};
    \node[symlink, right=\nodespace of run003](latest) {latest};

    \node[task, below=\levelspace of config, xshift=-0.75cm](config_file) {config.txt};
    \node[run, right=\nodespace of config_file](arch1) {\textless arch1\textgreater };
    \node[run, right=\nodespace of arch1](arch2) {\textless arch2\textgreater };
    \node[run, right=\nodespace of arch2](archX) {...};

    \node[run, below=\levelspace of arch1, xshift=-2.5cm](cct1) {\textless circuit1\textgreater };
    \node[run, right=\nodespace of cct1](cct2) {\textless circuit2\textgreater };
    \node[run, right=\nodespace of cct2](cctX) {...};

    \node[run, below=\levelspace of cct1, xshift=-0.75cm](params1) {\textless params1\textgreater };
    \node[run, right=\nodespace of params1](params2) {\textless params2\textgreater };
    \node[run, right=\nodespace of params2](paramsX) {...};

    \node[run_result, below=\levelspace of params1] (result1) {\makecell{odin.out \\ abc.out \\ vpr.out}};
    \node[run_result, below=\levelspace of params2] (result2) {\makecell{odin.out \\ abc.out \\ vpr.out}};

    \draw[arrow] (task) -- (config);
    \draw[arrow] (task) -- (run001);
    \draw[arrow] (task) -- (run002);
    \draw[arrow] (task) -- (run003);
    \draw[arrow] (task) -- (latest);

    \draw[arrow, gray] (latest) -- (run003);

    \draw[arrow] (config) -- (config_file);
    \draw[arrow] (run001) -- (arch1);
    \draw[arrow] (run001) -- (arch2);
    \draw[arrow] (run001) -- (archX);

    \draw[arrow] (arch1) -- (cct1);
    \draw[arrow] (arch1) -- (cct2);
    \draw[arrow] (arch1) -- (cctX);

    \draw[arrow] (cct1) -- (params1);
    \draw[arrow] (cct1) -- (params2);
    \draw[arrow] (cct1) -- (paramsX);

    \draw[arrow] (params1) -- (result1);
    \draw[arrow] (params2) -- (result2);

\end{tikzpicture}
\end{document}

### 2019-06-27 ODIN_II/SRC/ast_util.cpp
initial conditional evaluation

@@ -852,6 +852,11 @@ ast_node_t *resolve_node(STRING_CACHE *local_param_table_sc, char *module_name, 
 				newNode = fold_binary(&node);
 				break;
 
+			case IF_Q:
+			{
+				newNode = fold_conditional(&node);
+				break;
+			}
 			default:
 				break;
 		}
@@ -1327,6 +1332,47 @@ ast_node_t * fold_binary(ast_node_t **node)
 		oassert(!((*node)->type == BINARY_OPERATION && (*node)->types.operation.op == POWER));
 	}
 	return NULL;
+}
+
+/*---------------------------------------------------------------------------------------------
+ * (function: calculate_ternary)
+ *-------------------------------------------------------------------------------------------*/
+ast_node_t * fold_conditional(ast_node_t **node)
+{
+	if(!node || !(*node))
+		return NULL;
+
+	operation_list op_id = (*node)->types.operation.op;
+	ast_node_t *to_return = NULL;
+
+	switch (op_id)
+	{
+		case IF_Q:
+		{
+			ast_node_t *child_condition = (*node)->children[0];
+			if(node_is_constant(child_condition))
+			{
+				VNumber condition = *(child_condition->types.vnumber);
+
+				if(V_TRUE(condition))
+				{
+					to_return = ast_node_deep_copy((*node)->children[1]);
+				}
+				else if(V_FALSE(condition))
+				{
+					to_return = ast_node_deep_copy((*node)->children[2]);
+				}
+				// otherwise we keep it as is to build the circuitry
+			}
+			break;
+		}
+		default:
+		{
+			break;
+		}
+	}
+
+	return to_return;
 }
 
 /*---------------------------------------------------------------------------------------------

### 2019-06-27 ODIN_II/SRC/include/ast_util.h

@@ -39,4 +39,6 @@ ast_node_t *node_is_constant(ast_node_t *node);
 ast_node_t * fold_binary(ast_node_t **node);
 ast_node_t *fold_unary(ast_node_t *node);
 
+ast_node_t * fold_conditional(ast_node_t **node);
+
 long clog2(long value_in, int length);

### 2024-07-16 doc/src/vtr/vtr_flow_fig.tex

@@
         \node (arch) [device] {FPGA \\ Architecture}; & & \node (hdl) [design] {HDL}; & \\
-        & \node (odin_synth) [process] {Odin II}; & \node (quartus_map) [process] {Quartus \\ Map}; & \node (yosys_frontend) [] {Yosys}; \\
+        & \node (parmys_synth) [process] {Parmys}; & \node (quartus_map) [process] {Quartus \\ Map}; & \node (yosys_frontend) [] {Yosys}; \\
         & & \node (vqm2blif) [process] {VQM to BLIF}; & \node (abc_yosys) [process_no_shadow] {ABC}; \\
@@
 
-    \draw [arrow] (arch) |- (odin_synth);
+    \draw [arrow] (arch) |- (parmys_synth);
     \draw [arrow] (arch) |- (vpr_pack.west);
@@
 
-    \draw [arrow] (hdl) -- (odin_synth.north);
+    \draw [arrow] (hdl) -- (parmys_synth.north);
     \draw [arrow] (hdl) -- (quartus_map);
     \draw [arrow] (hdl) -- (yosys.north);
-    \draw[arrow] (odin_synth) -- (abc_synth);
+    \draw[arrow] (parmys_synth) -- (abc_synth);
     \draw[arrow] (quartus_map) -- (vqm2blif);
@@
         \draw[vtr_callout]
-            ([xshift=-\vtroffset,yshift=\vtroffset]odin_synth.north west)
+            ([xshift=-\vtroffset,yshift=\vtroffset]parmys_synth.north west)
          -- ([xshift=-\vtroffset,yshift=-\vtroffset]vpr_analysis.south west)
@@
          -- ([xshift=\vtroffset,yshift=\vtroffset]abc_synth.south east)
-         -- ([xshift=\vtroffset,yshift=\vtroffset]odin_synth.north east)
+         -- ([xshift=\vtroffset,yshift=\vtroffset]parmys_synth.north east)
          -- cycle;

### 2024-07-16 doc/src/vtr/vtr_task_fig.tex

@@
 
-    \node[run_result, below=\levelspace of params1] (result1) {\makecell{odin.out \\ abc.out \\ vpr.out}};
-    \node[run_result, below=\levelspace of params2] (result2) {\makecell{odin.out \\ abc.out \\ vpr.out}};
+    \node[run_result, below=\levelspace of params1] (result1) {\makecell{parmys.out \\ abc.out \\ vpr.out}};
+    \node[run_result, below=\levelspace of params2] (result2) {\makecell{parmys.out \\ abc.out \\ vpr.out}};
 
